### 2021-05-28 2020-05-17_proposal-unexplained-heterogeneity.rmd
first project draft

---
title: "Project proposal: Explanations for unexplained heterogeneity in meta-analysis"
author: "Anton Olsson Collentine"
date: '2021-05-17'
output: word_document
---

-	When studies fail to replicate heterogeneity is often raised as a potential explanation. 
-	This explanation is supported by most meta-analyses in psychology reporting large heterogeneity in the studied effect (Stanley, van Erp). Even in medicine, where heterogeneity estimates tend to be smaller (Ioannidis), the majority of meta-analyses appear to find heterogeneity (between 50-75%; Viechtbauer referencing Field; Higgins; Lipsey).
-	Some authors have argued that heterogeneity is inevitable in meta-analysis (Higgins), an argument which finds support in recent research demonstrating that substantial heterogeneity between studies can arise even from apparently minor differences in data processing (multiverse papers),  
-	Moreover, most preregistered multi-lab projects in psychology, where data collection and processing is identical across studies, seem to find little or no heterogeneity (Olsson-Collentine et al.)
-	It seems like in most situations (the exception being multi-lab projects or meta-analyses with very strict inclusion criteria) we should expect meta-analyses to have heterogeneity 
-	However, heterogeneity estimates in meta-analysis are almost invariably discussed as representing differences in true effect sizes between included studies (e.g., refs)
-	That is, arising from moderator effects
-	However, heterogeneity estimates in meta-analysis may also arise from statistical artefacts.
-	These may derive from, for example, the estimation procedure (von Hippel), differential measurement error  (Olsson-Collentine et al), publication bias (Augusteijn), and even arbitrary decisions in the research process (multiverse papers)
-	Fundamentally, anything that affects effect size estimates may also affect statistical heterogeneity estimates.
-	As such, researchers should make a concerted effort to rule out artefactual explanations of statistical heterogeneity before examining moderators
-	In this paper, we describe several explanations for unexplained heterogeneity in meta-analysis, examine their general effect on heterogeneity estimates and how to rule them out. 
-	We first provide a short definition of heterogeneity and how it is quantified.

#### What is heterogeneity 
-	It may be useful to distinguish between clinical heterogeneity (differences in contextual factors between studies) and statistical heterogeneity (a statistical estimate of the sensitivity to differences in contextual factors between studies).  
-	A statistical definition of heterogeneity may make it more clear why ‘heterogeneity’ cannot be assumed to directly represent the differences in true effects between included studies.
-	We can statistically define heterogeneity as: the residual variance between effect sizes after taking into account sampling error. 
-	Mention something about estimation, but probably don’t go into details (provide references).
-	Mention different estimators (I2, Q, Tau2), but discuss them further when talking about explanations

## Methods

#### Potential empirical aspects to add
a)	Strictness of inclusions medicine vs psychology
    - How many papers do meta-analysts exclude from their initial search?
      - Check PRISMA flowcharts in meta-analyses.
    - And/or compare the number of inclusion criteria
b)	Literature review of different QRPs that affect heterogeneity (“poor methodological quality”/ “small study effects”)
c)	Moderators (% explained variance vs. power)
    - Perhaps we can just do a smallish review, e.g., look at Robbie’s meta-meta data?
d) Check how often I2 used over tau (might already be some paper who has done this)
e) Check how often sample sizes in primary studies are large enought to lead to substantial I2 heterogeneity
f) Check how often reliability is available in primary studies (or refer to e.g., Flake et al)
g)	Use the meta-multiverse data to demonstrate the effect of arbitrary decisions on heterogeneity

## Explanations for heterogeneity
-	Here we first list potential artefactual explanations and how they might be examined. We then shortly discuss examining potential moderators of true heterogeneity, but primarily refer the reader to the several excellent papers available on this topic.

#### Explanation 1: estimation artefacts
a) I2 -> heterogeneity increases with primary study sample sizes
    - Rule out by using tau instead
b) Central chi-square distribution (Q->I2/tau) can lead to heterogeneity (von Hippel)
    - If heterogeneity is large enough, not a sufficient explanation = can rule out

#### Explanation 2: measurement artefacts
-	Differential measurement error
    - Can examine how large the effect would be on estimates if we know reliability in all primary studies? = rule out as sufficient explanation if heterogeneity too large
-	Range restrictions? (Schmidt & Hunter, 2015; cited in Olsson-Collentine et al.)
-	“poor methodological quality”

#### Explanation 3: arbitrary data processing and model decisions
-	Multiverse papers
    - Can never rule out as explanation (maybe with access to raw data?)

#### Explanation 4: QRPs
-	Publication bias (Hilde)
    - Can try to rule out with inclusion of grey literature, publication bias tests
-	Selective reporting (including p-hacking) ((“small study bias”)
    - Different forms of p-hacking? 
    - Can rule out only by checking preregistration

#### Explanation 5: systematic variance (moderators)
-	Short explanation

## Summary & Conclusion
-	There are many potential explanations for why the purportedly same true effect size differs between studies and results in heterogeneity in meta-analysis.
-	Before examining moderators as an explanation, researchers should make a concerted effort to rule out artefactual explanations such as those discussed in this paper.

### 2021-06-01 2020-05-17_proposal-unexplained-heterogeneity.rmd
Proposal v2

@@
 ---
-title: "Project proposal: Explanations for unexplained heterogeneity in meta-analysis"
+title: "Project proposal: Explanations for unexplained effect size heterogeneity"
 author: "Anton Olsson Collentine"
-date: '2021-05-17'
+date: '2021-06-01'
 output: word_document
@@
 
--	When studies fail to replicate heterogeneity is often raised as a potential explanation. 
--	This explanation is supported by most meta-analyses in psychology reporting large heterogeneity in the studied effect (Stanley, van Erp).
-Even in medicine, where heterogeneity estimates tend to be smaller (Ioannidis), the majority of meta-analyses appear to find heterogeneity (between 50-75%; Viechtbauer referencing Field; Higgins; Lipsey).
--	Some authors have argued that heterogeneity is inevitable in meta-analysis (Higgins), an argument which finds support in recent research demonstrating that substantial heterogeneity between studies can arise even from apparently minor differences in data processing (multiverse papers),
--	Moreover, most preregistered multi-lab projects in psychology, where data collection and processing is identical across studies, seem to find little or no heterogeneity (Olsson-Collentine et al.)
--	It seems like in most situations (the exception being multi-lab projects or meta-analyses with very strict inclusion criteria) we should expect meta-analyses to have heterogeneity 
--	However, heterogeneity estimates in meta-analysis are almost invariably discussed as representing differences in true effect sizes between included studies (e.g., refs)
--	That is, arising from moderator effects
--	However, heterogeneity estimates in meta-analysis may also arise from statistical artefacts.
--	These may derive from, for example, the estimation procedure (von Hippel), differential measurement error  (Olsson-Collentine et al), publication bias (Augusteijn), and even arbitrary decisions in the research process (multiverse papers)
--	Fundamentally, anything that affects effect size estimates may also affect statistical heterogeneity estimates.
--	As such, researchers should make a concerted effort to rule out artefactual explanations of statistical heterogeneity before examining moderators
--	In this paper, we describe several explanations for unexplained heterogeneity in meta-analysis, examine their general effect on heterogeneity estimates and how to rule them out. 
--	We first provide a short definition of heterogeneity and how it is quantified.
+## Short summary
 
-#### What is heterogeneity 
--	It may be useful to distinguish between clinical heterogeneity (differences in contextual factors between studies) and statistical heterogeneity (a statistical estimate of the sensitivity to differences in contextual factors between studies).
--	A statistical definition of heterogeneity may make it more clear why ‘heterogeneity’ cannot be assumed to directly represent the differences in true effects between included studies.
--	We can statistically define heterogeneity as: the residual variance between effect sizes after taking into account sampling error. 
--	Mention something about estimation, but probably don’t go into details (provide references).
--	Mention different estimators (I2, Q, Tau2), but discuss them further when talking about explanations
+-	Theoretical (mainly) paper consisting of introduction + 7 explanations + empirical section + conclusion
+-	Sections are:
+    - Introduction (1-2 pages)
+        - Content: Comparable studies can have different outcome (heterogeneity), we present 7 explanations, argue that artefactual explanations should be examined before moderators, and summarize the content of the paper.
+    1) Model and sampling error (1.5 – 2 pages)
+        - Present meta-analytic heterogeneity model
+        - Explain that it accounts for sampling error
+        - Discuss estimation artefact that can increase heterogeneity estimate (I2 vs. tau, central chi-square distribution of Q)
+    2) Multiverse effect (2 pages)
+        - Discuss based on multi-analyst papers
+        - Illustrate using data from meta-multiverse project
+        - Highlight that even though multiverse does not affect average estimates in absence of selective reporting, it can affect heterogeneity estimates
+    3) Publication bias (max 1 page)
+        - Literature review: Hilde’s paper + others
+    4) QRPs (minimum 2 pages with illustrations)
+        - should usually decreases heterogeneity, assuming people select for significance
+        - Literature review on different types of QRPs (talk to Esther)
+    5) “Incompetence”/ “poor methodological quality” (1/3 of a page)
+        - Common critique when replications ‘fail’
+        - This critique is not detailed enough as a critique
+        - If an “incompetent” researcher uses a less effective intervention, they should be seen as a design moderator (but must first be better specified)
+        - Cochrane recommends talking about ‘risk of bias’ rather than ‘poor methodological quality’
+    6) Measurement artefacts (1 – 2 pages)
+        - Differential measurement error leads to heterogeneity (Flake et al., examined availability of reliability estimates in primary studies)
+        - Range restrictions? Other measurement factors? (see Schmidt & Hunter, 2015)
+    7) Moderators (2 pages)
+        - Short discussion, extensively described in other papers
+        - Types of moderators (contextual factors): settings, sample population, treatment and measurement variables
+        - Meta-regression, power, IPD
+    - Methods & results (2  pages)
+        - Description of empirical data collection
+        - Discuss findings in the context of the section(s) 1- 7 that they affect
+    - Conclusion (1/2 page)
+-	Total: ~15 pages
 
-## Methods
+## Detailed summary
+Below I present a detailed summary of
 
-#### Potential empirical aspects to add
-a)	Strictness of inclusions medicine vs psychology
-    - How many papers do meta-analysts exclude from their initial search?
-      - Check PRISMA flowcharts in meta-analyses.
-    - And/or compare the number of inclusion criteria
-b)	Literature review of different QRPs that affect heterogeneity (“poor methodological quality”/ “small study effects”)
-c)	Moderators (% explained variance vs. power)
-    - Perhaps we can just do a smallish review, e.g., look at Robbie’s meta-meta data?
-d) Check how often I2 used over tau (might already be some paper who has done this)
-e) Check how often sample sizes in primary studies are large enought to lead to substantial I2 heterogeneity
-f) Check how often reliability is available in primary studies (or refer to e.g., Flake et al)
-g)	Use the meta-multiverse data to demonstrate the effect of arbitrary decisions on heterogeneity
+a) the introduction: for a better description of the proposed angle of the paper. 
+b) Methods & results section: my proposals for the empirical data collection. Needs to be discussed to set the scope of the paper.
 
-## Explanations for heterogeneity
--	Here we first list potential artefactual explanations and how they might be examined.
-We then shortly discuss examining potential moderators of true heterogeneity, but primarily refer the reader to the several excellent papers available on this topic.
+### Introduction
 
-#### Explanation 1: estimation artefacts
-a) I2 -> heterogeneity increases with primary study sample sizes
-    - Rule out by using tau instead
-b) Central chi-square distribution (Q->I2/tau) can lead to heterogeneity (von Hippel)
-    - If heterogeneity is large enough, not a sufficient explanation = can rule out
+**Comparable studies always have some design heterogeneity, whether they differ in the studies’ settings, their sample populations, the measures they use, or their treatment variables (e.g., Campbell & Stanley, 2015).** In turn, effect size estimates from comparable studies always have some statistical heterogeneity, one potential explanation of which is the studies’ design heterogeneity (i.e., the effect size estimates arising from different true effects).
+Other explanations for differences in statistical estimates from comparable studies include sampling error, multiverse variability, publication bias, Questionable Research Practices such as selective reporting (including p-hacking), ‘poor methodological quality’ or supposed incompetence of experimenters, and measurement artefacts.
 
-#### Explanation 2: measurement artefacts
--	Differential measurement error
-    - Can examine how large the effect would be on estimates if we know reliability in all primary studies? = rule out as sufficient explanation if heterogeneity too large
--	Range restrictions? (Schmidt & Hunter, 2015; cited in Olsson-Collentine et al.)
--	“poor methodological quality”
+**It can be challenging to disentangle the reasons for why effect size estimates differ.** From a theoretical perspective we are most interested in finding out to what extent these differences arise from design heterogeneity, to which end other explanations are nuisances or artefacts we must first control for.
+Perhaps this explains why heterogeneity estimates in meta-analysis are almost invariably discussed as representing differences in true effect sizes between included studies, ignoring the many alternative explanations.
 
-#### Explanation 3: arbitrary data processing and model decisions
--	Multiverse papers
-    - Can never rule out as explanation (maybe with access to raw data?)
+**Generally, when comparable studies show different results, there is an undue focus on design heterogeneity as an explanation for these differences.** Design heterogeneity is often one of the first explanations for any ‘failure’ to replicate a study (by the reader’s preferred definition), of which the ‘replication crisis’ in psychology and other domains continue to provide many examples.
+However, artefactual explanations for statistical heterogeneity by necessity should be partitioned out first before considering design heterogeneity as an explanation.
+In addition, design heterogeneity is often an unlikely explanation for statistical heterogeneity in at least in cognitive and social psychology (Olsson-Collentine et al, 2020).
+Fundamentally, anything that affects effect size estimates may also affect statistical heterogeneity estimates.As such, researchers should make a concerted effort to rule out artefactual explanations of statistical heterogeneity before examining design heterogeneity as an explanation.
 
-#### Explanation 4: QRPs
--	Publication bias (Hilde)
-    - Can try to rule out with inclusion of grey literature, publication bias tests
--	Selective reporting (including p-hacking) ((“small study bias”)
-    - Different forms of p-hacking? 
-    - Can rule out only by checking preregistration
+**The statistical results from comparable studies and their heterogeneity are usually examined through meta-analysis, and this is also the tool we use throughout this paper to discuss explanations for statistical heterogeneity.** Examining statistical heterogeneity and moderators thereof is usually considered one of the primary purposes of meta-analysis.
+To enhance the discussion of statistical heterogeneity , we will also collect empirical data relating to heterogeneity estimates in meta-analysis.
+We first describe and discuss all 7 explanations for statistical heterogeneity, then present a short methods section and discuss relevant sections in the context of our findings, before ending the paper with a summary and conclusion.
 
-#### Explanation 5: systematic variance (moderators)
--	Short explanation
+### Methods & Results
+- This section has a large impact on the scope of the project and needs discussion on what we should actually do. 
+- I first describe 3 research questions we might try to answer by collecting data, and then make different suggestions to limit the scope.
 
-## Summary & Conclusion
--	There are many potential explanations for why the purportedly same true effect size differs between studies and results in heterogeneity in meta-analysis.
--	Before examining moderators as an explanation, researchers should make a concerted effort to rule out artefactual explanations such as those discussed in this paper.
+#### Research questions and empirical data collection:
+  1) To what extent does meta-analysts’ usage of I2 instead of tau inflate heterogeneity estimates?
+        - Extract I2 and harmonic mean
+  2) To what extent does the smaller average number of studies in medical meta-analyses (compared to in psychology) represent stricter inclusion criteria? (as opposed to smaller population of studies)
+        - Extract proportion excluded studies from flowcharts
+        - Extract number of exclusion criteria 
+        - Extract type of exclusion criteria (settings, sample population, measurement or treatment variables, risk of bias)
+  3) To what extent and which type of moderators explains statistical heterogeneity in meta-analyses?
+        - Do the meta-analyses examine moderators? {yes/no}
+        - What % of statistical heterogeneity do moderators explain?
+        - How many moderators are examined, and what type? 
+            - 	Settings, sample population, measurement and treatment variables, risk of bias
+            - 	What % does each type explain?
+        - What is the power of the meta-analysis to find significant moderators?
+
+#### Scoping (suggestions for different levels of empirical data collection)
+- The moderator part above might be an entire paper by itself and we may not want to do everything. Below 5 suggestions for limiting the scope of data collection in this paper. Other suggestions?
+- My inclination would be towards suggestion 3/4 below, which would leave most of the analysis of moderators in meta-analysis to be part of a future project.
+
+Suggestions:
+
+  1) Extract:
+        - I2 + harmonic mean
+        - Proportion excluded studies (medicine vs. psychology)
+  2) Extract:
+        - I2 + harmonic mean
+        - Proportion excluded studies (medicine vs. psychology)
+        - PLUS: Number and type of exclusion criteria
+  3) Extract:
+        - Proportion excluded studies (medicine vs. psychology)
+        - Number and type of exclusion criteria
+        - PLUS: How many moderators examined and what type [e.g., psychology might have fewer exclusion criteria, but more moderators]
+  4) Extract:
+        - Proportion excluded studies (medicine vs. psychology)
+        - Number and type of exclusion criteria
+        - How many moderators examined and what type [e.g., psychology might have fewer exclusion criteria, but more moderators]
+        - PLUS: I2 + harmonic mean
+  5) Extract:
+        - Everything, including all the moderator stuff
+
+#### Sample 
+What would the sample look like?
+
+  -	A random sample of meta-analyses from medicine and psychology
+      - Already collected sample(s) from some project? Our inclusion/exclusion criteria? Sample size?

### 2023-02-21 proposal-unexplained-heterogeneity.rmd
added strucutred simulations

--- 2021-06-01 2020-05-17_proposal-unexplained-heterogeneity.rmd
@@
 
--	Theoretical (mainly) paper consisting of introduction + 7 explanations + empirical section + conclusion
+-	Theoretical (mainly) paper consisting of introduction + empirical data collection + 7 explanations + empirical illustration + conclusion
 -	Sections are:
@@
         - Content: Comparable studies can have different outcome (heterogeneity), we present 7 explanations, argue that artefactual explanations should be examined before moderators, and summarize the content of the paper.
+    - Empirical data collection (1 - 2 pages)
+        - RQ: How much do heterogeneity do moderators explain?
+        - we will extract Heterogeneity estimates before and after moderators in a sample of meta-analyses
+            - Sample: Perhaps van Erp's dataset, or set up search terms
+            - Extract: Q/I2/tau before/after moderators applied
     1) Model and sampling error (1.5 – 2 pages)
@@
         - Meta-regression, power, IPD
-    - Methods & results (2  pages)
-        - Description of empirical data collection
-        - Discuss findings in the context of the section(s) 1- 7 that they affect
+    - Empirical illustration (2-3 pages)
+        - Use a large meta-analysis to illustrate/examine all 7 points
+        - Good choice is Chris' cyberball meta-analysis (see final sentence)
+        - Preferably a smaller, homogenous subset from it
     - Conclusion (1/2 page)
--	Total: ~15 pages
+-	Total: ~18 pages
 
@@
 a) the introduction: for a better description of the proposed angle of the paper.
-b) Methods & results section: my proposals for the empirical data collection. Needs to be discussed to set the scope of the paper.
 
@@
 In addition, design heterogeneity is often an unlikely explanation for statistical heterogeneity in at least in cognitive and social psychology (Olsson-Collentine et al, 2020).
-Fundamentally, anything that affects effect size estimates may also affect statistical heterogeneity estimates.As such, researchers should make a concerted effort to rule out artefactual explanations of statistical heterogeneity before examining design heterogeneity as an explanation.
+Fundamentally, anything that affects effect size estimates may also affect statistical heterogeneity estimates.
+As such, researchers should make a concerted effort to rule out artefactual explanations of statistical heterogeneity before examining design heterogeneity as an explanation.
 
-**The statistical results from comparable studies and their heterogeneity are usually examined through meta-analysis, and this is also the tool we use throughout this paper to discuss explanations for statistical heterogeneity.** Examining statistical heterogeneity and moderators thereof is usually considered one of the primary purposes of meta-analysis.
-To enhance the discussion of statistical heterogeneity , we will also collect empirical data relating to heterogeneity estimates in meta-analysis.
-We first describe and discuss all 7 explanations for statistical heterogeneity, then present a short methods section and discuss relevant sections in the context of our findings, before ending the paper with a summary and conclusion.
-
-### Methods & Results
-- This section has a large impact on the scope of the project and needs discussion on what we should actually do. 
-- I first describe 3 research questions we might try to answer by collecting data, and then make different suggestions to limit the scope.
-
-#### Research questions and empirical data collection:
-  1) To what extent does meta-analysts’ usage of I2 instead of tau inflate heterogeneity estimates?
-        - Extract I2 and harmonic mean
-  2) To what extent does the smaller average number of studies in medical meta-analyses (compared to in psychology) represent stricter inclusion criteria? (as opposed to smaller population of studies)
-        - Extract proportion excluded studies from flowcharts
-        - Extract number of exclusion criteria 
-        - Extract type of exclusion criteria (settings, sample population, measurement or treatment variables, risk of bias)
-  3) To what extent and which type of moderators explains statistical heterogeneity in meta-analyses?
-        - Do the meta-analyses examine moderators? {yes/no}
-        - What % of statistical heterogeneity do moderators explain?
-        - How many moderators are examined, and what type? 
-            - 	Settings, sample population, measurement and treatment variables, risk of bias
-            - 	What % does each type explain?
-        - What is the power of the meta-analysis to find significant moderators?
-
-#### Scoping (suggestions for different levels of empirical data collection)
-- The moderator part above might be an entire paper by itself and we may not want to do everything. Below 5 suggestions for limiting the scope of data collection in this paper. Other suggestions?
-- My inclination would be towards suggestion 3/4 below, which would leave most of the analysis of moderators in meta-analysis to be part of a future project.
-
-Suggestions:
-
-  1) Extract:
-        - I2 + harmonic mean
-        - Proportion excluded studies (medicine vs. psychology)
-  2) Extract:
-        - I2 + harmonic mean
-        - Proportion excluded studies (medicine vs. psychology)
-        - PLUS: Number and type of exclusion criteria
-  3) Extract:
-        - Proportion excluded studies (medicine vs. psychology)
-        - Number and type of exclusion criteria
-        - PLUS: How many moderators examined and what type [e.g., psychology might have fewer exclusion criteria, but more moderators]
-  4) Extract:
-        - Proportion excluded studies (medicine vs. psychology)
-        - Number and type of exclusion criteria
-        - How many moderators examined and what type [e.g., psychology might have fewer exclusion criteria, but more moderators]
-        - PLUS: I2 + harmonic mean
-  5) Extract:
-        - Everything, including all the moderator stuff
-
-#### Sample 
-What would the sample look like?
-
-  -	A random sample of meta-analyses from medicine and psychology
-      - Already collected sample(s) from some project? Our inclusion/exclusion criteria? Sample size?
+**The statistical results from comparable studies and their heterogeneity are most usefully examined through meta-analysis.** Examining statistical heterogeneity and moderators thereof is usually considered one of the primary purposes of meta-analysis.
+We begin this paper with first collecting data on how much heterogeneity meta-analyses in psychology tend to attribute to moderators.
+We then describe and discuss all 7 explanations for statistical heterogeneity, followed by an illustration of these explanations in a large meta-analysis, before ending the paper with a summary and conclusion.

### 2023-02-21 simulate-reliability.r

@@ -1,0 +1,874 @@
+# Project: Unexplained heterogeneity
+# Code purpose: simulate the effect of reliability on heterogeneity
+# Author: Anton Olsson-Collentine (anton@olssoncollentine.com)
+
+#********************************************************************
+# fixed effect
+#********************************************************************
+
+# We use the classical test theory framework for this
+
+# Start with a single study
+# Step 1: simulate a true score
+
+# Latent score, translated into scale score?
+# Or start with a binary score?
+# Can also simulate multiple binary scores that sum to a scale score
+# but then why not simulate the scale score directly?
+
+#for reliability we must have more than one item..
+
+
+#variance is prob * (1 - prob) = 0.25 (maximal) for each item
+
+# Reliability is the proportion true score variance of the observed variance. Measurement error by definition has a mean of zero.
+
+# Let's first imagine we are measuring some unrestrained variable
+true_var <- 1
+true_mean <- 0
+n  <- 10
+error_var <- 0.5
+
+
+true <- rnorm(n = n, mean = true_mean, sd = sqrt(true_var))
+e <- rnorm(n = n, mean = 0, sd = sqrt(error_var))
+x <- true + e
+
+var_x <- true_var + error_var
+reliability <- true_var / var_x
+
+# We would probably prefer to go from reliability to measurement error
+#Assuming true var = 1 and
+reliability <- 0.8
+
+var_x <- true_var / reliability
+#and
+error_var <- var_x - true_var
+
+calc_error_var <- function(true_var, reliability){
+    var_x <- true_var / reliability
+    error_var <- var_x - true_var
+    data.frame(true_var, reliability, var_x, error_var)
+}
+
+reliability <- seq(from = 0, to = 1, by = 0.1)
+
+reference_values <- calc_error_var(true_var, reliability)
+
+#So for 0.8 reliability we have 0.25 measurement var and our observed data:
+true <- rnorm(n = n, mean = true_mean, sd = sqrt(true_var))
+e <- rnorm(n = n, mean = 0, sd = sqrt(error_var))
+x <- true + e
[…]

### 2023-02-21 supercomputer_simulator.r

@@ -1,0 +1,267 @@
+# Project: unexplained heterogeneity
+# Script purpose: run simulations on supercomputer
+# code: Anton Olsson-Collentine
+
+
+
+#****************************************
+# libraries and functions
+#****************************************
+library(parallel) #comes pre-installed with R
+
+simulate_rma <- function(k, sample_size, true_tau2, mu, reliability_min, reliability_max){
+ # Draw rho from mean rho
+    rho <- rnorm(n = k, mean = mu, sd = sqrt(true_tau2))
+    sampling_var_rho  <- (1-rho^2)^2 / (sample_size -1)
+
+    # given study rho, draw sample rho
+    r_se <- rnorm(k, mean = rho, sd = sqrt(sampling_var_rho))
+
+    #Compute observed r given measurement error
+    reliabilities <- runif(k, min = reliability_min, max = reliability_max) #draw reliabilities to avoid some strange draw
+    r_se_me <- r_se*sqrt(reliabilities)^2
+
+    r_var_estimate <- (1-r_se_me^2)^2 / (sample_size -1) #observed estimate
+
+    fit <- rma(yi = r_se_me, vi = r_var_estimate)
+
+    data.frame(intercept_b = fit$b,
+               intercept_p = fit$pval,
+               tau2_hat = fit$tau2,
+               tau2_p = fit$QEp)
+}
+
+#****************************************
+# 1) Mu and varying reliability, to show lack of effect
+#****************************************
+sample_size <- 150
+k <- 20
+true_tau2 <- 0.1
+mu <- seq(from = 0, to = 0.9, by = 0.1)
+reliability_min <- 0
+reliability_max <- 1
+reps <- 1e4
+
+cond_mu <- expand.grid(sample_size = sample_size,
+                    k = k,
+                    true_tau2 = true_tau2,
+                    mu = mu,
+                    reliability_min = reliability_min,
+                    reliability_max = reliability_max)
+
+
+reliability_min <- 0.5
+reliability_max <- 0.5
+
+cond_mu2 <- expand.grid(sample_size = sample_size,
+                    k = k,
+                    true_tau2 = true_tau2,
+                    mu = mu,
+                    reliability_min = reliability_min,
+                    reliability_max = reliability_max)
+
+
+
+mu  <- 0.3
+reliability_min <- seq(from = 0.5, to = 0, by = -0.1)
+reliability_max <- seq(from = 0.5, to = 1, by = 0.1)
+
[…]

### 2023-03-13 supplemental_data_max_tau2/empirical_tau2_values.r
estimating max reasonable tau2

@@ -1,0 +1,34 @@
+# Project: Unexplained heterogeneity
+# Script purpose: estimate common tau2 values for correlations in psychology
+# based on van Erp et al. (2017): 10.5334/jopd.33
+# Code: Anton Olsson-Collentine
+
+#**********************************************
+# Setup
+#**********************************************
+# Data downloaded from https://osf.io/p6ecw
+
+dat <- read.csv("supplemental_data_I2/Data 1990-2013 with tau values.csv")
+
+unique(dat$Type.of.ES)
+#also contains corrected Pearson's r
+
+dat_r <- dat[dat$Type.of.ES == "Pearson's r",]
+
+hist(dat_r$tau.2, breaks = 30)
+boxplot(dat_r$tau.2)
+# We see that 50% of tau2 values are between about ~0.02- 0.06,
+#with another 25% between 0.06- 0.125 and 25% below 0.02
+quantile(dat_r$tau.2)
+#Actually, the 25th quantile is 0.01, median 0.03, and the 75th is 0.0576
+#the maximum value is 0.2704, which gives tau = 0.52
+
+
+# For corrected pearson's rs
+dat_r_corrected <- dat[dat$Type.of.ES == "Corrected Pearson's r",]
+hist(dat_r_corrected$tau.2, breaks = 30)
+boxplot(dat_r_corrected$tau.2)
+quantile(dat_r_corrected$tau.2)
+# The corrected rs have a lower distribution
+# 25th quantiles = 0.001, median = 0.01, 75th = 0.0225,
+#The maximum is 0.2209

### 2023-03-13 supplemental_data_max_tau2/simulate-given-k-n.r

@@ -1,0 +1,104 @@
+# Project: unexplained heterogeneity
+# script purpose: Simulate tau2 values corresponding to an I2 of 90 - 95% for sample size 150 and number of studies equal to 20
+# code: Anton Olsson-Collentine
+
+#****************************************
+# functions and packages
+#****************************************
+
+simulate_rma_I2 <- function(effect_type = c("r", "r_z"), k, sample_size, true_tau2, mu){
+    #adjusted input and output compared to my primary function
+    #Also, only for perfect reliability, i.e., uses sampling_var_rho to estimate meta-analysis
+
+ # Draw rho from mean rho
+    rho <- rnorm(n = k, mean = mu, sd = sqrt(true_tau2))
+
+    if(effect_type == "r") {
+        sampling_var_rho  <- (1-rho^2)^2 / (sample_size -1)
+
+    }else if(effect_type == "r_z") { #fisher's z
+
+        sampling_var_rho  <- 1 / (sample_size - 3)
+    }else{
+        stop("specify effect type")
+    }
+
+    # given study rho, draw sample rho
+    r_se <- rnorm(k, mean = rho, sd = sqrt(sampling_var_rho))
+
+    fit <- rma(yi = r_se, vi = sampling_var_rho)
+
+    data.frame(I2 = fit$I2,
+               true_tau2 = true_tau2,
+               tau2_hat = fit$tau2,
+               tau2_p = fit$QEp)
+}
+
+library(parallel)
+#library(metafor) #is loaded through the simulation setup below
+
+#****************************************
+# run simulation
+#****************************************
+
+#NB! sample size and K
+sample_size <- 150
+k <- 20
+true_tau2 <- seq(from = 0.05, to = 0.2, by = 0.001)
+#Initial experimentation suggests 90% I2 is around 0.8 and 95% around 0.15
+#EDIT: for Fisher's z the above tau2 values are ok, for Pearson's r they need to be lower, about 0.03 to 0.1
+mu <- 0.3
+reps <- 1e3
+
+cond <- expand.grid(sample_size = sample_size,
+                    k = k,
+                    true_tau2 = true_tau2,
+                    mu = mu)
+
+
+out_list <- vector("list", length = nrow(cond))
+
+
+
+start <- Sys.time()
+
+ncores <- detectCores()
+cl <- makePSOCKcluster(ncores) # Create cluster based on nworkers. This because mclapply doesn't work on windows
+
[…]

### 2023-03-13 code/examining_REML_underestimate_perfect_reliability.r
cleaned up code

@@ -1,0 +1,151 @@
+# Project: unexplained heterogeneity
+# Script purpose: Explore the underestimate of tau2 by REML for large tau2 values
+
+
+#****************************************
+# Replicating figure 2 in Langan et al. (2019)
+#****************************************
+# langan: a comparison of heterogeneity variance estimators in simulated random-effects
+#https://onlinelibrary.wiley.com/doi/10.1002/jrsm.1316
+
+
+simulate_rma_langan <- function(version = c("original", "efficient"), k = 20, true_tau2 = 0.0444, mu = 0.5, sample_size = 40){ #see section 3.1
+
+    mu_i <- rnorm(n = k, mean = mu, sd = sqrt(true_tau2))
+    n_c = n_t = rep(sample_size, 20)
+
+    if(version == "original"){
+    n1 <- replicate(k, rnorm(n = sample_size, mean = 0, sd = 1), simplify = FALSE)
+    n2 <- lapply(1:k, function(i) rnorm(n = sample_size, mean = mu_i[i], sd = 1))
+
+    avg_c <- lapply(n1, mean) #function below from heterogeneity project #1
+    var_c <- lapply(n1, var)
+    avg_c <- do.call(rbind, avg_c)
+    var_c <- do.call(rbind, var_c)
+
+    avg_t <- lapply(n2, mean)
+    var_t <- lapply(n2, var)
+    avg_t <- do.call(rbind, avg_t)
+    var_t <- do.call(rbind, var_t)
+
+    fit <- rma(measure = "SMD",  m1i = avg_t, m2i = avg_c, sd1i = sqrt(var_t),
+                   sd2i = sqrt(var_c), n1i = n_t, n2i = n_c, method = "REML") #fit meta-analysis transforming into standardized mean difference (Hedge's g)
+
+    }else if(version == "efficient"){
+
+        avg_c <- rnorm(n = k, mean = 0, sd = 1 / sqrt(n_c)) #Draw means from sampling distribution control group
+        avg_t <- rnorm(n = k, mean = mu_i, sd = 1 / sqrt(n_t)) #Draw means from sampling distribution treatment group
+        var_c <- rchisq(n = k, df = n_c - 1) / (n_c - 1) #draw variances from sampling distribution control group
+        var_t <- rchisq(n = k, df = n_t - 1) / (n_t - 1) #draw variances from sampling distribution treatment group
+
+        fit <- rma(measure = "SMD",  m1i = avg_t, m2i = avg_c, sd1i = sqrt(var_t),
[…]

### 2023-03-13 code/supercomputer_simulator.r

@@ -8,6 +8,7 @@
 # libraries and functions
 #****************************************
 library(parallel) #comes pre-installed with R
+#library(metafor) #loaded when setting up simulations
 
 simulate_rma <- function(effect_type = c("r", "r_z"), k, sample_size, true_tau2, mu, reliability_min, reliability_max){
  # Draw rho from mean rho
@@ -40,52 +41,27 @@ simulate_rma <- function(effect_type = c("r", "r_z"), k, sample_size, true_tau2,
                tau2_p = fit$QEp)
 }
 
+
 #****************************************
-# 1) Mu and varying reliability, to show lack of effect
+# Plot 2) as reliability decreases the underestimate becomes worse
 #****************************************
+#Reliability is fixed in each study here
+
 sample_size <- 150
 k <- 20
-true_tau2 <- 0.1
-mu <- seq(from = 0, to = 0.9, by = 0.1)
-reliability_min <- 0
-reliability_max <- 1
-reps <- 1e4
+true_tau2 <- 0.078 #corresponds to 95% I2 according to my simulations for Pearson's r and the above sample size and k
+mu <- seq(from = 0, to = 0.6, by = 0.1)
+#Based on the 'upper median' (83.35% quantile) from Schäfer & Schwarz for non-preregistered studies.
+reliability_min <- seq(from = 0.1, to = 1, by = 0.1)
+reps <- 1e3
 
-cond_mu <- expand.grid(sample_size = sample_size,
+cond <- expand.grid(sample_size = sample_size,
                     k = k,
                     true_tau2 = true_tau2,
                     mu = mu,
-                    reliability_min = reliability_min,
-                    reliability_max = reliability_max)
+                    reliability_min = reliability_min)
 
-
-reliability_min <- 0.5
-reliability_max <- 0.5
-
-cond_mu2 <- expand.grid(sample_size = sample_size,
-                    k = k,
-                    true_tau2 = true_tau2,
-                    mu = mu,
-                    reliability_min = reliability_min,
-                    reliability_max = reliability_max)
-
-
-
-mu  <- 0.3
-reliability_min <- seq(from = 0.5, to = 0, by = -0.1)
-reliability_max <- seq(from = 0.5, to = 1, by = 0.1)
-
-cond_rel <- data.frame(sample_size = sample_size,
-                    k = k,
-                    true_tau2 = true_tau2,
-                    mu = mu,
[…]

### 2023-03-14 code/supercomputer_simulator.r
added plots code

@@ -10,8 +10,21 @@
 library(parallel) #comes pre-installed with R
 #library(metafor) #loaded when setting up simulations
 
-simulate_rma <- function(effect_type = c("r", "r_z"), k, sample_size, true_tau2, mu, reliability_min, reliability_max){
- # Draw rho from mean rho
+simulate_rma <- function(
+    effect_type = c("r", "r_z"),
+    reliability_distribution = c("uniform", "normal"),
+    k,
+    sample_size,
+    true_tau2,
+    mu,
+    reliability_min, #for uniform distribution
+    reliability_max,
+    reliability_mean = NULL, #for normal distribution
+    reliability_sd = NULL,
+    steplength = 1, #these are for controlling the fisher algorithm in rma
+    maxiter = 100){ #these are the default values in the function: https://www.metafor-project.org/doku.php/tips:convergence_problems_rma
+
+    # Draw rho from mean rho
     rho <- rnorm(n = k, mean = mu, sd = sqrt(true_tau2))
 
     if(effect_type == "r") {
@@ -27,13 +40,24 @@ simulate_rma <- function(effect_type = c("r", "r_z"), k, sample_size, true_tau2,
     # given study rho, draw sample rho
     r_se <- rnorm(k, mean = rho, sd = sqrt(sampling_var_rho))
 
+    #draw new reliabilities for each run to avoid some strange draw
+    if(reliability_distribution == "uniform"){
+    reliabilities <- runif(k, min = reliability_min, max = reliability_max)
+    } else if(reliability_distribution == "normal"){
+        reliabilities <- rnorm(k, mean = reliability_mean, sd = reliability_sd)
+        while(max(reliabilities > 1)){
+            print(max(reliabilities))
+        reliabilities <- rnorm(k, mean = reliability_mean, sd = reliability_sd)
+        }
+    } #note that in the second case we stop reliabilities from superceding 1, inducing some bias in the range of reliabilities
+
     #Compute observed r given measurement error
-    reliabilities <- runif(k, min = reliability_min, max = reliability_max) #draw reliabilities to avoid some strange draw
     r_se_me <- r_se*sqrt(reliabilities)^2
 
     r_var_estimate <- (1-r_se_me^2)^2 / (sample_size -1) #observed estimate
 
-    fit <- rma(yi = r_se_me, vi = r_var_estimate)
+    fit <- rma(yi = r_se_me, vi = r_var_estimate,
+              control=list(stepadj=steplength, maxiter=maxiter))
 
     data.frame(intercept_b = fit$b,
[…]

### 2023-03-14 code/plots.r
cleaned plot script

@@ -1,0 +1,190 @@
+# Project: Unexplained heterogeneity
+# Code purpose: Creating plots from simulated data
+# Author: Anton Olsson-Collentine (anton@olssoncollentine.com)
+
+
+#****************************************
+# Simulating data for plots
+#****************************************
+# see supercomputer_simulator.r
+
+
+#****************************************
+# plotting
+#****************************************
+
+library(ggplot2)
+
+
+# Plot perfect reliability
+
+dat2 <- readRDS("data/tau2_increasing_rel.RDS")
+
+dat2 <- lapply(dat2, function(x) as.data.frame(as.list(x)))
+
+dat_tau2 <- data.table::rbindlist(dat2, idcol = "condition")
+
+dat_tau2$true_tau2 <- gsub("&.*", "", dat_tau2$condition)
+dat_tau2$true_tau2 <- as.numeric(gsub("tau2 = ", "", dat_tau2$true_tau2))
+
+dat_tau2$reliability <- gsub(".* & reliability = ", "", dat_tau2$condition)
+dat_tau2 <- dat_tau2[dat_tau2$reliability == 1,]
+
+dat3 <- readRDS("data/tau2_increasing_fisherz.RDS")
+
+dat3 <- lapply(dat3, function(x) as.data.frame(as.list(x)))
+
+dat2_tau2 <- data.table::rbindlist(dat3, idcol = "condition")
+
+dat2_tau2$true_tau2 <- gsub("&.*", "", dat2_tau2$condition)
+dat2_tau2$true_tau2 <- as.numeric(gsub("tau2 = ", "", dat2_tau2$true_tau2))
+
+dat2_tau2$reliability <- gsub(".* & reliability = ", "", dat2_tau2$condition)
+
+dat4 <- readRDS("data/tau2_increasing_r_sampling_var_rho.RDS")
+
+dat4 <- lapply(dat4, function(x) as.data.frame(as.list(x)))
+
+dat3_tau2 <- data.table::rbindlist(dat4, idcol = "condition")
+
+dat3_tau2$true_tau2 <- gsub("&.*", "", dat3_tau2$condition)
+dat3_tau2$true_tau2 <- as.numeric(gsub("tau2 = ", "", dat3_tau2$true_tau2))
+
+dat3_tau2$reliability <- gsub(".* & reliability = ", "", dat3_tau2$condition)
+
[…]

### 2023-03-14 code/simulate-reliability.r

@@ -1,939 +1,0 @@
-# Project: Unexplained heterogeneity
-# Code purpose: simulate the effect of reliability on heterogeneity
-# Author: Anton Olsson-Collentine (anton@olssoncollentine.com)
-
-#********************************************************************
-# fixed effect
-#********************************************************************
-
-# We use the classical test theory framework for this
-
-# Start with a single study
-# Step 1: simulate a true score
-
-# Latent score, translated into scale score?
-# Or start with a binary score?
-# Can also simulate multiple binary scores that sum to a scale score
-# but then why not simulate the scale score directly?
-
-#for reliability we must have more than one item..
-
-
-#variance is prob * (1 - prob) = 0.25 (maximal) for each item
-
-# Reliability is the proportion true score variance of the observed variance. Measurement error by definition has a mean of zero.
-
-# Let's first imagine we are measuring some unrestrained variable
-true_var <- 1
-true_mean <- 0
-n  <- 10
-error_var <- 0.5
-
-
-true <- rnorm(n = n, mean = true_mean, sd = sqrt(true_var))
-e <- rnorm(n = n, mean = 0, sd = sqrt(error_var))
-x <- true + e
-
-var_x <- true_var + error_var
-reliability <- true_var / var_x
-
-# We would probably prefer to go from reliability to measurement error
-#Assuming true var = 1 and
-reliability <- 0.8
-
-var_x <- true_var / reliability
-#and
-error_var <- var_x - true_var
-
-calc_error_var <- function(true_var, reliability){
-    var_x <- true_var / reliability
-    error_var <- var_x - true_var
-    data.frame(true_var, reliability, var_x, error_var)
-}
-
-reliability <- seq(from = 0, to = 1, by = 0.1)
-
-reference_values <- calc_error_var(true_var, reliability)
-
-#So for 0.8 reliability we have 0.25 measurement var and our observed data:
-true <- rnorm(n = n, mean = true_mean, sd = sqrt(true_var))
-e <- rnorm(n = n, mean = 0, sd = sqrt(error_var))
-x <- true + e
[…]

### 2023-03-14 code/plots.r
plot code figure 4

@@ -188,3 +188,26 @@ ggsave("figures/plot3.png", width = 8.62, height = 9.93)
 
 
 # Plot 4
+dat <- readRDS("data/over_vs_underestimate.RDS")
+dat <- lapply(dat, function(x) as.data.frame(as.list(x)))
+
+dat <- data.table::rbindlist(dat, idcol = "mu")
+
+dat$reliability_sd <- gsub(".*& ", "", dat$mu)
+dat$reliability_sd <- gsub(".*= ", "", dat$reliability_sd)
+dat$mu <- gsub("&.*", "", dat$mu)
+dat$mu <- as.numeric(gsub("mu = ", "", dat$mu))
+
+line_labels <- dat[dat$mu == 0.6,]
+
+ggplot(dat,aes(x = mu, y = tau2_hat)) +
+geom_line(aes(linetype = reliability_sd), show.legend = FALSE) +
+#scale_linetype_manual(values = 10:1) +
+geom_text(data = line_labels, aes(x = mu, y = tau2_hat, group = reliability_sd), nudge_x = 0.025,
+label = line_labels$reliability_sd) +
+#annotate("text", x = 0.885, y = 0.086, label = "Reliability", hjust = 0) +
+geom_hline(yintercept = 0.00078, linetype = "dashed") +
+scale_x_continuous(breaks = seq(from = 0, to = 1, by = 0.2)) +
+theme_bw()
+
+ggsave("figures/plot4.png", width = 8.62, height = 9.93)

### 2023-03-14 manuscript/narrative.rmd
added skeleton manuscript

---
title: "Narrative"
author: "Anton Olsson-Collentne"
date: '2023-03-14'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.width = 6, fig.asp = 0.75)
```

In Olsson-Collentine et al., (2019) we demonstrate that variation in reliability between the studies included in a meta-analysis will lead to an overestimate of statistical heterogeneity. This point is also made by Hunter and Schmidt (year) whose claims are repeated in a recent paper published by Perspectives on Psychological Science (REF). However, in this article we demonstrate that this overestimate only holds in the absence of true heterogeneity (i.e., under a fixed effect meta-analytic framework). In the presence of true heterogeneity, imperfect reliability in the studies instead leads to an underestimate of the true heterogeneity.

## Method and results
Troughout this article we use sample size = 150 participants in all studies, and k = 20 studies in each meta-analysis. For simplicity we also use Pearson's correlations for the simulations. In Figure 1 we demonstrate the overrestimate that might be expected in the absence of true heterogeneity, across a range a empirical reliability estimates from Flake et al., (2017) and Sanchez-Meca et al (YEAR). Both reported an average alpha close to .8 and with an SD of .13 (Flake et al) or that ranged up to .13 (Sanche-Meca et al.). In Figure 1, we draw reliabilities for the K studies from a normal distribution centered on 0.8 and with varying SDs.

```{r plot1}
knitr::include_graphics("../figures/plot1.png")
```

In the presence of true heterogeneity, we underestimate it when study reliabilities are imperfect. The effect of sample size is minimal. We demonstrate this relationship for tau2 = 0.078, which corresponds to and I2 of 95% given our sample size, number of studies and effect size type. We show this across a range of empirically relevant effect sizes based on Schäfer & Schwarz (2019), with the upper limit corresponding to the 'upper median' (83.35% quantile) of non-preregistered effect sizes in their study.

```{r plot2}
knitr::include_graphics("../figures/plot2.png")
```

As the true heterogeneity increases, the underestimate due to imperfect reliability of included studies becomes worse in an absolute sense. Nonetheless, as Figure 3 demonstrates, this relationship is linear and the proportional underestimate stays stable regardless of true heterogeneity. Figure 2 demonstrated that effect size largely does not matter for the underestimate of heterogeneity due to imperfect reliabilities; here we have used an average effect size (r = 0.24) corresponding to the median found amongst correlational intelligence studies by Nuijten et al. (YEAR).

```{r plot3}
knitr::include_graphics("../figures/plot3.png")
```

Finally, figure 4 demonstrates that even under conditions generous to the overrestimate due to varying reliability (i.e., large effect size and large variance in reliabilities) the overrestimate due to these factors is not able to compensate for the underrestimate of heterogeneity due to imperfect reliability. In sum, imperfect reliability in studies included in meta-analysis will tend to lead to an underestimate of true heterogeneity. Nonetheless, this underestimate is quite small under realistic conditions and is unlikely to be of large concern.

### 2023-04-26 code/plots.r
plot updates

@@ -36,27 +36,27 @@ geom_line(aes(x = mu, y = sqrt(tau2_hat), color = reliability_sd))
 #ggsave("figures/plot1.png", width = 8.62, height = 9.93)
 
 
-# Plot 2
+# Plot 2 original
 
-line_labels <- dat[dat$r == "0.6",]
+# line_labels <- dat[dat$r == "0.6",]
 
-ggplot(dat,aes(x = r, y = percent_tau2)) +
-geom_line(aes(linetype = reliability), show.legend = FALSE) +
-scale_linetype_manual(values = 8:1) +
-geom_label(data = line_labels, aes(x = r, y = percent_tau2, group = reliability), nudge_x = 0.025,
-label = line_labels$reliability) +
-#annotate("text", x = 0.885, y = 0.086, label = "Reliability", hjust = 0) +
-geom_hline(yintercept = 100, linetype = "dashed") +
-theme_bw()
+# ggplot(dat,aes(x = r, y = percent_tau2)) +
+# geom_line(aes(linetype = reliability), show.legend = FALSE) +
+# scale_linetype_manual(values = 8:1) +
+# geom_label(data = line_labels, aes(x = r, y = percent_tau2, group = reliability), nudge_x = 0.025,
+# label = line_labels$reliability) +
+# #annotate("text", x = 0.885, y = 0.086, label = "Reliability", hjust = 0) +
+# geom_hline(yintercept = 100, linetype = "dashed") +
+# theme_bw()
 
-ggsave("figures/plot2.png", width = 8.62, height = 9.93)
+# ggsave("figures/plot2.png", width = 8.62, height = 9.93)
 
 ## Plot 2 v2
 # That is, with perfect reliability and tau2-values of
 # c(0.002, 0.0035, 0.0055, 0.0085, 0.012, 0.0185, 0.031, 0.069)
 # (i.e, 20 - 90% I2)
 
-dat <- readRDS("data/reliability_1.RDS")
+dat <- readRDS("data/k_20_over_vs_underestimate.RDS")
 dat <- lapply(dat, function(x) as.data.frame(as.list(x)))
 
 dat <- data.table::rbindlist(dat, idcol = "mu")
@@ -66,8 +66,13 @@ conditions <- unlist(strsplit(dat$mu, split = ";"))
 
[…]

### 2023-04-26 code/supercomputer_simulator.r

@@ -17,16 +17,16 @@ source("code/functions.r") #for rnorm_truncated and simulate_rma
 #****************************************
 sample_size <- 150
 k <- 20
-true_tau2 <- c(0.002, 0.0035, 0.0055, 0.0085, 0.012, 0.0185, 0.031, 0.069)
+true_tau2 <- c(0, 0.002, 0.0035, 0.0055, 0.0085, 0.012, 0.0185, 0.031, 0.069)
 #Using Pearson's r as the effect size. Correspond to I2 20 - 90%.
 # Below corresponding values for k = 5
 # k <- 5
-#true_tau2 <- c(0.0015, 0.0045, 0.008, 0.013, 0.0195, 0.0315, 0.0550, 0.1245)
+#true_tau2 <- c(0, 0.0015, 0.0045, 0.008, 0.013, 0.0195, 0.0315, 0.0550, 0.1245)
 mu <- seq(from = 0, to = 0.6, by = 0.1) #
 
 #Based on Flake et al., average alpha was .79, SD = .13, range .17 - .87
 #Based on Sanchez-Meca, mean across 5 meta-analysis was 0.767 - 0.891 and SD ranged between 0.034 - 0.133
-reliability_mean <- c(0.8, 0.9, 1)
+reliability_mean <- c(0.6, 0.7, 0.8, 0.9, 1)
 reliability_sd <- seq(from = 0, to = 0.15, by = 0.05)
 
 

### 2023-04-26 code/truncated_tau2.r

@@ -1,8 +1,8 @@
 
-
+source("./code/functions.r")
 es <- seq(0, 0.6, 0.1)
-vars <- c(0.002, 0.0035, 0.0055, 0.0085, 0.012, 0.0185, 0.031, 0.069)
-
+vars <- c(0, 0.002, 0.0035, 0.0055, 0.0085, 0.012, 0.0185, 0.031, 0.069)
+#vars <- c(0, 0.0015, 0.0045, 0.008, 0.013, 0.0195, 0.0315, 0.0550, 0.1245)
 a <- lapply(vars, function(x){
 
 lapply(es, function(y) {
@@ -15,4 +15,4 @@ data.frame(tau2 = var(res), mu = y, nominal_tau2 = x)
 
 b <- lapply(a, function(x) do.call(rbind, x))
 b <- do.call(rbind, b)
-saveRDS(b, "data/truncated_tau2.RDS")
+saveRDS(b, "data/k_5_truncated_tau2.RDS")

### 2023-04-26 manuscript/methods.rmd
renamed folder + added supplemental analyses

---
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Methods

All code and data for this project are available at (OSF LINK). For our simulations and analyses we used R (REF) and we used the R package 'metafor' (REF) to perform meta-analyses and estimate heterogeneity, For parallel computing we took advantage of the R packages 'parallel' (REF) and 'parabar' (REF).

## Simulation study design

Figure 1 provides an overview of the design of this simulation study. Our design was broadly the same for product-moment correlations and Fisher's $z$, with some minor differences we highlight in the text below detailing the procedure.

```{r Figure 1, echo=FALSE, out.width= "100%"}
knitr::include_graphics("../figures/flowchart-simulation-design.png")
```

(FIGURE 1: a flowchart summarizing the simulation design)

When generating data for meta-analysis we proceeded as follows. First,  we sampled $i = 1, ..., k$ true study effect sizes $\theta_i$ from $Ntrunc(\theta, \tau^2)$. That is, a normal distribution with a mean of $\theta$ and a variance of $\tau^2$ that we truncated at $\{-1, 1\}$ to stay within the bounds of the product-moment correlation. For Fisher' $z$ no truncation was applied.

Practically, we implemented the truncation as follows. First, we sampled $k$ probabilities from an uniform distribution $U(a, b)$ where $a$ and $b$ are the values given by the cumulative distribution function (CDF) of the normal distribution $N(\theta, \tau^2)$ at $\{-1, 1\}$. Second, we then back-transformed the $k$ sampled values onto the $N(\theta, \tau^2)$ distribution using the inverse of the CDF.

For each of the $k$ sampled true study effect sizes $\theta_i$ we then sampled one observed effect size $\gamma_i$ from a truncated normal distribution $Ntrunc(\theta_i, \sigma_i^2)$, using the same truncation procedure. The sampling variance $\sigma_i^2$ for each true effect size $\theta_i$ we computed using the standard formula for Pearson's $r$ $\sigma_i^2 =  \frac{(1 - \theta_i^2)^2}{N_i - 1}$ where $N_i$ is the total sample size for study $i = 1, ..., k$. For Fisher's $z$ no truncation was applied and we approximated the sampling variance as $\sigma^2_i = \frac{1}{N_i - 3}$. At this point we have $k$ sampled effect sizes $\gamma_i$ in Pearson's $r$ or Fisher's $z$ without measurement error.

To add measurement error we first sampled a reliability $R_i$ for each study from a truncated ($\{0, 1\}$) normal distribution $Ntrunc(\bar{R}, \delta)$, given some average reliability across studies $\bar{R}$ and standard deviation in reliability $\delta$. We assumed that both the dependent and independent variable within a study were measured with the same reliability such that $R_i = R_{iXX'} = R_{iYY'}$. We then computed observed effect sizes $r_i$ for each study given the attenuation formula $r_{i} = \gamma_i \times \sqrt{R_{XX'}} \times \sqrt{R_{YY'}}$. Because we assumed equal reliability in $X$ and $Y$ this simplifies to $r_i = \gamma_i \times R_i$.   Fisher's $z$ we transformed to product-moment correlations before adding measurement error and then back-transformed into Fisher's $z$ before the next step. At this point we have $k$ effect sizes $r_i$ in Pearson's $r$ or Fisher's $z$ with sampling error and measurement error, ready for meta-analysis.

We meta-analyzed the $k$ effect sizes using a random effects model with inverse variance weights and the Restricted Maximum Likelihood (REML) estimator of heterogeneity. We choose the REML estimator because it is a heterogeneity estimator that performs well and is widely implemented (e.g., REF, REF2), but also show our results using the classic DerSimonian-Laird (DL) estimator in supplement X. Similarly, inverse variance weights are widely used, perform similarly to other common weights (e.g., REF1 REF2) and we see no reason other implementations would change our results substantially. The procedure described in this section was replicated 10,000 times for each combination of parameter values.

## Parameter values

We vary the grand mean $\theta$ from 0 to 0.6 in steps of 0.1 to cover all realistic effect sizes and explore the boundaries of the the interaction between effect size and reliability variance. For Fisher's $z$ we transformed these value into equivalent effect sizes at the Fisher's $z$ scale. To put these values into context we turn to the empirical estimates of typical correlational effect sizes provided by Nuijten et al., (REF) and Schäfer & Scharz (REF). Nuijten et al. performed a meta-meta-analysis on the fields of RESEARCH and reports a median meta-analytic effect size of 0.24 for predictive validity and correlational studies. This estimate is likely positively biased due to publication bias and selective reporting. Schäfer & Scharz report a median $r$ of 0.16 amongst preregistered research in FIELD, with an 'upper median' (i.e., the 83.35% quantile) of 0.41. This estimate is likely less affected by positive bias but should not be expected to be unbiased (REF Scheel).

We defined the between-studies variance ($\tau^2$) to cover a wide range of variance in true effect sizes and empirical estimates reported in the psychological literature. Van Erp et al., (REF) provide empirical heterogeneity estimates from a large number of meta-analyses. The median Pearson's $r$ $\tau^2$ value in these data 
was 0.03 (interquartile range 0.01 - 0.0576). As the point of this study is that empirical heterogeneity estimates may not be accurately estimated we selected a range of $\tau^2$ estimates for the product-moment that we considered reasonable and that also covered most of the empirical estimates from van Erp et al. These were {0, ..., VALUES}.

Fisher's z is measured at a different scale and it is not possible to directly convert $\tau^2$-values expressed in Pearson's $r$ into corresponding $\tau^2$-values on the Fisher's $z$ scale. To be able to compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale. The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity (REF Higgins), an as such is measured from 0 - 100%. The $I^2$ index can be defined as $I^2 = \frac{\tau^2}{s^2 + \tau^2}$ where $s^2 = \frac{\Sigma w_i(k - 1)}{(\Sigma w_i)^2 - \Sigma w_i^2}$ and $w_i$ is the precision $w_i = 1 / \sigma^2$. This means that its value generally depends on both the number of studies $K$ and the sample size $N_i$ in these studies (REF: Higgins, I2 is not an absolute measure of heterogeneity). However, because we keep $N_i$ fixed across studies (see below), the value of $I^2$ depends only on $N = N_i$ and the heterogeneity. We thus compute the $I^2$ index for each sample size condition and $\tau^2$-values defined in Pearson's $r$, and then compute corresponding $\tau^2$-values on the Fisher's $z$ scale given these $I^2$-values and sample sizes.

We ran our analyses across a range of sample sizes $N$ and number of studies $K$. Within meta-analyses we used a fixed sample size across all studies such that $N = N_i$. We made this choice rather than treating sample size as a random variable to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance. We base our range of sample sizes on empirical estimates of typical sample size in correlational research in psychology by XXX (still needs lookup) and consider the following values {VALUES}.

When focusing on the bias of heterogeneity estimate, the value of $K$ would not matter for an unbiased estimator. However, both the heterogeneity estimator we use, restricted maximum likelihood, and most other heterogeneity estimators are truncated at zero (REF Viechtbauer). Lower values for $K$ then results in a larger proportion of truncated estimates below zero because of increased sampling variance. As such, the values of $K$ can affect bias, especially at lower levels of heterogeneity. In their review of meta-analyses in psychology, van Erp et al. (YEAR) find $K$ values ranging from XX to XX, and in our simulations we consider the following values {VALUES}.

Finally, we defined average reliability levels and their standard deviation based on empirical estimates from the literature. Flake et al. (YEAR) report that based on 245 estimates of Cronbach's alpha in psychology the average estimate was 0.79 with a standard deviation of 0.13. The interquartile range was approximately 0.68 - 0.87 for studies wherein the associated scale lacked a reference and 0.79 - 0.88 for scales that did have a reference. Sanchez-Meca et al. (REF) report the reliability estimates based on five 'reliability generalization studies'. These five studies summarized reliability in 25 - 51 primary studies (184 total), and the mean reliability ranged from 0.767 to 0.891 with standard deviations ranging between 0.034 - 0.133. Given some likely positive bias in such empirical values, we examined the following mean reliabilities $\{0.6, 0.7, 0.8, 0.9\}$, and standard deviations $\{0, 0.05, 0.1, 0.15\}$.

### 2023-04-26 manuscript/narrative.rmd
renamed folder + added supplemental analyses

@@
 ## Method and results
-Troughout this article we use sample size = 150 participants in all studies, and k = 20 studies in each meta-analysis.
-For simplicity we also use Pearson's correlations for the simulations.
-In Figure 1 we demonstrate the overrestimate that might be expected in the absence of true heterogeneity, across a range a empirical reliability estimates from Flake et al., (2017) and Sanchez-Meca et al (YEAR).
-Both reported an average alpha close to .8 and with an SD of .13 (Flake et al) or that ranged up to .13 (Sanche-Meca et al.).
-In Figure 1, we draw reliabilities for the K studies from a normal distribution centered on 0.8 and with varying SDs.
-
-```{r plot1}
-knitr::include_graphics("../figures/plot1.png")
-```
-
-In the presence of true heterogeneity, we underestimate it when study reliabilities are imperfect.
-The effect of sample size is minimal.
-We demonstrate this relationship for tau2 = 0.078, which corresponds to and I2 of 95% given our sample size, number of studies and effect size type.
-We show this across a range of empirically relevant effect sizes based on Schäfer & Schwarz (2019), with the upper limit corresponding to the 'upper median' (83.35% quantile) of non-preregistered effect sizes in their study.
-
-```{r plot2}
-knitr::include_graphics("../figures/plot2.png")
-```
-
-As the true heterogeneity increases, the underestimate due to imperfect reliability of included studies becomes worse in an absolute sense.
-Nonetheless, as Figure 3 demonstrates, this relationship is linear and the proportional underestimate stays stable regardless of true heterogeneity.
-Figure 2 demonstrated that effect size largely does not matter for the underestimate of heterogeneity due to imperfect reliabilities; here we have used an average effect size (r = 0.24) corresponding to the median found amongst correlational intelligence studies by Nuijten et al.
-(YEAR).
-
-```{r plot3}
-knitr::include_graphics("../figures/plot3.png")
-```
-
-Finally, figure 4 demonstrates that even under conditions generous to the overrestimate due to varying reliability (i.e., large effect size and large variance in reliabilities) the overrestimate due to these factors is not able to compensate for the underrestimate of heterogeneity due to imperfect reliability.
-In sum, imperfect reliability in studies included in meta-analysis will tend to lead to an underestimate of true heterogeneity.
-Nonetheless, this underestimate is quite small under realistic conditions and is unlikely to be of large concern.

### 2023-04-26 supplement_reliability_Maassen/extract_reliabilities-maassen.r

@@ -1,0 +1,32 @@
+# Project: Unexplained heterogeneity
+# Code purpose: estimate typical reliability and its SD from E. Maassen's data
+# Author: Anton Olsson-Collentine (anton@olssoncollentine.com)
+
+
+#******************************************
+library(openxlsx)
+
+
+dat <- openxlsx::read.xlsx("supplement_reliability_Maassen/codebook-main-step4_rel.xlsx")
+
+#From Esther:
+#Your columns of interest are rel_rep and rel_est.
+#(i.e., reported and estimated reliabilities)
+
+#Note that this reliability may mean nothing, given that it’s on all the items, and therefore one factor is assumed, even though there may be more factors (f_total) underlying the data.
+
+mean(dat$rel_rep)
+#[1] 0.8392653
+sd(dat$rel_rep)
+#[1] 0.0997923
+
+mean(dat$rel_est, na.rm = TRUE)
+#[1] 0.8657143
+sd(dat$rel_est, na.rm = TRUE)
+#[1] 0.07702407
+
+# Higher estimates and with less SD when recomputed. Likely a because of a correlation with who made their data openly available.
+
+
+dat$comp_rel <- dat$rel_rep - dat$rel_est
+mean(dat$comp_rel, na.rm = TRUE)

### 2023-04-26 supplemental_data/check-sample-sizes.r

@@ -1,0 +1,17 @@
+# Check sample sizes in https://journals.sagepub.com/doi/full/10.1177/2515245919838781
+
+
+
+library(haven)
+
+dat <- haven::read_sav("supplemental_data_max_tau2/data.sav")
+quantile(dat$Sample)
+#1300 rows
+
+dat2 <- dat[dat$Sample < 1e4,]
+#1291 rows
+quantile(dat2$Sample, na.rm = TRUE)
+
+dat3 <- dat2[dat2$Sample < 5e3,]
+#1290 rows
+quantile(dat3$Sample)

### 2023-04-26 supplemental_data/empirical_tau2_values.r

@@ -8,7 +8,7 @@
 #**********************************************
 # Data downloaded from https://osf.io/p6ecw
 
-dat <- read.csv("supplemental_data_I2/Data 1990-2013 with tau values.csv")
+dat <- read.csv("supplemental_data_max_tau2/Data 1990-2013 with tau values.csv")
 
 unique(dat$Type.of.ES)
 #also contains corrected Pearson's r
@@ -19,7 +19,7 @@ hist(dat_r$tau.2, breaks = 30)
 boxplot(dat_r$tau.2)
 # We see that 50% of tau2 values are between about ~0.02- 0.06,
 #with another 25% between 0.06- 0.125 and 25% below 0.02
-quantile(dat_r$tau.2)
+quantile(dat_r$tau)
 #Actually, the 25th quantile is 0.01, median 0.03, and the 75th is 0.0576
 #the maximum value is 0.2704, which gives tau = 0.52
 
@@ -28,7 +28,24 @@ quantile(dat_r$tau.2)
 dat_r_corrected <- dat[dat$Type.of.ES == "Corrected Pearson's r",]
 hist(dat_r_corrected$tau.2, breaks = 30)
 boxplot(dat_r_corrected$tau.2)
-quantile(dat_r_corrected$tau.2)
+quantile(dat_r_corrected$tau)
 # The corrected rs have a lower distribution
 # 25th quantiles = 0.001, median = 0.01, 75th = 0.0225,
 #The maximum is 0.2209
+
+quantile(dat_r_corrected$X..of.effect.sizes)
+dat_r_corrected2 <- dat_r_corrected[dat_r_corrected$X..of.effect.sizes < 1000,]
+quantile(dat_r_corrected2$X..of.effect.sizes)
+
+
+# Check K
+dat_r_any <- dat[grepl("Pearson", dat$Type.of.ES),]
+unique(dat_r_any$Type.of.ES)
+quantile(dat_r_any$X..of.effect.sizes)
+hist(dat_r_any$X..of.effect.sizes)
+
+quantile(dat_r$X..of.effect.sizes)
+hist(dat_r$X..of.effect.sizes)
+
+dat_r_any2 <- dat_r_any[dat_r_any$X..of.effect.sizes < 1000,]
+quantile(dat_r_any2$X..of.effect.sizes)

### 2023-04-26 code/supercomputer_simulator.r
added control over steplength to adjust for lower N not converging

@@ -15,9 +15,10 @@ source("code/functions.r") #for rnorm_truncated and simulate_rma
 #****************************************
 # Conditions
 #****************************************
-sample_size <- 150
-k <- 20
-true_tau2 <- c(0, 0.002, 0.0035, 0.0055, 0.0085, 0.012, 0.0185, 0.031, 0.069)
+sample_size <- c(50, 100, 150, 200)
+k <- c(5, 20, 40, 200)
+true_tau <- c(0, 0.1, 0.15, 0.2)
+true_tau2 <- true_tau^2
 #Using Pearson's r as the effect size. Correspond to I2 20 - 90%.
 # Below corresponding values for k = 5
 # k <- 5
@@ -26,8 +27,8 @@ mu <- seq(from = 0, to = 0.6, by = 0.1) #
 
 #Based on Flake et al., average alpha was .79, SD = .13, range .17 - .87
 #Based on Sanchez-Meca, mean across 5 meta-analysis was 0.767 - 0.891 and SD ranged between 0.034 - 0.133
-reliability_mean <- c(0.6, 0.7, 0.8, 0.9, 1)
-reliability_sd <- seq(from = 0, to = 0.15, by = 0.05)
+reliability_mean <- c(0.6, 0.7, 0.8, 0.9)
+reliability_sd <- seq(from = 0.15, to = 0.15, by = 0.05)
 
 
 cond <- expand.grid(sample_size = sample_size,
@@ -41,16 +42,18 @@ cond <- expand.grid(sample_size = sample_size,
 remove_cond <- cond[["reliability_mean"]] == 1 & cond[["reliability_sd"]] > 0
 cond <- cond[!remove_cond,]
 
-#these below are just appendices to makes sure the function runs as expected
+#these below are higher level control input to the function
 cond$reliability_max <- cond$reliability_mean
 cond$reliability_min <- cond$reliability_mean
 cond$effect_type <- "r"
 cond$reliability_distribution <- "normal"
+cond$step_length  <-  0.5 #decrease from 1 to 0.5 to improve convergence for low N
+cond$maxiterations  <-  100 #default values are steplength = 1, and maxiter = 100
 
 #****************************************
 # Setup parallel simulation
 #****************************************
-reps <- 1e4
+reps <- 1e2
 out_list <- vector("list", length = nrow(cond))
 
 ncores <- detectCores()
@@ -83,7 +86,9 @@ task <- function(iteration){ #anonymous function needed when using for replicati
                     reliability_min = cond_r$reliability_min,
[…]

### 2023-05-08 supplemental_data/compute_tau2.r
analytical computations of tau2 for fisher z

@@ -1,0 +1,89 @@
+#****************************************
+# Project: unexplained heterogeneity
+# script purpose: Compute I2 values for set tau values and sample sizes
+# for Pearson's r. Then use these I2 values and sample sizes to compute
+# corresponding tau values for Fisher's z.
+# code: Anton Olsson-Collentine
+#****************************************
+
+#****************************************
+# Background
+#****************************************
+
+# Given N and tau, compute I2
+#I2 = tau2 / s2 + tau2
+#s2 = sum(w_i)*(k - 1) / (sum(wi_i)^2) - sum(w_i^2)
+
+# Because of using Pearson's r, the sampling variance depends on the
+# effect size and N. Consequently, I2 depends on rho, N, tau
+# But because we have fixed sample size across studies, not on K
+
+# We only want one tau-value for each N and tau, not for each ES as well
+# Option 1: Compute I2 for ES = 0 and use that
+# Option 2: Compute I2 for all ES and take average/median.
+# I prefer option 1 as the simpler option, and also because I2 is not linear
+
+#****************************************
+# Functions
+#****************************************
+compute_s2 <- function(sigma2, k){
+
+    wi <- rep((1 / sigma2), k)
+
+    s2_num <- sum(wi)*(k - 1)
+    s2_denom <- (sum(wi)^2) - sum(wi^2)
+    s2_num / s2_denom #s2 out
+}
+
+compute_I2 <- function(vec = c(rho, tau, sample_size)){
+    # #rho = average true effect size
+    # #tau = SD in true effect sizes
+    # #sample_size = fixed within-study sample size
+
+    rho <- vec[["rho"]]
+    tau <- vec[["tau"]]
+    sample_size <- vec[["sample_size"]]
+
+    tau2 <- tau^2
+    k <- 20 #constant, random value, doesn't matter when N is fixed across studies
+
+    sigma2  <- (1-rho^2)^2 / (sample_size -1)
+    s2 <- compute_s2(sigma2, k)
+
+    tau2 / (s2 + tau2) #I2
+
+}
+
+compute_tau2_z <- function(vec = c(I2, sample_size)){
+    #I2 = I2 from pearson R
+    #sample_size = fixed within-study sample size
+    I2 <- vec[["I2"]]
+    sample_size <- vec[["sample_size"]]
[…]

### 2023-05-08 supplemental_data/compute_tau2.r
compute I2 analytically

@@ -26,6 +26,26 @@
 #****************************************
 # Functions
 #****************************************
+sampling_var <- function(rho, sample_size){
+    (1-rho^2)^2 / (sample_size -1)
+}
+
+# Apply the law of the unconscious statistician
+# to compute the expected value of the within study variance
+# so that we can account for tau
+# https://en.wikipedia.org/wiki/Law_of_the_unconscious_statistician
+
+internal <- function(x, rho, tau, sample_size){
+
+    g_x <- sampling_var(rho, sample_size)
+    f_x <- dnorm(x, mean = rho, sd = tau)
+
+    g_x * f_x #to be integrated over, see https://en.wikipedia.org/wiki/Law_of_the_unconscious_statistician
+}
+
+integrate(internal, lower = -0.99, upper = 0.99,
+           rho = rho, tau = tau, sample_size = sample_size)
+
 compute_s2 <- function(sigma2, k){
 
     wi <- rep((1 / sigma2), k)
@@ -47,8 +67,13 @@ compute_I2 <- function(vec = c(rho, tau, sample_size)){
     tau2 <- tau^2
     k <- 20 #constant, random value, doesn't matter when N is fixed across studies
 
-    sigma2  <- (1-rho^2)^2 / (sample_size -1)
+    sigma2 <- integrate(internal, #function g(x)f(x)
+                        lower = -0.99, upper = 0.99,
+                        rho = rho, tau = tau, sample_size = sample_size)
+    sigma2 <- sigma2[["value"]]
     s2 <- compute_s2(sigma2, k)
+    #this last step is superfluous, because all studies have the same weights
+    #I leave it in for clarity and in case the function is used otherwise
 
     tau2 / (s2 + tau2) #I2
 
@@ -66,6 +91,8 @@ compute_tau2_z <- function(vec = c(I2, sample_size)){
 
     I2*s2 / (1 - I2) #tau2
 }
+
+
 
 #****************************************
 # Computations

### 2023-05-08 supplemental_data/empirical_tau2_values.r

@@ -8,7 +8,7 @@
 #**********************************************
 # Data downloaded from https://osf.io/p6ecw
 
-dat <- read.csv("supplemental_data_max_tau2/Data 1990-2013 with tau values.csv")
+dat <- read.csv("supplemental_data/Data 1990-2013 with tau values.csv")
 
 unique(dat$Type.of.ES)
 #also contains corrected Pearson's r
@@ -49,3 +49,19 @@ hist(dat_r$X..of.effect.sizes)
 
 dat_r_any2 <- dat_r_any[dat_r_any$X..of.effect.sizes < 1000,]
 quantile(dat_r_any2$X..of.effect.sizes)
+
+# Check correction or not
+dat_r = dat[dat$Type.of.ES %in% c("Pearson's r", "Weighted Pearson's r", "Unweighted Pearson's r"),]
+unique(dat_r$Reference) #13
+
+dat_r_corrected = dat[dat$Type.of.ES %in% c("Corrected Pearson's r", "Weighted corrected Pearson's r"),]
+unique(dat_r_corrected$Reference) #13
+
+# Corrected generally
+dat_cor <- dat[dat$Type.of.ES %in% c("Corrected Pearson's r",
+       "Weighted corrected Pearson's r",
+       "Corrected Cohen's d"),]
+unique(dat_cor$Reference) #14
+
+length(unique(dat$Reference)) - length(unique(dat))
+#49 uncorrected and 14 corrected

### 2023-05-08 supplemental_data/simulate-given-k-n.r

@@ -1,76 +1,0 @@
-# Project: unexplained heterogeneity
-# script purpose: Simulate tau2 values corresponding to an I2 of 90 - 95% for sample size 150 and number of studies equal to 20
-# code: Anton Olsson-Collentine
-
-#****************************************
-# functions and packages
-#****************************************
-
-
-library(parallel)
-#library(metafor) #is loaded through the simulation setup below
-
-source("code/functions.r") #for rnorm_truncated and simulate_rma_I2
-#****************************************
-# run simulation
-#****************************************
-
-#NB! sample size and K
-sample_size <- 150
-k <- 20
-# To try to find low tau2 values for Pearson's r.
-true_tau2 <- seq(from = 0, to = 0.14, by = 0.0005)
-mu <- 0 #based on my simulations mu does not matter. Set to zero to avoid truncation.
-reps <- 1e4
-
-cond <- expand.grid(sample_size = sample_size,
-                    k = k,
-                    true_tau2 = true_tau2,
-                    mu = mu)
-
-
-out_list <- vector("list", length = nrow(cond))
-
-
-
-start <- Sys.time()
-
-ncores <- detectCores()
-cl <- makePSOCKcluster(ncores) # Create cluster based on nworkers. This because mclapply doesn't work on windows
-
-clusterEvalQ(cl, library(metafor))
-clusterExport(cl, c("simulate_rma_I2", "rnorm_truncated"))
-
-for(r in 1:nrow(cond)){ #gives us a list of lists
-
-    mes <- paste0("\n now on condition ", r)
-    cat(mes)
-
-cond_r <- cond[r,]
-#need to add because otherwise the nodes don't find r. Since this happens outside the nodes.
-
-clusterExport(cl, "cond_r") #exported everything in environment to each node, otherwise they don't have access to all functions in the master global environment
-
-    out_list[[r]] <- parallel::parLapply(
-                    cl = cl, # cluster
-                    1:reps, #looping over
-                    function(iteration){ #anonymous function needed when using for replications
-
-                        simulate_rma_I2(effect_type = "r",
-                                    k = cond_r$k,
-                                    sample_size = cond_r$sample_size,
-                                    true_tau2 = cond_r$true_tau2,
-                                    mu = cond_r$mu)
-        }
-    )
-}
-stopCluster(cl) # Shut down the nodes
-
-end <- Sys.time()
-end - start
-
-
[…]

### 2023-05-08 code/plots.r
minor code adjustments

@@ -132,7 +132,7 @@ ggsave("figures/plot3.png", width = 8.62, height = 9.93)
 
 # Plot 4
 
-dat <- readRDS("data/k_20_over_vs_underestimate.RDS")
+dat <- readRDS("data/k_5_k20_tau2_over_vs_underestimate.RDS")
 dat <- lapply(dat, function(x) as.data.frame(as.list(x)))
 
 dat <- data.table::rbindlist(dat, idcol = "mu")
@@ -155,7 +155,7 @@ dat <- dat[dat$mean_reliability < 1,]
 line_labels <- dat[dat$mu == "0.6",]
 
 # Add real truncated values
-trunc <- readRDS("data/truncated_tau2.RDS")
+trunc <- readRDS("data/k_20_truncated_tau2.RDS")
 trunc$true_tau2 <- trunc$nominal_tau2
 trunc$tau2_hat <- trunc$tau2
 
@@ -212,7 +212,7 @@ facet_wrap(~true_tau, scales = "free") +
 theme_bw()
 
 
-ggsave("figures/k_20_R_0.6-0.8_tau_sd_0.15.png", width = 8.62, height = 9.93)
+ggsave("figures/k_5_k20tau_R_0.6-0.9_tau_sd_0.15.png", width = 8.62, height = 9.93)
 
 
 

### 2023-05-08 code/supercomputer_simulator.r

@@ -109,8 +109,13 @@ end - start
 
 e <- lapply(out_list, function(x) do.call(rbind, x))
 e_means <- lapply(e, colMeans)
-names(e_means) <- c(paste0("mu = ", cond$mu, ";reliability_sd = ", cond$reliability_sd,
-                           ";mean_rel = ", cond$reliability_mean, ";true_tau2 = ", cond$true_tau2))
+#e_medians <- lapply(e, function(x) apply(x, 2, median))
+names(e) <- c(paste0("mu = ", cond$mu,
+                     ";k = ", cond$k,
+                     ";N = ", cond$sample_size,
+                     ";reliability_sd = ", cond$reliability_sd,
+                     ";mean_rel = ", cond$reliability_mean,
+                     ";true_tau2 = ", cond$true_tau2))
 
 lapply(e_means, round, 3)
-#saveRDS(e_means, "../data_new/over_vs_underestimate.RDS")
+#saveRDS(e, "../data_new/over_vs_underestimate.RDS")

### 2023-05-10 code/supercomputer_simulator.r
removed export cond_r

@@ -19,15 +19,12 @@ sample_size <- c(50, 100, 150, 200)
 k <- c(5, 20, 40, 200)
 true_tau <- c(0, 0.1, 0.15, 0.2)
 true_tau2 <- true_tau^2
-#Using Pearson's r as the effect size. Correspond to I2 20 - 90%.
-# Below corresponding values for k = 5
-# k <- 5
-#true_tau2 <- c(0, 0.0015, 0.0045, 0.008, 0.013, 0.0195, 0.0315, 0.0550, 0.1245)
-mu <- seq(from = 0, to = 0.6, by = 0.1) #
+#Using Pearson's r as the effect size.
+mu <- seq(from = 0, to = 0.6, by = 0.1)
 
 #Based on Flake et al., average alpha was .79, SD = .13, range .17 - .87
 #Based on Sanchez-Meca, mean across 5 meta-analysis was 0.767 - 0.891 and SD ranged between 0.034 - 0.133
-reliability_mean <- c(0.6, 0.7, 0.8, 0.9)
+reliability_mean <- c(0.6, 0.7, 0.8, 0.9, 1)
 reliability_sd <- seq(from = 0.15, to = 0.15, by = 0.05)
 
 
@@ -53,7 +50,7 @@ cond$maxiterations  <-  100 #default values are steplength = 1, and maxiter = 10
 #****************************************
 # Setup parallel simulation
 #****************************************
-reps <- 1e2
+reps <- 1e4
 out_list <- vector("list", length = nrow(cond))
 
 ncores <- detectCores()
@@ -65,18 +62,14 @@ clusterExport(cl, c("simulate_rma", "rnorm_truncated"))
 #****************************************
 # Run simulation
 #****************************************
+
 start <- Sys.time()
 for(r in 1:nrow(cond)){ #gives us a list of lists
 
     mes <- paste0("\n now on condition ", r)
     cat(mes)
 
-cond_r <- cond[r,]
-#need to add because otherwise the nodes don't find r. Since this happens outside the nodes.
-
-clusterExport(cl, "cond_r") #exported each row to the nodes otherwise they don't have access
-
-task <- function(iteration){ #anonymous function needed when using for replications
+task <- function(iteration, cond_r){ #anonymous function needed when using for replications
         simulate_rma(effect_type = cond_r$effect_type,
                     reliability_distribution = cond_r$reliability_distribution,
                     k = cond_r$k,
[…]

### 2023-05-10 code/compute_tau_fishers_z.r
updated fishers z tau values and truncation

@@ -37,14 +37,14 @@ sampling_var <- function(rho, sample_size){
 
 internal <- function(x, rho, tau, sample_size){
 
-    g_x <- sampling_var(rho, sample_size)
+    g_x <- sampling_var(x, sample_size)
     f_x <- dnorm(x, mean = rho, sd = tau)
 
     g_x * f_x #to be integrated over, see https://en.wikipedia.org/wiki/Law_of_the_unconscious_statistician
 }
 
-integrate(internal, lower = -0.99, upper = 0.99,
-           rho = rho, tau = tau, sample_size = sample_size)
+# integrate(internal, lower = -0.999, upper = 0.999,
+#            rho = rho, tau = tau, sample_size = sample_size)
 
 compute_s2 <- function(sigma2, k){
 
@@ -55,17 +55,16 @@ compute_s2 <- function(sigma2, k){
     s2_num / s2_denom #s2 out
 }
 
-compute_I2 <- function(vec = c(rho, tau, sample_size)){
-    # #rho = average true effect size
-    # #tau = SD in true effect sizes
-    # #sample_size = fixed within-study sample size
+compute_I2 <- function(vec = c(rho, tau, sample_size, k = 20)){
+    # rho = average true effect size
+    # tau = SD in true effect sizes
+    # sample_size = fixed within-study sample size
+    # k = 20: constant, random value, doesn't matter when N is fixed across studies
 
     rho <- vec[["rho"]]
     tau <- vec[["tau"]]
     sample_size <- vec[["sample_size"]]
-
     tau2 <- tau^2
-    k <- 20 #constant, random value, doesn't matter when N is fixed across studies
 
     sigma2 <- integrate(internal, #function g(x)f(x)
                         lower = -0.99, upper = 0.99,
@@ -79,15 +78,18 @@ compute_I2 <- function(vec = c(rho, tau, sample_size)){
 
 }
 
-compute_tau2_z <- function(vec = c(I2, sample_size)){
+compute_tau2_z <- function(vec = c(I2, sample_size, k = 20)){
     #I2 = I2 from pearson R
     #sample_size = fixed within-study sample size
+    # k <- 20: constant, random value, doesn't matter when N is fixed across studies
+
     I2 <- vec[["I2"]]
     sample_size <- vec[["sample_size"]]
-    k <- 20 #constant, random value, doesn't matter when N is fixed across studies
 
     sigma2  <- 1 / (sample_size - 3)
     s2 <- compute_s2(sigma2, k)
[…]

### 2023-05-10 code/supercomputer_simulator.r

@@ -7,7 +7,8 @@
 #****************************************
 # libraries and functions
 #****************************************
-library(parallel) #comes pre-installed with R
+library(parabar) #for a progress-bar in parallel computing
+# library(parallel) #comes pre-installed with R
 #library(metafor) #loaded when setting up simulations
 
 source("code/functions.r") #for rnorm_truncated and simulate_rma
@@ -17,10 +18,18 @@ source("code/functions.r") #for rnorm_truncated and simulate_rma
 #****************************************
 sample_size <- c(50, 100, 150, 200)
 k <- c(5, 20, 40, 200)
+mu <- seq(from = 0, to = 0.6, by = 0.1)
+#Using Pearson's r as the effect size.
 true_tau <- c(0, 0.1, 0.15, 0.2)
 true_tau2 <- true_tau^2
-#Using Pearson's r as the effect size.
-mu <- seq(from = 0, to = 0.6, by = 0.1)
+
+#Using Fisher's z as the effect size (see 'compute_tau_fisher_z.r')
+true_tau_50 <- c(0.1031263, 0.1566007, 0.2123514) #For N = 50)
+true_tau_100 <- c(0.1020357, 0.1549445, 0.2101057) #For N = 100
+true_tau_150 <- c(0.1016845, 0.1544113, 0.2093826) #For N = 150
+true_tau_200 <- c(0.1015112, 0.1541480, 0.2090256) #For N = 200
+true_tau <- c(0, true_tau_50, true_tau_100, true_tau_150, true_tau_200)
+true_tau2 <- true_tau^2
 
 #Based on Flake et al., average alpha was .79, SD = .13, range .17 - .87
 #Based on Sanchez-Meca, mean across 5 meta-analysis was 0.767 - 0.891 and SD ranged between 0.034 - 0.133
@@ -39,25 +48,34 @@ cond <- expand.grid(sample_size = sample_size,
 remove_cond <- cond[["reliability_mean"]] == 1 & cond[["reliability_sd"]] > 0
 cond <- cond[!remove_cond,]
 
+#drop incorrect tau-values for fisher's z
+keep_cond <- cond[["true_tau2"]] == 0 |
+               (cond[["sample_size"]] == 50 & cond[["true_tau2"]] %in% true_tau_50^2) |
+               (cond[["sample_size"]] == 100 & cond[["true_tau2"]] %in% true_tau_100^2) |
[…]

### 2023-05-10 code/truncated_tau.r

@@ -1,0 +1,19 @@
+
+source("./code/functions.r")
+es <- seq(0, 0.6, 0.1)
+tau <- c(0, 0.1, 0.15, 0.2)
+
+a <- lapply(tau, function(x){
+
+lapply(es, function(y) {
+res <- rnorm_truncated(n = 1e6, mean = y, sd = x,
+lower_bound = -1, upper_bound = 1)
+
+data.frame(tau = sd(res), mu = y, nominal_tau = x)
+})
+
+})
+
+b <- lapply(a, function(x) do.call(rbind, x))
+b <- do.call(rbind, b)
+saveRDS(b, "data/truncated_tau.RDS")

### 2023-05-10 code/truncated_tau2.r

@@ -1,18 +1,0 @@
-
-source("./code/functions.r")
-es <- seq(0, 0.6, 0.1)
-vars <- c(0, 0.002, 0.0035, 0.0055, 0.0085, 0.012, 0.0185, 0.031, 0.069)
-#vars <- c(0, 0.0015, 0.0045, 0.008, 0.013, 0.0195, 0.0315, 0.0550, 0.1245)
-a <- lapply(vars, function(x){
-
-lapply(es, function(y) {
-res <- rnorm_truncated(n = 1e6, mean = y, sd = sqrt(x),
-lower_bound = -1, upper_bound = 1)
-data.frame(tau2 = var(res), mu = y, nominal_tau2 = x)
-})
-
-})
-
-b <- lapply(a, function(x) do.call(rbind, x))
-b <- do.call(rbind, b)
-saveRDS(b, "data/k_5_truncated_tau2.RDS")

### 2023-05-10 code/plots.r
updated data na dplots

@@ -258,4 +258,113 @@ geom_line(aes(color = condition), show.legend = TRUE) +
 geom_hline(yintercept = sqrt(0.078), linetype = "dashed") +
 ylab("prop of true tau2") +
 theme_bw()
-2
+
+
+#****************************************
+# supplemental plots + fisher's z
+#****************************************
+
+# Plots with many different N and K
+# Either for Pearson's r or for Fisher's z
+# Tau and mu are equal in both cases
+
+#1) Create all Pearson's r plots
+
+
+dat <- readRDS("data/means_r_z_over_vs_underestimate.RDS")
+dat <- lapply(dat, function(x) as.data.frame(as.list(x)))
+
+dat <- data.table::rbindlist(dat, idcol = "mu")
+
+#create condition ids
+grep_strip <- function(string, vec){
+    #Uses a common string like "k = " to extract "k = 50" or "k = 200"
+    #Then removes the "k = " part and returns as a numeric value
+    a <- grep(string, vec, value = TRUE)
+    b <- gsub(string, "", a)
+}
+
+conditions <- unlist(strsplit(dat$mu, split = ";"))
+
+dat$k <- grep_strip("k = ", conditions)
+dat$N <- grep_strip("N = ", conditions)
+dat$reliability_sd <- grep_strip("reliability_sd = ", conditions)
+dat$true_tau2 <- as.numeric(grep_strip("true_tau2 =", conditions)) #numeric for plotting later
+dat$mean_reliability <- grep_strip("mean_rel = ", conditions)
+dat$mu <- as.numeric(grep_strip("mu = ", conditions))
+dat$effect_type <- grep_strip("effect_tpe = ", conditions) #nb type in name!
+
+# Add real truncated values
+trunc <- readRDS("data/truncated_tau.RDS")
+trunc$true_tau <- trunc$nominal_tau
+trunc$tau_hat <- trunc$tau
+
+#Prepping plot
+dat$tau_hat <- sqrt(dat$tau2_hat)
+dat$true_tau <- sqrt(dat$true_tau2)
+dat$mean_reliability <- factor(dat$mean_reliability)
+dat$k <- factor(dat$k, levels = c("5", "20", "40", "200"))
+dat$N <- factor(dat$N, levels = c("50", "100", "150", "200"))
+
[…]

### 2023-05-10 supplemental_data/brannik_raw-data-simulation.r
doublechecked brannik method

@@ -1,0 +1,100 @@
+
+
+
+dat <- matrix(c("University_1","130","0.24","0.75",
+"University_2","90","0.11","0.75",
+"Private_1","30","0.05","0.60",
+"Private_2","25","0.17","0.60",
+"Volunteer_1","50","0.38","0.90",
+"Volunteer_2","65","0.50","0.90"), nrow = 6, byrow = TRUE)
+
+dat <- as.data.frame(dat)
+names(dat) <- c("study", "n", "r", "rel")
+
+dat$r <- as.numeric(dat$r)
+dat$rel <- as.numeric(dat$rel)
+dat$n <- as.numeric(dat$n)
+dat$r_var <- (1 - dat$r^2)^2 / (dat$n - 1)
+
+dat$rho <- dat$r / sqrt(dat$rel)
+dat$rho_var <- (1 - dat$rho^2)^2 / (dat$n - 1)
+
+dat$rho_var2 <- dat$r_var / sqrt(dat$rel)
+#estimate rho var based on observed r
+
+library(metafor)
+rma(yi = dat$r, vi = dat$r_var)
+rma(yi = dat$rho, vi = dat$rho_var)
+
+rma(yi = dat$rho, vi = dat$rho_var2)
+# If I use the variance based on the estimated r
+# I get smaller heterogeneity with true effect sizes!
+
+# 2 ideas that might be the difference
+# a) add measurement error before sampling error
+#   - This does not explain it, I just tried with my
+#     function and it still underestimates tau by about 0.03
+# b) Compute observed correlations from a sample
+# (NB. b) interacts with a))
+#   - This also doesn't explain it, I just used Brannick's
+# code below and there is still an underestimate when using
+# meta-analyzing unreliabile data
+
+# Brannick et al., simulation
+TxTy <- rnorm(10, mean = .26, sd = .13)
+rel_x <- runif(10, min = 0, max = 1)
+rel_y <- runif(10, min = 0, max = 1)
+
+TxX <- sqrt(rel_X)
+TyY <- sqrt(rel_y)
+xy <- TxTy*TxX*TyY
+
+#Compute attenuated correlations
+n <- 142
+
+rhobox <- matrix(0,4,4)+diag(4) # placeholder for correlation matrix for sampling
+mu <- rep(0,4)          # multivariate means of zero
+names.sim <- cbind("Tx", "Ty", "X", "Y")
[…]

### 2023-05-10 manuscript/introduction.qmd
intro and methods as .qmd files

---
format: docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Estimating heterogeneity of effect sizes is usually considered the second main purpose of meta-analysis, after estimating an average effect size. Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size's sensitivity to variability in study design, or more specifically to sensitivity to differences in four study factors: 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables (e.g., Campbell & Stanley, 2015). From a more statistical perspective, heterogeneity is defined as the variance between 'true' subpopulation effect sizes. However, true subpopulation effect sizes can refer to two different entities. Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as measurement error, restriction of range, and dichotomization; Hunter & Schmidt, YEAR). Of these measurement artifacts, measurement error is present in any empirical effect size estimate and systematically attenuates effect sizes. It is usually measured as reliability (defined in classical test theory as the ratio of true score variance to measurement error variance; REF) and is relatively common to report in primary studies in psychology (REF: Flake). However, whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for imperfect reliability is implicit only in the models of Hunter & Schmidt (YEAR).

One reason could be that correcting for unreliability is controversial and has been so since it was proposed (for some arguments spanning the last century, see Table 2 of LeBreton et al, YEAR). On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., REF1, REF2). Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1). Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; REF Flake) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas). Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2, REF3), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer. On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures. As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest. From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald, Hunter & Schmidt).

Meta-analyses in psychology rarely correct for unreliability in primary study measurements. Wiernick and Dahlke (REF) report that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability. Similarly, Schmidt, Oh and Hayes (REF) report that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts. The exception is the subfield of Industrial-Organizational studies. Although at the meta-analysis level rather than study level, Cortina (REF) showed that about 80% of the 1,294 meta-analyses published across 59 meta-analytic studies in the Journal of Applied Psychology between 1978 - 1997 corrected for measurement artifacts, and Anguinis (REF) that around 50% of meta-analyses identified in studies published across several management journal between 1989 - 2009 corrected for unreliability of measurement. In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses should be aware of how this decision affects heterogeneity estimates.

Imperfect reliability in primary studies imbues heterogeneity estimates with both a positive and a negative bias. The positive bias is discussed extensively by Hunter and Schmidt (REF) who introduced the concept of psychometric meta-analysis (i.e., meta-analysis corrected for measurement artifacts in primary studies, such as unreliability), as well as discussed in other highly cited meta-analysis handbooks (Borenstein, 2009, p.342; Card, 2012, p. 126) and recent literature (Wiernick and Dahlke, 2020). However, there appears to be a lack of acknowledgement that imperfect reliability can also lead to a negative bias in heterogeneity estimates. The only mention we found thereof was in the 2nd edition of _Methods of Meta-analysis_ (Hunter & Schmidt, p. 142) in a paragraph later absent from the 3rd edition (REF). Below we detail the argument for both a positive and negative bias in heterogeneity estimates.

## Bias in heterogeneity estimates due to unreliability

In classical test theory (CTT), reliability of measurement ($R_{xx}$) is defined as the proportion true score variance ($\sigma^2_T$) to observed score variance ($\sigma^2_X$) and is presumed to be constant for all participants in a sample $R_{xx} = \sigma^2_T / \sigma^2_X$. We use unreliability of measurement to refer to $1 - R_{xx}$. Imperfect reliability in measurements leads to attenuation of observed effect sizes. For product-moment correlations the attenuation is simply computed as $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ where $r_{xy}$ is the observed correlation between the variables $X$ and $Y$, $\rho_{xy}$ is the true correlation (with or without sampling error) and $R_{xx'}$ and $R_{yy'}$ are the measurement reliabilities for $X$ and $Y$. As such, correcting for unreliability is straightforward for product-moment correlations (although the sampling variance must then also be corrected: REF), but can also be done for other effect size types (REF: Wiernick).

A positive bias can be expected in heterogeneity estimates of uncorrected effect sizes due to the variance in reliability between primary studies. This is because the attenuation in effect size will then differ from study to study, which results in differences across studies in observed effect sizes beyond their true variability. This is most easily seen by assuming infinite sample size (i.e., no within-study variance) and in the context of a fixed effect model (i.e., with a common underlying mean across studies and no true heterogeneity). Table 1 illustrates how under these conditions three studies with differing reliability will observe variability in effect sizes beyond their true variability (i.e., zero). Table 1 also shows how this effect increases as the true effect size increases.

(TABLE 1; reproduced with permission from Olsson-Collentine, 2020)

A negative bias can be expected in heterogeneity estimates of uncorrected effect sizes due to larger attenuation of larger effect sizes for the same level of unreliability. That is, although the attenuation formula $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ implies that any effect size $\rho_{xy}$ will be attenuated by the same proportion for fixed reliabilities $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$, this proportion corresponds to a larger absolute value for larger effect sizes. As a consequence, in the presence of heterogeneity larger true effect sizes will move further towards zero than smaller ones, decreasing heterogeneity in observed effect sizes.

Table 2 illustrates this relationship for infinite sample size (i.e., no sampling variance) and several levels of average reliability across three studies.

(Table 2)

Attenuation of effect size can also be understood as a downwards shift of the distribution of true effect sizes. To illustrate, assume an infinite sample size (i.e., no within-study sampling variance), some heterogeneity $\tau^2$, equal reliability in outcome and measurement such that $R_{xx'} = R_{yy'} = R$, and that all studies within a population of studies have the same reliability. Observed effect sizes given measurement error then corresponds to multiplying all true effect sizes $\theta_i$ by a factor equal to the reliability $R$. This is equal to multiplying the distribution of effect sizes by a constant $R$. If the distribution of effect sizes is normally distributed $\theta_i \sim N(\theta, \tau^2)$ with a mean $\theta$ (i.e., the assumption of the random effects model), then $R \times N(\theta, \tau) = N(R\theta, R^2  \tau^2) = N(R\theta, R\tau)$. Because reliability is always a value < 1, it can be seen that the square root of the observed heterogeneity for uncorrected effect sizes is less than its true value by a factor $R$. This depiction is not entirely correct for Pearson's $r$ as it is bounded at {-1, 1}, but holds approximately as long as the mean effect size $\theta$ and heterogeneity $\tau^2$ are not too large.

To make this concrete, consider an average reliability of 0.8 (the median reported reliability in psychology; REF). Then, $0.8 \times N(\theta, \tau) = N(0.8\theta, 0.8 \tau)$, resulting in 20% less heterogeneity among observed effect sizes than in true effect sizes. We can consider the positive bias in the same vein. Assuming a common true effect size $\mu$ across studies and infinite sample size within studies we can view $\mu$ as a constant. In this case, attenuation consists of a distribution of reliabilities $R \sim N(\bar{R}, \sigma_R)$ with some mean $\bar{R}$ and standard deviation $\sigma_R$ across studies, that is multiplied by the constant $\mu$ to yield a distribution of observed effect sizes $r \sim N(\mu \bar{R}, \mu \sigma_R)$. To make this concrete, consider some values $\mu = 0.2$, $\bar{R} = 0.8$, $\sigma_R = 0.1$ which then yield $r \sim N(0.16, 0.01)$. Again, also this depiction is not entirely correct as reliability is bounded at {0, 1}, but holds approximately as long as the mean reliability $\bar{R}$ and its standard deviation $\sigma_R$ are not too large.

It is difficult to predict the total effect of the negative and positive bias in heterogeneity due to unreliability. This depends on average effect size, the amount of heterogeneity, mean reliability across studies, the variability in reliabilities, and as the sampling variance for Pearson's $r$ is affected by its effect size, sample size within studies may also matter. We thus perform a Monte-Carlo simulation study to examine the expected bias in heterogeneity estimates when meta-analyzing primary studies with imperfect measurement reliabilities.

### 2023-05-11 manuscript/introduction.qmd
updated introduction and added tables

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
+```
+
+```{r load data and prep, include = FALSE}
+doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to') #to define the type of document, then format tables using an ifelse statement, see also: https://stackoverflow.com/questions/58999720/scientific-journal-table-of-variable-mean-per-year-with-s-d
+
+tables <- readRDS("tables.RDS")
+library(kableExtra) #must be loaded to print saved kableextra tables
+
 ```
@@
 However, true subpopulation effect sizes can refer to two different entities.
-Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as measurement error, restriction of range, and dichotomization; Hunter & Schmidt, YEAR).
+Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as measurement error, restriction of range, and dichotomization; Hunter & Schmidt, 2015).
 Of these measurement artifacts, measurement error is present in any empirical effect size estimate and systematically attenuates effect sizes.
-It is usually measured as reliability (defined in classical test theory as the ratio of true score variance to measurement error variance; REF) and is relatively common to report in primary studies in psychology (REF: Flake).
-However, whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for imperfect reliability is implicit only in the models of Hunter & Schmidt (YEAR).
+It is usually measured as reliability (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., REF) and is relatively common to report in primary studies in psychology (Flake et al., 2017).
+However, whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for imperfect reliability is implicit only in the models of Hunter & Schmidt (2015).
+This has consequences for heterogeneity estimates reported in many areas of psychology and in this article we use Monte-Carlo simulations to explore the consequences for heterogeneity estimates in meta-analysis from not correcting for unreliability.
 
-One reason could be that correcting for unreliability is controversial and has been so since it was proposed (for some arguments spanning the last century, see Table 2 of LeBreton et al, YEAR).
-On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., REF1, REF2).
+Meta-analyses in psychology rarely correct for unreliability in primary study measurements.
+Wiernick and Dahlke (2020) report that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability.
+Similarly, Schmidt (2010) report that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts.
+The exception is the subfield of Industrial-Organizational studies.
+Although at the meta-analysis level rather than study level, Cortina (2003) showed that about 80% of the 1,294 meta-analyses published across 59 meta-analytic studies in the Journal of Applied Psychology between 1978 - 1997 corrected for measurement artifacts, and Anguinis (2011) that around 50% of meta-analyses identified in studies published across several management journal between 1989 - 2009 corrected for unreliability of measurement.
+
+One reason that few meta-analysts correct for unreliability could be that correcting for unreliability is generally controversial and has been so since it was proposed (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017).
+On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017).
 Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1).
-Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; REF Flake) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas).
+Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas).
 Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2, REF3), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
@@
 As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest.
-From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald, Hunter & Schmidt).
-
-Meta-analyses in psychology rarely correct for unreliability in primary study measurements.
-Wiernick and Dahlke (REF) report that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability.
-Similarly, Schmidt, Oh and Hayes (REF) report that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts.
-The exception is the subfield of Industrial-Organizational studies.
-Although at the meta-analysis level rather than study level, Cortina (REF) showed that about 80% of the 1,294 meta-analyses published across 59 meta-analytic studies in the Journal of Applied Psychology between 1978 - 1997 corrected for measurement artifacts, and Anguinis (REF) that around 50% of meta-analyses identified in studies published across several management journal between 1989 - 2009 corrected for unreliability of measurement.
+From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015).
 In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses should be aware of how this decision affects heterogeneity estimates.
@@
 Imperfect reliability in primary studies imbues heterogeneity estimates with both a positive and a negative bias.
-The positive bias is discussed extensively by Hunter and Schmidt (REF) who introduced the concept of psychometric meta-analysis (i.e., meta-analysis corrected for measurement artifacts in primary studies, such as unreliability), as well as discussed in other highly cited meta-analysis handbooks (Borenstein, 2009, p.342; Card, 2012, p.
-126) and recent literature (Wiernick and Dahlke, 2020).
+The positive bias is discussed by several highly cited meta-analytic handbooks, most prominently by Hunter and Schmidt (2015; 12,235 citations on google scholar as of 2023-05-11) who introduced the concept of psychometric meta-analysis (i.e., meta-analysis corrected for measurement artifacts such as unreliability), but also by Borenstein et al.
+(2009, p.
+342; 19,020 citations), Card (2015, p.126; 1,716 citations) as well as in recent literature (Wiernick and Dahlke, 2020).
 However, there appears to be a lack of acknowledgement that imperfect reliability can also lead to a negative bias in heterogeneity estimates.
-The only mention we found thereof was in the 2nd edition of _Methods of Meta-analysis_ (Hunter & Schmidt, p.
-142) in a paragraph later absent from the 3rd edition (REF).
+The only one of the above sources to refer to a negative bias is Hunter and Schmidt who mention it shortly in their 2nd edition of *Methods of Meta-analysis* (2004, p.
+142) in a paragraph later absent from the 3rd edition (Hunter & Schmidt, 2015).
 Below we detail the argument for both a positive and negative bias in heterogeneity estimates.
@@
 Imperfect reliability in measurements leads to attenuation of observed effect sizes.
-For product-moment correlations the attenuation is simply computed as $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ where $r_{xy}$ is the observed correlation between the variables $X$ and $Y$, $\rho_{xy}$ is the true correlation (with or without sampling error) and $R_{xx'}$ and $R_{yy'}$ are the measurement reliabilities for $X$ and $Y$.
-As such, correcting for unreliability is straightforward for product-moment correlations (although the sampling variance must then also be corrected: REF), but can also be done for other effect size types (REF: Wiernick).
+For product-moment correlations the observed correlation can be computed as $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ where $r_{xy}$ is the observed correlation between the variables $X$ and $Y$, $\rho_{xy}$ is the true correlation (with or without sampling error) and $R_{xx'}$ and $R_{yy'}$ are the measurement reliabilities for $X$ and $Y$.
+As such, correcting for unreliability is straightforward for product-moment correlations, but can also be done for other effect size types (Wiernick and Dahlke, 2020).
+A corrected correlation should also have its sampling variance corrected.
+This is inflated by a factor equal to the square of the reliabilities, such that if $a = \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ then the corrected sampling variance $V_\rho = V_{r} / a^2$ (e.g., Borenstein, 2009, p.
+343).
 
@@
 
-(TABLE 1; reproduced with permission from Olsson-Collentine, 2020)
+Table 1.
 
-A negative bias can be expected in heterogeneity estimates of uncorrected effect sizes due to larger attenuation of larger effect sizes for the same level of unreliability.
-That is, although the attenuation formula $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ implies that any effect size $\rho_{xy}$ will be attenuated by the same proportion for fixed reliabilities $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$, this proportion corresponds to a larger absolute value for larger effect sizes.
+*Variance in reliability leads to positive bias in heterogeneity estimates which increases with average effect size*
+
+```{r positive bias table}
+
+if (doc_type == "docx") {
+  print("Table 1 approximately here. Cannot be printed nicely to Word, see tables.pdf")
+} else{
+  tables$positive_bias_table
+}
+
+```
+
+A negative bias can be expected in heterogeneity estimates of uncorrected effect sizes because given the same unreliability, the attenuation in absolute value increases with effect size.
+That is, although the attenuation formula $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ implies that any effect size $\rho_{xy}$ will be attenuated by the same proportion given the same reliabilities $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$, this proportion corresponds to a larger absolute value for larger effect sizes.
 As a consequence, in the presence of heterogeneity larger true effect sizes will move further towards zero than smaller ones, decreasing heterogeneity in observed effect sizes.
+Table 2 illustrates this relationship for infinite sample size (i.e., no sampling variance) and three levels of reliability across three studies.
 
-Table 2 illustrates this relationship for infinite sample size (i.e., no sampling variance) and several levels of average reliability across three studies.
+Table 2.
 
-(Table 2)
+*Imperfect reliability leads to negative bias in heterogeneity estimates*
+
+```{r negative bias table}
+
+if (doc_type == "docx") {
+  print("Table 2 approximately here. Cannot be printed nicely to Word, see tables.pdf")
+} else{
+  tables$negative_bias_table
+}
+
+```
 
@@
 To illustrate, assume an infinite sample size (i.e., no within-study sampling variance), some heterogeneity $\tau^2$, equal reliability in outcome and measurement such that $R_{xx'} = R_{yy'} = R$, and that all studies within a population of studies have the same reliability.
-Observed effect sizes given measurement error then corresponds to multiplying all true effect sizes $\theta_i$ by a factor equal to the reliability $R$.
-This is equal to multiplying the distribution of effect sizes by a constant $R$.
+Observed effect sizes given measurement error then corresponds to multiplying all true effect sizes $\theta_i$ by a factor equal to the reliability $R$ (i.e., multiplying the distribution of effect sizes by a constant $R$).
 If the distribution of effect sizes is normally distributed $\theta_i \sim N(\theta, \tau^2)$ with a mean $\theta$ (i.e., the assumption of the random effects model), then $R \times N(\theta, \tau) = N(R\theta, R^2 \tau^2) = N(R\theta, R\tau)$.
-Because reliability is always a value < 1, it can be seen that the square root of the observed heterogeneity for uncorrected effect sizes is less than its true value by a factor $R$.
+Because reliability is always a value \< 1, it can be seen that the square root of the observed heterogeneity for uncorrected effect sizes is less than its true value by a factor $R$.
 This depiction is not entirely correct for Pearson's $r$ as it is bounded at {-1, 1}, but holds approximately as long as the mean effect size $\theta$ and heterogeneity $\tau^2$ are not too large.
 
-To make this concrete, consider an average reliability of 0.8 (the median reported reliability in psychology; REF).
-Then, $0.8 \times N(\theta, \tau) = N(0.8\theta, 0.8 \tau)$, resulting in 20% less heterogeneity among observed effect sizes than in true effect sizes.
+To make this concrete, consider an average reliability of 0.8 (the median reported reliability in psychology; Flake et al., 2017).
+Then, $0.8 \times N(\theta, \tau) = N(0.8\theta, 0.8 \tau)$, resulting in 20% less standard deviation among observed effect sizes than in true effect sizes.
 We can consider the positive bias in the same vein.
@@
 To make this concrete, consider some values $\mu = 0.2$, $\bar{R} = 0.8$, $\sigma_R = 0.1$ which then yield $r \sim N(0.16, 0.01)$.
-Again, also this depiction is not entirely correct as reliability is bounded at {0, 1}, but holds approximately as long as the mean reliability $\bar{R}$ and its standard deviation $\sigma_R$ are not too large.
+Again, this depiction is not entirely correct as reliability is bounded at {0, 1}, but holds approximately as long as the mean reliability $\bar{R}$ and its standard deviation $\sigma_R$ are not too large.
 
 It is difficult to predict the total effect of the negative and positive bias in heterogeneity due to unreliability.
-This depends on average effect size, the amount of heterogeneity, mean reliability across studies, the variability in reliabilities, and as the sampling variance for Pearson's $r$ is affected by its effect size, sample size within studies may also matter.
-We thus perform a Monte-Carlo simulation study to examine the expected bias in heterogeneity estimates when meta-analyzing primary studies with imperfect measurement reliabilities.
+This depends on average effect size, the amount of heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies.
+Hence, we perform a Monte-Carlo simulation study to explore the expected bias in heterogeneity estimates when meta-analyzing primary studies with imperfect measurement reliabilities.

### 2023-05-11 code/tables.qmd
table code file..

---
format: pdf
---

# Project: Unreliable heterogeneity

```{r setup, include=FALSE}
# Script purpose: create tables for manuscript
# code: Anton Olsson-Collentine

knitr::opts_chunk$set(echo = FALSE, warning = FALSE)

library(kableExtra) #For creating good-looking pdf tables
```

## Table 1
### Variance in reliability leads to positive bias in heterogeneity estimates

```{r effect-size-heterogeneity-table}
#library(kableExtra)
rooted_relabilities <- matrix(rep(c(.6, .70,.80), 3),
                              ncol = 3, byrow = TRUE)
rho <- c(0.0, .30, .50)
ma <- c("I", "II", "III")

rooted_relabilities <- sweep(rooted_relabilities, 1, rho, "*")
devs <- apply(rooted_relabilities, 1, sd)
table1 <- data.frame(ma, rho, rooted_relabilities, devs)

positive_bias_table <- table1 %>%
         knitr::kable(booktabs = T, escape = FALSE, digits = 2,
          col.names = c("Meta-Analysis",
                        "$\\rho_{xy}$",
                        "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .6",
                        "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .7",
                        "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .8",
                        "SD(ES)")) %>%
  kable_styling(position = "left") %>%
  add_header_above(c(" " = 2, "Study 1"= 1, "Study 2" = 1, "Study 3" = 1)) %>%
  add_header_above(c(" " = 2, "Observed Effect Sizes" = 3)) %>%
  footnote(general = "Reproduced from Olsson Collentine et al. (2020). The values under Study 1, 2 and 3 are observed effect sizes for that study given its measurement reliability $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ and the true effect size $\\\\rho_{xy}$ when within-study sample size is infinite. SD (ES) is the standard deviation of the observed effect sizes for meta-analysis I, II and III, equivalent to heterogeneity given infinite within-study sample sizes. Code to reproduce table: LINK",
  threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)

positive_bias_table
```

## Table 2 (version 1)
### Imperfect reliability leads to negative bias in heterogeneity estimates

```{r negative-bias-heterogeneity}
library(kableExtra)

theta  <- matrix(rep(c(0.1, 0.15, 0.2), 3), ncol = 3, byrow = TRUE)
R <- c(1, 0.8, 0.6)
r <- sweep(theta, 1, R, "*")
sds <- apply(r, 1, sd)

dat <- cbind(as.character(R), r, sds)

dat %>% knitr::kable(booktabs = TRUE, escape = FALSE, digits = 2,
                    align=rep('c', 5),
 col.names = c("$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$",
               "Study 1",
               "Study 2",
               "Study 3",
               "SD(ES)")) %>%
 kable_styling(position = "left") %>%
 add_header_above(c(" " = 1, "Observed Effect Sizes" = 3)) %>%
 footnote(general = "The values under Study 1, 2 and 3 are observed effect sizes for that study given the measurement reliability $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ and the true effect size of that study (first row) when within-study sample size is infinite. SD (ES) is the standard deviation of the observed effect sizes, equivalent to heterogeneity given infinite within-study sample sizes. Code to reproduce table: LINK",
  threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)
```

## Table 2 (version 2)
### Imperfect reliability leads to negative bias in heterogeneity estimates

```{r, alternative negative bias table}
#Alternative table
rel <- c("$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = 1",
         "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = 0.8",
         "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = 0.6")

dat <- cbind(rel, r, sds)

negative_bias_table <- dat %>%
            knitr::kable(booktabs = TRUE, escape = FALSE, digits = 2,
            col.names = c("",
                        "Study 1",
                        "Study 2",
                        "Study 3",
                        "SD(ES)")) %>%
            kable_styling(position = "left") %>%
            add_header_above(c(" " = 1, "Observed Effect Sizes" = 3)) %>%
            column_spec (1, border_right = FALSE) %>%
            footnote(general = "The values under Study 1, 2 and 3 are observed effect sizes for that study given the measurement reliability $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ and the true effect size of that study (first row) when within-study sample size is infinite. SD (ES) is the standard deviation of the observed effect sizes, equivalent to heterogeneity given infinite within-study sample sizes. Code to reproduce table: LINK",
  threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)

negative_bias_table
```

```{r output tables}
#Saving the output of all tables so that they can be imported to the main manuscript easily

saveRDS(list(positve_bias_table = positive_bias_table,
             negative_bias_table = negative_bias_table),
        "../manuscript/tables.RDS")
```

### 2023-05-11 code/compute_tau_fishers_z.r
fixed problem with k = 20 in function

@@ -55,7 +55,7 @@ compute_s2 <- function(sigma2, k){
     s2_num / s2_denom #s2 out
 }
 
-compute_I2 <- function(vec = c(rho, tau, sample_size, k = 20)){
+compute_I2 <- function(vec = c(rho, tau, sample_size), k = 20){
     # rho = average true effect size
     # tau = SD in true effect sizes
     # sample_size = fixed within-study sample size
@@ -78,7 +78,7 @@ compute_I2 <- function(vec = c(rho, tau, sample_size, k = 20)){
 
 }
 
-compute_tau2_z <- function(vec = c(I2, sample_size, k = 20)){
+compute_tau2_z <- function(vec = c(I2, sample_size), k = 20){
     #I2 = I2 from pearson R
     #sample_size = fixed within-study sample size
     # k <- 20: constant, random value, doesn't matter when N is fixed across studies

### 2023-05-12 manuscript/methods.qmd
intro and methods as .qmd files
added fisher z discussion to methods

--- 2023-04-26 manuscript/methods.rmd
@@
 ---
-output: word_document
+format: docx
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
 
+## Choice of effect size type for simulations
+
+We use correlations as our effect size type in these simulations because correcting them for unreliability is straightforward and because they are the effect size type most commonly corrected in practice.
+However, unreliability attenuates all types of effect sizes and all can be corrected, either directly (standardized mean differences; Wiernick & Dahlke, 2020) or by first transforming them into correlations (with associated assumptions).
+
+Perhaps the most common way to meta-analyze correlations is to first transform them into Fisher's z correlations (Borenstein et al., 2009, p.
+41 - 43).
+The Fisher z transformation is the inverse hyperbolic tangent of the product moment correlation and can be computed as $equation$ and its variance approximated as $equation2$.
+There is some controversy in the use of the Fisher z transformation for meta-analysis (for a good summary see Brannick et al., 2019, p.
+492), in particular among proponents of Hunter and Schmidt's (2015) meta-anlytic models.
+
+We perform simulations both on product-moment correlations and Fisher z transformed correlations.
+For our purpose of estimating heterogeneity, Fisher z correlations have several advantages: 1) their sampling variance is not dependent on the size of the correlation, 2) they are not bounded to {-1, 1}.
+However, they also have two disadvantages: 1) that the heterogeneity estimate is measure on the Fisher z scale, which makes it hard to interpret.
+A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (REF) for an indirect estimate of the heterogeneity.
+Second, adding measurement error to the Fisher z correlations requires first transforming them into product-moment correlations, adding measurement error, and then backtransforming to Fisher z for meta-analysis.
+
+The primary advantage to using product-moment correlations in the meta-analysis when interested in heterogeneity is that values on this scale are more interpretable that on the Fisher z scale.
+The disadvantage is that the sampling variance is dependent on the effect size since it is computed as $equation$.
+Given the common use of both effect size types and their mix of advantages and disadvantages we report results for both.
+
 ## Simulation study design
@@
 
-```{r Figure 1, echo=FALSE, out.width= "100%"}
+```{r Figure 1, out.width= "100%"}
 knitr::include_graphics("../figures/flowchart-simulation-design.png")
@@
 That is, a normal distribution with a mean of $\theta$ and a variance of $\tau^2$ that we truncated at $\{-1, 1\}$ to stay within the bounds of the product-moment correlation.
+We implemented the truncation in R using inverse transform sampling (link: functions.r).
 For Fisher' $z$ no truncation was applied.
-
-Practically, we implemented the truncation as follows.
-First, we sampled $k$ probabilities from an uniform distribution $U(a, b)$ where $a$ and $b$ are the values given by the cumulative distribution function (CDF) of the normal distribution $N(\theta, \tau^2)$ at $\{-1, 1\}$.
-Second, we then back-transformed the $k$ sampled values onto the $N(\theta, \tau^2)$ distribution using the inverse of the CDF.
 
@@
 We meta-analyzed the $k$ effect sizes using a random effects model with inverse variance weights and the Restricted Maximum Likelihood (REML) estimator of heterogeneity.
-We choose the REML estimator because it is a heterogeneity estimator that performs well and is widely implemented (e.g., REF, REF2), but also show our results using the classic DerSimonian-Laird (DL) estimator in supplement X.
-Similarly, inverse variance weights are widely used, perform similarly to other common weights (e.g., REF1 REF2) and we see no reason other implementations would change our results substantially.
+We choose the REML estimator because it is a heterogeneity estimator that performs well and is widely implemented (e.g., REF, REF2), but also show our results using the classic DerSimonian-Laird (DL) estimator in supplement XX.
+We used inverse variance weights due to their widespread use, but as we used a fixed sample size across studies within a meta-analysis (see next section) the weights are of little importance.
+As a sensitivity analysis, we also implemented the Hunter and Schmidt (2015) model which uses a different heterogeneity estimator and found qualitatively similar results (supplement XX).
 The procedure described in this section was replicated 10,000 times for each combination of parameter values.
@@
 Van Erp et al., (REF) provide empirical heterogeneity estimates from a large number of meta-analyses.
-The median Pearson's $r$ $\tau^2$ value in these data
-was 0.03 (interquartile range 0.01 - 0.0576).
-As the point of this study is that empirical heterogeneity estimates may not be accurately estimated we selected a range of $\tau^2$ estimates for the product-moment that we considered reasonable and that also covered most of the empirical estimates from van Erp et al.
-These were {0, ..., VALUES}.
+The median (uncorrected)Pearson's $r$ $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24).
+As the point of this study is that empirical heterogeneity estimates may not be accurately estimated we selected a range of $\tau$ estimates for the product-moment that we considered reasonable and that also covered most of the empirical estimates from van Erp et al.
+These were {0, 0.1, 0.15, 0.2}.
+We selected the largest heterogeneity level to avoid excissive truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1.
 
-Fisher's z is measured at a different scale and it is not possible to directly convert $\tau^2$-values expressed in Pearson's $r$ into corresponding $\tau^2$-values on the Fisher's $z$ scale.
+Fisher's z is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson's $r$ into corresponding $\tau$-values on the Fisher's $z$ scale.
 To be able to compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale.
@@
 However, because we keep $N_i$ fixed across studies (see below), the value of $I^2$ depends only on $N = N_i$ and the heterogeneity.
-We thus compute the $I^2$ index for each sample size condition and $\tau^2$-values defined in Pearson's $r$, and then compute corresponding $\tau^2$-values on the Fisher's $z$ scale given these $I^2$-values and sample sizes.
+
+We thus computed the $I^2$ index for each sample size condition and $\tau$-values defined in Pearson's $r$, and then computed corresponding $\tau$-values on the Fisher's $z$ scale given these $I^2$-values and sample sizes.
+Two complications were that the value for $\sigma^2$ varies with effect size for Pearson's $r$ and that effect size varies across studies with $\tau$.
+We resolved this by setting $\theta = 0$ and computing the expected $\sigma^2$ given $\tau$ and sample size and then using this expected value to compute $I^2$.
+We used the 'law of the unconscious statistician' to compute the expected value of $\sigma^2$ which says that the expected value of a function $g(X)$ of a random variable can be expressed in terms of the probability distribution of X: $E[g(X)] = \int_{-\infty}^\infty g(x) f(x) dx$.
+In our case, $g(x)$ corresponds to the sampling variance of the product-moment correlation, $f(x) \sim N(\theta, \tau^2)$ and because the product-moment correlation is bounded at {-1, 1} we set $\infty$ instead to 0.999.
+The resulting $\tau$ values for fisher z are only minimally different from {0, 0.1, 0.15, 0.2}, but can be found in supplement XX.
 
@@
 Within meta-analyses we used a fixed sample size across all studies such that $N = N_i$.
-We made this choice rather than treating sample size as a random variable to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance.
+We made this choice rather than treating sample size as a random variable (as done by e.g., REF1, REF2) to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance.
 We base our range of sample sizes on empirical estimates of typical sample size in correlational research in psychology by XXX (still needs lookup) and consider the following values {VALUES}.

### 2023-05-12 code/supercomputer_simulator.r
transformed mu r to z

@@ -19,22 +19,30 @@ source("code/functions.r") #for rnorm_truncated and simulate_rma
 sample_size <- c(50, 100, 150, 200)
 k <- c(5, 20, 40, 200)
 mu <- seq(from = 0, to = 0.6, by = 0.1)
-#Using Pearson's r as the effect size.
-true_tau <- c(0, 0.1, 0.15, 0.2)
-true_tau2 <- true_tau^2
-
-#Using Fisher's z as the effect size (see 'compute_tau_fisher_z.r')
-true_tau_50 <- c(0.1031263, 0.1566007, 0.2123514) #For N = 50)
-true_tau_100 <- c(0.1020357, 0.1549445, 0.2101057) #For N = 100
-true_tau_150 <- c(0.1016845, 0.1544113, 0.2093826) #For N = 150
-true_tau_200 <- c(0.1015112, 0.1541480, 0.2090256) #For N = 200
-true_tau <- c(0, true_tau_50, true_tau_100, true_tau_150, true_tau_200)
-true_tau2 <- true_tau^2
 
 #Based on Flake et al., average alpha was .79, SD = .13, range .17 - .87
 #Based on Sanchez-Meca, mean across 5 meta-analysis was 0.767 - 0.891 and SD ranged between 0.034 - 0.133
 reliability_mean <- c(0.6, 0.7, 0.8, 0.9, 1)
 reliability_sd <- seq(from = 0.15, to = 0.15, by = 0.05)
+
+effect_type  <- "r" # or r_z
+
+#Using Pearson's r as the effect size.
+true_tau <- c(0, 0.1, 0.15, 0.2)
+true_tau2 <- true_tau^2
+
+if(effect_type == "r_z"){
+    #Using Fisher's z as the effect size (see 'compute_tau_fisher_z.r')
+
+    true_tau_50 <- c(0.1031263, 0.1566007, 0.2123514) #For N = 50
+    true_tau_100 <- c(0.1020357, 0.1549445, 0.2101057) #For N = 100
+    true_tau_150 <- c(0.1016845, 0.1544113, 0.2093826) #For N = 150
+    true_tau_200 <- c(0.1015112, 0.1541480, 0.2090256) #For N = 200
+    true_tau <- c(0, true_tau_50, true_tau_100, true_tau_150, true_tau_200)
+    true_tau2 <- true_tau^2
+
+    mu <- atanh(mu)
+}
 
 
 cond <- expand.grid(sample_size = sample_size,
[…]

### 2023-05-12 code/plots.r
dropped old z-values from r-data

@@ -271,7 +271,9 @@ theme_bw()
 #1) Create all Pearson's r plots
 
 
-dat <- readRDS("data/means_r_z_over_vs_underestimate.RDS")
+dat_r <- readRDS("data/means_r_over_vs_underestimate.RDS")
+dat_z <- readRDS("data/means_z_over_vs_underestimate.RDS")
+dat <- c(dat_r, dat_z)
 dat <- lapply(dat, function(x) as.data.frame(as.list(x)))
 
 dat <- data.table::rbindlist(dat, idcol = "mu")
@@ -292,7 +294,7 @@ dat$reliability_sd <- grep_strip("reliability_sd = ", conditions)
 dat$true_tau2 <- as.numeric(grep_strip("true_tau2 =", conditions)) #numeric for plotting later
 dat$mean_reliability <- grep_strip("mean_rel = ", conditions)
 dat$mu <- as.numeric(grep_strip("mu = ", conditions))
-dat$effect_type <- grep_strip("effect_tpe = ", conditions) #nb type in name!
+dat$effect_type <- grep_strip("effect_type = ", conditions)
 
 # Add real truncated values
 trunc <- readRDS("data/truncated_tau.RDS")

### 2023-05-15 code/functions.r
add Hunter & Schmidt method to simulation

@@ -27,6 +27,7 @@ rnorm_truncated <- function(
 simulate_rma <- function(
 
     effect_type = c("r", "r_z"),
+    method = c("Hedges", "HS"),
     reliability_distribution = c("uniform", "normal"),
     k, #number of studies
     sample_size, #sample size, fixed across studies
@@ -103,8 +104,19 @@ simulate_rma <- function(
     }
 
     # fit meta-analysis on observed values
-    fit <- rma(yi = r_se_me, vi = r_var_estimate,
+    if(method == "Hedges"){
+        fit <- rma(yi = r_se_me, vi = r_var_estimate,
+                  control=list(stepadj=steplength, maxiter=maxiter))
+
+    else if(method == "HS"){ #https://www.metafor-project.org/doku.php/tips:hunter_schmidt_method
+        r_var_estimate <- escalc(measure="COR", ri=r_se_me,
+                      ni=rep(sample_size, k), vtype="AV")[["vi"]]
+
+        fit <- rma(yi = r_se_me, vi = r_var_estimate, weights = rep(sample_size, k),
               control=list(stepadj=steplength, maxiter=maxiter))
+
+    }
+
 
     data.frame(intercept_b = fit$b,
                intercept_p = fit$pval,

### 2023-05-15 code/supercomputer_simulator.r

@@ -26,6 +26,7 @@ reliability_mean <- c(0.6, 0.7, 0.8, 0.9, 1)
 reliability_sd <- seq(from = 0.15, to = 0.15, by = 0.05)
 
 effect_type  <- "r" # or r_z
+meta_method  <- "Hedges" #or HS
 
 #Using Pearson's r as the effect size.
 true_tau <- c(0, 0.1, 0.15, 0.2)
@@ -69,6 +70,7 @@ if(effect_type == "r_z"){
 
 #these below are higher level control input to the function
 cond$effect_type <- effect_type
+cond$method  <- meta_method
 cond$reliability_distribution <- "normal"
 cond$step_length  <-  0.5 #decrease from 1 to 0.5 to improve convergence for low N
 cond$maxiterations  <-  100 #default values are steplength = 1, and maxiter = 100
@@ -133,15 +135,16 @@ end - start
 #****************************************
 
 e <- lapply(out_list, data.table::rbindlist)
-e_means <- lapply(e, colMeans)
-#e_medians <- lapply(e, function(x) apply(x, 2, median))
 names(e) <- c(paste0("mu = ", cond$mu,
                      ";k = ", cond$k,
                      ";N = ", cond$sample_size,
                      ";reliability_sd = ", cond$reliability_sd,
                      ";mean_rel = ", cond$reliability_mean,
                      ";true_tau2 = ", cond$true_tau2,
-                     "; effect_type = ", cond$effect_type))
+                     ";effect_type = ", cond$effect_type))
 
-#lapply(e_means, round, 3)
-#saveRDS(e, "../data_new/over_vs_underestimate.RDS")
+e_means <- lapply(e, colMeans)
+#e_medians <- lapply(e, function(x) apply(x, 2, median))
+
+#saveRDS(e, "../data_new/raw_over_vs_underestimate.RDS")
+#saveRDS(e_means, "../data_new/means_over_vs_underestimate.RDS")

### 2023-05-15 code/functions.r
fixed minor error in last commit

@@ -108,7 +108,7 @@ simulate_rma <- function(
         fit <- rma(yi = r_se_me, vi = r_var_estimate,
                   control=list(stepadj=steplength, maxiter=maxiter))
 
-    else if(method == "HS"){ #https://www.metafor-project.org/doku.php/tips:hunter_schmidt_method
+    }else if(method == "HS"){ #https://www.metafor-project.org/doku.php/tips:hunter_schmidt_method
         r_var_estimate <- escalc(measure="COR", ri=r_se_me,
                       ni=rep(sample_size, k), vtype="AV")[["vi"]]
 

### 2023-05-15 code/supercomputer_simulator.r

@@ -141,7 +141,8 @@ names(e) <- c(paste0("mu = ", cond$mu,
                      ";reliability_sd = ", cond$reliability_sd,
                      ";mean_rel = ", cond$reliability_mean,
                      ";true_tau2 = ", cond$true_tau2,
-                     ";effect_type = ", cond$effect_type))
+                     ";effect_type = ", cond$effect_type,
+                     ";method = ", cond$method))
 
 e_means <- lapply(e, colMeans)
 #e_medians <- lapply(e, function(x) apply(x, 2, median))

### 2023-05-15 supplemental_data/brannik_raw-data-simulation.r
check other's het. estimates

@@ -1,34 +1,6 @@
 
 
-
-dat <- matrix(c("University_1","130","0.24","0.75",
-"University_2","90","0.11","0.75",
-"Private_1","30","0.05","0.60",
-"Private_2","25","0.17","0.60",
-"Volunteer_1","50","0.38","0.90",
-"Volunteer_2","65","0.50","0.90"), nrow = 6, byrow = TRUE)
-
-dat <- as.data.frame(dat)
-names(dat) <- c("study", "n", "r", "rel")
-
-dat$r <- as.numeric(dat$r)
-dat$rel <- as.numeric(dat$rel)
-dat$n <- as.numeric(dat$n)
-dat$r_var <- (1 - dat$r^2)^2 / (dat$n - 1)
-
-dat$rho <- dat$r / sqrt(dat$rel)
-dat$rho_var <- (1 - dat$rho^2)^2 / (dat$n - 1)
-
-dat$rho_var2 <- dat$r_var / sqrt(dat$rel)
-#estimate rho var based on observed r
-
-library(metafor)
-rma(yi = dat$r, vi = dat$r_var)
-rma(yi = dat$rho, vi = dat$rho_var)
-
-rma(yi = dat$rho, vi = dat$rho_var2)
-# If I use the variance based on the estimated r
-# I get smaller heterogeneity with true effect sizes!
+# Brannick et al., simulation
 
 # 2 ideas that might be the difference
 # a) add measurement error before sampling error
@@ -40,7 +12,6 @@ rma(yi = dat$rho, vi = dat$rho_var2)
 # code below and there is still an underestimate when using
 # meta-analyzing unreliabile data
 
-# Brannick et al., simulation
 TxTy <- rnorm(10, mean = .26, sd = .13)
 rel_x <- runif(10, min = 0, max = 1)
 rel_y <- runif(10, min = 0, max = 1)

### 2023-05-15 supplemental_data/checking-metafor-psychmeta-borenstein.r

@@ -1,0 +1,113 @@
+# Project: Unexplained heterogeneity
+# Code purpose: Checking other's corrections for unreliability
+# Author: Anton Olsson-Collentine (anton@olssoncollentine.com)
+
+
+
+
+# Metafor example
+# https://www.metafor-project.org/doku.php/tips:hunter_schmidt_method
+
+library(metafor)
+dat <- dat.mcdaniel1994
+dat <- escalc(measure="COR", ri=ri, ni=ni, data=dat, vtype="AV")
+res <- rma(yi, vi, weights=ni, data=dat, method="HS")
+res
+
+# NB! Wolfgang shows increase in heterogeneity after corrected
+# for unreliability!
+
+
+# Psych-meta example
+
+# Data from: https://cran.r-project.org/web/packages/psychmeta/vignettes/ma_r.html
+# But only for X and Y (not Z)
+dat <- matrix(
+    c("1", "1", "X", "Y", "416", "0.49", "0.79", "0.77", "Watson2005",
+    "2", "1", "X", "Y", "241", "0.54", "0.82", "0.84", "Watson2005",
+    "3", "1", "X", "Y", "479", "0.34", "0.73", "0.87", "Zellars2006",
+    "4", "1", "X", "Y", "167", "0.37", "0.78", "0.79", "Bandura1980"),
+ ncol = 9, byrow = TRUE)
+
+dat <- as.data.frame(dat)
+names(dat) <- c("sample_id", "moderator","x_name","y_name","n","rxyi", "rxxi","ryyi", "citekey")
+dat[,5] <- as.numeric(dat[,5])
+dat[,6] <- as.numeric(dat[,6])
+dat[,7] <- as.numeric(dat[,7])
+dat[,8] <- as.numeric(dat[,8])
+
+
+install.packages("psychmeta")
+library(psychmeta)
+
+# Without correction for unreliability
+fit1 <- ma_r(rxyi = rxyi,
+               n = n,
+               construct_x = x_name,
+               construct_y = y_name,
+               rxx = rxxi,
+               ryy = ryyi,
+               sample_id = sample_id,
+               data = dat
+               )
+
+summary(fit1)
+# tau  0.0842 (sd_res)
+
+# https://cran.r-project.org/web/packages/psychmeta/vignettes/overview.html+
+# section "Individual-correction meta-analyses"
+
+fit2 <- ma_r(ma_method = "ic",
+               rxyi = rxyi,
[…]

### 2023-05-15 manuscript/methods.qmd
minor polish of methods

@@
 
-We vary the grand mean $\theta$ from 0 to 0.6 in steps of 0.1 to cover all realistic effect sizes and explore the boundaries of the the interaction between effect size and reliability variance.
-For Fisher's $z$ we transformed these value into equivalent effect sizes at the Fisher's $z$ scale.
-To put these values into context we turn to the empirical estimates of typical correlational effect sizes provided by Nuijten et al., (REF) and Schäfer & Scharz (REF).
-Nuijten et al.
-performed a meta-meta-analysis on the fields of RESEARCH and reports a median meta-analytic effect size of 0.24 for predictive validity and correlational studies.
-This estimate is likely positively biased due to publication bias and selective reporting.
-Schäfer & Scharz report a median $r$ of 0.16 amongst preregistered research in FIELD, with an 'upper median' (i.e., the 83.35% quantile) of 0.41.
-This estimate is likely less affected by positive bias but should not be expected to be unbiased (REF Scheel).
-
-We defined the between-studies variance ($\tau^2$) to cover a wide range of variance in true effect sizes and empirical estimates reported in the psychological literature.
-Van Erp et al., (REF) provide empirical heterogeneity estimates from a large number of meta-analyses.
-The median (uncorrected)Pearson's $r$ $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24).
-As the point of this study is that empirical heterogeneity estimates may not be accurately estimated we selected a range of $\tau$ estimates for the product-moment that we considered reasonable and that also covered most of the empirical estimates from van Erp et al.
-These were {0, 0.1, 0.15, 0.2}.
-We selected the largest heterogeneity level to avoid excissive truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1.
-
-Fisher's z is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson's $r$ into corresponding $\tau$-values on the Fisher's $z$ scale.
-To be able to compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale.
-The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity (REF Higgins), an as such is measured from 0 - 100%.
-The $I^2$ index can be defined as $I^2 = \frac{\tau^2}{s^2 + \tau^2}$ where $s^2 = \frac{\Sigma w_i(k - 1)}{(\Sigma w_i)^2 - \Sigma w_i^2}$ and $w_i$ is the precision $w_i = 1 / \sigma^2$.
-This means that its value generally depends on both the number of studies $K$ and the sample size $N_i$ in these studies (REF: Higgins, I2 is not an absolute measure of heterogeneity).
-However, because we keep $N_i$ fixed across studies (see below), the value of $I^2$ depends only on $N = N_i$ and the heterogeneity.
-
-We thus computed the $I^2$ index for each sample size condition and $\tau$-values defined in Pearson's $r$, and then computed corresponding $\tau$-values on the Fisher's $z$ scale given these $I^2$-values and sample sizes.
-Two complications were that the value for $\sigma^2$ varies with effect size for Pearson's $r$ and that effect size varies across studies with $\tau$.
-We resolved this by setting $\theta = 0$ and computing the expected $\sigma^2$ given $\tau$ and sample size and then using this expected value to compute $I^2$.
-We used the 'law of the unconscious statistician' to compute the expected value of $\sigma^2$ which says that the expected value of a function $g(X)$ of a random variable can be expressed in terms of the probability distribution of X: $E[g(X)] = \int_{-\infty}^\infty g(x) f(x) dx$.
-In our case, $g(x)$ corresponds to the sampling variance of the product-moment correlation, $f(x) \sim N(\theta, \tau^2)$ and because the product-moment correlation is bounded at {-1, 1} we set $\infty$ instead to 0.999.
-The resulting $\tau$ values for fisher z are only minimally different from {0, 0.1, 0.15, 0.2}, but can be found in supplement XX.
-
-We ran our analyses across a range of sample sizes $N$ and number of studies $K$.
+We ran our analyses across a range of within-study sample sizes $N$ and number of studies $K$.
 Within meta-analyses we used a fixed sample size across all studies such that $N = N_i$.
@@
 
+We vary the grand mean $\theta$ from 0 to 0.6 in steps of 0.1 to cover all realistic effect sizes and explore the boundaries of the interaction between effect size and reliability variance.
+For Fisher's $z$ we transformed these value into equivalent effect sizes at the Fisher's $z$ scale.
+For context to  these values consider the empirical estimates of typical correlational effect sizes provided by Nuijten et al., (REF) and Schäfer & Scharz (REF).
+Nuijten et al.
+performed a meta-meta-analysis on the fields of RESEARCH and reports a median meta-analytic effect size of 0.24 for predictive validity and correlational studies.
+This estimate is likely positively biased due to publication bias and selective reporting.
+Schäfer & Scharz report a median $r$ of 0.16 amongst preregistered research in FIELD, with an 'upper median' (i.e., the 83.35% quantile) of 0.41.
+This estimate is likely less affected by positive bias but should not be expected to be unbiased (REF Scheel).
+
+We defined the between-studies standard deviation ($\tau$) to cover a wide range of variability in true effect sizes and empirical estimates reported in the psychological literature.
+Van Erp et al., (REF) provide empirical heterogeneity estimates from XX meta-analyses in XX article published in the journal Psychological Bulletine between 1990 - 2013.
+The median (uncorrected)Pearson's $r$ $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24).
+The point of this study is that empirical heterogeneity estimates may not be accurately estimated.
+Nonetheless, we selected approximately the interquartile range of $\tau$-values reported in van Erp et al.
+for the product-moment correlation, both to make sure we covered the empirically reported range of values and because we considered them reasonable variability on the product-moment correlation scale.
+The values we selected were {0, 0.1, 0.15, 0.2}.
+We selected the largest heterogeneity level to avoid excissive truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1.
+Our values are similar to those used by Brannick et al.
+(2019) in their simulation study on the performance of different heterogeneity estimators after corrections for unreliability (they used $\tau = $ 0, 0.08, 0.13, 0.2).
+
+Fisher's $z$ is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson's $r$ into corresponding $\tau$-values on the Fisher's $z$ scale.
+To be able to compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale.
+The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity (REF Higgins), and as such is measured from 0 - 100%.
+The $I^2$ index can be defined as $I^2 = \frac{\tau^2}{s^2 + \tau^2}$ where $s^2 = \frac{\Sigma w_i(k - 1)}{(\Sigma w_i)^2 - \Sigma w_i^2}$ and $w_i$ is the precision $w_i = 1 / \sigma^2$.
+This means that the $I^2$ value generally depends on both the number of studies $K$ and the sample size $N_i$ in these studies (REF: Higgins, I2 is not an absolute measure of heterogeneity).
+However, because we keep $N_i$ fixed across studies (see below), the value of $I^2$ depends only on $N = N_i$ and the heterogeneity.
+
+We thus computed the $I^2$ index for each sample size condition and $\tau$-values defined in Pearson's $r$, and then computed corresponding $\tau$-values on the Fisher's $z$ scale given these $I^2$-values and sample sizes.
+Two complications were that the value for $\sigma^2$ varies with effect size for Pearson's $r$ and that effect size varies across studies because of $\tau$.
+We resolved this by setting $\theta = 0$ and computing the expected $\sigma^2$ given $\tau$ and sample size and then using this expected value to compute $I^2$.
+We used the 'law of the unconscious statistician' to compute the expected value of $\sigma^2$, which says that the expected value of a function $g(X)$ of a random variable can be expressed in terms of the probability distribution of X: $E[g(X)] = \int_{-\infty}^\infty g(x) f(x) dx$.
+In our case, $g(x)$ corresponds to the sampling variance of the product-moment correlation, $f(x) \sim N(\theta, \tau^2)$ and because the product-moment correlation is bounded at {-1, 1} we set $\infty$ instead to 0.999.
+The resulting $\tau$ values for Fisher $z$ are only minimally different from {0, 0.1, 0.15, 0.2}.
+For example, for $N = 150$ the corresponding Fisher $z$ $\tau$-values were {VALUES} and as $N$ increased heterogeneity values were more similar (see supplement XX).
+
 Finally, we defined average reliability levels and their standard deviation based on empirical estimates from the literature.

### 2023-05-15 code/plots.r
added fisher z plot

@@ -273,6 +273,7 @@ theme_bw()
 
 dat_r <- readRDS("data/means_r_over_vs_underestimate.RDS")
 dat_z <- readRDS("data/means_z_over_vs_underestimate.RDS")
+
 dat <- c(dat_r, dat_z)
 dat <- lapply(dat, function(x) as.data.frame(as.list(x)))
 
@@ -370,3 +371,46 @@ facet_grid(N~true_tau, scales = "free") +
 theme_bw()
 
 ggsave("figures/z_k_20_0.6-0.9_tau_sd_0.15.png", width = 8.62, height = 9.93)
+
+
+
+#****************************************
+# supplemental plots + fisher's z
+#****************************************
+dat_20_150 <- dat[dat$k == 20 & dat$N == 150,]
+
+# For fisher's z there is no truncation,
+# hack the trunc dataframe for the plot
+trunc$effect_type  <- "Pearson's r"
+trunc <- rbind(trunc, trunc)
+trunc[29:56, "effect_type"] <- "Fisher's z"
+
+dat_z <- dat_20_150[dat_20_150$effect_type == "r_z"]
+trunc$tau_hat[29:56] <- rep(unique(dat_z$true_tau), each = 7)
+
+dat_20_150$true_tau <- ifelse(dat_20_150$effect_type == "r_z",
+                              trunc(dat_20_150$true_tau *100) / 100,
+                              dat_20_150$true_tau)
+
+
+# What I've done in the above to lines of code is to
+# 1) put the 'truncated' true_tau for r_z to fisher z nominal value
+# 2) put the 'true tau' in the main data object to the nominal r value
+# This helps with the facet_grid to make the plot look nice
+# But changes the interpretation of the black line for fisher z
+# Does it? It is still the true heterogeneity
+
+# More plot prep
+dat_20_150$effect_type <- ifelse(dat_20_150$effect_type == "r_z",
+                                "Fisher's z",
+                                "Pearson's r")
+
+
+ggplot(dat_20_150, aes(x = mu, y = tau_hat)) +
+geom_line(aes(linetype = mean_reliability), show.legend = TRUE) +
+geom_line(data = trunc, linetype = 1) +
+scale_linetype_manual(values = 2:5) +
+guides(linetype = guide_legend(reverse = TRUE)) +
+expand_limits(y = 0) +
+facet_grid(effect_type~true_tau, scales = "free", switch = "y") +
+theme_bw()

### 2023-05-15 code/plots.r
removed old plotting code

@@ -16,251 +16,6 @@
 library(ggplot2)
 
 #****************************************
-# Plots 1 - 4
-#****************************************
-
-dat <- readRDS("data/reliability_overestimate.RDS")
-dat <- lapply(dat, function(x) as.data.frame(as.list(x)))
-
-dat <- data.table::rbindlist(dat, idcol = "mu")
-
-dat$reliability_sd <- gsub(".*& ", "", dat$mu)
-dat$reliability_sd <- gsub(".*= ", "", dat$reliability_sd)
-dat$mu <- gsub("&.*", "", dat$mu)
-dat$mu <- as.numeric(gsub("mu = ", "", dat$mu))
-
-
-ggplot(data = dat) +
-geom_line(aes(x = mu, y = sqrt(tau2_hat), color = reliability_sd))
-
-#ggsave("figures/plot1.png", width = 8.62, height = 9.93)
-
-
-# Plot 2 original
-
-# line_labels <- dat[dat$r == "0.6",]
-
-# ggplot(dat,aes(x = r, y = percent_tau2)) +
-# geom_line(aes(linetype = reliability), show.legend = FALSE) +
-# scale_linetype_manual(values = 8:1) +
-# geom_label(data = line_labels, aes(x = r, y = percent_tau2, group = reliability), nudge_x = 0.025,
-# label = line_labels$reliability) +
-# #annotate("text", x = 0.885, y = 0.086, label = "Reliability", hjust = 0) +
-# geom_hline(yintercept = 100, linetype = "dashed") +
-# theme_bw()
-
-# ggsave("figures/plot2.png", width = 8.62, height = 9.93)
-
-## Plot 2 v2
-# That is, with perfect reliability and tau2-values of
-# c(0.002, 0.0035, 0.0055, 0.0085, 0.012, 0.0185, 0.031, 0.069)
-# (i.e, 20 - 90% I2)
-
-dat <- readRDS("data/k_20_over_vs_underestimate.RDS")
-dat <- lapply(dat, function(x) as.data.frame(as.list(x)))
-
-dat <- data.table::rbindlist(dat, idcol = "mu")
-
-#create condition ids
-conditions <- unlist(strsplit(dat$mu, split = ";"))
-
-dat$true_tau2 <- grep("true_tau2 =", conditions, value = TRUE)
-dat$true_tau2 <- as.numeric(gsub("true_tau2 = ", "", dat$true_tau2))
[…]

### 2023-05-16 code/plots.r
Add hunter schmidt data and plot

@@ -4,11 +4,39 @@
 
 
 #****************************************
-# Simulating data for plots
+# functions
 #****************************************
-# see supercomputer_simulator.r
+grep_strip <- function(string, vec){
+    #Uses a common string like "k = " to extract "k = 50" or "k = 200"
+    #Then removes the "k = " part and returns as character
+    a <- grep(string, vec, value = TRUE)
+    gsub(string, "", a)
+}
 
+create_condition_cols <- function(data){
 
+    #takes a data frame as input with a column named "mu"
+    #in which all conditions are separated by a semicolon
+    # Then creates new columns named after each condition
+
+    #split all conditions into a character vector
+    conditions <- unlist(strsplit(data$mu, split = ";"))
+    #select all unique condition factors (e.g., mu, k, N)
+    category <- unique(gsub("=.*", "= ", conditions))
+    # Remove equal sign to column names (kept above in case some condition name include 'k')
+    col_name <- gsub(" = ", "", category)
+
+    for(g in 1:length(category)){
+
+        #This loop is equivalent to the line below, but applied to all conditions
+        # data$k <- grep_strip("k = ", conditions)
+
+        data[[col_name[g]]] <- grep_strip(category[g], conditions)
+
+    }
+
+    data #out
+}
 #****************************************
 # plotting
 #****************************************
@@ -16,7 +44,7 @@
 library(ggplot2)
 
 #****************************************
-# supplemental plots + fisher's z
+# General plot prep
 #****************************************
 
 # Plots with many different N and K
@@ -28,29 +56,16 @@ library(ggplot2)
 
 dat_r <- readRDS("data/means_r_over_vs_underestimate.RDS")
 dat_z <- readRDS("data/means_z_over_vs_underestimate.RDS")
+dat_hs <- readRDS("data/means_r_HS_over_vs_underestimate.RDS")
 
-dat <- c(dat_r, dat_z)
+dat <- c(dat_r, dat_z, dat_hs)
[…]

### 2023-05-16 supplemental_data/checking-metafor-psychmeta-borenstein.r
folder cleanup

@@ -84,8 +84,7 @@ dat <- matrix(c("University_1","130","0.24","0.75",
 
 dat <- as.data.frame(dat)
 names(dat) <- c("study", "n", "r", "ryy")
-dat <- cbind(dat, 1)
-names(dat)[5] <- "rxx"
+dat$rxx <-  1
 
 dat$r <- as.numeric(dat$r)
 dat$ryy <- as.numeric(dat$ryy)

### 2023-05-16 code/plots.r
updated code for z-r-hs plot

@@ -53,7 +53,6 @@ library(ggplot2)
 
 #1) Create all Pearson's r plots
 
-
 dat_r <- readRDS("data/means_r_over_vs_underestimate.RDS")
 dat_z <- readRDS("data/means_z_over_vs_underestimate.RDS")
 dat_hs <- readRDS("data/means_r_HS_over_vs_underestimate.RDS")
@@ -66,18 +65,16 @@ dat <- data.table::rbindlist(dat, idcol = "mu")
 #create condition ids
 dat <- create_condition_cols(dat)
 
-
 # Add real truncated values
 trunc <- readRDS("data/truncated_tau.RDS")
-trunc$true_tau <- trunc$nominal_tau
-trunc$tau_hat <- trunc$tau
+trunc$true_tau <- trunc$tau
 
 #Prepping plot
 dat$true_tau2 <- as.numeric(dat$true_tau2)
 dat$mu <- as.numeric(dat$mu)
 
 dat$tau_hat <- sqrt(dat$tau2_hat)
-dat$true_tau <- sqrt(dat$true_tau2)
+dat$nominal_tau <- sqrt(dat$true_tau2)
 
 dat$mean_rel <- factor(dat$mean_rel)
 dat$k <- factor(dat$k, levels = c("5", "20", "40", "200"))
@@ -90,40 +87,66 @@ dat$N <- factor(dat$N, levels = c("50", "100", "150", "200"))
 dat_20_150 <- dat[dat$k == 20 & dat$N == 150,]
 
 # For fisher's z there is no truncation,
-# hack the trunc dataframe for the plot
+# create dataframe with all true heterogeneity levels
 trunc$effect_type  <- "Pearson's r"
-trunc <- rbind(trunc, trunc)
-trunc[29:56, "effect_type"] <- "Fisher's z"
+hs_true <- trunc
+hs_true$effect_type <- "Hunter & Schmidt" #effect type is r, but named this way for plotting
 
+z_true <- trunc
+z_true$effect_type <- "Fisher's z"
 dat_z <- dat_20_150[dat_20_150$effect_type == "r_z"]
-trunc$tau_hat[29:56] <- rep(unique(dat_z$true_tau), each = 7)
+z_tau <- rep(unique(dat_z$nominal_tau), each = 7) #4 tau-values across 4*7 conditions
+z_true$true_tau <- z_tau
+#NB! the z_true object has mu expressed in Pearson's r, not fisher z (this is useful for plotting later)
 
[…]

### 2023-05-17 manuscript/methods.qmd
added model to methods

@@
 All code and data for this project are available at (OSF LINK).
-For our simulations and analyses we used R (REF) and we used the R package 'metafor' (REF) to perform meta-analyses and estimate heterogeneity, For parallel computing we took advantage of the R packages 'parallel' (REF) and 'parabar' (REF).
+For our simulations and analyses we used R (REF) and we used the R package 'metafor' (REF) to perform meta-analyses and estimate heterogeneity.
+For parallel computing we took advantage of the R packages 'parallel' (REF) and 'parabar' (REF).
 
@@
 We use correlations as our effect size type in these simulations because correcting them for unreliability is straightforward and because they are the effect size type most commonly corrected in practice.
-However, unreliability attenuates all types of effect sizes and all can be corrected, either directly (standardized mean differences; Wiernick & Dahlke, 2020) or by first transforming them into correlations (with associated assumptions).
+{see also argument in Field that all effect size types are pretty similar} However, unreliability attenuates all types of effect sizes and all can be corrected, either directly (standardized mean differences; Wiernick & Dahlke, 2020) or by first transforming them into correlations (with associated assumptions).
 
@@
 We perform simulations both on product-moment correlations and Fisher z transformed correlations.
-For our purpose of estimating heterogeneity, Fisher z correlations have several advantages: 1) their sampling variance is not dependent on the size of the correlation, 2) they are not bounded to {-1, 1}.
-However, they also have two disadvantages: 1) that the heterogeneity estimate is measure on the Fisher z scale, which makes it hard to interpret.
+For our purpose of estimating heterogeneity, Fisher z correlations have two advantages: 1) their sampling variance is not dependent on the size of the correlation, 2) they are not bounded to {-1, 1}.
+However, they also have two disadvantages: 1) that the heterogeneity estimate is measured on the Fisher z scale, which makes it hard to interpret.
 A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (REF) for an indirect estimate of the heterogeneity.
@@
 Given the common use of both effect size types and their mix of advantages and disadvantages we report results for both.
+
+## Meta-analytic model
+
+We apply two meta-analytic models, the Hedges and Vevea (1998) random-effects  model and the Hunter and Schmidt (2015) 'bare-bones' random-effects model.
+We implement these two models as a sensitivity analysis rather than because of an interest in comparing the two in detail.
+The Hedges and Vevea (HV) model uses inverse variance weight computed as $w_i = 1 / (V_i + T^2)$ (e.g., Borenstein et al., 2010) where $V_i$ is the sampling variance of study $i$ and $T^2$ is the estimated between-studies variance.
+The Hunter and Schmidt (HS) model weighs studies by sample size $w_i = N_i$.
+The differences between these weights are relatively minor (Borenstein et al., 2010; Hunter and Schmidt, 2015, p.
+XX).
+Although the HV model is typically applied to Fisher z transformed correlations (e.g., Borenstein et al., p.
+41-43), we apply it to both transformed and untransformed correlations for the reasons described in the previous section.
+The HS model typically includes correcting for unreliability or other measurement artifacts.
+However, we implement it without corrections ('bare-bones'; Hunter & Schmidt, 2015) because this allows us to examine the bias in heterogeneity estimates compared to a perfectly implemented correction of unreliability.
+
+Estimating heterogeneity in both the HV and bare-bones HS model essentially consists of comparing the observed variance in effect sizes with what would be expected from sampling error alone and then ascribing any excess observed variance to heterogeneity (Borenstein et al., 2010, Box 1; Hunter & Schmidt, 2015, p.
+XX).
+However, the process for doing so differs somewhat between methods.
+We apply the Restricted Maximum Likelihood (REML) estimator of heterogeneity when estimating heterogeneity in the Hedges and Vevea (1998) model, because it is a heterogeneity estimator that performs well and is widely implemented (e.g., Viechtbauer, 2005; REF2), and is also the default estimator in the R package metafor.
+Describing this REML estimation in detail is beyond the scope of this paper and we refer interested readers to the description in Viechtbauer (2005) {this is also what Brannick et al do}.
+The Hunter and Schmidt (2015) model estimates heterogeneity $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $ \frac{\Sigma[N_i \sigma^2_{ei}]}{\Sigma N_i}$ (p.
+87, 2nd edition) and $\sigma^2_{ei}$ is the usual sampling variance for the Pearson correlation but with $\rho$ replaced with the sample size weighted average correlation $\bar{r} = \frac{\Sigma n_i r_i}{\Sigma n_i}$ instead of $r_i$ as is common.
+A generalized version of the HS estimator for effect sizes other than correlations can be found in Viechtbauer (2005; 2015) (see ?rma).
 
@@
 Fisher's $z$ we transformed to product-moment correlations before adding measurement error and then back-transformed into Fisher's $z$ before the next step.
-At this point we have $k$ effect sizes $r_i$ in Pearson's $r$ or Fisher's $z$ with sampling error and measurement error, ready for meta-analysis.
-
-We meta-analyzed the $k$ effect sizes using a random effects model with inverse variance weights and the Restricted Maximum Likelihood (REML) estimator of heterogeneity.
-We choose the REML estimator because it is a heterogeneity estimator that performs well and is widely implemented (e.g., REF, REF2), but also show our results using the classic DerSimonian-Laird (DL) estimator in supplement XX.
-We used inverse variance weights due to their widespread use, but as we used a fixed sample size across studies within a meta-analysis (see next section) the weights are of little importance.
-As a sensitivity analysis, we also implemented the Hunter and Schmidt (2015) model which uses a different heterogeneity estimator and found qualitatively similar results (supplement XX).
-The procedure described in this section was replicated 10,000 times for each combination of parameter values.
+At this point we have $k$ effect sizes $r_i$ in Pearson's $r$ or Fisher's $z$ with sampling error and measurement error.
+We then applied either the HV meta-analytic model or the HS model, as described in the section 'the meta-analytic model'.
+The procedure described in the current section was replicated 10,000 times for each combination of parameter values.
 

### 2023-05-17 code/functions.r
analytical solution variance of truncated normal

@@ -163,3 +163,25 @@ simulate_rma_I2 <- function(effect_type = c("r", "r_z"), k, sample_size, true_ta
                tau2_hat = fit$tau2,
                tau2_p = fit$QEp)
 }
+
+
+# Compute variance of a truncated normal distribution
+compute_var_truncated <- function(mu, nominal_tau, a = -1, b = 1){
+# Formula from wikipedia (double-check): https://en.wikipedia.org/wiki/Truncated_normal_distribution
+
+    #mu = mean of distribution before truncation
+    #nominal_tau = standard deviation of distribution before truncation
+    #a  = lower bound of truncation
+    #b = upper bound of truncation
+
+    alpha <- (a - mu) / nominal_tau
+    beta <-  (b - mu) / nominal_tau
+
+    denom <- pnorm(beta) - pnorm(alpha)
+    num1 <- beta*dnorm(beta) - alpha*dnorm(alpha)
+    num2 <- dnorm(beta) - dnorm(alpha)
+    true_tau2 <- nominal_tau^2 * (1 - num1/denom - (num2 / denom)^2)
+
+    sqrt(true_tau2) #out true_tau
+
+}

### 2023-05-17 code/functions.r
implemented analytical truncation

@@ -164,6 +164,41 @@ simulate_rma_I2 <- function(effect_type = c("r", "r_z"), k, sample_size, true_ta
                tau2_p = fit$QEp)
 }
 
+#****************************************
+# Plot prep functions
+#****************************************
+
+grep_strip <- function(string, vec){
+    #Uses a common string like "k = " to extract "k = 50" or "k = 200"
+    #Then removes the "k = " part and returns as character
+    a <- grep(string, vec, value = TRUE)
+    gsub(string, "", a)
+}
+
+create_condition_cols <- function(data){
+
+    #takes a data frame as input with a column named "mu"
+    #in which all conditions are separated by a semicolon
+    # Then creates new columns named after each condition
+
+    #split all conditions into a character vector
+    conditions <- unlist(strsplit(data$mu, split = ";"))
+    #select all unique condition factors (e.g., mu, k, N)
+    category <- unique(gsub("=.*", "= ", conditions))
+    # Remove equal sign to column names (kept above in case some condition name include 'k')
+    col_name <- gsub(" = ", "", category)
+
+    for(g in 1:length(category)){
+
+        #This loop is equivalent to the line below, but applied to all conditions
+        # data$k <- grep_strip("k = ", conditions)
+
+        data[[col_name[g]]] <- grep_strip(category[g], conditions)
+
+    }
+
+    data #out
+}
 
 # Compute variance of a truncated normal distribution
 compute_var_truncated <- function(mu, nominal_tau, a = -1, b = 1){
@@ -173,6 +208,10 @@ compute_var_truncated <- function(mu, nominal_tau, a = -1, b = 1){
     #nominal_tau = standard deviation of distribution before truncation
     #a  = lower bound of truncation
     #b = upper bound of truncation
+
+    if(nominal_tau == 0){
+        return(rep(0, length(mu)))
+    }
 
     alpha <- (a - mu) / nominal_tau
     beta <-  (b - mu) / nominal_tau

### 2023-05-17 code/plots.r

@@ -6,37 +6,9 @@
 #****************************************
 # functions
 #****************************************
-grep_strip <- function(string, vec){
-    #Uses a common string like "k = " to extract "k = 50" or "k = 200"
-    #Then removes the "k = " part and returns as character
-    a <- grep(string, vec, value = TRUE)
-    gsub(string, "", a)
-}
 
-create_condition_cols <- function(data){
 
-    #takes a data frame as input with a column named "mu"
-    #in which all conditions are separated by a semicolon
-    # Then creates new columns named after each condition
-
-    #split all conditions into a character vector
-    conditions <- unlist(strsplit(data$mu, split = ";"))
-    #select all unique condition factors (e.g., mu, k, N)
-    category <- unique(gsub("=.*", "= ", conditions))
-    # Remove equal sign to column names (kept above in case some condition name include 'k')
-    col_name <- gsub(" = ", "", category)
-
-    for(g in 1:length(category)){
-
-        #This loop is equivalent to the line below, but applied to all conditions
-        # data$k <- grep_strip("k = ", conditions)
-
-        data[[col_name[g]]] <- grep_strip(category[g], conditions)
-
-    }
-
-    data #out
-}
+source("./code/functions.r") # for function compute_var_truncated
 #****************************************
 # plotting
 #****************************************
@@ -47,11 +19,7 @@ library(ggplot2)
 # General plot prep
 #****************************************
 
-# Plots with many different N and K
-# Either for Pearson's r or for Fisher's z
-# Tau and mu are equal in both cases
 
-#1) Create all Pearson's r plots
 
 dat_r <- readRDS("data/means_r_over_vs_underestimate.RDS")
 dat_z <- readRDS("data/means_z_over_vs_underestimate.RDS")
@@ -66,8 +34,15 @@ dat <- data.table::rbindlist(dat, idcol = "mu")
 dat <- create_condition_cols(dat)
 
 # Add real truncated values
-trunc <- readRDS("data/truncated_tau.RDS")
-trunc$true_tau <- trunc$tau
+es <- seq(0, 0.6, 0.1)
+tau <- c(0, 0.1, 0.15, 0.2)
+
+tau_trunc <- lapply(tau, compute_var_truncated, mu = es)
+tau_trunc <- unlist(tau_trunc)
+trunc <- data.frame(mu = es,
[…]

### 2023-05-17 code/truncated_tau.r

@@ -1,19 +1,0 @@
-
-source("./code/functions.r")
-es <- seq(0, 0.6, 0.1)
-tau <- c(0, 0.1, 0.15, 0.2)
-
-a <- lapply(tau, function(x){
-
-lapply(es, function(y) {
-res <- rnorm_truncated(n = 1e6, mean = y, sd = x,
-lower_bound = -1, upper_bound = 1)
-
-data.frame(tau = sd(res), mu = y, nominal_tau = x)
-})
-
-})
-
-b <- lapply(a, function(x) do.call(rbind, x))
-b <- do.call(rbind, b)
-saveRDS(b, "data/truncated_tau.RDS")

### 2023-05-17 code/functions.r
moved all functions to functions.r file

@@ -3,7 +3,7 @@
 # code: Anton Olsson-Collentine
 
 #****************************************
-# Functions
+# Simulation functions
 #****************************************
 
 rnorm_truncated <- function(
@@ -125,44 +125,6 @@ simulate_rma <- function(
 }
 
 
-# The below function is an adapted version of above, specifically for
-# estimating tau2 values that correspond to I2 levels
-# Hence, it assumes perfect reliability has adapted output
-
-simulate_rma_I2 <- function(effect_type = c("r", "r_z"), k, sample_size, true_tau2, mu){
-    #adjusted input and output compared to my primary function
-    #Also, only for perfect reliability, i.e., uses sampling_var_rho to estimate meta-analysis
-
- # Draw rho from mean rho and compute sampling variances
-
-    if(effect_type == "r") {
-        rho <- rnorm_truncated(n = k, mean = mu, sd = sqrt(true_tau2),
-                              lower_bound = -1, upper_bound = 1)
-        sampling_var_rho  <- (1-rho^2)^2 / (sample_size -1)
-
-        # given study rho, draw sample rho
-        r_se <- rnorm_truncated(k, mean = rho, sd = sqrt(sampling_var_rho),
-                                lower_bound = -1, upper_bound = 1)
-
-    }else if(effect_type == "r_z") { #fisher's z
-        rho <- rnorm(n = k, mean = mu, sd = sqrt(true_tau2))
-        sampling_var_rho  <- 1 / (sample_size - 3)
-
-    # given study rho, draw sample rho
-    r_se <- rnorm(k, mean = rho, sd = sqrt(sampling_var_rho))
-
-    }else{
-        stop("specify effect type")
-    }
-
-
-    fit <- rma(yi = r_se, vi = sampling_var_rho)
-
-    data.frame(I2 = fit$I2,
-               true_tau2 = true_tau2,
-               tau2_hat = fit$tau2,
-               tau2_p = fit$QEp)
-}
 
 #****************************************
 # Plot prep functions

### 2023-05-17 code/plots.r

@@ -4,16 +4,12 @@
 
 
 #****************************************
-# functions
+# functions and packages
 #****************************************
 
+source("./code/functions.r") # for plot prep functions
 
-source("./code/functions.r") # for function compute_var_truncated
-#****************************************
-# plotting
-#****************************************
-
-library(ggplot2)
+library(ggplot2) #for plotting
 
 #****************************************
 # General plot prep

### 2023-05-19 manuscript/introduction.qmd
improved first 2 paragraphs of introduction

@@
 
-Estimating heterogeneity of effect sizes is usually considered the second main purpose of meta-analysis, after estimating an average effect size.
+Estimating heterogeneity of effect sizes is usually considered the main purpose of meta-analysis in addition to estimating an average effect size.
 Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size's sensitivity to variability in study design, or more specifically to sensitivity to differences in four study factors: 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables (e.g., Campbell & Stanley, 2015).
-From a more statistical perspective, heterogeneity is defined as the variance between 'true' subpopulation effect sizes.
+Heterogeneity is a primary outcome in meta-analysis because its presence signals the existence of moderators and so can be seen as an opportunity for theoretical development (Simons et al., 2017), because it affects the implementation of research by indicating that an intervention may not be equally effective under all conditions or for all people, and because its presence changes the interpretation of the average effect sized derived in meta-analysis from the true population effect size (under homogeneity) to the average of true subpopulation effect sizes (under heterogeneity).
+As such, estimating heterogeneity without bias is of central importance to meta-analysis (for a more extensive discussion on the importance of heterogeneity, see Olsson Collentine et al., 2020; Simons, 2017).
+
+Meta-analytic estimates are highly dependent on the quality of the data in the primary studies.
+Unfortunately, there is concern that researchers in psychology tend to neglect proper measurement (ref measurement-schmesurement) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021).
+Such inattention to measurement can bias heterogeneity estimates.
+From a statistical perspective, heterogeneity is defined as the variance between 'true' subpopulation effect sizes.
 However, true subpopulation effect sizes can refer to two different entities.
 Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as measurement error, restriction of range, and dichotomization; Hunter & Schmidt, 2015).
-Of these measurement artifacts, measurement error is present in any empirical effect size estimate and systematically attenuates effect sizes.
-It is usually measured as reliability (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., REF) and is relatively common to report in primary studies in psychology (Flake et al., 2017).
+Of these measurement artifacts, measurement error (imperfect reliability of measurement) is present in any empirical effect size estimate and systematically attenuates effect sizes.
 However, whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for imperfect reliability is implicit only in the models of Hunter & Schmidt (2015).
@@
 Similarly, Schmidt (2010) report that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts.
-The exception is the subfield of Industrial-Organizational studies.
-Although at the meta-analysis level rather than study level, Cortina (2003) showed that about 80% of the 1,294 meta-analyses published across 59 meta-analytic studies in the Journal of Applied Psychology between 1978 - 1997 corrected for measurement artifacts, and Anguinis (2011) that around 50% of meta-analyses identified in studies published across several management journal between 1989 - 2009 corrected for unreliability of measurement.
+The exception is the subfield of Industrial-Organizational studies where corrections tend to be more common (e.g., Cortina, 2003; Anguinis, 2011).
 
@@
 Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas).
-Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2, REF3), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
+Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2) due to factors such as publication bias (REF1) and _p_-hacking (REF2), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
 On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures.

### 2023-05-20 manuscript/introduction.qmd
intro: improved first 3 paragraphs

@@
 
-Meta-analytic estimates are highly dependent on the quality of the data in the primary studies.
-Unfortunately, there is concern that researchers in psychology tend to neglect proper measurement (ref measurement-schmesurement) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021).
-Such inattention to measurement can bias heterogeneity estimates.
-From a statistical perspective, heterogeneity is defined as the variance between 'true' subpopulation effect sizes.
-However, true subpopulation effect sizes can refer to two different entities.
-Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as measurement error, restriction of range, and dichotomization; Hunter & Schmidt, 2015).
-Of these measurement artifacts, measurement error (imperfect reliability of measurement) is present in any empirical effect size estimate and systematically attenuates effect sizes.
-However, whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for imperfect reliability is implicit only in the models of Hunter & Schmidt (2015).
-This has consequences for heterogeneity estimates reported in many areas of psychology and in this article we use Monte-Carlo simulations to explore the consequences for heterogeneity estimates in meta-analysis from not correcting for unreliability.
+Meta-analytic estimates are highly dependent on the quality of the data in the primary studies that the method summarizes.
+However, there is concern that researchers in psychology tend to neglect proper measurement (ref measurement-schmesurement) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021).
+For example, Flake et al.
+(2017) report that although latent constructs are widely examined in personality and social psychology, researchers barely report any of the validity evidence required to ascertain the extent to which implemented scales measure the constructs of interest.
+Such inattention to measurement can also bias heterogeneity estimates.
+We will focus here on reliability of measurement (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., REF), which is the only measurement validity evidence researchers report consistently (Flake et al., 2017), and which can be corrected for in meta-analysis (Hunter & Schmidt; 2015).
+
+Measurement error (imperfect reliability) is present in any empirical effect size estimate and systematically attenuates observed effect sizes compared to true underlying effect sizes.
+This affects heterogeneity estimates, which from a statistical perspective estimate the variance between true effect sizes.
+However, reliability is not necessarily accounted for in heterogeneity estimates, because ‘true’ underlying effect sizes can refer to two different entities.
+Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as imperfect reliability, restriction of range, and dichotomization; Hunter & Schmidt, 2015).
+Whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for measurement artifacts is implicit only in the models of Hunter & Schmidt (2015).
+This has consequences for heterogeneity estimates reported in many areas of psychology.
+Inattention to measurement reliability and how it affects heterogeneity estimates can lead researchers to misinterpret their meta-analytic average effect size estimate, ignore the presence of moderators or search for (and discover) non-existent moderators, and to implement research interventions inappropriately.
 
@@
 From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015).
-In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses should be aware of how this decision affects heterogeneity estimates.
+In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses need to be aware of how this decision affects heterogeneity estimates.
 

### 2023-05-20 manuscript/introduction.qmd
intro: dropped last paragraph of no-one talking about it

@@
 
-Imperfect reliability in primary studies imbues heterogeneity estimates with both a positive and a negative bias.
-The positive bias is discussed by several highly cited meta-analytic handbooks, most prominently by Hunter and Schmidt (2015; 12,235 citations on google scholar as of 2023-05-11) who introduced the concept of psychometric meta-analysis (i.e., meta-analysis corrected for measurement artifacts such as unreliability), but also by Borenstein et al.
-(2009, p.
-342; 19,020 citations), Card (2015, p.126; 1,716 citations) as well as in recent literature (Wiernick and Dahlke, 2020).
-However, there appears to be a lack of acknowledgement that imperfect reliability can also lead to a negative bias in heterogeneity estimates.
-The only one of the above sources to refer to a negative bias is Hunter and Schmidt who mention it shortly in their 2nd edition of *Methods of Meta-analysis* (2004, p.
-142) in a paragraph later absent from the 3rd edition (Hunter & Schmidt, 2015).
-Below we detail the argument for both a positive and negative bias in heterogeneity estimates.
-
 ## Bias in heterogeneity estimates due to unreliability
@@
 343).
+Imperfect reliability in primary studies imbues heterogeneity estimates with both a positive and a negative bias.
 

### 2023-05-20 manuscript/introduction.qmd
intro: updated bias explanations

@@
 
-A positive bias can be expected in heterogeneity estimates of uncorrected effect sizes due to the variance in reliability between primary studies.
+A positive bias can be expected in heterogeneity estimates of uncorrected effect sizes due to the variability in reliability between primary studies.
 This is because the attenuation in effect size will then differ from study to study, which results in differences across studies in observed effect sizes beyond their true variability.
-This is most easily seen by assuming infinite sample size (i.e., no within-study variance) and in the context of a fixed effect model (i.e., with a common underlying mean across studies and no true heterogeneity).
-Table 1 illustrates how under these conditions three studies with differing reliability will observe variability in effect sizes beyond their true variability (i.e., zero).
-Table 1 also shows how this effect increases as the true effect size increases.
+This is most easily seen by assuming a common true effect size $\mu$ across studies (i.e., a fixed effect and no true heterogeneity) and infinite sample size within studies.
+Any observed effect size $\hat{\mu}$ can then be seen as the true effect size $\mu$ multiplied by a study’s reliability $R_i$ (assuming $R_i = R_{xx’} = R_{yy’}$).
+Hence, the distribution of observed effect sizes is the true effect size $\mu$ multiplied by the distribution of reliabilities across studies.
+That is, $\hat{mu} \sim \mu \times N(\bar{R}, \sigma_R) = N(\mu \bar{R}, \mu \sigma_R$, where $\bar{R}$ is the average reliability across studies and $\sigma_R$ its standard deviation.
+Because reliability lies between 0 – 1, under these conditions imperfect reliability implies that the observed average effect size will be less than the true effect size by a factor $\bar{R}$ and estimated heterogeneity will be larger than true heterogeneity (i.e., zero) by the product of $\mu \times \sigma_R$ (if heterogeneity is expressed as standard deviation rather than variance).
+That is, variability in reliability across studies leads to a positive bias in heterogeneity that depends on the size of the true effect size.
+Table 1 illustrates this effect for three studies with differing reliability and zero true heterogeneity.
 
@@
 
-A negative bias can be expected in heterogeneity estimates of uncorrected effect sizes because given the same unreliability, the attenuation in absolute value increases with effect size.
+A negative bias can be expected in heterogeneity estimates of uncorrected effect sizes because in absolute terms larger effect sizes are more attenuated by a given reliability.
 That is, although the attenuation formula $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ implies that any effect size $\rho_{xy}$ will be attenuated by the same proportion given the same reliabilities $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$, this proportion corresponds to a larger absolute value for larger effect sizes.
-As a consequence, in the presence of heterogeneity larger true effect sizes will move further towards zero than smaller ones, decreasing heterogeneity in observed effect sizes.
+Consequently, in the presence of heterogeneity attenuation will move larger true effect sizes further towards zero than smaller ones, decreasing heterogeneity in observed effect sizes.
+This negative bias can be observed directly by assuming that an average effect size $\theta$ has some heterogeneity $\tau^2$, all studies have infinite sample size (i.e., no within-study sampling variance), and equal reliability ($R = R_i = R_{xx'} = R_{yy'}$).
+Any observed effect size $\hat{\theta_i}$ can then be seen as a study’s true effect size $\theta_i$ multiplied by the study reliability $R$.
+Hence, the distribution of observed effect sizes is the distribution of true effect sizes $N(\theta, \tau^2$ multiplied by the reliability $R$, such that $R \times N(\theta, \tau^2) = N(R \theta, R^2 \tau^2) = N(R \theta, R \tau)$.
+Because reliability lies between 0 – 1, under these conditions imperfect reliability implies that both the observed average effect size and the estimated heterogeneity will be less than their true value by a factor $\bar{R}$ (if heterogeneity is expressed as standard deviation rather than variance).
+This depiction is not entirely correct for Pearson's $r$ as it is bounded at {-1, 1}, but holds approximately if the mean effect size $\theta$ and heterogeneity $\tau^2$ are not too large.
+Hence, imperfect reliability in studies can be expected to create a negative bias in heterogeneity estimates whenever true heterogeneity is present.
+To make this concrete, consider an average reliability of 0.8 (the median reported reliability in psychology; Flake et al., 2017).
+Then, $0.8 \times N(\theta, \tau) = N(0.8\theta, 0.8 \tau)$, resulting in 20% less standard deviation among observed effect sizes than in true effect sizes.
 Table 2 illustrates this relationship for infinite sample size (i.e., no sampling variance) and three levels of reliability across three studies.
@@
 
-Attenuation of effect size can also be understood as a downwards shift of the distribution of true effect sizes.
-To illustrate, assume an infinite sample size (i.e., no within-study sampling variance), some heterogeneity $\tau^2$, equal reliability in outcome and measurement such that $R_{xx'} = R_{yy'} = R$, and that all studies within a population of studies have the same reliability.
-Observed effect sizes given measurement error then corresponds to multiplying all true effect sizes $\theta_i$ by a factor equal to the reliability $R$ (i.e., multiplying the distribution of effect sizes by a constant $R$).
-If the distribution of effect sizes is normally distributed $\theta_i \sim N(\theta, \tau^2)$ with a mean $\theta$ (i.e., the assumption of the random effects model), then $R \times N(\theta, \tau) = N(R\theta, R^2 \tau^2) = N(R\theta, R\tau)$.
-Because reliability is always a value \< 1, it can be seen that the square root of the observed heterogeneity for uncorrected effect sizes is less than its true value by a factor $R$.
-This depiction is not entirely correct for Pearson's $r$ as it is bounded at {-1, 1}, but holds approximately as long as the mean effect size $\theta$ and heterogeneity $\tau^2$ are not too large.
-
-To make this concrete, consider an average reliability of 0.8 (the median reported reliability in psychology; Flake et al., 2017).
-Then, $0.8 \times N(\theta, \tau) = N(0.8\theta, 0.8 \tau)$, resulting in 20% less standard deviation among observed effect sizes than in true effect sizes.
-We can consider the positive bias in the same vein.
-Assuming a common true effect size $\mu$ across studies and infinite sample size within studies we can view $\mu$ as a constant.
-In this case, attenuation consists of a distribution of reliabilities $R \sim N(\bar{R}, \sigma_R)$ with some mean $\bar{R}$ and standard deviation $\sigma_R$ across studies, that is multiplied by the constant $\mu$ to yield a distribution of observed effect sizes $r \sim N(\mu \bar{R}, \mu \sigma_R)$.
-To make this concrete, consider some values $\mu = 0.2$, $\bar{R} = 0.8$, $\sigma_R = 0.1$ which then yield $r \sim N(0.16, 0.01)$.
-Again, this depiction is not entirely correct as reliability is bounded at {0, 1}, but holds approximately as long as the mean reliability $\bar{R}$ and its standard deviation $\sigma_R$ are not too large.
-
 It is difficult to predict the total effect of the negative and positive bias in heterogeneity due to unreliability.

### 2023-05-20 manuscript/introduction.qmd
intro: improved final paragraph

@@
 
-It is difficult to predict the total effect of the negative and positive bias in heterogeneity due to unreliability.
-This depends on average effect size, the amount of heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies.
+It is difficult to predict the total effect of the negative and positive bias in heterogeneity due to unreliability, and existing literature tends to focus on the positive bias (e.g., Borenstein, 2009, p.
+342; Card, 2015, p.
+126; Hunter & Schmidt, 2015; Wiernick & Dahlke, 2020).
+However, given that the negative bias could often be 20% or more and the value of accurate  heterogeneity estimates for many research outcomes in psychology, we consider it important to have insight into the total bias in heterogeneity estimates that should be expected due to unreliability.
+This depends on average effect size, true heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies.
 Hence, we perform a Monte-Carlo simulation study to explore the expected bias in heterogeneity estimates when meta-analyzing primary studies with imperfect measurement reliabilities.

### 2023-05-20 manuscript/introduction.qmd
intro: minor edits final paragraph

@@
 
-Meta-analytic estimates are highly dependent on the quality of the data in the primary studies that the method summarizes.
-However, there is concern that researchers in psychology tend to neglect proper measurement (ref measurement-schmesurement) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021).
+Meta-analytic estimates are highly dependent on the quality of data in primary studies.
+However, there is concern that researchers in psychology tend to neglect proper measurement (Flake & Fried, 2020) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021).
 For example, Flake et al.
@@
 Consequently, in the presence of heterogeneity attenuation will move larger true effect sizes further towards zero than smaller ones, decreasing heterogeneity in observed effect sizes.
-This negative bias can be observed directly by assuming that an average effect size $\theta$ has some heterogeneity $\tau^2$, all studies have infinite sample size (i.e., no within-study sampling variance), and equal reliability ($R = R_i = R_{xx'} = R_{yy'}$).
+This negative bias can be observed directly by assuming an average effect size $\theta$ with some heterogeneity $\tau^2$, infinite sample size within studies (i.e., no within-study sampling variance), and equal reliability across studies ($R = R_i = R_{xx'} = R_{yy'}$).
 Any observed effect size $\hat{\theta_i}$ can then be seen as a study’s true effect size $\theta_i$ multiplied by the study reliability $R$.
-Hence, the distribution of observed effect sizes is the distribution of true effect sizes $N(\theta, \tau^2$ multiplied by the reliability $R$, such that $R \times N(\theta, \tau^2) = N(R \theta, R^2 \tau^2) = N(R \theta, R \tau)$.
+Hence, the distribution of observed effect sizes is the distribution of true effect sizes $N(\theta, \tau^2)$ multiplied by the reliability $R$, such that $R \times N(\theta, \tau^2) = N(R \theta, R^2 \tau^2) = N(R \theta, R \tau)$.
 Because reliability lies between 0 – 1, under these conditions imperfect reliability implies that both the observed average effect size and the estimated heterogeneity will be less than their true value by a factor $\bar{R}$ (if heterogeneity is expressed as standard deviation rather than variance).
@@
 Hence, imperfect reliability in studies can be expected to create a negative bias in heterogeneity estimates whenever true heterogeneity is present.
-To make this concrete, consider an average reliability of 0.8 (the median reported reliability in psychology; Flake et al., 2017).
+Consider an average reliability of 0.8 (the median reported reliability in psychology; Flake et al., 2017).
 Then, $0.8 \times N(\theta, \tau) = N(0.8\theta, 0.8 \tau)$, resulting in 20% less standard deviation among observed effect sizes than in true effect sizes.
@@
 126; Hunter & Schmidt, 2015; Wiernick & Dahlke, 2020).
-However, given that the negative bias could often be 20% or more and the value of accurate  heterogeneity estimates for many research outcomes in psychology, we consider it important to have insight into the total bias in heterogeneity estimates that should be expected due to unreliability.
+However, given the value of accurate  heterogeneity estimates for many research outcomes in psychology and that the negative bias could often be 20% or more, we consider it important for meta-analysts to have insight into the total bias in heterogeneity estimates that should be expected due to unreliability.
 This depends on average effect size, true heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies.

### 2023-05-21 manuscript/introduction.qmd
intro: added references to selective reporting part

@@
 Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas).
-Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2) due to factors such as publication bias (REF1) and _p_-hacking (REF2), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
+Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated in psychology (e.g., Kvarven et al., 2019; Schäfer & Scharz, 2019; Scheel et al., 2020) due to selective reporting based on the significance of outcomes (Simmons et al., 2011) and publication bias (REF?), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
 On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures.
@@
 This depends on average effect size, true heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies.
-Hence, we perform a Monte-Carlo simulation study to explore the expected bias in heterogeneity estimates when meta-analyzing primary studies with imperfect measurement reliabilities.
+Hence, we perform a Monte-Carlo simulation study to explore the expected bias in heterogeneity estimates due to unreliability in primary studies.

### 2023-05-21 manuscript/methods.qmd
method: improved fisher z discussion

@@
 
-We use correlations as our effect size type in these simulations because correcting them for unreliability is straightforward and because they are the effect size type most commonly corrected in practice.
-{see also argument in Field that all effect size types are pretty similar} However, unreliability attenuates all types of effect sizes and all can be corrected, either directly (standardized mean differences; Wiernick & Dahlke, 2020) or by first transforming them into correlations (with associated assumptions).
+We focus on correlations as our effect size type in these simulations because correcting them for unreliability is straightforward and because they are the effect size type most commonly corrected in practice.
+However, unreliability attenuates all types of effect sizes and all can be corrected, either directly (standardized mean differences; Wiernick & Dahlke, 2020) or by first transforming them into correlations.
+Although effect sizes can generally be transformed into each other (e.g., Borenstein, 2009, p.
+46) this should be done with care as it can sometimes affect conclusions (e.g., we found a violation of monotonicity in heterogeneity estimates when converting effect sizes into correlations in a previous project; Olsson-Collentine, 2020, supplement A).
 
-Perhaps the most common way to meta-analyze correlations is to first transform them into Fisher's z correlations (Borenstein et al., 2009, p.
+Perhaps the most common way to meta-analyze correlations is to first transform them into Fisher $z$ correlations (Borenstein et al., 2009, p.
 41 - 43).
-The Fisher z transformation is the inverse hyperbolic tangent of the product moment correlation and can be computed as $equation$ and its variance approximated as $equation2$.
-There is some controversy in the use of the Fisher z transformation for meta-analysis (for a good summary see Brannick et al., 2019, p.
-492), in particular among proponents of Hunter and Schmidt's (2015) meta-anlytic models.
+The Fisher $z$ transformation is the inverse hyperbolic tangent of the product moment correlation and can be computed as $z = 0.5 \times ln\left(\frac{1+r}{1-r}\right)$ and its variance approximated as $V_z = 1 / (n -3)$.
+There is controversy in the use of the Fisher $z$ transformation for meta-analysis (e.g., Hunter & Schmidt, 2015, p.
+XX, ch.
+5 "Use of Fisher’s z in Meta-Analysis of Correlations"), largely related to whether Pearson's $r$ or Fisher's $z$ leads to less bias in average correlations (Schulze, 2007).
+We observe that estimating heterogeneity does not seem to have been the focus of the literature's discussion of the appropriateness of the Fisher $z$ transformation (Brannick et al., 2019; Field, 2005; Hafdahl & Williams, 2009, Hunter & Schmidt, 2015) and that the Fisher $z$ transformation is often applied in practice.
 
-We perform simulations both on product-moment correlations and Fisher z transformed correlations.
-For our purpose of estimating heterogeneity, Fisher z correlations have two advantages: 1) their sampling variance is not dependent on the size of the correlation, 2) they are not bounded to {-1, 1}.
-However, they also have two disadvantages: 1) that the heterogeneity estimate is measured on the Fisher z scale, which makes it hard to interpret.
-A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (REF) for an indirect estimate of the heterogeneity.
-Second, adding measurement error to the Fisher z correlations requires first transforming them into product-moment correlations, adding measurement error, and then backtransforming to Fisher z for meta-analysis.
-
-The primary advantage to using product-moment correlations in the meta-analysis when interested in heterogeneity is that values on this scale are more interpretable that on the Fisher z scale.
-The disadvantage is that the sampling variance is dependent on the effect size since it is computed as $equation$.
+Using Pearson's $r$ for meta-analysis when interested in heterogeneity has the advantage that resulting estimates on this scale are more interpretable that on the Fisher $z$ scale.
+Although Fisher $z$ correlations can be back-transformed into Pearson's $r$ for interpretation, its heterogeneity cannot.
+A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (REF).
+The width of the interval can then be used as an estimate of the heterogeneity.
+We do not believe this increases interpretability in our case and so prefer to directly estimate the heterogeneity.
+However, the disadvantage to using Pearson's $r$ when estimating heterogeneity is that 1) the effect size is bounded to {-1, 1} and 2) the sampling variance $$ depends on the effect size.
+The bounded nature of the Pearson correlation means that large levels of heterogeneity will lead to substantial truncation of effect sizes if the average effect size is also large.
+That the sampling variance covaries with effect size leads to a small positive bias in heterogeneity estimates (supplement XX).
+This can be solved by replacing $\rho$ with the average correlation across studies $\bar{r}$ (Hunter & Schmidt, 2015), and removing this covariance is also the primary intention behind the Fisher $z$ transformation.
+Fisher $z$ correlations also have the advantage that they are not bounded to {-1, 1}.
+A disadvantage to using Fisher $z$ correlations, in addition to interpretability of heterogeneity estimates, is that they require first backtransforming into product-moment correlations before adding measurement error, and then transforming to Fisher $z$ again for meta-analysis.
+Especially for large Fisher $z$ values (which can happen if the average effect size and heterogeneity are large) this transformation process may introduce inaccuracies in the estimates.
 Given the common use of both effect size types and their mix of advantages and disadvantages we report results for both.
@@
 
-We apply two meta-analytic models, the Hedges and Vevea (1998) random-effects  model and the Hunter and Schmidt (2015) 'bare-bones' random-effects model.
-We implement these two models as a sensitivity analysis rather than because of an interest in comparing the two in detail.
+We apply two meta-analytic models to ensure that our results are not model-dependent: the Hedges and Vevea (1998) random-effects  model and the Hunter and Schmidt (2015) 'bare-bones' random-effects model.
 The Hedges and Vevea (HV) model uses inverse variance weight computed as $w_i = 1 / (V_i + T^2)$ (e.g., Borenstein et al., 2010) where $V_i$ is the sampling variance of study $i$ and $T^2$ is the estimated between-studies variance.
@@
 The HS model typically includes correcting for unreliability or other measurement artifacts.
-However, we implement it without corrections ('bare-bones'; Hunter & Schmidt, 2015) because this allows us to examine the bias in heterogeneity estimates compared to a perfectly implemented correction of unreliability.
+However, we implement it without corrections ('bare-bones'; Hunter & Schmidt, 2015) because this allows us to focus on the bias in heterogeneity estimates rather than the performance of corrections.
 
@@
 However, the process for doing so differs somewhat between methods.
-We apply the Restricted Maximum Likelihood (REML) estimator of heterogeneity when estimating heterogeneity in the Hedges and Vevea (1998) model, because it is a heterogeneity estimator that performs well and is widely implemented (e.g., Viechtbauer, 2005; REF2), and is also the default estimator in the R package metafor.
-Describing this REML estimation in detail is beyond the scope of this paper and we refer interested readers to the description in Viechtbauer (2005) {this is also what Brannick et al do}.
-The Hunter and Schmidt (2015) model estimates heterogeneity $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $ \frac{\Sigma[N_i \sigma^2_{ei}]}{\Sigma N_i}$ (p.
+We apply the Restricted Maximum Likelihood (REML) estimator of heterogeneity when estimating heterogeneity in the Hedges and Vevea (1998) model, because it is the generally recommended heterogeneity estimator (Veroniki et al., 2016; Viechtbauer, 2005).
+Describing the REML estimator in detail is beyond the scope of this paper and we refer interested readers to the description in Viechtbauer (2005).
+The Hunter and Schmidt (2015) model estimates heterogeneity as $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $ \frac{\Sigma[N_i \sigma^2_{ei}]}{\Sigma N_i}$ (p.
 87, 2nd edition) and $\sigma^2_{ei}$ is the usual sampling variance for the Pearson correlation but with $\rho$ replaced with the sample size weighted average correlation $\bar{r} = \frac{\Sigma n_i r_i}{\Sigma n_i}$ instead of $r_i$ as is common.

### 2023-05-21 manuscript/methods.qmd
method: fisher z + model sections improved

@@
 5 "Use of Fisher’s z in Meta-Analysis of Correlations"), largely related to whether Pearson's $r$ or Fisher's $z$ leads to less bias in average correlations (Schulze, 2007).
-We observe that estimating heterogeneity does not seem to have been the focus of the literature's discussion of the appropriateness of the Fisher $z$ transformation (Brannick et al., 2019; Field, 2005; Hafdahl & Williams, 2009, Hunter & Schmidt, 2015) and that the Fisher $z$ transformation is often applied in practice.
+However, estimating heterogeneity does not seem to have been the focus of most of the literature's discussion of the appropriateness of the Fisher $z$ transformation (Field, 2005; Hafdahl & Williams, 2009, Hunter & Schmidt, 2015).
+Hall and Brannick (2002) do report heterogeneity estimates but focus on the coverage of prediction intervals ('credibility intervals' in Hunter & Schmidt terminology) which combine the heterogeneity estimate and point estimate, and in Brannick et al., (2019) they examine results only after corrections for attenuation.
+Given the inconclusive evidence and that the Fisher $z$ transformation is often applied in practice, we use it as one of the effect size types in our simulations.
 
-Using Pearson's $r$ for meta-analysis when interested in heterogeneity has the advantage that resulting estimates on this scale are more interpretable that on the Fisher $z$ scale.
+The other effect size type we us is Pearson's $r$.
+Pearson's $r$ has the advantage that resulting heterogeneity estimates on this scale are more interpretable that on the Fisher $z$ scale.
 Although Fisher $z$ correlations can be back-transformed into Pearson's $r$ for interpretation, its heterogeneity cannot.
-A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (REF).
+A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (Hedges & Vevea, 1998).
 The width of the interval can then be used as an estimate of the heterogeneity.
-We do not believe this increases interpretability in our case and so prefer to directly estimate the heterogeneity.
-However, the disadvantage to using Pearson's $r$ when estimating heterogeneity is that 1) the effect size is bounded to {-1, 1} and 2) the sampling variance $$ depends on the effect size.
+We prefer to compute comparable heterogeneity values for Pearson's $r$ and Fisher's $z$ (see the methods section 'Parameter values') and report the bias in heterogeneity estimates, as we believe this is more interpretable for applied meta-analysts.
+
+The disadvantage to using Pearson's $r$ when estimating heterogeneity is that 1) the effect size is bounded to {-1, 1} and 2) the sampling variance $\sigma^2_r = \frac{(1-\rho^2)^2}{n - 1}$ depends on the true effect size $\rho$.
 The bounded nature of the Pearson correlation means that large levels of heterogeneity will lead to substantial truncation of effect sizes if the average effect size is also large.
 That the sampling variance covaries with effect size leads to a small positive bias in heterogeneity estimates (supplement XX).
-This can be solved by replacing $\rho$ with the average correlation across studies $\bar{r}$ (Hunter & Schmidt, 2015), and removing this covariance is also the primary intention behind the Fisher $z$ transformation.
-Fisher $z$ correlations also have the advantage that they are not bounded to {-1, 1}.
-A disadvantage to using Fisher $z$ correlations, in addition to interpretability of heterogeneity estimates, is that they require first backtransforming into product-moment correlations before adding measurement error, and then transforming to Fisher $z$ again for meta-analysis.
+This can be solved by replacing $\rho$ with the average correlation across studies $\bar{r}$ (Hunter & Schmidt, 2015), or using the Fisher $z$ transformation.
+Fisher $z$ correlations have the additional advantage that they are not bounded to {-1, 1}.
+A disadvantage to using Fisher $z$ correlations, in addition to interpretability of heterogeneity estimates, is that they require backtransforming into product-moment correlations before adding measurement error, and then transforming to Fisher $z$ again for meta-analysis.
 Especially for large Fisher $z$ values (which can happen if the average effect size and heterogeneity are large) this transformation process may introduce inaccuracies in the estimates.
@@
 XX).
-Although the HV model is typically applied to Fisher z transformed correlations (e.g., Borenstein et al., p.
-41-43), we apply it to both transformed and untransformed correlations for the reasons described in the previous section.
+Although the HV model is typically applied to Fisher $z$ transformed correlations (e.g., Borenstein et al., p.
+41-43), we apply it to both transformed and untransformed correlations to enable a more direct comparison with the HS model and facilitate interpretation.
 The HS model typically includes correcting for unreliability or other measurement artifacts.
@@
 We apply the Restricted Maximum Likelihood (REML) estimator of heterogeneity when estimating heterogeneity in the Hedges and Vevea (1998) model, because it is the generally recommended heterogeneity estimator (Veroniki et al., 2016; Viechtbauer, 2005).
-Describing the REML estimator in detail is beyond the scope of this paper and we refer interested readers to the description in Viechtbauer (2005).
+Describing the REML estimator in detail is beyond the scope of this paper and we refer interested readers to the description in Veroniki et al., (2016).
 The Hunter and Schmidt (2015) model estimates heterogeneity as $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $ \frac{\Sigma[N_i \sigma^2_{ei}]}{\Sigma N_i}$ (p.
 87, 2nd edition) and $\sigma^2_{ei}$ is the usual sampling variance for the Pearson correlation but with $\rho$ replaced with the sample size weighted average correlation $\bar{r} = \frac{\Sigma n_i r_i}{\Sigma n_i}$ instead of $r_i$ as is common.
-A generalized version of the HS estimator for effect sizes other than correlations can be found in Viechtbauer (2005; 2015) (see ?rma).
+A generalized version of the HS estimator for effect sizes other than correlations can be found in Viechtbauer (2005; 2015).
 

### 2023-05-21 manuscript/methods.qmd
method: improved parameter section

@@
 Within meta-analyses we used a fixed sample size across all studies such that $N = N_i$.
-We made this choice rather than treating sample size as a random variable (as done by e.g., REF1, REF2) to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance.
-We base our range of sample sizes on empirical estimates of typical sample size in correlational research in psychology by XXX (still needs lookup) and consider the following values {VALUES}.
+We made this choice rather than treating sample size as a random variable (as done by e.g., Field, 2005; Brannick et al., 2019) to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance.
+We base our range of sample sizes on empirical estimates of typical sample size in correlational research in psychology.
+Fraley et al., report median sample sizes for between-person studies in 9 psychology journals between 2011 - 2018 which ranged between 69 - 496 depending on journal and year.
+Sassenber and Ditrich (2019) find a median sample size of 110 (interquartile range, 71 - 195) across 4 journals and years (2009, 2011, 2016, and 2018) in social psychology, and Bakker et al.
+(personal communication, 2023) find median sample sizes between 24 - 184 for 6 journals in psychology for the years 1995, 2006, 2019.
+Given these empirical findings we consider the following sample size values {50, 100, 150, 200}.
 
-When focusing on the bias of heterogeneity estimate, the value of $K$ would not matter for an unbiased estimator.
-However, both the heterogeneity estimator we use, restricted maximum likelihood, and most other heterogeneity estimators are truncated at zero (REF Viechtbauer).
-Lower values for $K$ then results in a larger proportion of truncated estimates below zero because of increased sampling variance.
-As such, the values of $K$ can affect bias, especially at lower levels of heterogeneity.
-In their review of meta-analyses in psychology, van Erp et al.
-(YEAR) find $K$ values ranging from XX to XX, and in our simulations we consider the following values {VALUES}.
+When focusing on the bias of heterogeneity estimate, the number of meta-analyzed studies, $K$, would not matter for an unbiased estimator.
+However, both the heterogeneity estimator we use, restricted maximum likelihood, and most other heterogeneity estimators are truncated at zero (Viechtbauer, 2005).
+Lower values for $K$ then results in a larger proportion of truncated estimates below zero because of increased sampling variance in $\tau^2$.
+As such, values of $K$ can affect bias, especially at lower levels of heterogeneity.
+We consider the following values of $K$: {5, 20, 40, 100}.
 
@@
 For Fisher's $z$ we transformed these value into equivalent effect sizes at the Fisher's $z$ scale.
-For context to  these values consider the empirical estimates of typical correlational effect sizes provided by Nuijten et al., (REF) and Schäfer & Scharz (REF).
+For context to  these values consider the empirical estimates of typical correlational effect sizes provided by Nuijten et al.
+(2020) and Schäfer & Scharz (2019).
 Nuijten et al.
-performed a meta-meta-analysis on the fields of RESEARCH and reports a median meta-analytic effect size of 0.24 for predictive validity and correlational studies.
+performed a meta-meta-analysis on the fields of intelligence research and reports a median meta-analytic effect size of 0.24 for predictive validity and correlational studies.
 This estimate is likely positively biased due to publication bias and selective reporting.
-Schäfer & Scharz report a median $r$ of 0.16 amongst preregistered research in FIELD, with an 'upper median' (i.e., the 83.35% quantile) of 0.41.
-This estimate is likely less affected by positive bias but should not be expected to be unbiased (REF Scheel).
+Schäfer & Scharz report a median $r$ of 0.16 amongst preregistered research in psychology in general, with an 'upper median' (i.e., the 83.35% quantile) of 0.41.
+This estimate is likely less affected by positive bias but should not be expected to be unbiased as there appears to be some positive bias also in preregistered research on average (Scheel, 2021).
 
 We defined the between-studies standard deviation ($\tau$) to cover a wide range of variability in true effect sizes and empirical estimates reported in the psychological literature.
-Van Erp et al., (REF) provide empirical heterogeneity estimates from XX meta-analyses in XX article published in the journal Psychological Bulletine between 1990 - 2013.
+Van Erp et al.
+(2017) provide empirical heterogeneity estimates from 747 meta-analyses in 61 article published in the journal Psychological Bulletin between 1990 - 2013.
 The median (uncorrected)Pearson's $r$ $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24).
@@
 Nonetheless, we selected approximately the interquartile range of $\tau$-values reported in van Erp et al.
-for the product-moment correlation, both to make sure we covered the empirically reported range of values and because we considered them reasonable variability on the product-moment correlation scale.
+for the product-moment correlation, both to make sure we covered the empirically reported range of values and because we considered them to represent reasonable variability on the product-moment correlation scale.
 The values we selected were {0, 0.1, 0.15, 0.2}.
@@
 Fisher's $z$ is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson's $r$ into corresponding $\tau$-values on the Fisher's $z$ scale.
-To be able to compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale.
-The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity (REF Higgins), and as such is measured from 0 - 100%.
+To compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale.
+The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity (Higgins, 2002; 2003), and as such is measured from 0 - 100%.
 The $I^2$ index can be defined as $I^2 = \frac{\tau^2}{s^2 + \tau^2}$ where $s^2 = \frac{\Sigma w_i(k - 1)}{(\Sigma w_i)^2 - \Sigma w_i^2}$ and $w_i$ is the precision $w_i = 1 / \sigma^2$.
-This means that the $I^2$ value generally depends on both the number of studies $K$ and the sample size $N_i$ in these studies (REF: Higgins, I2 is not an absolute measure of heterogeneity).
-However, because we keep $N_i$ fixed across studies (see below), the value of $I^2$ depends only on $N = N_i$ and the heterogeneity.
+This means that the $I^2$ value generally depends on both the number of studies $K$ and the sample size $N_i$ in these studies (Borenstein, 2017).
+However, because we keep $N_i$ fixed across studies, the value of $I^2$ depends only on $N = N_i$ and the heterogeneity.
 
@@
 In our case, $g(x)$ corresponds to the sampling variance of the product-moment correlation, $f(x) \sim N(\theta, \tau^2)$ and because the product-moment correlation is bounded at {-1, 1} we set $\infty$ instead to 0.999.
-The resulting $\tau$ values for Fisher $z$ are only minimally different from {0, 0.1, 0.15, 0.2}.
-For example, for $N = 150$ the corresponding Fisher $z$ $\tau$-values were {VALUES} and as $N$ increased heterogeneity values were more similar (see supplement XX).
+The resulting $\tau$ values for Fisher $z$ were only minimally different from {0, 0.1, 0.15, 0.2}.
+For example, for $N = 50$ the corresponding Fisher $z$ $\tau$-values were {0, 0.1031263, 0.1566007, 0.2123514} and as $N$ increased heterogeneity values were more similar (see supplement XX).
 
@@
 Flake et al.
-(YEAR) report that based on 245 estimates of Cronbach's alpha in psychology the average estimate was 0.79 with a standard deviation of 0.13.
+(2017) report that based on 245 estimates of Cronbach's alpha in psychology the average estimate was 0.79 with a standard deviation of 0.13.
 The interquartile range was approximately 0.68 - 0.87 for studies wherein the associated scale lacked a reference and 0.79 - 0.88 for scales that did have a reference.
-Sanchez-Meca et al.
-(REF) report the reliability estimates based on five 'reliability generalization studies'.
+Sánchez-Meca et al.
+(2003) report the reliability estimates based on five 'reliability generalization studies'.
 These five studies summarized reliability in 25 - 51 primary studies (184 total), and the mean reliability ranged from 0.767 to 0.891 with standard deviations ranging between 0.034 - 0.133.
-Given some likely positive bias in such empirical values, we examined the following mean reliabilities $\{0.6, 0.7, 0.8, 0.9\}$, and standard deviations $\{0, 0.05, 0.1, 0.15\}$.
+Given some likely positive bias in such empirical values (Hussey et al., 2023), we examined the following mean reliabilities $\{0.6, 0.7, 0.8, 0.9\}$, and standard deviations $\{0, 0.05, 0.1, 0.15\}$.

### 2023-05-21 code/tables.qmd
tables: removed table 2 version 2

@@
 SD (ES) is the standard deviation of the observed effect sizes for meta-analysis I, II and III, equivalent to heterogeneity given infinite within-study sample sizes.
+As true heterogeneity is absent, any SD(ES) values above zero represent bias.
 Code to reproduce table: LINK",
@@
 
-## Table 2 (version 1)
+## Table 2
 ### Imperfect reliability leads to negative bias in heterogeneity estimates
@@
 
-dat %>% knitr::kable(booktabs = TRUE, escape = FALSE, digits = 2,
+negative_bias_table <- dat %>%
+ knitr::kable(booktabs = TRUE, escape = FALSE, digits = 2,
                     align=rep('c', 5),
@@
   threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)
-```
-
-## Table 2 (version 2)
-### Imperfect reliability leads to negative bias in heterogeneity estimates
-
-```{r, alternative negative bias table}
-#Alternative table
-rel <- c("$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = 1",
-         "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = 0.8",
-         "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = 0.6")
-
-dat <- cbind(rel, r, sds)
-
-negative_bias_table <- dat %>%
-            knitr::kable(booktabs = TRUE, escape = FALSE, digits = 2,
-            col.names = c("",
-                        "Study 1",
-                        "Study 2",
-                        "Study 3",
-                        "SD(ES)")) %>%
-            kable_styling(position = "left") %>%
-            add_header_above(c(" " = 1, "Observed Effect Sizes" = 3)) %>%
-            column_spec (1, border_right = FALSE) %>%
-            footnote(general = "The values under Study 1, 2 and 3 are observed effect sizes for that study given the measurement reliability $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ and the true effect size of that study (first row) when within-study sample size is infinite.
-SD (ES) is the standard deviation of the observed effect sizes, equivalent to heterogeneity given infinite within-study sample sizes.
-Code to reproduce table: LINK",
-  threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)
 
@@
 
-saveRDS(list(positve_bias_table = positive_bias_table,
+saveRDS(list(positive_bias_table = positive_bias_table,
              negative_bias_table = negative_bias_table),

### 2023-05-21 manuscript/manuscript.qmd
manuscript: combined intro+methods+results

--- 2023-05-21 manuscript/methods.qmd
@@
 ```
+
+```{r load data and prep, include = FALSE}
+doc_type <- knitr::opts_knit$get('rmarkdown.pandoc.to') #to define the type of document, then format tables using an ifelse statement, see also: https://stackoverflow.com/questions/58999720/scientific-journal-table-of-variable-mean-per-year-with-s-d
+
+tables <- readRDS("tables.RDS")
+library(kableExtra) #must be loaded to print saved kableextra tables
+
+```
+
+Estimating heterogeneity of effect sizes is usually considered the main purpose of meta-analysis in addition to estimating an average effect size.
+Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size's sensitivity to variability in study design, or more specifically to sensitivity to differences in four study factors: 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables (e.g., Campbell & Stanley, 2015).
+Heterogeneity is a primary outcome in meta-analysis because its presence signals the existence of moderators and so can be seen as an opportunity for theoretical development (Simons et al., 2017), because it affects the implementation of research by indicating that an intervention may not be equally effective under all conditions or for all people, and because its presence changes the interpretation of the average effect sized derived in meta-analysis from the true population effect size (under homogeneity) to the average of true subpopulation effect sizes (under heterogeneity).
+As such, estimating heterogeneity without bias is of central importance to meta-analysis (for a more extensive discussion on the importance of heterogeneity, see Olsson Collentine et al., 2020; Simons, 2017).
+
+Meta-analytic estimates are highly dependent on the quality of data in primary studies.
+However, there is concern that researchers in psychology tend to neglect proper measurement (Flake & Fried, 2020) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021).
+For example, Flake et al.
+(2017) report that although latent constructs are widely examined in personality and social psychology, researchers barely report any of the validity evidence required to ascertain the extent to which implemented scales measure the constructs of interest.
+Such inattention to measurement can also bias heterogeneity estimates.
+We will focus here on reliability of measurement (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., REF), which is the only measurement validity evidence researchers report consistently (Flake et al., 2017), and which can be corrected for in meta-analysis (Hunter & Schmidt; 2015).
+
+Measurement error (imperfect reliability) is present in any empirical effect size estimate and systematically attenuates observed effect sizes compared to true underlying effect sizes.
+This affects heterogeneity estimates, which from a statistical perspective estimate the variance between true effect sizes.
+However, reliability is not necessarily accounted for in heterogeneity estimates, because ‘true’ underlying effect sizes can refer to two different entities.
+Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as imperfect reliability, restriction of range, and dichotomization; Hunter & Schmidt, 2015).
+Whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for measurement artifacts is implicit only in the models of Hunter & Schmidt (2015).
+This has consequences for heterogeneity estimates reported in many areas of psychology.
+Inattention to measurement reliability and how it affects heterogeneity estimates can lead researchers to misinterpret their meta-analytic average effect size estimate, ignore the presence of moderators or search for (and discover) non-existent moderators, and to implement research interventions inappropriately.
+
+Meta-analyses in psychology rarely correct for unreliability in primary study measurements.
+Wiernick and Dahlke (2020) report that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability.
+Similarly, Schmidt (2010) report that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts.
+The exception is the subfield of Industrial-Organizational studies where corrections tend to be more common (e.g., Cortina, 2003; Anguinis, 2011).
+
+One reason that few meta-analysts correct for unreliability could be that correcting for unreliability is generally controversial and has been so since it was proposed (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017).
+On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017).
+Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1).
+Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas).
+Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated in psychology (e.g., Kvarven et al., 2019; Schäfer & Scharz, 2019; Scheel et al., 2020) due to selective reporting based on the significance of outcomes (Simmons et al., 2011) and publication bias (REF?), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
+On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures.
+As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest.
+From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015).
+In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses need to be aware of how this decision affects heterogeneity estimates.
+
+## Bias in heterogeneity estimates due to unreliability
+
+In classical test theory (CTT), reliability of measurement ($R_{xx}$) is defined as the proportion true score variance ($\sigma^2_T$) to observed score variance ($\sigma^2_X$) and is presumed to be constant for all participants in a sample $R_{xx} = \sigma^2_T / \sigma^2_X$.
+We use unreliability of measurement to refer to $1 - R_{xx}$.
+Imperfect reliability in measurements leads to attenuation of observed effect sizes.
+For product-moment correlations the observed correlation can be computed as $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ where $r_{xy}$ is the observed correlation between the variables $X$ and $Y$, $\rho_{xy}$ is the true correlation (with or without sampling error) and $R_{xx'}$ and $R_{yy'}$ are the measurement reliabilities for $X$ and $Y$.
+As such, correcting for unreliability is straightforward for product-moment correlations, but can also be done for other effect size types (Wiernick and Dahlke, 2020).
+A corrected correlation should also have its sampling variance corrected.
+This is inflated by a factor equal to the square of the reliabilities, such that if $a = \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ then the corrected sampling variance $V_\rho = V_{r} / a^2$ (e.g., Borenstein, 2009, p.
+343).
+Imperfect reliability in primary studies imbues heterogeneity estimates with both a positive and a negative bias.
+
+A positive bias can be expected in heterogeneity estimates of uncorrected effect sizes due to the variability in reliability between primary studies.
+This is because the attenuation in effect size will then differ from study to study, which results in differences across studies in observed effect sizes beyond their true variability.
+This is most easily seen by assuming a common true effect size $\mu$ across studies (i.e., a fixed effect and no true heterogeneity) and infinite sample size within studies.
+Any observed effect size $\hat{\mu}$ can then be seen as the true effect size $\mu$ multiplied by a study’s reliability $R_i$ (assuming $R_i = R_{xx’} = R_{yy’}$).
+Hence, the distribution of observed effect sizes is the true effect size $\mu$ multiplied by the distribution of reliabilities across studies.
+That is, $\hat{mu} \sim \mu \times N(\bar{R}, \sigma_R) = N(\mu \bar{R}, \mu \sigma_R$, where $\bar{R}$ is the average reliability across studies and $\sigma_R$ its standard deviation.
+Because reliability lies between 0 – 1, under these conditions imperfect reliability implies that the observed average effect size will be less than the true effect size by a factor $\bar{R}$ and estimated heterogeneity will be larger than true heterogeneity (i.e., zero) by the product of $\mu \times \sigma_R$ (if heterogeneity is expressed as standard deviation rather than variance).
+That is, variability in reliability across studies leads to a positive bias in heterogeneity that depends on the size of the true effect size.
+Table 1 illustrates this effect for three studies with differing reliability and zero true heterogeneity.
+
+Table 1.
+
+*Variance in reliability leads to positive bias in heterogeneity estimates which increases with average effect size*
+
+```{r positive bias table}
+
+if (doc_type == "docx") {
+  print("Table 1 approximately here. Cannot be printed nicely to Word, see tables.pdf")
+} else{
+  tables$positive_bias_table
+}
+
+```
+
+A negative bias can be expected in heterogeneity estimates of uncorrected effect sizes because in absolute terms larger effect sizes are more attenuated by a given reliability.
+That is, although the attenuation formula $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ implies that any effect size $\rho_{xy}$ will be attenuated by the same proportion given the same reliabilities $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$, this proportion corresponds to a larger absolute value for larger effect sizes.
+Consequently, in the presence of heterogeneity attenuation will move larger true effect sizes further towards zero than smaller ones, decreasing heterogeneity in observed effect sizes.
+This negative bias can be observed directly by assuming an average effect size $\theta$ with some heterogeneity $\tau^2$, infinite sample size within studies (i.e., no within-study sampling variance), and equal reliability across studies ($R = R_i = R_{xx'} = R_{yy'}$).
+Any observed effect size $\hat{\theta_i}$ can then be seen as a study’s true effect size $\theta_i$ multiplied by the study reliability $R$.
+Hence, the distribution of observed effect sizes is the distribution of true effect sizes $N(\theta, \tau^2)$ multiplied by the reliability $R$, such that $R \times N(\theta, \tau^2) = N(R \theta, R^2 \tau^2) = N(R \theta, R \tau)$.
+Because reliability lies between 0 – 1, under these conditions imperfect reliability implies that both the observed average effect size and the estimated heterogeneity will be less than their true value by a factor $\bar{R}$ (if heterogeneity is expressed as standard deviation rather than variance).
+This depiction is not entirely correct for Pearson's $r$ as it is bounded at {-1, 1}, but holds approximately if the mean effect size $\theta$ and heterogeneity $\tau^2$ are not too large.
+Hence, imperfect reliability in studies can be expected to create a negative bias in heterogeneity estimates whenever true heterogeneity is present.
+Consider an average reliability of 0.8 (the median reported reliability in psychology; Flake et al., 2017).
+Then, $0.8 \times N(\theta, \tau) = N(0.8\theta, 0.8 \tau)$, resulting in 20% less standard deviation among observed effect sizes than in true effect sizes.
+Table 2 illustrates this relationship for infinite sample size (i.e., no sampling variance) and three levels of reliability across three studies.
+
+Table 2.
+
+*Imperfect reliability leads to negative bias in heterogeneity estimates*
+
+```{r negative bias table}
+
+if (doc_type == "docx") {
+  print("Table 2 approximately here. Cannot be printed nicely to Word, see tables.pdf")
+} else{
+  tables$negative_bias_table
+}
+
+```
+
+It is difficult to predict the total effect of the negative and positive bias in heterogeneity due to unreliability, and existing literature tends to focus on the positive bias (e.g., Borenstein, 2009, p.
+342; Card, 2015, p.
+126; Hunter & Schmidt, 2015; Wiernick & Dahlke, 2020).
+However, given the value of accurate  heterogeneity estimates for many research outcomes in psychology and that the negative bias could often be 20% or more, we consider it important for meta-analysts to have insight into the total bias in heterogeneity estimates that should be expected due to unreliability.
+This depends on average effect size, true heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies.
+Hence, we perform a Monte-Carlo simulation study to explore the expected bias in heterogeneity estimates due to unreliability in primary studies.
 
@@
 Given some likely positive bias in such empirical values (Hussey et al., 2023), we examined the following mean reliabilities $\{0.6, 0.7, 0.8, 0.9\}$, and standard deviations $\{0, 0.05, 0.1, 0.15\}$.
+
+# Results
+
+The overall bias in heterogeneity estimates due to imperfect reliability in primary studies tended to be negative.
+We found that the positive bias due to variance in reliability across studies was less than the negative bias due to average unreliability across all levels of heterogeneity in our primary analysis.
+As such, we report here only the condition with maximum variance in reliability (SD = 0.15).
+Analyses with lower levels of reliability variance can be found in supplement XX.
+In addition, although a smaller number of studies in a meta-analysis (k) leads to a larger positive bias in the absence of heterogeneity, as would be expected, there was no practical difference in the bias once heterogeneity was above 0.1.
+As such, we report only results for k = 20 in the main manuscript.
+Analyses with other values for k (5, 40, 200) can be found in supplement XX.
+Finally, the effect on bias in heterogeneity estimates of varying within-study sample size was relatively small, and we similarly report only results for N = 150 in the main manuscript and present results for other sample sizes in supplement XX.
+
+Figure 1 show the absolute and relative bias in heterogeneity estimates for different levels of heterogeneity (measured in $\tau$) and meta-analytic models.
+Each column of Figure 1 corresponds to one nominal level of heterogeneity ($\tau$ $\epsilon$ {0, 0.1, 0.15, 0.2}) and each row to a different analysis (HV with Fisher $z$, HV with Pearson's $r$, and HS).
+The black unbroken lines indicate actual heterogeneity, which may differ from nominal heterogeneity because of truncation or in the case of Fisher $z$ because the comparable heterogeneity levels were slightly different (see methods).
+The different dashed lines in Figure 1 shows the $\hat{\tau}$ estimates for four different levels of average reliability across studies, and the x-axis shows the average (superpopulation) effect size.
+As can be seen, the bias is marginally worse for Fisher z than the other two methods (although Fisher z is on a different scale) but the results are approximately the same for all three methods.
+As such, we will focus in the remainder of the results on Pearson's r (row two).
+
+```{r Figure 2, echo=FALSE, out.width = '100%'}
+knitr::include_graphics("../figures/z-r-hs-plot.png")
+```
+
+_Figure 1._ EXPLANATION. Code to reproduce figure: LINK.
+
+As expected, for zero heterogeneity (leftmost column Figure 1) we see an overestimate of heterogeneity that increases with effect size.
+There are two sources of positive bias in the absence of heterogeneity.
+First, and the only one that applies when $\mu = 0$, there is positive bias due to the truncation of negative heterogeneity estimates (Viectbauer, 2005).
+Second, there is bias due to the variance in reliability.
+For example, for $mu = 0.2$ (the closest value to the median correlation reported by Schäfer & Schwarz) and $\bar{R} = 0.8$ the bias for Pearson's $r$ is XX.
+For $\mu = 0.4$ (the upper median reported by Schäfer & Schwarz) and $\bar{R} = 0.8$ the bias is XX.
+For the highest average reliability (0.9) a somewhat less inclined slope can be seen.
+This is because the standard deviation for this reliability level is severely truncated when the reliability is bounded at 1.
+The increase in positive bias with effect size can also be seen in the presence of heterogeneity (all other columns).
+However, it is superseded by the negative bias.
+
+For all heterogeneity levels in Figure 1 above zero we see an overall underestimate of heterogeneity due to unreliability in primary studies.
+This negative bias can be relatively severe, is worse with lower average reliability, and is worse in an absolute sense but better in a relative sense as heterogeneity increases.
+For example, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.1$, the bias is XX (XX%) for Pearson's r, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.15$ the bias is XX (XX%), and for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.2$ the bias is XX (XX%).
+
+To explore further under what conditions the positive and negative bias in heterogeneity estimates would reach an equilibrium, we added an analysis for several conditions with $\tau < 0.1$.
+We used the following values {0.02, 0.04, 0.06, 0.08}.
+Figure 2 presents the results of this analysis for Pearson's r.
+Because these lower heterogeneity levels can be expected to be affected by truncation to a larger extent (which results in a positive bias), we present results for four different numbers of studies in the meta-analysis (rows Figure 2, lower $k$ is associated with larger positive bias).
+The standard deviation in reliabilities is again 0.15.
+Figure 2 shows that even when only little heterogeneity is present, it will generally be underestimated.
+If $\mu < 0.3$ then imperfect reliability leads to a negative bias in heterogeneity estimates for all $\tau < 0.06$ in Pearson's $r$.
+Even for lower levels of heterogeneity, heterogeneity will generally be positively biased only when average effect size is large or the number of meta-analyzed studies is small.
+
+```{r Figure 3, echo=FALSE, out.width = '100%'}
+knitr::include_graphics("../figures/r_tau_0.02-0.08.png")
+```
+
+_Figure 2._ EXPLANATION. Code to reproduce figure: LINK.

### 2023-05-23 manuscript/manuscript.qmd
intro: improved long sentence in first paragraph

@@
 Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size's sensitivity to variability in study design, or more specifically to sensitivity to differences in four study factors: 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables (e.g., Campbell & Stanley, 2015).
-Heterogeneity is a primary outcome in meta-analysis because its presence signals the existence of moderators and so can be seen as an opportunity for theoretical development (Simons et al., 2017), because it affects the implementation of research by indicating that an intervention may not be equally effective under all conditions or for all people, and because its presence changes the interpretation of the average effect sized derived in meta-analysis from the true population effect size (under homogeneity) to the average of true subpopulation effect sizes (under heterogeneity).
-As such, estimating heterogeneity without bias is of central importance to meta-analysis (for a more extensive discussion on the importance of heterogeneity, see Olsson Collentine et al., 2020; Simons, 2017).
+Heterogeneity is a primary outcome in meta-analysis because of several reasons.
+First, its presence signals the existence of moderators and so can be seen as an opportunity for theoretical development (Simons et al., 2017).
+Second, heterogeneity affects the implementation of research by indicating that an intervention may not be equally effective under all conditions or for all people.
+Third, the presence of heterogeneity changes the interpretation of the average effect sized derived in meta-analysis from the true population effect size (under homogeneity) to the average of true subpopulation effect sizes (under heterogeneity).
+As such, estimating heterogeneity with minimal bias is of central importance to meta-analysis (for a more extensive discussion on the importance of heterogeneity, see Olsson Collentine et al., 2020; Simons, 2017).
 
@@
 We apply the Restricted Maximum Likelihood (REML) estimator of heterogeneity when estimating heterogeneity in the Hedges and Vevea (1998) model, because it is the generally recommended heterogeneity estimator (Veroniki et al., 2016; Viechtbauer, 2005).
-Describing the REML estimator in detail is beyond the scope of this paper and we refer interested readers to the description in Veroniki et al., (2016).
+Describing the REML estimator in detail is beyond the scope of this paper and we refer interested readers to the description in Veroniki et al.
+(2016).
 The Hunter and Schmidt (2015) model estimates heterogeneity as $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $ \frac{\Sigma[N_i \sigma^2_{ei}]}{\Sigma N_i}$ (p.

### 2023-05-23 manuscript/manuscript.qmd
methods: copyedits

@@
 
-The other effect size type we us is Pearson's $r$.
-Pearson's $r$ has the advantage that resulting heterogeneity estimates on this scale are more interpretable that on the Fisher $z$ scale.
+The other effect size type we use is Pearson's $r$.
+Pearson's $r$ has the advantage that resulting heterogeneity estimates on this scale are more interpretable than on the Fisher $z$ scale.
 Although Fisher $z$ correlations can be back-transformed into Pearson's $r$ for interpretation, its heterogeneity cannot.
@@
 The values we selected were {0, 0.1, 0.15, 0.2}.
-We selected the largest heterogeneity level to avoid excissive truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1.
+We selected the largest heterogeneity level to avoid excessive truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1.
 Our values are similar to those used by Brannick et al.

### 2023-05-23 manuscript/manuscript.qmd
result: copyedits

@@
 
-_Figure 1._ EXPLANATION. Code to reproduce figure: LINK.
+_Figure 2._ EXPLANATION. Code to reproduce figure: LINK.
 
@@
 For all heterogeneity levels in Figure 1 above zero we see an overall underestimate of heterogeneity due to unreliability in primary studies.
-This negative bias can be relatively severe, is worse with lower average reliability, and is worse in an absolute sense but better in a relative sense as heterogeneity increases.
+This negative bias can be relatively severe, is worse with lower effect size and lower average reliability, and worsens in an absolute sense but improves in a relative sense as heterogeneity increases.
 For example, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.1$, the bias is XX (XX%) for Pearson's r, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.15$ the bias is XX (XX%), and for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.2$ the bias is XX (XX%).
@@
 
-_Figure 2._ EXPLANATION. Code to reproduce figure: LINK.
+_Figure 3._ EXPLANATION. Code to reproduce figure: LINK.

### 2023-05-24 code/clean_raw_data.r
code: separated out cleaning from plots

@@ -1,0 +1,74 @@
+# Project: Unexplained heterogeneity
+# Code purpose: cleaning raw data
+# Author: Anton Olsson-Collentine (anton@olssoncollentine.com)
+
+
+#****************************************
+# functions and packages
+#****************************************
+
+source("./code/functions.r") # for create_condition_cols function
+
+#****************************************
+# Clean data
+#****************************************
+
+#With r
+dat_r <- readRDS("data/raw_r_tau_0-0.2.RDS")
+
+#Turn list into dataframe using list-names as idcol
+dat_r <- data.table::rbindlist(dat_r, idcol = "mu")
+
+#create condition ids
+dat_r <- create_condition_cols(dat_r)
+
+saveRDS(dat_r, "data/cleaned_r_tau_0-0.2.RDS")
+
+e_means <- lapply(dat_r, colMeans)
+saveRDS(e_means, "data/means_r_tau_0-0.2.RDS")
+
+# With z
+dat_z <- readRDS("data/raw_z_tau_0-0.2.RDS")
+
+#Turn list into dataframe using list-names as idcol
+dat_z <- data.table::rbindlist(dat_z, idcol = "mu")
+
+#create condition ids
+dat_z <- create_condition_cols(dat_z)
+
+saveRDS(dat_z, "data/cleaned_z_tau_0-0.2.RDS")
+
+e_means <- lapply(dat_z, colMeans)
+saveRDS(e_means, "data/means_z_tau_0-0.2.RDS")
+
+#e_medians <- lapply(e, function(x) apply(x, 2, median))
+
+# For HS method
+
+dat_hs <- readRDS("data/raw_r_HS_tau_0-0.2.RDS")
+
+#Turn list into dataframe using list-names as idcol
+dat_hs <- data.table::rbindlist(dat_hs, idcol = "mu")
+
+#create condition ids
+dat_hs <- create_condition_cols(dat_hs)
+
+saveRDS(dat_hs, "data/cleaned_hs_tau_0-0.2.RDS")
+
+e_means <- lapply(dat_hs, colMeans)
+saveRDS(e_means, "data/means_hs_tau_0-0.2.RDS")
+
+# For low tau
+
+dat_low <- readRDS("data/raw_r_tau_0.02-0.08.RDS")
+
+#Turn list into dataframe using list-names as idcol
+dat_low <- data.table::rbindlist(dat_low, idcol = "mu")
+
[…]

### 2023-05-24 code/plots.r

@@ -15,19 +15,11 @@ library(ggplot2) #for plotting
 # General plot prep
 #****************************************
 
-
-
 dat_r <- readRDS("data/means_r_over_vs_underestimate.RDS")
 dat_z <- readRDS("data/means_z_over_vs_underestimate.RDS")
 dat_hs <- readRDS("data/means_r_HS_over_vs_underestimate.RDS")
 
 dat <- c(dat_r, dat_z, dat_hs)
-dat <- lapply(dat, function(x) as.data.frame(as.list(x)))
-
-dat <- data.table::rbindlist(dat, idcol = "mu")
-
-#create condition ids
-dat <- create_condition_cols(dat)
 
 # Add real truncated values
 es <- seq(0, 0.6, 0.1)
@@ -122,8 +114,10 @@ ggsave("figures/z-r-hs-plot.png", width = 8.62, height = 9.93)
 #****************************************
 # Main plot 2 (r, N = 150, k variable, tau low)
 #****************************************
-dat_r_low <- readRDS("data/means_r_borderline_estimate_tau0.02-0.08.RDS")
-dat_r <- readRDS("data/means_r_over_vs_underestimate.RDS")
+dat_r_low <- readRDS("data/means_r_tau_0.02-0.08.RDS")
+dat_r <- readRDS("data/means_r_tau_0-0.2.RDS")
+
+#The following line needs to be updated!
 dat_r_zero <- dat_r[grep("true_tau2 = 0;", names(dat_r))]
 
 dat <- c(dat_r_zero, dat_r_low)

### 2023-05-24 code/supercomputer_simulator.r

@@ -144,8 +144,5 @@ names(e) <- c(paste0("mu = ", cond$mu,
                      ";effect_type = ", cond$effect_type,
                      ";method = ", cond$method))
 
-e_means <- lapply(e, colMeans)
-#e_medians <- lapply(e, function(x) apply(x, 2, median))
 
-#saveRDS(e, "../data_new/raw_over_vs_underestimate.RDS")
-#saveRDS(e_means, "../data_new/means_over_vs_underestimate.RDS")
+#saveRDS(e, "../data_new/raw_r_tau_0-0.2.RDS")

### 2023-05-24 code/clean_raw_data.r
code: updated to create new means data

@@ -10,7 +10,7 @@
 source("./code/functions.r") # for create_condition_cols function
 
 #****************************************
-# Clean data
+# Clean data raw data
 #****************************************
 
 #With r
@@ -24,8 +24,6 @@ dat_r <- create_condition_cols(dat_r)
 
 saveRDS(dat_r, "data/cleaned_r_tau_0-0.2.RDS")
 
-e_means <- lapply(dat_r, colMeans)
-saveRDS(e_means, "data/means_r_tau_0-0.2.RDS")
 
 # With z
 dat_z <- readRDS("data/raw_z_tau_0-0.2.RDS")
@@ -72,3 +70,61 @@ saveRDS(dat_low, "data/cleaned_r_tau_0.02-0.08.RDS")
 
 e_means <- lapply(dat_low, colMeans)
 saveRDS(e_means, "data/means_r_tau_0.02-0.08.RDS")
+
+
+#****************************************
+# Create means
+#****************************************
+
+#With r
+dat_r <- readRDS("data/raw_r_tau_0-0.2.RDS")
+r_means <- lapply(dat_r, colMeans)
+
+#Turn list into dataframe using list-names as idcol
+r_means <- lapply(r_means, function(x) as.data.frame(as.list(x)))
+r_means <- data.table::rbindlist(r_means, idcol = "mu")
+
+#create condition ids
+r_means <- create_condition_cols(r_means)
+
+saveRDS(r_means, "data/means_r_tau_0-0.2.RDS")
+
+#With z
+dat_z <- readRDS("data/raw_z_tau_0-0.2.RDS")
+z_means <- lapply(dat_z, colMeans)
+
+#Turn list into dataframe using list-names as idcol
+z_means <- lapply(z_means, function(x) as.data.frame(as.list(x)))
+z_means <- data.table::rbindlist(z_means, idcol = "mu")
+
+#create condition ids
+z_means <- create_condition_cols(z_means)
+
+saveRDS(z_means, "data/means_z_tau_0-0.2.RDS")
+
+
+#With hs
+dat_hs <- readRDS("data/raw_r_hs_tau_0-0.2.RDS")
+hs_means <- lapply(dat_hs, colMeans)
+
+#Turn list into dataframe using list-names as idcol
+hs_means <- lapply(hs_means, function(x) as.data.frame(as.list(x)))
+hs_means <- data.table::rbindlist(hs_means, idcol = "mu")
+
+#create condition ids
[…]

### 2023-05-24 manuscript/manuscript.qmd
manuscript: added in-text-values for dynamic values

@@
 
+in_text <- readRDS("in-text-values.RDS")
 tables <- readRDS("tables.RDS")
@@
 The different dashed lines in Figure 1 shows the $\hat{\tau}$ estimates for four different levels of average reliability across studies, and the x-axis shows the average (superpopulation) effect size.
-As can be seen, the bias is marginally worse for Fisher z than the other two methods (although Fisher z is on a different scale) but the results are approximately the same for all three methods.
-As such, we will focus in the remainder of the results on Pearson's r (row two).
+As can be seen, the bias is marginally worse for Fisher $z$ than the other two methods (although Fisher $z$ is on a different scale) but the results are approximately the same for all three methods.
+As such, we will focus in the remainder of the results on Pearson's $r$ (row two).
 
@@
 As expected, for zero heterogeneity (leftmost column Figure 1) we see an overestimate of heterogeneity that increases with effect size.
-There are two sources of positive bias in the absence of heterogeneity.
-First, and the only one that applies when $\mu = 0$, there is positive bias due to the truncation of negative heterogeneity estimates (Viectbauer, 2005).
+Although Pearson's $r$ has a small positive bias due to the dependance between effect size and sampling variance (supplement XX), more generally there are two sources of positive bias in the absence of heterogeneity.
+First, and the only one that applies when $\mu = 0$, there is positive bias due to the truncation of negative heterogeneity estimates (Viechtbauer, 2005).
 Second, there is bias due to the variance in reliability.
-For example, for $mu = 0.2$ (the closest value to the median correlation reported by Schäfer & Schwarz) and $\bar{R} = 0.8$ the bias for Pearson's $r$ is XX.
-For $\mu = 0.4$ (the upper median reported by Schäfer & Schwarz) and $\bar{R} = 0.8$ the bias is XX.
+For example, for $mu = 0.2$ (the closest value to the median correlation reported by Schäfer & Schwarz) and $\bar{R} = 0.8$ the bias for Pearson's $r$ compared to true zero heterogeneity is `r in_text$bias_tau_zero$mu_0.2`.
+For $\mu = 0.4$ (the upper median reported by Schäfer & Schwarz) and $\bar{R} = 0.8$ the bias is `r in_text$bias_tau_zero$mu_0.4`.
 For the highest average reliability (0.9) a somewhat less inclined slope can be seen.
@@
 This negative bias can be relatively severe, is worse with lower effect size and lower average reliability, and worsens in an absolute sense but improves in a relative sense as heterogeneity increases.
-For example, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.1$, the bias is XX (XX%) for Pearson's r, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.15$ the bias is XX (XX%), and for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.2$ the bias is XX (XX%).
+For example, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.1$, the bias is `r in_text$bias_increasing_tau$tau_0.1` for Pearson's $r$, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.15$ the bias is `r in_text$bias_increasing_tau$tau_0.15`, and for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.2$ the bias is `r in_text$bias_increasing_tau$tau_0.2`.
 
@@
 We used the following values {0.02, 0.04, 0.06, 0.08}.
-Figure 2 presents the results of this analysis for Pearson's r.
+Figure 2 presents the results of this analysis for Pearson's $r$.
 Because these lower heterogeneity levels can be expected to be affected by truncation to a larger extent (which results in a positive bias), we present results for four different numbers of studies in the meta-analysis (rows Figure 2, lower $k$ is associated with larger positive bias).

### 2023-05-24 code/in-text-values.r

@@ -1,0 +1,71 @@
+# Project: unexplained heterogeneity
+# Script purpose: functions for simulations
+# code: Anton Olsson-Collentine
+
+#****************************************
+# functions and packages
+#****************************************
+
+source("./code/functions.r") # for create_condition_cols function
+
+
+#****************************************
+# [1] Result section
+#****************************************
+dat_r <- readRDS("data/means_r_tau_0-0.2.RDS")
+
+dat_r$tau_hat <- sqrt(dat_r$tau2_hat)
+dat_r$true_tau <- sqrt(as.numeric(dat_r$true_tau2))
+
+dat <- dat_r[dat_r$k == "20" & dat_r$N == "150",]
+
+#*[1.1] bias when tau =0, R = 0.8 ----
+#****************************************
+
+dat_1.1 <- dat[dat$true_tau == 0 & dat$mean_rel == 0.6 ,]
+#mu = 0.2
+mu_0.2 <- dat_1.1$tau_hat[dat_1.1$mu == 0.2] - dat_1.1$true_tau[dat_1.1$mu == 0.2]
+mu_0.2 <- round(mu_0.2, 3)
+
+#mu = 0.4
+mu_0.4 <- dat_1.1$tau_hat[dat_1.1$mu == 0.4] - dat_1.1$true_tau[dat_1.1$mu == 0.4]
+mu_0.4 <- round(mu_0.4, 3)
+
+bias_tau_zero <- data.frame(mu_0.2, mu_0.4)
+
+#*[1.2] bias when tau increasing, mu = 0.2, R = 0.8 ----
+#****************************************
+dat_1.2 <- dat[dat$mu == 0.2 & dat$mean_rel == 0.8,]
+#tau = 0.1
+tau_hat <- dat_1.2$tau_hat[dat_1.2$true_tau == 0.1]
+true_tau <- dat_1.2$true_tau[dat_1.2$true_tau == 0.1]
+bias_0.1 <-  tau_hat - true_tau
+percent_0.1 <- 100*(tau_hat / true_tau)
+
+res_0.1 <- paste0(round(bias_0.1, 3), " (", round(percent_0.1, 0), "%)")
+
+#tau = 0.15
+tau_hat <- dat_1.2$tau_hat[dat_1.2$true_tau == 0.15]
+true_tau <- dat_1.2$true_tau[dat_1.2$true_tau == 0.15]
+bias_0.15 <-  tau_hat - true_tau
[…]

### 2023-05-25 code/summarize_raw_data.r
code: simpliefied data cleaning code

@@ -9,72 +9,11 @@
 
 source("./code/functions.r") # for create_condition_cols function
 
+
 #****************************************
-# Clean data raw data
+# Create dataset with means
 #****************************************
-
-#With r
-dat_r <- readRDS("data/raw_r_tau_0-0.2.RDS")
-
-#Turn list into dataframe using list-names as idcol
-dat_r <- data.table::rbindlist(dat_r, idcol = "mu")
-
-#create condition ids
-dat_r <- create_condition_cols(dat_r)
-
-saveRDS(dat_r, "data/cleaned_r_tau_0-0.2.RDS")
-
-
-# With z
-dat_z <- readRDS("data/raw_z_tau_0-0.2.RDS")
-
-#Turn list into dataframe using list-names as idcol
-dat_z <- data.table::rbindlist(dat_z, idcol = "mu")
-
-#create condition ids
-dat_z <- create_condition_cols(dat_z)
-
-saveRDS(dat_z, "data/cleaned_z_tau_0-0.2.RDS")
-
-e_means <- lapply(dat_z, colMeans)
-saveRDS(e_means, "data/means_z_tau_0-0.2.RDS")
-
 #e_medians <- lapply(e, function(x) apply(x, 2, median))
-
-# For HS method
-
-dat_hs <- readRDS("data/raw_r_HS_tau_0-0.2.RDS")
-
-#Turn list into dataframe using list-names as idcol
-dat_hs <- data.table::rbindlist(dat_hs, idcol = "mu")
-
-#create condition ids
-dat_hs <- create_condition_cols(dat_hs)
-
-saveRDS(dat_hs, "data/cleaned_hs_tau_0-0.2.RDS")
-
-e_means <- lapply(dat_hs, colMeans)
-saveRDS(e_means, "data/means_hs_tau_0-0.2.RDS")
-
-# For low tau
-
-dat_low <- readRDS("data/raw_r_tau_0.02-0.08.RDS")
-
-#Turn list into dataframe using list-names as idcol
-dat_low <- data.table::rbindlist(dat_low, idcol = "mu")
-
-#create condition ids
-dat_low <- create_condition_cols(dat_low)
-
-saveRDS(dat_low, "data/cleaned_r_tau_0.02-0.08.RDS")
-
-e_means <- lapply(dat_low, colMeans)
-saveRDS(e_means, "data/means_r_tau_0.02-0.08.RDS")
-
-
-#****************************************
-# Create means
[…]

### 2023-05-25 manuscript/manuscript.qmd
results: added some dynamic results

@@
 For example, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.1$, the bias is `r in_text$bias_increasing_tau$tau_0.1` for Pearson's $r$, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.15$ the bias is `r in_text$bias_increasing_tau$tau_0.15`, and for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.2$ the bias is `r in_text$bias_increasing_tau$tau_0.2`.
+For an effect size of $\mu = 0.2$ true heterogeneity between 0.1 - 0.2 is underestimated by between `r round(100*in_text$underestimate_mu02$underestimate_min$underestimate, 0)`% ($\tau = 0.2$, $\bar{R} = 0.9$) and `r round(100*in_text$underestimate_mu02$underestimate_max$underestimate, 0)`% ($\tau = 0.1$, $\bar{R} = 0.6$).
 

### 2023-05-25 code/in-text-values.r

@@ -64,8 +64,34 @@ bias_increasing_tau <- list(tau_0.1 = res_0.1,
                             tau_0.15 = res_0.15,
                             tau_0.2 = res_0.2)
 
+
+#*[1.3] bias when mu = 0.2 and tau => 0.1 ----
+#****************************************
+dat02 <- dat[dat$mu == 0.2 & dat$true_tau > 0,]
+dat02$prop <- dat02$tau_hat / dat02$true_tau
+
+dat02$underestimate <- 1 - dat02$prop
+
+#"heterogeneity can be expected to be underestimated by XX%"
+underestimate_min <- data.frame(dat02[which.min(dat02$underestimate),
+ c("true_tau", "mean_rel", "underestimate")])
+
+underestimate_max <- data.frame(dat02[which.max(dat02$underestimate),
+ c("true_tau", "mean_rel", "underestimate")])
+
+underestimate_mu02 <- list(underestimate_min = underestimate_min,
+                           underestimate_max = underestimate_max)
+
+
+#*[1.3] bias when mu = 0.2 and tau => 0.1 ----
+#****************************************
+#****************************************
+# [2] Output
+#****************************************
+
 saveRDS(list(
     bias_tau_zero = bias_tau_zero,
-    bias_increasing_tau = bias_increasing_tau),
+    bias_increasing_tau = bias_increasing_tau,
+    underestimate_mu02 = underestimate_mu02),
     file = "./manuscript/in-text-values.RDS"
 )

### 2023-05-25 manuscript/manuscript.qmd
intro: added Borsboom

@@
 Such inattention to measurement can also bias heterogeneity estimates.
-We will focus here on reliability of measurement (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., REF), which is the only measurement validity evidence researchers report consistently (Flake et al., 2017), and which can be corrected for in meta-analysis (Hunter & Schmidt; 2015).
+We will focus here on reliability of measurement (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., Lord & Novick, 1968), which is the only measurement validity evidence researchers report consistently (Flake et al., 2017), and which can be corrected for in meta-analysis (Hunter & Schmidt; 2015).
 
@@
 Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1).
-Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas).
+Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (Sijtsma, 2008).
 Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated in psychology (e.g., Kvarven et al., 2019; Schäfer & Scharz, 2019; Scheel et al., 2020) due to selective reporting based on the significance of outcomes (Simmons et al., 2011) and publication bias (REF?), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
@@
 As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest.
-From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015).
+From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015), although  treating disattenuated effect sizes as directly comparable to latent scores requires certain assumptions (Borsboom & Mellenbergh, 2002).
 In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses need to be aware of how this decision affects heterogeneity estimates.

### 2023-05-25 manuscript/manuscript.qmd
manuscript: added discussion to qmd

@@
 _Figure 3._ EXPLANATION. Code to reproduce figure: LINK.
+
+# Discussion
+
+We used Monte-Carlo simulations to study the total bias that should be expected in meta-analytic heterogeneity estimates due to unreliability in primary studies.
+Our results indicate that uncontrolled unreliability in primary studies should under almost all circumstances be expected to lead to an severe underestimate of heterogeneity in meta-analysis.
+We observed a positive bias in heterogeneity estimates due to unreliability only under zero heterogeneity, or under exceptional circumstances (i.e., unusual combinations of low heterogeneity, large variance in primary study reliabilities, large effect sizes, and low number of meta-analyzed studies).
+This generally negative bias was observed under conditions that should have been favorable to a positive bias (i.e., large variance in primary study reliability), indicating that negative bias should probably be expected in all cases except for zero true heterogeneity.
+Our results were approximately the same whether one used a classic random-effects meta-analytic model (Hedges & Vevea, 1998) or a 'bare-bones' psychometric meta-analytic model (Hunter & Schmidt, 2015), and whether one used Pearson's $r$ as the effect size or Fisher's $z$, although the negative bias was slightly larger with Fisher's $z$.
+The severity of the underestimation of heterogeneity suggests meta-analyses in psychology often miss moderators and a risk that theoretical development in psychology has been misguided by meta-analyses.
+
+The negative bias in heterogeneity estimates can be extreme if reliability is low.
+For a typical effect size around $r$ = 0.2, larger heterogeneity ($\tau > 0.1$) can be expected to be underestimated by about 15% - 60% primarily depending on average reliability in primary studies (higher reliability leads to a smaller underestimate).
+The underestimate is worse with lower average effect size, but although it improves with larger effect sizes it is likely to remain negative, and less variability in primary study reliabilities will lead to larger negative bias.
+As true heterogeneity increases, the proportional underestimate decreases to some degree, but the underestimate in absolute terms will worsen.
+Although small degrees of true heterogeneity ($\tau < .08$) can be overestimated under certain conditions, this would typically require a combination of a small number of studies, large variance in study reliabilities, and a large average effect size.
+The severity and generality of the underestimation means that meta-analysts cannot ignore the effect of reliability on their heterogeneity estimates.
+
+If true heterogeneity is zero, there is a positive bias in observed heterogeneity estimates, but for even small degrees of true heterogeneity there tends to be a negative bias in observed estimates.
+As such, it is not possible to know whether a small degree of observed heterogeneity is due to overestimated true zero heterogeneity or underestimated small heterogeneity.
+However, for observed heterogeneity above $\tau > 0.025$ one can usually be confident that the estimate is an underestimate of true heterogeneity rather than an overestimate of zero heterogeneity if the average effect sizes is below approximately $r = 0.2$.
+
+Similarly, if researchers find zero heterogeneity in their observed estimates, true heterogeneity is most likely to be zero or small.
+Larger true heterogeneity (e.g., $\tau > 0.08$) has a higher expected observed heterogeneity value even when average reliability is as low as 0.6, and as such is expected to result in zero estimates less often than zero or smaller true heterogeneity for fixed conditions.
+For example, for the conditions reported in the main article (i.e., K = 20, N = 150, $\sigma_R$ = 0.15), an assuming a typical (reported) average reliability of 0.8 and a zero average correlation, then an observed zero heterogeneity estimate is XX times more likely to correspond to true zero heterogeneity than $\tau = 0.1$.
+That said, an observed zero heterogeneity estimate is still possible to observe under larger true heterogeneity levels as variability in heterogeneity estimates is often large in practice (Ioannidis; ref2).
+In addition, separating between zero and small heterogeneity is unfeasible considering realistic sample sizes in meta-analyses in psychology (Olsson-Collentine et al., 2020).
+In our view, the distinction between zero and small heterogeneity in meta-analysis is primarily a hypothetical one, as we consider low heterogeneity and meta-analysis of a large number of studies unlikely to cooccur in practice.
+
+We found that using Pearson's $r$ or Fisher's $z$ as our effect size resulted in very similar degrees of bias in heterogeneity estimates due to uncorrected unreliability in primary studies.
+The underestimate is slightly worse for Fisher's $z$ than for Pearson's $r$.
+The usage of Pearson's $r$ adds a slight positive bias to heterogeneity estimates (due to its bounded nature and effect size dependent sampling variance) which apparently is beneficial for heterogeneity estimates when not correcting for unreliability.
+However, the additional negative bias when using Fisher's $z$ is small and we consider the difference negligible in the context of the severe underestimates generated by unreliability.
+
+## Should meta-analysts correct for unreliability in primary studies?
+
+Despite the severity of underestimation due to unreliability, it is neither straightforward to correct for unreliability using the formulas of classical test theory nor self-evident that doing so is the optimal solution.
+Although our simulation results are not affected by whether researchers are interested in observed measures or latent constructs, it is safe to assume that the majority of researcher in psychology are interested in latent constructs.
+A central assumption behind applying the classical test theory correction to examine latent constructs is that disattenuated effect sizes (“true scores”) correspond approximately to latent construct scores (Hunter & Schmidt, 2015; Borsboom & Mellenbergh, 2002).
+Given that these assumptions hold, or one is satisfied with correcting for measurement error regardless of the possibility of a latent construct interpretation, there are several complexities that arise when attempting to correct for unreliability in practice.
+First, reliability estimates are lower bound estimates, meaning systematically correcting for unreliability can be expected to lead to overestimated effect sizes and heterogeneity.
+Unfortunately, the most reported reliability estimate (Chronbach’s alpha; Flake et al., 2017) is known to also underestimate the lower bound (Sijtsma, YEAR).
+Second, estimates of reliability (e.g., those based on a single measurement occasion such as Cronbach’s alpha) do not necessarily account for all types of measurement error (Hunter & Schmidt, 2015, p.
+XX, section “Using the appropriate reliability coefficient”).
+If so, they overestimate reliability.
+Hence, from a meta-analytic perspective, most reliability estimates available for correcting effect sizes are likely to be affected by both a positive and a negative bias compared to the true reliability, but what the total effect is likely to be is uncertain (and hence how accurate a heterogeneity estimate based on these corrections would be).
+Even so, in view of these two problems one could argue that an imperfect correction of unreliability is likely to be preferable over no correction (e.g., Oswald, YEAR), especially given the severe underestimation of heterogeneity.
+Further study would be useful to provide evidence to whether this is truly the case.
+
+Another concrete problem is that reliability estimates are sometimes not reported in primary studies.
+Missing reliability estimates must then be imputed, which is usually done by taking the average of observed reliabilities (e.g., Gnambs & Sengewald.
+2023) or imputing some ‘typical’ reliability (Hunter & Schmidt, 2015, p.
+XX).
+Both approaches have problems.
+There can be much debate about what a ‘typical’ reliability is (LeBreton et al., 2014) and selecting different values can have large effect on estimates.
+Mean imputation is viewed as inappropriate by experts in missing data (e.g., Enders, 2010, p.
+42-43; van Buuren, 2018, p.
+12; Schafer, 2002, p.
+159) partly because the pooled estimate will be biased unless observed values coincide with the population mean, particularly so if missing values are related to the value of the statistic (and it seems highly likely that missingness is more common with low reliability).
+However, it is not clear that more complex imputation procedures make sense except with substantially higher number of studies in than is common in meta-analyses in psychology.
+
+These are practical concerns with applying reliability corrections, but a more pernicious challenge to treating disattenuated effect sizes as equivalent to construct scores is the problem of validity (Borsboom & Mellenbergh, 2002; Hunter & Schmidt, 2015, p.
+41 – 53, 2nd edition).
+First, doing so requires assuming that the constructs underlying measurements are unidimensional in nature (Borsboom & Mellenbergh, 2002), which in many cases is a strong assumption.
+Second, to treat disattenuated effect sizes as latent construct scores requires assuming that the applied measurement instruments measure the construct of interest directly rather than through some proxy variable.
+Although corrections akin to the classic unreliability correction can be applied for simple causal models, these methods quickly become insufficient as the relationship between latent constructs grows in complexity (Hunter & Schmidt, 2015, p.
+41 – 53, 2nd edition).
+As such, some argue that simple corrections for unreliability under the classical test theory framework are untenable and potentially misleading if one’s interest lies in latent constructs (Borsboom & Mellenbergh, 2002).
+From this perspective, researchers should consider modern psychometric models such as Structural Equation Models (SEM; e.g., REF) which permit delving deeper into measurements and assumptions about latent construct relationships.
+
+Meta-analysis can be conceptualized as a SEM model, which holds promise that the benefits of modern psychometric theory might be applied also when the interest is in summarizing research across multiple studies.
+In Meta-Analytic Structural Equation Modeling (MASEM; e.g., Cheung & Chan, 2005) unreliability in primary studies can be corrected in three ways (Gnambs & Sengewald, 2023): 1) by estimating a SEM model in each primary study and then pooling estimates (parameter-based MASEM: Cheung, 2015, p.
+241), 2) by directly modeling latent variables in the MASEM, or 3) by first correcting correlations before introducing them into the MASEM framework (i.e., fundamentally identical to the classical test theory approach).
+The first option is limited because it is rarely the case that the same SEM model can be fit across multiple independent datasets, except in highly structure cases (e.g., Brunner et al., 2022).
+The second option requires item-level statistics rather than summary values, which are rarely available to meta-analysts.
+The third option, naturally, suffers the same challenges and assumptions as corrections under classical test theory.
+As such, the potential to use MASEM to take advantages of modern psychometric theory when correcting for unreliability is challenging in practice.
+That said, multi-lab replication projects, which are increasingly common in psychology (e.g., REF1, REF2), are well-suited to take advantage of MASEM.
+
+In the end, meta-analysis depends on the quality of the primary studies that make up its data.
+As long as researchers in psychology underappreciate the importance of measurement, meta-analytic estimates will be poor.
+For meta-analysts that consider acceptable the challenges and assumptions of corrections to unreliability, the many caveats inherent in doing so should be acknowledged and estimates based on both corrected and uncorrected estimates reported.
+Meta-analysts who consider corrections untenable need to consider that heterogeneity estimates may be severely underestimated.
+In either case, meta-analysts should get used to extracting reliability information whenever available from primary studies.
+The good news is that large true heterogeneity ($\tau > 0.1$) is on average distinguishable from zero or small estimates even in observed heterogeneity estimates.
+
+## Limitations
+
+In our simulations we assumed equal reliability for the measures $X$ and $Y$ (i.e., $R_i = R_{xx'} = R_{yy'}$) within a study, which is not realistic.
+However, as we never use the reliabilities individually (i.e., only expressed as $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$) we do not expect major impact on our results.
+Consider that if $R_i = 0.7$ then $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}} = 0.7$ whereas if if $R_{xx'}$ and $R_{yy'}$ took other values from a distribution centered on 0.7 this value would be lower (e.g., if $R_{xx'} = 0.5$ and $R_{yy'} = 0.9$ then $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}} \approx 0.67$).
+As we use the results of $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ in our simulations to generating observed effect sizes from true effect sizes (i.e., $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$, where $\rho_{xy}$ is the true effect size after sampling error), letting the reliability for $X$ and $Y$ vary would lead to slightly larger attenuation of observed effect sizes on average.
+That is, letting the reliability of $X$ and $Y$ vary within studies would lead to larger negative bias in observed heterogeneity estimates than in our simulations, indicating that the effect might even be somewhat more severe than in our estimates.
+
+We also only considered bias in heterogeneity due to unreliability in primary studies, but there exist other measurement artifacts such as restriction of range, dichotomization and more (Hunter & Schmidt, 2015).
+Although measurement error is the only measurement artifact that is always present, many of these other measurement artifacts may impact heterogeneity estimates when they are present.
+Based on Hunter & Schmidt’s arguments that measurement artifacts all tend to attenuate effect sizes in a similar way to unreliability, we expect that their net effect on heterogeneity estimates may similarly be a negative bias.
+The extent to which this prediction is true (and whether these measurement artifacts can be corrected for in MASEM; Gnambs & Sengewalds, 2023) awaits further research.
+
+We also used a fixed sample size (N) across studies within a meta-analysis in our simulations, which is unrealistic.
+We showed in Supplement XX that increasing the fixed sample size led to slightly better observed heterogeneity estimates for low reliability.
+Implementing a distribution of sample sizes from which within meta-analyses sample sizes was drawn (e.g., as done by REF1; ref2) could be one way to increase realism in sample sizes, although to the detriment of interpretability.
+However, given a sufficiently large number of repetitions in our simulations, this would not change estimates of bias, assuming the same mean sample sizes were used.
+That said, moving from fixed sample sizes to a distribution of sample sizes would affect the variability of any estimates, but this was not the focus of our study.
+
+We did not examine variance in heterogeneity estimates, or their root mean squared error.
+Our focus was on to what extent unreliability leads to bias in heterogeneity estimates and should be a concern for substantive researcher attempting to use meta-analysis to learn more about the contextual sensitivity of their research topic.
+As such, although we apply different meta-analytic models and effect sizes, the purpose was not to compare the efficiency of these different modes of estimation, but to show our results with the dominant meta-analytic approaches for correlations and demonstrate that they are not to any large degree dependent on the meta-analytic approach.
+As the effect of unreliability in primary studies is simply to change the heterogeneity of effect sizes the results of previous studies on the comparable performance of different heterogeneity estimators applies also here (e.g., REF1, REF2, REF3).
+That said, examining the variance of heterogeneity estimates could be useful to gain a better idea of how likely a particular observed heterogeneity estimate is to belong to different distributions of true heterogeneity.
+As this would depend on N, K, mean reliability, variability in reliability, average effect size, observed heterogeneity, and the true heterogeneity levels being compared, implementing such an analysis would probably be best done on a case-by-case basis (e.g., through a webapp such as developed for sensitivity analyses of the effect of publication bias on heterogeneity by Augusteijn et al., YEAR).
+Given the challenges in correcting for unreliability, such a sensitivity analysis could be a promising tool to reason around the impact of unreliability in applied meta-analysis.
+
+Our simulations assume reliabilities to be known.
+This is never the case, and any estimate of reliability is at best a lower bound estimate.
+By assuming reliabilities to be known, we were able to show the bias of observed heterogeneity estimates compared to the ideal case of no measurement error.
+However, this may mislead readers to believe that a simple correction for unreliability will leave them with an unbiased estimate.
+As discussed in the section ‘should meta-analysts correct for unreliability?’ correcting for reliability based on estimated reliability is not straightforward and whether the resulting estimate will be unbiased is contingent on many factors and in many cases unresolved.
+As such, examining the effect of corrections based on estimated reliabilities was beyond the scope of this study.
+
+## Conclusion
+
+Imperfect measurement reliability in primary studies generally leads to a severe underestimate of observed meta-analytic heterogeneity.
+As few meta-analyses in psychology correct for unreliability in primary studies, heterogeneity is likely underestimated in a large proportion of meta-analyses in psychology.
+Consequently, variability in effects has been ignored, moderators potentially neglected and theoretical development in psychology been misguided by meta-analyses.
+Sophisticated methods (MASEM) for correcting measurement unreliability are only applicable in exceptional cases and corrections based on classical test theory come with caveats and assumptions.
+Accurate estimation of meta-analytic heterogeneity is difficult and will remain so unless measurement concerns (REF) are taken seriously in primary research.
+The good news is that study designs (multi-lab replication studies) which can apply more sophisticated versions of MASEM are becoming increasingly common in psychology, and that large true heterogeneity can on average be distinguished from zero or small heterogeneity even in the absence of corrections to unreliability in primary studies.

### 2023-05-25 manuscript/manuscript.qmd
manuscript: added figure captions

@@
 
-Figure 1 show the absolute and relative bias in heterogeneity estimates for different levels of heterogeneity (measured in $\tau$) and meta-analytic models.
-Each column of Figure 1 corresponds to one nominal level of heterogeneity ($\tau$ $\epsilon$ {0, 0.1, 0.15, 0.2}) and each row to a different analysis (HV with Fisher $z$, HV with Pearson's $r$, and HS).
+Figure 2 show the absolute and relative bias in heterogeneity estimates for different levels of heterogeneity (measured in $\tau$) and meta-analytic models.
+Each column of Figure 2 corresponds to one nominal level of heterogeneity ($\tau$ $\epsilon$ {0, 0.1, 0.15, 0.2}) and each row to a different analysis (HV with Fisher $z$, HV with Pearson's $r$, and HS).
 The black unbroken lines indicate actual heterogeneity, which may differ from nominal heterogeneity because of truncation or in the case of Fisher $z$ because the comparable heterogeneity levels were slightly different (see methods).
-The different dashed lines in Figure 1 shows the $\hat{\tau}$ estimates for four different levels of average reliability across studies, and the x-axis shows the average (superpopulation) effect size.
+The different dashed lines in Figure 2 shows the $\hat{\tau}$ estimates for four different levels of average reliability across studies, and the x-axis shows the average (superpopulation) effect size.
 As can be seen, the bias is marginally worse for Fisher $z$ than the other two methods (although Fisher $z$ is on a different scale) but the results are approximately the same for all three methods.
@@
 
-_Figure 2._ EXPLANATION. Code to reproduce figure: LINK.
+_Figure 2._ Unreliability in primary studies leads to a net negative bias when heterogeneity is positive.
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
+Columns indicate the nominal true heterogeneity standard deviation $\tau$.
+Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Each dashed line correspond to an average reliability in primary studies.
+Standard deviation in reliability was 0.15.
+Code to reproduce figure: LINK.
 
-As expected, for zero heterogeneity (leftmost column Figure 1) we see an overestimate of heterogeneity that increases with effect size.
+As expected, for zero heterogeneity (leftmost column Figure 2) we see an overestimate of heterogeneity that increases with effect size.
 Although Pearson's $r$ has a small positive bias due to the dependance between effect size and sampling variance (supplement XX), more generally there are two sources of positive bias in the absence of heterogeneity.
@@
 
-For all heterogeneity levels in Figure 1 above zero we see an overall underestimate of heterogeneity due to unreliability in primary studies.
+For all heterogeneity levels in Figure 2 above zero we see an overall underestimate of heterogeneity due to unreliability in primary studies.
 This negative bias can be relatively severe, is worse with lower effect size and lower average reliability, and worsens in an absolute sense but improves in a relative sense as heterogeneity increases.
@@
 We used the following values {0.02, 0.04, 0.06, 0.08}.
-Figure 2 presents the results of this analysis for Pearson's $r$.
-Because these lower heterogeneity levels can be expected to be affected by truncation to a larger extent (which results in a positive bias), we present results for four different numbers of studies in the meta-analysis (rows Figure 2, lower $k$ is associated with larger positive bias).
+Figure 3 presents the results of this analysis for Pearson's $r$.
+Because these lower heterogeneity levels can be expected to be affected by truncation to a larger extent (which results in a positive bias), we present results for four different numbers of studies in the meta-analysis (rows Figure 3, lower $k$ is associated with larger positive bias).
 The standard deviation in reliabilities is again 0.15.
-Figure 2 shows that even when only little heterogeneity is present, it will generally be underestimated.
+Figure 3 shows that even when only little heterogeneity is present, it will generally be underestimated.
 If $\mu < 0.3$ then imperfect reliability leads to a negative bias in heterogeneity estimates for all $\tau < 0.06$ in Pearson's $r$.
@@
 
-_Figure 3._ EXPLANATION. Code to reproduce figure: LINK.
+_Figure 3._ Unreliability in primary studies typically leads to a net negative bias even for small degrees of heterogeneity.
+Effect size is Pearson's $r$ and rows indicate the number of studies in the meta-analysis.
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
+Columns indicate the nominal true heterogeneity standard deviation $\tau$.
+Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Each dashed line correspond to an average reliability in primary studies.
+Standard deviation in reliability was 0.15.
+Code to reproduce figure: LINK.
 
@@
 That said, an observed zero heterogeneity estimate is still possible to observe under larger true heterogeneity levels as variability in heterogeneity estimates is often large in practice (Ioannidis; ref2).
-In addition, separating between zero and small heterogeneity is unfeasible considering realistic sample sizes in meta-analyses in psychology (Olsson-Collentine et al., 2020).
+In addition, separating between zero and small heterogeneity is infeasible considering realistic sample sizes in meta-analyses in psychology (Olsson-Collentine et al., 2020).
 In our view, the distinction between zero and small heterogeneity in meta-analysis is primarily a hypothetical one, as we consider low heterogeneity and meta-analysis of a large number of studies unlikely to cooccur in practice.
@@
 Sophisticated methods (MASEM) for correcting measurement unreliability are only applicable in exceptional cases and corrections based on classical test theory come with caveats and assumptions.
-Accurate estimation of meta-analytic heterogeneity is difficult and will remain so unless measurement concerns (REF) are taken seriously in primary research.
+Accurate estimation of meta-analytic heterogeneity is difficult and will remain so unless measurement concerns (Flake & Fried, 2020; Flake et al., 2017; Kane et al., 2021) are taken seriously in primary research.
 The good news is that study designs (multi-lab replication studies) which can apply more sophisticated versions of MASEM are becoming increasingly common in psychology, and that large true heterogeneity can on average be distinguished from zero or small heterogeneity even in the absence of corrections to unreliability in primary studies.

### 2023-06-09 code/functions.r
added intermediate save + Hedge's -> HV

@@ -27,7 +27,7 @@ rnorm_truncated <- function(
 simulate_rma <- function(
 
     effect_type = c("r", "r_z"),
-    method = c("Hedges", "HS"),
+    method = c("HV", "HS"), #Hedges & Vevea or Hunter & Schmidt
     k, #number of studies
     sample_size, #sample size, fixed across studies
     true_tau2, #variance of superpopulation
@@ -96,7 +96,7 @@ simulate_rma <- function(
     }
 
     # fit meta-analysis on observed values
-    if(method == "Hedges"){
+    if(method == "HV"){
         fit <- rma(yi = r_se_me, vi = r_var_estimate,
                   control=list(stepadj=steplength, maxiter=maxiter))
 

### 2023-06-09 code/supercomputer_simulator.r

@@ -26,10 +26,10 @@ mu <- seq(from = 0, to = 0.6, by = 0.1)
 #Based on Flake et al., average alpha was .79, SD = .13, range .17 - .87
 #Based on Sanchez-Meca, mean across 5 meta-analysis was 0.767 - 0.891 and SD ranged between 0.034 - 0.133
 reliability_mean <- c(0.6, 0.7, 0.8, 0.9, 1)
-reliability_sd <- seq(from = 0.05, to = 0.15, by = 0.05)
+reliability_sd <- seq(from = 0, to = 0.15, by = 0.05)
 
 effect_type  <- "r" # or r_z
-meta_method  <- "Hedges" #or HS
+meta_method  <- "HV" #or HS
 
 #Using Pearson's r as the effect size.
 true_tau <- c(0, 0.1, 0.15, 0.2)
@@ -128,6 +128,17 @@ for(r in 1:nrow(cond)){ #gives us a list of lists
     out_list[[r]] <- out_list[[r]][, lapply(.SD, mean)] #data.table colMeans but returns a dataframe (well, data.table)
     out_list[[r]] <- cbind(out_list[[r]], cond[r,])
 
+    if(r %% 1000 == 0 | r == nrow(cond)){ #if even thousand or simulation finished
+
+        e <- rbindlist(out_list[r-999:r])
+
+        file_name <- paste0("../data/", effect_type, "_",
+                            meta_method, "_means_cond_", r-999,"-", r, ".csv")
+        fwrite(x = e, file = file_name)
+
+        rm(e)
+    }
+
 
 }
 
@@ -141,16 +152,6 @@ end - start
 # Clean up and save results
 #****************************************
 
-e <- rbindlist(out_list)
+#e <- rbindlist(out_list)
 
 #saveRDS(e, "../data_new/raw_r_tau_0-0.2.RDS")
-
-    if(r %% 1000 == 0 | r == nrow(cond)){ #if even thousand or simulation finished
-
-        e <- rbindlist(out_list[r-999:r])
-
-        file_name <- paste0("../data/simulation_means_cond_", r-999,"-", r, ".csv")
-        fwrite(x = e, file = file_name)
-
-        rm(e)
-    }

### 2023-06-09 code/supercomputer_simulator.r
minor code clarification

@@ -128,14 +128,16 @@ for(r in 1:nrow(cond)){ #gives us a list of lists
     out_list[[r]] <- out_list[[r]][, lapply(.SD, mean)] #data.table colMeans but returns a dataframe (well, data.table)
     out_list[[r]] <- cbind(out_list[[r]], cond[r,])
 
-    if(r %% 1000 == 0 | r == nrow(cond)){ #if even thousand or simulation finished
+    if(r %% 1000 == 0 | r == nrow(cond)){
+         #if even thousand or simulation finished
+         #turn results into dataframe and save as csv
 
         e <- rbindlist(out_list[r-999:r])
 
         file_name <- paste0("../data/", effect_type, "_",
                             meta_method, "_means_cond_", r-999,"-", r, ".csv")
-        fwrite(x = e, file = file_name)
 
+        fwrite(x = e, file = file_name)
         rm(e)
     }
 

### 2023-06-09 code/in-text-values.r
typical CI range

@@ -83,8 +83,38 @@ underestimate_mu02 <- list(underestimate_min = underestimate_min,
                            underestimate_max = underestimate_max)
 
 
-#*[1.3] bias when mu = 0.2 and tau => 0.1 ----
 #****************************************
+# [2] Discussion
+#****************************************
+k  <- 12
+
+sample_size = 150
+mu <- 0
+true_tau2 <- 0.17^2
+
+quick_sim <- function(k, sample_size, mu, true_tau2){
+    rho <- rnorm_truncated(n = k, mean = mu, sd = sqrt(true_tau2),
+                            lower_bound = -1, upper_bound = 1)
+    sampling_var_rho  <- (1-rho^2)^2 / (sample_size -1)
+
+    # given study rho, draw sample rho
+    r_se <- rnorm_truncated(k, mean = rho, sd = sqrt(sampling_var_rho),
+                            lower_bound = -1, upper_bound = 1)
+    r_var_estimate <- (1-r_se^2)^2 / (sample_size -1)
+    fit <- metafor::rma(yi = r_se, vi = r_var_estimate)
+
+    cis <- confint(fit)
+
+    CI_range <- cis$random[2, 3] - cis$random[2, 2]
+
+    CI_range
+}
+
+quick_sim(k = k, sample_size = sample_size, mu = mu, true_tau2 = true_tau2)
+
+set.seed(1532)
+res <- replicate(1e4, quick_sim(k = k, sample_size = sample_size, mu = mu, true_tau2 = true_tau2))
+median(res)
 #****************************************
 # [2] Output
 #****************************************

### 2023-06-10 code/functions.r
code: added error catching

@@ -24,6 +24,12 @@ rnorm_truncated <- function(
     qnorm(trunc_norm, mean = mean, sd = sd)
 }
 
+try_run <- function(code, silent = TRUE) {
+  tryCatch(code, error = function(c) {
+    return() #return null
+  })
+}
+
 simulate_rma <- function(
 
     effect_type = c("r", "r_z"),
@@ -97,27 +103,40 @@ simulate_rma <- function(
 
     # fit meta-analysis on observed values
     if(method == "HV"){
-        fit <- rma(yi = r_se_me, vi = r_var_estimate,
-                  control=list(stepadj=steplength, maxiter=maxiter))
+        fit <- try_run(rma(yi = r_se_me, vi = r_var_estimate,
+                  control=list(stepadj=steplength, maxiter=maxiter)))
 
     }else if(method == "HS"){ #https://www.metafor-project.org/doku.php/tips:hunter_schmidt_method
         r_var_estimate <- escalc(measure="COR", ri=r_se_me,
                       ni=rep(sample_size, k), vtype="AV")[["vi"]]
 
-        fit <- rma(yi = r_se_me, vi = r_var_estimate, weights = rep(sample_size, k),
-              control=list(stepadj=steplength, maxiter=maxiter))
+        fit <- try_run(rma(yi = r_se_me, vi = r_var_estimate, weights = rep(sample_size, k),
+              control=list(stepadj=steplength, maxiter=maxiter)))
 
     }
 
-
-    data.frame(intercept_b = fit$b,
-               intercept_p = fit$pval,
-               tau2_hat = fit$tau2,
-               tau2_p = fit$QEp)
+    if(is.null(fit)){ #i.e., non-convergence of Fisher scoring algorithm
+        return()
+    } else{
+        data.frame(intercept_b = fit$b,
+                   intercept_p = fit$pval,
+                   tau2_hat = fit$tau2,
+                   tau2_p = fit$QEp)
+    }
 }
 
-
-
+do_this <- function(input){
+    if(input > 0){
+        a = 3
+    }else{
+        a = NULL
+    }
+if(is.null(a)){
+    return()
+} else{
+    data.frame(1, 2)
+}
+}
 #****************************************
 # Plot prep functions
 #****************************************

### 2023-06-10 code/supercomputer_simulator.r

@@ -20,7 +20,7 @@ source("code/functions.r") #for rnorm_truncated and simulate_rma
 # Conditions
 #****************************************
 sample_size <- c(50, 100, 150, 200)
-k <- c(5, 20, 40, 200)
+k <- c(5, 20, 40)
 mu <- seq(from = 0, to = 0.6, by = 0.1)
 
 #Based on Flake et al., average alpha was .79, SD = .13, range .17 - .87
@@ -87,7 +87,7 @@ ncores <-parallel::detectCores()
 cl <- parabar::start_backend(ncores) # Create cluster based on nworkers.
 
 parabar::evaluate(cl, library(metafor))
-parabar::export(cl, c("simulate_rma", "rnorm_truncated"))
+parabar::export(cl, c("simulate_rma", "rnorm_truncated", "try_run"))
 
 #****************************************
 # Run simulation
@@ -125,21 +125,22 @@ for(r in 1:nrow(cond)){ #gives us a list of lists
 
     #Compute the mean across replications for the condition and add condition identifiers
     out_list[[r]] <- rbindlist(out_list[[r]]) #function from data.table
+    non_converged <- reps - nrow(out_list[[r]])
     out_list[[r]] <- out_list[[r]][, lapply(.SD, mean)] #data.table colMeans but returns a dataframe (well, data.table)
-    out_list[[r]] <- cbind(out_list[[r]], cond[r,])
+    out_list[[r]] <- cbind(out_list[[r]], cond[r,], non_converged)
 
-    if(r %% 1000 == 0 | r == nrow(cond)){
-         #if even thousand or simulation finished
-         #turn results into dataframe and save as csv
+    # if(r %% 1000 == 0 | r == nrow(cond)){
+    #      #if even thousand or simulation finished
+    #      #turn results into dataframe and save as csv
 
-        e <- rbindlist(out_list[r-999:r])
+    #     e <- rbindlist(out_list[r-999:r])
 
-        file_name <- paste0("../data/", effect_type, "_",
-                            meta_method, "_means_cond_", r-999,"-", r, ".csv")
+    #     file_name <- paste0("../data/", effect_type, "_",
+    #                         meta_method, "_means_cond_", r-999,"-", r, ".csv")
 
-        fwrite(x = e, file = file_name)
-        rm(e)
-    }
+    #     fwrite(x = e, file = file_name)
+    #     rm(e)
+    # }
 
 
 }

### 2023-06-10 code/supercomputer_simulator.r
code: improved intermittent saving

@@ -82,6 +82,7 @@ cond$maxiterations  <-  100 #default values are steplength = 1, and maxiter = 10
 #****************************************
 reps <- 1e1
 out_list <- vector("list", length = nrow(cond))
+last_save <- 0 #last condition row that was saved, see end of simulation below
 
 ncores <-parallel::detectCores()
 cl <- parabar::start_backend(ncores) # Create cluster based on nworkers.
@@ -129,18 +130,23 @@ for(r in 1:nrow(cond)){ #gives us a list of lists
     out_list[[r]] <- out_list[[r]][, lapply(.SD, mean)] #data.table colMeans but returns a dataframe (well, data.table)
     out_list[[r]] <- cbind(out_list[[r]], cond[r,], non_converged)
 
-    # if(r %% 1000 == 0 | r == nrow(cond)){
-    #      #if even thousand or simulation finished
-    #      #turn results into dataframe and save as csv
+    if(r %% 1000 == 0 | r == nrow(cond)){
+         #if even thousand or simulation finished
+         #turn results into dataframe and save as csv
 
-    #     e <- rbindlist(out_list[r-999:r])
+        save_range <- last_save + 1:r
 
-    #     file_name <- paste0("../data/", effect_type, "_",
-    #                         meta_method, "_means_cond_", r-999,"-", r, ".csv")
+        e <- rbindlist(out_list[save_range])
 
-    #     fwrite(x = e, file = file_name)
-    #     rm(e)
-    # }
+        file_name <- paste0("../data/", effect_type, "_",
+                            meta_method, "_means_cond_",
+                            save_range[1],"-", save_range[r],
+                            ".csv")
+
+        fwrite(x = e, file = file_name)
+        rm(e)
+        last_save <- r
+    }
 
 
 }

### 2023-06-10 code/supercomputer_simulator.r
code: fixed saving location

@@ -138,7 +138,7 @@ for(r in 1:nrow(cond)){ #gives us a list of lists
 
         e <- rbindlist(out_list[save_range])
 
-        file_name <- paste0("../data/", effect_type, "_",
+        file_name <- paste0("./data/", effect_type, "_",
                             meta_method, "_means_cond_",
                             save_range[1],"-", save_range[r],
                             ".csv")

### 2023-06-10 code/supercomputer_simulator.r
code: start simulation from last save

@@ -95,7 +95,7 @@ parabar::export(cl, c("simulate_rma", "rnorm_truncated", "try_run"))
 #****************************************
 
 start <- Sys.time()
-for(r in 1:nrow(cond)){ #gives us a list of lists
+for(r in last_save+1:nrow(cond)){ #gives us a list of lists
 
     progress_bar_format <- paste0( #for parabar
         "Condition ", r, "/", nrow(cond), ". [:bar] :percent [:elapsed]"

### 2023-06-10 code/supercomputer_simulator.r
combined all conditions into one run

@@ -20,67 +20,74 @@ source("code/functions.r") #for rnorm_truncated and simulate_rma
 # Conditions
 #****************************************
 sample_size <- c(50, 100, 150, 200)
-k <- c(5, 20, 40)
-mu <- seq(from = 0, to = 0.6, by = 0.1)
-
-#Based on Flake et al., average alpha was .79, SD = .13, range .17 - .87
-#Based on Sanchez-Meca, mean across 5 meta-analysis was 0.767 - 0.891 and SD ranged between 0.034 - 0.133
+k <- c(5, 20, 40, 200)
 reliability_mean <- c(0.6, 0.7, 0.8, 0.9, 1)
 reliability_sd <- seq(from = 0, to = 0.15, by = 0.05)
 
-effect_type  <- "r" # or r_z
-meta_method  <- "HV" #or HS
+mu <- seq(from = 0, to = 0.6, by = 0.2)
 
 #Using Pearson's r as the effect size.
+effect_type  <- "r" #
+meta_method  <- c("HV", "HS") #Hedges & Vevea and Hunter & Schmidt
+
 true_tau <- c(0, 0.1, 0.15, 0.2)
 true_tau2 <- true_tau^2
 
-if(effect_type == "r_z"){
-    #Using Fisher's z as the effect size (see 'compute_tau_fisher_z.r')
-
-    true_tau_50 <- c(0.1031263, 0.1566007, 0.2123514) #For N = 50
-    true_tau_100 <- c(0.1020357, 0.1549445, 0.2101057) #For N = 100
-    true_tau_150 <- c(0.1016845, 0.1544113, 0.2093826) #For N = 150
-    true_tau_200 <- c(0.1015112, 0.1541480, 0.2090256) #For N = 200
-    true_tau <- c(0, true_tau_50, true_tau_100, true_tau_150, true_tau_200)
-    true_tau2 <- true_tau^2
-
-    mu <- atanh(mu)
-}
-
-
-cond <- expand.grid(sample_size = sample_size,
+r_cond <- expand.grid(sample_size = sample_size,
                     k = k,
                     true_tau2 = true_tau2,
                     mu = mu,
                     reliability_mean = reliability_mean,
-                    reliability_sd = reliability_sd)
+                    reliability_sd = reliability_sd,
+                    effect_type = effect_type,
+                    method = meta_method)
+
+# Using Fisher' z as the effect size
[…]

### 2023-06-10 code/supercomputer_simulator.r
run subset of conditions only

@@ -17,18 +17,18 @@ library(data.table) #to bind lists into dataframe efficiently
 source("code/functions.r") #for rnorm_truncated and simulate_rma
 
 #****************************************
-# Conditions
+# Full conditions
 #****************************************
 sample_size <- c(50, 100, 150, 200)
 k <- c(5, 20, 40, 200)
 reliability_mean <- c(0.6, 0.7, 0.8, 0.9, 1)
 reliability_sd <- seq(from = 0, to = 0.15, by = 0.05)
 
-mu <- seq(from = 0, to = 0.6, by = 0.2)
+mu <- seq(from = 0, to = 0.6, by = 0.1)
 
 #Using Pearson's r as the effect size.
 effect_type  <- "r" #
-meta_method  <- c("HV", "HS") #Hedges & Vevea and Hunter & Schmidt
+r_method  <- c("HV", "HS") #Hedges & Vevea and Hunter & Schmidt
 
 true_tau <- c(0, 0.1, 0.15, 0.2)
 true_tau2 <- true_tau^2
@@ -40,11 +40,11 @@ r_cond <- expand.grid(sample_size = sample_size,
                     reliability_mean = reliability_mean,
                     reliability_sd = reliability_sd,
                     effect_type = effect_type,
-                    method = meta_method)
+                    method = r_method)
 
 # Using Fisher' z as the effect size
 effect_type  <- "r_z"
-meta_method  <- "HV"
+z_method  <- "HV"
 
 mu <- atanh(mu)
 
@@ -63,7 +63,7 @@ z_cond <- expand.grid(sample_size = sample_size,
                     reliability_mean = reliability_mean,
                     reliability_sd = reliability_sd,
                     effect_type = effect_type,
-                    method = meta_method)
+                    method = z_method)
 
 #drop incorrect tau-values for fisher's z
 keep_cond <- z_cond[["true_tau2"]] == 0 |
@@ -85,11 +85,56 @@ cond$step_length  <-  0.5 #decrease from 1 to 0.5 to improve convergence for low
 cond$maxiterations  <-  100 #default values are steplength = 1, and maxiter = 100
 
 #****************************************
+# Subset conditions
+#****************************************
+#1) Figure 1 main manuscript
+cond1 <- cond[cond$sample_size == 150 &
+              cond$k == 20 &
+              cond$reliability_sd ==0.15,]
+
+#2) Figure 2 main manuscript
[…]

### 2023-06-10 code/supercomputer_simulator.r
fixed typo

@@ -134,7 +134,7 @@ reps <- 1e4
 out_list <- vector("list", length = nrow(cond))
 last_save <- 0 #last condition row that was saved, see end of simulation below
 # save_points <- 1000 #save simulation results after how many conditions have run
-save_points <- save_point_subset
+save_points <- save_points_subset
 
 ncores <-parallel::detectCores()
 cl <- parabar::start_backend(ncores) # Create cluster based on nworkers.

### 2023-06-10 code/functions.r
removed old functions

@@ -125,53 +125,9 @@ simulate_rma <- function(
     }
 }
 
-do_this <- function(input){
-    if(input > 0){
-        a = 3
-    }else{
-        a = NULL
-    }
-if(is.null(a)){
-    return()
-} else{
-    data.frame(1, 2)
-}
-}
 #****************************************
 # Plot prep functions
 #****************************************
-
-grep_strip <- function(string, vec){
-    #Uses a common string like "k = " to extract "k = 50" or "k = 200"
-    #Then removes the "k = " part and returns as character
-    a <- grep(string, vec, value = TRUE)
-    gsub(string, "", a)
-}
-
-create_condition_cols <- function(data){
-
-    #takes a data frame as input with a column named "mu"
-    #in which all conditions are separated by a semicolon
-    # Then creates new columns named after each condition
-
-    #split all conditions into a character vector
-    conditions <- unlist(strsplit(data$mu, split = ";"))
-    #select all unique condition factors (e.g., mu, k, N)
-    category <- unique(gsub("=.*", "= ", conditions))
-    # Remove equal sign to column names (kept above in case some condition name include 'k')
-    col_name <- gsub(" = ", "", category)
-
-    for(g in 1:length(category)){
-
-        #This loop is equivalent to the line below, but applied to all conditions
-        # data$k <- grep_strip("k = ", conditions)
-
-        data[[col_name[g]]] <- grep_strip(category[g], conditions)
-
-    }
-
-    data #out
-}
 
 # Compute variance of a truncated normal distribution
 compute_var_truncated <- function(mu, nominal_tau, a = -1, b = 1){

### 2023-06-11 code/supercomputer_simulator.r
fixed mistake in save_range

@@ -186,7 +186,7 @@ for(r in last_save+1:nrow(cond)){ #gives us a list of lists
          #if reached a save point or simulation finished
          #turn results into dataframe and save as csv
 
-        save_range <- last_save + 1:r
+        save_range <- (last_save + 1):r
 
         e <- rbindlist(out_list[save_range])
 

### 2023-06-11 code/plots.r
updated main figures

@@ -15,11 +15,9 @@ library(ggplot2) #for plotting
 # General plot prep
 #****************************************
 
-dat_r <- readRDS("data/means_r_over_vs_underestimate.RDS")
-dat_z <- readRDS("data/means_z_over_vs_underestimate.RDS")
-dat_hs <- readRDS("data/means_r_HS_over_vs_underestimate.RDS")
 
-dat <- c(dat_r, dat_z, dat_hs)
+dat <- read.csv("data/means_combined.csv")
+
 
 # Add real truncated values
 es <- seq(0, 0.6, 0.1)
@@ -33,21 +31,24 @@ trunc <- data.frame(mu = es,
 
 
 #Prepping plot
-dat$true_tau2 <- as.numeric(dat$true_tau2)
-dat$mu <- as.numeric(dat$mu)
-
 dat$tau_hat <- sqrt(dat$tau2_hat)
 dat$nominal_tau <- sqrt(dat$true_tau2)
 
-dat$mean_rel <- factor(dat$mean_rel)
 dat$k <- factor(dat$k, levels = c("5", "20", "40", "200"))
-dat$N <- factor(dat$N, levels = c("50", "100", "150", "200"))
-
+dat$N <- factor(dat$sample_size, levels = c("50", "100", "150", "200"))
 
 #****************************************
-# Main plot 1 (N = 150, k = 20, r, r_z, and HS)
+# Main manuscript plots
 #****************************************
-dat_20_150 <- dat[dat$k == 20 & dat$N == 150,]
+
+# Figure 1 (N = 150, k = 20, r, r_z, and HS)
+#****************************************
+dat_20_150 <- dat[dat$k == 20 & dat$N == 150 &
+                 dat$reliability_mean < 1 &
+                 dat$reliability_sd == 0.15,]
+
+dat_20_150$reliability_mean <- factor(dat_20_150$reliability_mean)
+
 
 # For fisher's z there is no truncation,
 # create dataframe with all true heterogeneity levels
@@ -57,7 +58,7 @@ hs_true$effect_type <- "Hunter & Schmidt" #effect type is r, but named this way 
 
 z_true <- trunc
 z_true$effect_type <- "Fisher's z"
-dat_z <- dat_20_150[dat_20_150$effect_type == "r_z"]
+dat_z <- dat_20_150[dat_20_150$effect_type == "r_z",]
 z_tau <- rep(unique(dat_z$nominal_tau), each = 7) #4 tau-values across 4*7 conditions
[…]

### 2023-06-11 code/supercomputer_simulator.r
added supplement D simulation code

@@ -105,6 +105,7 @@ cond2$true_tau2 <- rep(true_tau2, each = 4)
 # Supplement A (variance in reliabilities)
 cond3 <- cond[cond$k == 20 &
               cond$sample_size == 150 &
+              cond$reliability_mean < 1 &
               cond$method == "HV" &
               cond$effect_type == "r",]
 
@@ -120,12 +121,18 @@ cond5 <- cond[cond$k == 20 &
               cond$method == "HV" &
               cond$effect_type == "r",]
 
+# Supplement D (Perfect reliability)
+cond6 <- cond[cond$k == 20 &
+              cond$sample_size == 150 &
+              cond$reliability_mean == 1,]
+
+
 #****************************************
 # Conditions run
 #****************************************
-cond <- rbind(cond1, cond2, cond3, cond4, cond5)
+cond <- rbind(cond1, cond2, cond3, cond4, cond5, cond6)
 
-save_points_subset <- cumsum(sapply(list(cond1, cond2, cond3, cond4, cond5), nrow))
+save_points_subset <- cumsum(sapply(list(cond1, cond2, cond3, cond4, cond5, cond6), nrow))
 
 #****************************************
 # Setup parallel simulation

### 2023-06-11 code/supercomputer_simulator.r
added option to not do intermediate saves

@@ -140,6 +140,7 @@ save_points_subset <- cumsum(sapply(list(cond1, cond2, cond3, cond4, cond5, cond
 reps <- 1e4
 out_list <- vector("list", length = nrow(cond))
 last_save <- 0 #last condition row that was saved, see end of simulation below
+intermediate_save <- TRUE
 # save_points <- 1000 #save simulation results after how many conditions have run
 save_points <- save_points_subset
 
@@ -189,23 +190,24 @@ for(r in last_save+1:nrow(cond)){ #gives us a list of lists
     out_list[[r]] <- out_list[[r]][, lapply(.SD, mean)] #data.table colMeans but returns a dataframe (well, data.table)
     out_list[[r]] <- cbind(out_list[[r]], cond[r,], non_converged)
 
-    if(r %in% save_points | r == nrow(cond)){
-         #if reached a save point or simulation finished
-         #turn results into dataframe and save as csv
+    if(intermediate_save){
+        if(r %in% save_points | r == nrow(cond)){
+            #if reached a save point or simulation finished
+            #turn results into dataframe and save as csv
 
-        save_range <- (last_save + 1):r
+            save_range <- (last_save + 1):r
 
-        e <- rbindlist(out_list[save_range])
+            e <- rbindlist(out_list[save_range])
 
-        file_name <- paste0("./data/means_cond_",
-                            save_range[1],"-", save_range[r],
-                            ".csv")
+            file_name <- paste0("./data/means_cond_",
+                                save_range[1],"-", save_range[r],
+                                ".csv")
 
-        fwrite(x = e, file = file_name)
-        rm(e)
-        last_save <- r
+            fwrite(x = e, file = file_name)
+            rm(e)
+            last_save <- r
+        }
     }
-
 
 }
 
@@ -221,4 +223,4 @@ end - start
 
 #e <- rbindlist(out_list)
 
-#saveRDS(e, "../data_new/raw_r_tau_0-0.2.RDS")
+#data.table::fwrite(x = e, file = "/data/means_combined.csv")

### 2023-06-11 code/supercomputer_simulator.r
code clarification

@@ -132,17 +132,18 @@ cond6 <- cond[cond$k == 20 &
 #****************************************
 cond <- rbind(cond1, cond2, cond3, cond4, cond5, cond6)
 
-save_points_subset <- cumsum(sapply(list(cond1, cond2, cond3, cond4, cond5, cond6), nrow))
+#save_points_subset <- cumsum(sapply(list(cond1, cond2, cond3, cond4, cond5, cond6), nrow))
 
 #****************************************
 # Setup parallel simulation
 #****************************************
 reps <- 1e4
 out_list <- vector("list", length = nrow(cond))
+
 last_save <- 0 #last condition row that was saved, see end of simulation below
-intermediate_save <- TRUE
-# save_points <- 1000 #save simulation results after how many conditions have run
-save_points <- save_points_subset
+intermediate_save <- FALSE
+save_points <- 1000 #save simulation results after how many conditions have run
+#save_points <- save_points_subset
 
 ncores <-parallel::detectCores()
 cl <- parabar::start_backend(ncores) # Create cluster based on nworkers.

### 2023-06-11 code/plots.r
added supplement A + D figures

@@ -127,6 +127,8 @@ dat_low <- dat[dat$N == 150 &
                dat$reliability_mean < 1 &
                dat$reliability_sd == 0.15,]
 
+dat_low$reliability_mean <- factor(dat_low$reliability_mean)
+
 # Truncated values (different from Figure 1 as tau is different)
 es <- seq(0, 0.6, 0.1)
 tau <- c(0.02, 0.04, 0.06, 0.08)
@@ -140,7 +142,6 @@ trunc_low <- data.frame(mu = es,
 trunc_low$true_tau <- trunc_low$nominal_tau
 trunc_low$tau_hat <- trunc_low$tau
 
-dat_low$reliability_mean <- factor(dat_low$reliability_mean)
 
 
 #Plot with fixed N and varying k, and small tau-values
@@ -158,8 +159,233 @@ theme_bw()
 
 #ggsave("figures/manuscript/r_tau_0.02-0.08.png", width = 8.62, height = 9.93)
 
+#****************************************
+# Supplement A plots
+#****************************************
+
+dat_a <- dat[dat$k == 20 &
+             dat$sample_size == 150 &
+             dat$reliability_mean < 1 &
+             dat$method == "HV" &
+             dat$effect_type == "r" &
+             dat$nominal_tau %in% c(0, 0.1, 0.15, 0.2),]
+
+dat_a$reliability_mean <- factor(dat_a$reliability_mean)
+dat_a$reliability_sd <- factor(dat_a$reliability_sd,
+                               levels = c(0.15, 0.1, 0.05, 0))
+
+# Truncated values (same as in Figure 1, row Pearson's r)
+es <- seq(0, 0.6, 0.1)
+tau <- c(0, 0.1, 0.15, 0.2)
+
+tau_trunc <- lapply(tau, compute_var_truncated, mu = es)
+tau_trunc <- unlist(tau_trunc)
+trunc_a <- data.frame(mu = es,
+                    nominal_tau = rep(tau, each = length(es)),
+                    true_tau = tau_trunc)
+
+trunc_a$true_tau <- trunc_a$nominal_tau
+trunc_a$tau_hat <- trunc_a$tau
+
+
+# NB! the object true_tau_line below comes from Figure 1 main manuscript section
+
+ggplot(dat_a, aes(x = mu, y = tau_hat)) +
+geom_line(aes(linetype = reliability_mean), show.legend = TRUE) +
+scale_linetype_manual(values = 2:5) +
[…]

### 2023-06-11 code/plots.r
condensed supplements A-C into A + created supplemental plots

@@ -163,6 +163,10 @@ theme_bw()
 # Supplement A plots
 #****************************************
 
+
+# Supplement A Figure 1 ((variance in reliabilities))
+#****************************************
+
 dat_a <- dat[dat$k == 20 &
              dat$sample_size == 150 &
              dat$reliability_mean < 1 &
@@ -201,13 +205,96 @@ xlab(expression("Average effect size "~mu)) +
 facet_grid(reliability_sd~nominal_tau, scales = "free") +
 theme_bw()
 
-ggsave("figures/supplement/supplement_A_fig1.png", width = 8.62, height = 9.93)
+#ggsave("figures/supplement/supplement_A_fig1.png", width = 8.62, height = 9.93)
+
+
+# Supplement A figure 2 (variable k)
+#****************************************
+dat_b <- dat[dat$sample_size == 150 &
+             dat$reliability_sd == 0.15 &
+             dat$nominal_tau %in% c(0, 0.1, 0.15, 0.2) &
+             dat$method == "HV" &
+             dat$effect_type == "r",]
+
+dat_b$reliability_mean <- factor(dat_b$reliability_mean)
+dat_b$k <- factor(dat_b$k)
+
+# Truncated values (same as in supplement A)
+es <- seq(0, 0.6, 0.1)
+tau <- c(0, 0.1, 0.15, 0.2)
+
+tau_trunc <- lapply(tau, compute_var_truncated, mu = es)
+tau_trunc <- unlist(tau_trunc)
+trunc_b <- data.frame(mu = es,
+                    nominal_tau = rep(tau, each = length(es)),
+                    true_tau = tau_trunc)
+
+trunc_b$true_tau <- trunc_b$nominal_tau
+trunc_b$tau_hat <- trunc_b$tau
+
+
+# NB! the object true_tau_line below comes from Figure 1 main manuscript section
+
+ggplot(dat_b, aes(x = mu, y = tau_hat)) +
+geom_line(aes(linetype = reliability_mean), show.legend = TRUE) +
+scale_linetype_manual(values = 2:5) +
+geom_line(aes(x = mu, y = true_tau), data = trunc_b, linetype = 1) +
+guides(linetype = guide_legend(reverse = TRUE, title = "Mean Reliability")) +
+expand_limits(y = 0) +
+ylab(expression("Between-studies standard deviation "~tau)) +
+xlab(expression("Average effect size "~mu)) +
[…]

### 2023-06-11 code/supercomputer_simulator.r

@@ -102,26 +102,26 @@ cond2 <- cond[cond$sample_size == 150 &
               cond$effect_type == "r",]
 cond2$true_tau2 <- rep(true_tau2, each = 4)
 
-# Supplement A (variance in reliabilities)
+# Supplement A1 (variance in reliabilities)
 cond3 <- cond[cond$k == 20 &
               cond$sample_size == 150 &
               cond$reliability_mean < 1 &
               cond$method == "HV" &
               cond$effect_type == "r",]
 
-# Supplement B (variable K)
+# Supplement A2 (variable K)
 cond4 <- cond[cond$sample_size == 150 &
               cond$reliability_sd == 0.15 &
               cond$method == "HV" &
               cond$effect_type == "r",]
 
-# Supplement C (variable sample size)
+# Supplement A3 (variable sample size)
 cond5 <- cond[cond$k == 20 &
               cond$reliability_sd == 0.15 &
               cond$method == "HV" &
               cond$effect_type == "r",]
 
-# Supplement D (Perfect reliability)
+# Supplement B (Perfect reliability)
 cond6 <- cond[cond$k == 20 &
               cond$sample_size == 150 &
               cond$reliability_mean == 1,]

### 2023-06-11 supplement/supplement_A.qmd
draft of supplement A + B

---
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Supplement A: varying reliability SD and the effect of N and k

```{r Figure A1, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_A_fig1.png")
```

```{r Figure A2, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_A_fig2.png")
```

```{r Figure A3, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_A_fig3.png")
```

### 2023-06-11 supplement/supplement_B.qmd
draft of supplement A + B

---
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Supplement B: perfect reliability and Pearson's $r$

Figure 1 shows the bias in heterogeneity estimates in the absence of measurement unreliability in primary studies (i.e., reliability = 1).

```{r Figure 1, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_B_fig1.png")
```

_Figure 1._ Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1). The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).

As can be seen, all three models show a positive bias when heterogeneity is zero, as heterogeneity estimates are truncated at zero. However, we also see a positive bias when heterogeneity is above zero for the Pearson's correlation using the HV method (second row). This is due to a dependancy between effect size and sampling variance for Pearson's $r$ as normally computed. That is, the sampling variance is estimated as $\hat{\sigma}^2_i = \frac{(1-r_i^2)^2}{n_i - 1}$. We do not see this effect for the HS method or HV method using Fisher's $z$. In the HS method $r_i$ has been replaced by $\bar{r}$ and for Fisher's $z$ the sampling variance is approximated as $\hat{\sigma}^2_i = 1 / (n_i -3)$, in both cases removing the dependency between effect size and sampling variance. To demonstrate that this is indeed the case, Figure 2 demonstrates how the bias in Pearson's $r$ disappears if we reshuffle the sampling variances among primary studies so that the dependency between effect size and sampling variance disappears.

```{r Figure 2, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_B_fig2.png")
```

_Figure 2._ Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 1e3 repetitions. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).

### 2023-06-11 code/examining_REML_underestimate_perfect_reliability.r
removed unused folders

@@ -1,151 +1,0 @@
-# Project: unexplained heterogeneity
-# Script purpose: Explore the underestimate of tau2 by REML for large tau2 values
-
-
-#****************************************
-# Replicating figure 2 in Langan et al. (2019)
-#****************************************
-# langan: a comparison of heterogeneity variance estimators in simulated random-effects
-#https://onlinelibrary.wiley.com/doi/10.1002/jrsm.1316
-
-
-simulate_rma_langan <- function(version = c("original", "efficient"), k = 20, true_tau2 = 0.0444, mu = 0.5, sample_size = 20){ #see section 3.1
-
-    mu_i <- rnorm(n = k, mean = mu, sd = sqrt(true_tau2))
-    n_c = n_t = rep(sample_size, k)
-
-    if(version == "original"){
-    n1 <- replicate(k, rnorm(n = sample_size, mean = 0, sd = 1), simplify = FALSE)
-    n2 <- lapply(1:k, function(i) rnorm(n = sample_size, mean = mu_i[i], sd = 1))
-
-    avg_c <- lapply(n1, mean) #function below from heterogeneity project #1
-    var_c <- lapply(n1, var)
-    avg_c <- do.call(rbind, avg_c)
-    var_c <- do.call(rbind, var_c)
-
-    avg_t <- lapply(n2, mean)
-    var_t <- lapply(n2, var)
-    avg_t <- do.call(rbind, avg_t)
-    var_t <- do.call(rbind, var_t)
-
-    fit <- rma(measure = "SMD",  m1i = avg_t, m2i = avg_c, sd1i = sqrt(var_t),
-                   sd2i = sqrt(var_c), n1i = n_t, n2i = n_c, method = "REML") #fit meta-analysis transforming into standardized mean difference (Hedge's g)
-
-    }else if(version == "efficient"){
-
-        avg_c <- rnorm(n = k, mean = 0, sd = 1 / sqrt(n_c)) #Draw means from sampling distribution control group
-        avg_t <- rnorm(n = k, mean = mu_i, sd = 1 / sqrt(n_t)) #Draw means from sampling distribution treatment group
-        var_c <- rchisq(n = k, df = n_c - 1) / (n_c - 1) #draw variances from sampling distribution control group
-        var_t <- rchisq(n = k, df = n_t - 1) / (n_t - 1) #draw variances from sampling distribution treatment group
-
-        fit <- rma(measure = "SMD",  m1i = avg_t, m2i = avg_c, sd1i = sqrt(var_t),
[…]

### 2023-06-11 code/parabar_demonstration.r

@@ -1,91 +1,0 @@
-# Project: unexplained heterogeneity
-# Code purpose: demonstrate how to add a progress bar in parallel computing using the 'parabar' package
-
-
-sample_size <- 150
-k <- 20
-true_tau2 <- c(0.002, 0.0035, 0.0055, 0.0085, 0.012, 0.0185, 0.031, 0.069) #corresponds to 20 -90% I2 according to my simulations for Pearson's r and the above sample size and k
-mu <- seq(from = 0, to = 0.6, by = 0.1)
-reliability_min <- seq(from = 1, to = 1, by = 0.1)
-reps <- 1e4
-
-cond <- expand.grid(sample_size = sample_size,
-                    k = k,
-                    true_tau2 = true_tau2,
-                    mu = mu,
-                    reliability_min = reliability_min)
-
-cond$reliability_max <- cond$reliability_min
-
-out_list <- vector("list", length = nrow(cond))
-
-
-library(parabar)
-
-start <- Sys.time()
-
-ncores <- detectCores()
-# cl <- makePSOCKcluster(ncores) # Create cluster based on nworkers.
-cl <- start_backend(ncores)
-
-# clusterEvalQ(cl, library(metafor))
-# clusterExport(cl, c("simulate_rma", "rnorm_truncated"))
-
-evaluate(cl, library(metafor))
-export(cl, c("simulate_rma", "rnorm_truncated"))
-peek(cl)
-
-for(r in 1:nrow(cond)){ #gives us a list of lists
-
-    progress_bar_format <- paste0(
-        "Condition ", r, "/", nrow(cond), ". [:bar] :percent [:elapsed]"
-    )
-
-    configure_bar(type = "modern", format = progress_bar_format)
-
-    # mes <- paste0("\n now on condition ", r)
-    # cat(mes)
-
-    cond_r <- cond[r,]
-    #need to add because otherwise the nodes don't find r. Since this happens outside the nodes.
-
-    # clusterExport(cl, "cond_r") #exported row to each node
-    export(cl, "cond_r")
-
-    task <- function(iteration){ #anonymous function needed when using for replications
-        simulate_rma(effect_type = "r",
-            reliability_distribution = "uniform",
-            k = cond_r$k,
-            sample_size = cond_r$sample_size,
-            true_tau2 = cond_r$true_tau2,
-            mu = cond_r$mu,
[…]

### 2023-06-11 code/summarize_raw_data.r

@@ -1,69 +1,0 @@
-# Project: Unexplained heterogeneity
-# Code purpose: cleaning raw data
-# Author: Anton Olsson-Collentine (anton@olssoncollentine.com)
-
-
-#****************************************
-# functions and packages
-#****************************************
-
-source("./code/functions.r") # for create_condition_cols function
-
-
-#****************************************
-# Create dataset with means
-#****************************************
-#e_medians <- lapply(e, function(x) apply(x, 2, median))
-
-#With r
-dat_r <- readRDS("data/raw_r_tau_0-0.2.RDS")
-r_means <- lapply(dat_r, colMeans)
-
-#Turn list into dataframe using list-names as idcol
-r_means <- lapply(r_means, function(x) as.data.frame(as.list(x)))
-r_means <- data.table::rbindlist(r_means, idcol = "mu")
-
-#create condition ids
-r_means <- create_condition_cols(r_means)
-
-saveRDS(r_means, "data/means_r_tau_0-0.2.RDS")
-
-#With z
-dat_z <- readRDS("data/raw_z_tau_0-0.2.RDS")
-z_means <- lapply(dat_z, colMeans)
-
-#Turn list into dataframe using list-names as idcol
-z_means <- lapply(z_means, function(x) as.data.frame(as.list(x)))
-z_means <- data.table::rbindlist(z_means, idcol = "mu")
-
-#create condition ids
-z_means <- create_condition_cols(z_means)
-
-saveRDS(z_means, "data/means_z_tau_0-0.2.RDS")
-
-
-#With hs
-dat_hs <- readRDS("data/raw_r_hs_tau_0-0.2.RDS")
-hs_means <- lapply(dat_hs, colMeans)
-
-#Turn list into dataframe using list-names as idcol
-hs_means <- lapply(hs_means, function(x) as.data.frame(as.list(x)))
-hs_means <- data.table::rbindlist(hs_means, idcol = "mu")
-
-#create condition ids
-hs_means <- create_condition_cols(hs_means)
-
-saveRDS(hs_means, "data/means_hs_tau_0-0.2.RDS")
-
-#With low tau
-dat_low <- readRDS("data/raw_r_tau_0.02-0.08.RDS")
-low_means <- lapply(dat_low, colMeans)
-
-#Turn list into dataframe using list-names as idcol
-low_means <- lapply(low_means, function(x) as.data.frame(as.list(x)))
[…]

### 2023-06-11 supplement_reliability_Maassen/extract_reliabilities-maassen.r

@@ -1,32 +1,0 @@
-# Project: Unexplained heterogeneity
-# Code purpose: estimate typical reliability and its SD from E. Maassen's data
-# Author: Anton Olsson-Collentine (anton@olssoncollentine.com)
-
-
-#******************************************
-library(openxlsx)
-
-
-dat <- openxlsx::read.xlsx("supplement_reliability_Maassen/codebook-main-step4_rel.xlsx")
-
-#From Esther:
-#Your columns of interest are rel_rep and rel_est.
-#(i.e., reported and estimated reliabilities)
-
-#Note that this reliability may mean nothing, given that it’s on all the items, and therefore one factor is assumed, even though there may be more factors (f_total) underlying the data.
-
-mean(dat$rel_rep)
-#[1] 0.8392653
-sd(dat$rel_rep)
-#[1] 0.0997923
-
-mean(dat$rel_est, na.rm = TRUE)
-#[1] 0.8657143
-sd(dat$rel_est, na.rm = TRUE)
-#[1] 0.07702407
-
-# Higher estimates and with less SD when recomputed. Likely a because of a correlation with who made their data openly available.
-
-
-dat$comp_rel <- dat$rel_rep - dat$rel_est
-mean(dat$comp_rel, na.rm = TRUE)

### 2023-06-11 supplemental_data/brannik_raw-data-simulation.r

@@ -1,71 +1,0 @@
-
-
-# Brannick et al., simulation
-
-# 2 ideas that might be the difference
-# a) add measurement error before sampling error
-#   - This does not explain it, I just tried with my
-#     function and it still underestimates tau by about 0.03
-# b) Compute observed correlations from a sample
-# (NB. b) interacts with a))
-#   - This also doesn't explain it, I just used Brannick's
-# code below and there is still an underestimate when using
-# meta-analyzing unreliabile data
-
-TxTy <- rnorm(10, mean = .26, sd = .13)
-rel_x <- runif(10, min = 0, max = 1)
-rel_y <- runif(10, min = 0, max = 1)
-
-TxX <- sqrt(rel_X)
-TyY <- sqrt(rel_y)
-xy <- TxTy*TxX*TyY
-
-#Compute attenuated correlations
-n <- 142
-
-rhobox <- matrix(0,4,4)+diag(4) # placeholder for correlation matrix for sampling
-mu <- rep(0,4)          # multivariate means of zero
-names.sim <- cbind("Tx", "Ty", "X", "Y")
-colnames(rhobox)<- names.sim
-rownames(rhobox)<- names.sim
-
-obs <- vector("double", 10)
-res <- vector("double", 1e3)
-
-for(rep in 1:1e3){
-
-for(j in 1:10){
-
-ryyi <- 0.8
-rxxi <- 0.8
-rho.r <- rnorm(1, mean = .26, sd = .13)
-att.rho.r <- rho.r*sqrt(rxxi*ryyi)              # attenuate rho for reliability
-
-# find the population correlation matrix from which to sample observations
-rhobox[2,1] <- rho.r                            # correlation between true scores
-rhobox[1,2] <- rho.r
-rhobox[3,1] <- sqrt(rxxi)                       # true x and observed X
-rhobox[1,3] <- sqrt(rxxi)
-rhobox[3,2] <- rho.r*sqrt(rxxi)                 # true Y and observed X
-rhobox[2,3] <- rho.r*sqrt(rxxi)
-rhobox[4,2] <- sqrt(ryyi)                       # true Y and observed Y
-rhobox[2,4] <- sqrt(ryyi)
-rhobox[4,1] <- rho.r*sqrt(ryyi)                 # true X and observed Y
-rhobox[1,4] <- rho.r*sqrt(ryyi)
[…]

### 2023-06-11 supplemental_data/check-sample-sizes.r

@@ -1,17 +1,0 @@
-# Check sample sizes in https://journals.sagepub.com/doi/full/10.1177/2515245919838781
-
-
-
-library(haven)
-
-dat <- haven::read_sav("supplemental_data_max_tau2/data.sav")
-quantile(dat$Sample)
-#1300 rows
-
-dat2 <- dat[dat$Sample < 1e4,]
-#1291 rows
-quantile(dat2$Sample, na.rm = TRUE)
-
-dat3 <- dat2[dat2$Sample < 5e3,]
-#1290 rows
-quantile(dat3$Sample)

### 2023-06-11 supplemental_data/checking-metafor-psychmeta-borenstein.r

@@ -1,112 +1,0 @@
-# Project: Unexplained heterogeneity
-# Code purpose: Checking other's corrections for unreliability
-# Author: Anton Olsson-Collentine (anton@olssoncollentine.com)
-
-
-
-
-# Metafor example
-# https://www.metafor-project.org/doku.php/tips:hunter_schmidt_method
-
-library(metafor)
-dat <- dat.mcdaniel1994
-dat <- escalc(measure="COR", ri=ri, ni=ni, data=dat, vtype="AV")
-res <- rma(yi, vi, weights=ni, data=dat, method="HS")
-res
-
-# NB! Wolfgang shows increase in heterogeneity after corrected
-# for unreliability!
-
-
-# Psych-meta example
-
-# Data from: https://cran.r-project.org/web/packages/psychmeta/vignettes/ma_r.html
-# But only for X and Y (not Z)
-dat <- matrix(
-    c("1", "1", "X", "Y", "416", "0.49", "0.79", "0.77", "Watson2005",
-    "2", "1", "X", "Y", "241", "0.54", "0.82", "0.84", "Watson2005",
-    "3", "1", "X", "Y", "479", "0.34", "0.73", "0.87", "Zellars2006",
-    "4", "1", "X", "Y", "167", "0.37", "0.78", "0.79", "Bandura1980"),
- ncol = 9, byrow = TRUE)
-
-dat <- as.data.frame(dat)
-names(dat) <- c("sample_id", "moderator","x_name","y_name","n","rxyi", "rxxi","ryyi", "citekey")
-dat[,5] <- as.numeric(dat[,5])
-dat[,6] <- as.numeric(dat[,6])
-dat[,7] <- as.numeric(dat[,7])
-dat[,8] <- as.numeric(dat[,8])
-
-
-install.packages("psychmeta")
-library(psychmeta)
-
-# Without correction for unreliability
-fit1 <- ma_r(rxyi = rxyi,
-               n = n,
-               construct_x = x_name,
-               construct_y = y_name,
-               rxx = rxxi,
-               ryy = ryyi,
-               sample_id = sample_id,
-               data = dat
-               )
-
-summary(fit1)
-# tau  0.0842 (sd_res)
-
-# https://cran.r-project.org/web/packages/psychmeta/vignettes/overview.html+
-# section "Individual-correction meta-analyses"
-
-fit2 <- ma_r(ma_method = "ic",
-               rxyi = rxyi,
[…]

### 2023-06-11 supplemental_data/empirical_tau2_values.r

@@ -1,67 +1,0 @@
-# Project: Unexplained heterogeneity
-# Script purpose: estimate common tau2 values for correlations in psychology
-# based on van Erp et al. (2017): 10.5334/jopd.33
-# Code: Anton Olsson-Collentine
-
-#**********************************************
-# Setup
-#**********************************************
-# Data downloaded from https://osf.io/p6ecw
-
-dat <- read.csv("supplemental_data/Data 1990-2013 with tau values.csv")
-
-unique(dat$Type.of.ES)
-#also contains corrected Pearson's r
-
-dat_r <- dat[dat$Type.of.ES == "Pearson's r",]
-
-hist(dat_r$tau.2, breaks = 30)
-boxplot(dat_r$tau.2)
-# We see that 50% of tau2 values are between about ~0.02- 0.06,
-#with another 25% between 0.06- 0.125 and 25% below 0.02
-quantile(dat_r$tau)
-#Actually, the 25th quantile is 0.01, median 0.03, and the 75th is 0.0576
-#the maximum value is 0.2704, which gives tau = 0.52
-
-
-# For corrected pearson's rs
-dat_r_corrected <- dat[dat$Type.of.ES == "Corrected Pearson's r",]
-hist(dat_r_corrected$tau.2, breaks = 30)
-boxplot(dat_r_corrected$tau.2)
-quantile(dat_r_corrected$tau)
-# The corrected rs have a lower distribution
-# 25th quantiles = 0.001, median = 0.01, 75th = 0.0225,
-#The maximum is 0.2209
-
-quantile(dat_r_corrected$X..of.effect.sizes)
-dat_r_corrected2 <- dat_r_corrected[dat_r_corrected$X..of.effect.sizes < 1000,]
-quantile(dat_r_corrected2$X..of.effect.sizes)
-
-
-# Check K
-dat_r_any <- dat[grepl("Pearson", dat$Type.of.ES),]
-unique(dat_r_any$Type.of.ES)
-quantile(dat_r_any$X..of.effect.sizes)
-hist(dat_r_any$X..of.effect.sizes)
-
-quantile(dat_r$X..of.effect.sizes)
-hist(dat_r$X..of.effect.sizes)
-
-dat_r_any2 <- dat_r_any[dat_r_any$X..of.effect.sizes < 1000,]
-quantile(dat_r_any2$X..of.effect.sizes)
-
-# Check correction or not
[…]

### 2023-06-11 code/03_supercomputer_simulator.r
improved script names

@@ -14,7 +14,7 @@ library(data.table) #to bind lists into dataframe efficiently
 #library(metafor) #loaded when setting up simulations
 
 
-source("code/functions.r") #for rnorm_truncated and simulate_rma
+source("code/02_functions.r") #for rnorm_truncated and simulate_rma
 
 #****************************************
 # Full conditions

### 2023-06-11 code/04_in-text-values.r

@@ -6,7 +6,7 @@
 # functions and packages
 #****************************************
 
-source("./code/functions.r") # for create_condition_cols function
+source("./code/02_functions.r") # for create_condition_cols function
 
 
 #****************************************

### 2023-06-11 code/05_plots.r

@@ -7,7 +7,7 @@
 # functions and packages
 #****************************************
 
-source("./code/functions.r") # for plot prep functions
+source("./code/02_functions.r") # for plot prep functions
 
 library(ggplot2) #for plotting
 
@@ -462,7 +462,6 @@ trunc <- data.frame(mu = es,
 
 
 # Plot D2.
-
 
 ggplot(e) +
 geom_line(aes(x = mu, y = tau_hat), linetype = 2) +

### 2023-06-11 code/03_supercomputer_simulator.r
removed duplicated conditions

@@ -132,6 +132,7 @@ cond6 <- cond[cond$k == 20 &
 #****************************************
 cond <- rbind(cond1, cond2, cond3, cond4, cond5, cond6)
 
+cond <- cond[!duplicated(cond),]
 #save_points_subset <- cumsum(sapply(list(cond1, cond2, cond3, cond4, cond5, cond6), nrow))
 
 #****************************************

### 2023-06-11 code/04_in-text-values.r
updated in-text-value code

@@ -12,17 +12,23 @@ source("./code/02_functions.r") # for create_condition_cols function
 #****************************************
 # [1] Result section
 #****************************************
-dat_r <- readRDS("data/means_r_tau_0-0.2.RDS")
+dat <- read.csv("data/means_combined.csv")
 
-dat_r$tau_hat <- sqrt(dat_r$tau2_hat)
-dat_r$true_tau <- sqrt(as.numeric(dat_r$true_tau2))
+dat$tau_hat <- sqrt(dat$tau2_hat)
+dat$true_tau <- sqrt(as.numeric(dat$true_tau2))
+dat$k <- factor(dat$k, levels = c("5", "20", "40", "200"))
+dat$N <- factor(dat$sample_size, levels = c("50", "100", "150", "200"))
 
-dat <- dat_r[dat_r$k == "20" & dat_r$N == "150",]
+dat <- dat[dat$k == "20" &
+           dat$N == "150" &
+           dat$reliability_sd == 0.15 &
+           dat$method == "HV" &
+           dat$effect_type == "r",]
 
 #*[1.1] bias when tau =0, R = 0.8 ----
 #****************************************
 
-dat_1.1 <- dat[dat$true_tau == 0 & dat$mean_rel == 0.6 ,]
+dat_1.1 <- dat[dat$true_tau == 0 & dat$reliability_mean == 0.8 ,]
 #mu = 0.2
 mu_0.2 <- dat_1.1$tau_hat[dat_1.1$mu == 0.2] - dat_1.1$true_tau[dat_1.1$mu == 0.2]
 mu_0.2 <- round(mu_0.2, 3)
@@ -35,30 +41,30 @@ bias_tau_zero <- data.frame(mu_0.2, mu_0.4)
 
 #*[1.2] bias when tau increasing, mu = 0.2, R = 0.8 ----
 #****************************************
-dat_1.2 <- dat[dat$mu == 0.2 & dat$mean_rel == 0.8,]
+dat_1.2 <- dat[dat$mu == 0.2 & dat$reliability_mean == 0.8,]
 #tau = 0.1
 tau_hat <- dat_1.2$tau_hat[dat_1.2$true_tau == 0.1]
 true_tau <- dat_1.2$true_tau[dat_1.2$true_tau == 0.1]
 bias_0.1 <-  tau_hat - true_tau
-percent_0.1 <- 100*(tau_hat / true_tau)
+prop_0.1 <- 1 - (tau_hat / true_tau)
 
[…]

### 2023-06-13 supplement/supplement_A.qmd
sample size in supplement A

@@
 
+Lower standard deviation in reliability leads to a more severe underestimate for large effect sizes.
+Figure A1 demonstrates this effect for four levels of between-studies standard deviation (columns Figure A1) and four levels of reliability standard deviation (rows Figure A1) when sample size (N) is set to 150 within studies and the number of meta-analyzed studies (K) is 20.
+All results in this manuscript are based on the outcome variable Pearsons's $r$ estimated with Hedges & Vevea (1998) approach and Restricted Maximum Likelihood (REML) and averaged across 10,000 replications.
+As can be seen in Figure A1, once reliability standard deviation reaches 0.05 (third row) there is practically no upward inflation of heterogeneity remaining, only just sufficient to compensate for the suppression due to the truncation of Pearson's $r$ at {-1, 1}.
+A such, especially in the case of large effect sizes heterogeneity can be expected to be somewhat more negatively biased as reliability standard deviation decreases.
+
 ```{r Figure A1, echo=FALSE, out.width = '100%'}
@@
 ```
+
+_Figure A1._ Inflation of effect sizes decreases with lower reliability SD leading to larger net negative bias in heterogeneity, especially for large effect sizes and moderate heterogeneity.
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
+Columns indicate nominal true between-studies standard deviation and rows indicate reliability standard deviation.
+Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Each dashed line correspond to an average reliability in primary studies.
+
+Lower number of meta-analyzed studies leads to a larger inflation of heterogeneity at low levels of true heterogeneity (Figure A2).
+In Figure A2, columns are again nominal true heterogeneity, whereas rows in Figure A2 are the number of studies in each meta-analysis.
+Generally, we would expect a lower number of studies to lead to larger sampling variance of heterogeneity estimates, hence a larger proportion of estimates truncated at zero, and consequently more inflated average heterogeneity estimates when number of studies is low (e.g., first row Figure A2) and heterogeneity is small or absent (first column Figure A2).
+This is also what we see in Figure A2.
+In addition, the number of studies mostly affects heterogeneity estimates when average effect size is small.
 
@@
 
+_Figure A2._ Expected inflation of effect sizes decreases with larger number of meta-analyzed when heterogeneity is absent or small and average effect size is small.
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
+Columns indicate nominal true between-studies standard deviation and rows indicate number of meta-analyzed studies.
+Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Each dashed line correspond to an average reliability in primary studies.
+
+```{r, include = FALSE}
+
+dat <- read.csv("data/means_combined.csv")
+dat$tau_hat <- sqrt(dat$tau2_hat)
+dat$nominal_tau <- sqrt(dat$true_tau2)
+dat$k <- factor(dat$k, levels = c("5", "20", "40", "200"))
+dat$N <- factor(dat$sample_size, levels = c("50", "100", "150", "200"))
+
+# differences are largest when mu = 0 and reliability is low
+dat_a3 <- dat_c[dat_c$sample_size %in% c(50, 200) &
+                dat_c$reliability_mean == 0.6 &
+                dat_c$mu == 0,]
+
+res1 <- split(dat_a3, dat_a3$nominal_tau)
+res1 <- lapply(res1, function(x) {
+    x[1, "tau_hat"] - x[2, "tau_hat"]
+})
+res1 <- abs(unlist(res1))
+
+#For standard reliability sample size can matter more
+dat_a3 <- dat_c[dat_c$sample_size %in% c(50, 200) &
+                dat_c$reliability_mean == 0.8 &
+                dat_c$mu == 0,]
+
+res2 <- split(dat_a3, dat_a3$nominal_tau)
+res2 <- lapply(res2, function(x) {
+    x[1, "tau_hat"] - x[2, "tau_hat"]
+})
+res2 <- abs(unlist(res2))
+
+```
+
+A smaller sample size in primary studies tends exacerbate the bias in observed heterogeneity estimates due to unreliability (Figure A3).
+In Figure A3, columns are again nominal true heterogeneity, whereas rows in Figure A3 are the (fixed) sample size within primary studies.
+Figure A3 shows that when true heterogeneity is zero (leftmost column) a smaller sample size tends to lead to a more positive bias in estimated heterogenetiy, although in the exceptional case when average effect size is also large ($\mu > 0.5$), large reliability variability increases heterogeneity less than when sample size is larger.
+When true heterogeneity is non-zero, smaller sample size corresponds with a larger negative bias in heterogenetiy estimates.
+For higher levels of reliability differences are small, but can be more serious when average reliability is low.
+For example, for $\bar{R} = 0.8$, the difference in estimate is at most $\hat{\tau} = 0.01$ for true heterogeneity above 0.1 when comparing a sample size of 50 versus 200 (column second from right, topmost row versus bottom row), whereas when $\bar{R} = 0.6$, the difference can be as much as $\hat{\tau} = 0.03$ (column second from right, topmost row versus bottom row).
+
 ```{r Figure A3, echo=FALSE, out.width = '100%'}
@@
 ```
+
+_Figure A3._ A smaller sample size in primary studies tends exacerbate the bias in observed heterogeneity estimates due to unreliability.
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
+Columns indicate nominal true between-studies standard deviation and rows indicate number of meta-analyzed studies.
+Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Each dashed line correspond to an average reliability in primary studies.

### 2023-06-13 code/05_plots.r
corrected truncation lines figure A3

@@ -271,9 +271,6 @@ trunc_c <- data.frame(mu = es,
                     nominal_tau = rep(tau, each = length(es)),
                     true_tau = tau_trunc)
 
-trunc_c$true_tau <- trunc_c$nominal_tau
-trunc_c$tau_hat <- trunc_c$tau
-
 
 # NB! the object true_tau_line below comes from Figure 1 main manuscript section
 

### 2023-06-14 supplement/supplement_A.qmd
changed supplement names to match manuscript appearance order

--- 2023-06-11 supplement/supplement_A.qmd
@@
 ---
+
 format: pdf
+
 ---
+# Supplement A: converting heterogeneity between Pearson's r and Fisher's z
 
@@
 
-# Supplement A: varying reliability SD and the effect of N and k
+The table below shows the results of our procedure to convert between studies variance expressed in Pearson's $r$ to between-studies variance expressed in Fisher's $z$ through the use of $I^2$ as described in the main text.
 
-```{r Figure A1, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/supplement/supplement_A_fig1.png")
+_Table A1_.
+```{r}
+
+source("../code/01_compute_tau_fishers_z.r") #all computations were made here
+
+#Gives us the object 'cond'
+cond_table <- cond[,c(1, 3, 2, 4, 6)]
+
+knitr::kable(cond_table, booktabs = TRUE, col.names = c("$\\rho$", "N", "$\\tau_r$", "$I^2$", "$\\tau_z$"), escape = FALSE)
 ```
 
-```{r Figure A2, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/supplement/supplement_A_fig2.png")
-```
-
-```{r Figure A3, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/supplement/supplement_A_fig3.png")
-```
+_Note._ $\rho$ = average effect size, N = (fixed) within-study sample size, $\tau_r$ = between-studies standard deviation in Pearson's $r$, $\tau_z$ between-studies standard deviation in Fisher's $z$.
+The $I^2$ statistic is here expressed as a proportion.

### 2023-06-14 supplement/supplement_B.qmd
changed supplement names to match manuscript appearance order

@@
 
-```{r Figure 1, echo=FALSE, out.width = '100%'}
+```{r Figure B1, echo=FALSE, out.width = '100%'}
 knitr::include_graphics("../figures/supplement/supplement_B_fig1.png")
@@
 
-_Figure 1._ Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1).
+_Figure B1._ Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1).
 The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
@@
 
-```{r Figure 2, echo=FALSE, out.width = '100%'}
+```{r Figure B2, echo=FALSE, out.width = '100%'}
 knitr::include_graphics("../figures/supplement/supplement_B_fig2.png")
@@
 
-_Figure 2._ Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$.
+_Figure B2._ Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$.
 The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 1e3 repetitions.

### 2023-06-14 supplement/supplement_C.qmd
changed supplement names to match manuscript appearance order

---
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Supplement C: varying reliability SD and the effect of K

Lower standard deviation in reliability leads to a more severe underestimate for large effect sizes. Figure C1 demonstrates this effect for four levels of between-studies standard deviation (columns Figure C1) and four levels of reliability standard deviation (rows Figure C1) when sample size (N) is set to 150 within studies and the number of meta-analyzed studies (K) is 20. All results in this manuscript are based on the outcome variable Pearsons's $r$ estimated with Hedges & Vevea (1998) approach and Restricted Maximum Likelihood (REML) and averaged across 10,000 replications. As can be seen in Figure C1, once reliability standard deviation reaches 0.05 (third row) there is practically no upward inflation of heterogeneity remaining, only just sufficient to compensate for the suppression due to the truncation of Pearson's $r$ at {-1, 1}. A such, especially in the case of large effect sizes heterogeneity can be expected to be somewhat more negatively biased as reliability standard deviation decreases.

```{r Figure C1, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_C_fig1.png")
```

_Figure C1._ Inflation of effect sizes decreases with lower reliability SD leading to larger net negative bias in heterogeneity, especially for large effect sizes and moderate heterogeneity. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations. Columns indicate nominal true between-studies standard deviation and rows indicate reliability standard deviation. Due to truncation in Pearson's $r$ these values may differ from actual true heterogeneity standard deviation (black solid lines). Each dashed line correspond to an average reliability in primary studies.

Lower number of meta-analyzed studies leads to a larger inflation of heterogeneity at low levels of true heterogeneity (Figure C2). In Figure C2, columns are again nominal true heterogeneity, whereas rows in Figure C2 are the number of studies in each meta-analysis. Generally, we would expect a lower number of studies to lead to larger sampling variance of heterogeneity estimates, hence a larger proportion of estimates truncated at zero, and consequently more inflated average heterogeneity estimates when number of studies is low (e.g., first row Figure C2) and heterogeneity is small or absent (first column Figure C2). This is also what we see in Figure C2. In addition, the number of studies mostly affects heterogeneity estimates when average effect size is small.

```{r Figure C2, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_C_fig2.png")
```

_Figure C2._ Expected inflation of effect sizes decreases with larger number of meta-analyzed when heterogeneity is absent or small and average effect size is small. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations. Columns indicate nominal true between-studies standard deviation and rows indicate number of meta-analyzed studies. Due to truncation in Pearson's $r$ these values may differ from actual true heterogeneity standard deviation (black solid lines). Each dashed line correspond to an average reliability in primary studies.

### 2023-06-14 code/05_plots.r

@@ -160,11 +160,11 @@ theme_bw()
 #ggsave("figures/manuscript/r_tau_0.02-0.08.png", width = 8.62, height = 9.93)
 
 #****************************************
-# Supplement A plots
+# supplement C plots
 #****************************************
 
 
-# Supplement A Figure 1 ((variance in reliabilities))
+# supplement C Figure 1 (variance in reliabilities)
 #****************************************
 
 dat_a <- dat[dat$k == 20 &
@@ -205,10 +205,10 @@ xlab(expression("Average effect size "~mu)) +
 facet_grid(reliability_sd~nominal_tau, scales = "free") +
 theme_bw()
 
-#ggsave("figures/supplement/supplement_A_fig1.png", width = 8.62, height = 9.93)
+#ggsave("figures/supplement/supplement_C_fig1.png", width = 8.62, height = 9.93)
 
 
-# Supplement A figure 2 (variable k)
+# supplement C figure 2 (variable k)
 #****************************************
 dat_b <- dat[dat$sample_size == 150 &
              dat$reliability_sd == 0.15 &
@@ -219,7 +219,7 @@ dat_b <- dat[dat$sample_size == 150 &
 dat_b$reliability_mean <- factor(dat_b$reliability_mean)
 dat_b$k <- factor(dat_b$k)
 
-# Truncated values (same as in supplement A)
+# Truncated values (same as in supplement C)
 es <- seq(0, 0.6, 0.1)
 tau <- c(0, 0.1, 0.15, 0.2)
 
@@ -246,10 +246,10 @@ xlab(expression("Average effect size "~mu)) +
 facet_grid(k~nominal_tau, scales = "free") +
 theme_bw()
 
-#ggsave("figures/supplement/supplement_A_fig2.png", width = 8.62, height = 9.93)
+#ggsave("figures/supplement/supplement_C_fig2.png", width = 8.62, height = 9.93)
 
 
-# Supplement A Figure 3 (variable sample size)
+# supplement C Figure 3 (variable sample size)
 #****************************************
 dat_c <- dat[dat$k == 20 &
              dat$reliability_sd == 0.15 &
@@ -261,7 +261,7 @@ dat_c <- dat[dat$k == 20 &
 dat_c$reliability_mean <- factor(dat_c$reliability_mean)
 dat_c$N <- factor(dat_c$N)
 
-# Truncated values (same as in supplement A)
[…]

### 2023-06-14 supplement/supplement_A.qmd
correction to supplement names...

--- 2023-06-11 supplement/supplement_A.qmd
@@
 
-# Supplement A: varying reliability SD and the effect of N and k
+# Supplement A: perfect reliability and Pearson's $r$
+
+Figure A1 shows the bias in heterogeneity estimates in the absence of measurement unreliability in primary studies (i.e., reliability = 1).
 
@@
 
+_Figure A1._ Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1).
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
+Columns indicate the nominal true heterogeneity standard deviation $\tau$.
+Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+
+As can be seen, all three models show a positive bias when heterogeneity is zero, as heterogeneity estimates are truncated at zero.
+However, we also see a positive bias when heterogeneity is above zero for the Pearson's correlation using the HV method (second row).
+This is due to a dependancy between effect size and sampling variance for Pearson's $r$ as normally computed.
+That is, the sampling variance is estimated as $\hat{\sigma}^2_i = \frac{(1-r_i^2)^2}{n_i - 1}$.
+We do not see this effect for the HS method or HV method using Fisher's $z$.
+In the HS method $r_i$ has been replaced by $\bar{r}$ and for Fisher's $z$ the sampling variance is approximated as $\hat{\sigma}^2_i = 1 / (n_i -3)$, in both cases removing the dependency between effect size and sampling variance.
+To demonstrate that this is indeed the case, Figure 2 demonstrates how the bias in Pearson's $r$ disappears if we reshuffle the sampling variances among primary studies so that the dependency between effect size and sampling variance disappears.
+
 ```{r Figure A2, echo=FALSE, out.width = '100%'}
@@
 
-```{r Figure A3, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/supplement/supplement_A_fig3.png")
-```
+_Figure A2._ Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$.
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 1e3 repetitions.
+Columns indicate the nominal true heterogeneity standard deviation $\tau$.
+Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).

### 2023-06-14 supplement/supplement_B.qmd
correction to supplement names...

--- 2023-06-11 supplement/supplement_B.qmd
@@
 ---
+
 format: pdf
+
 ---
+# Supplement B: converting heterogeneity between Pearson's r and Fisher's z
 
@@
 
-# Supplement B: perfect reliability and Pearson's $r$
+The table below shows the results of our procedure to convert between studies variance expressed in Pearson's $r$ to between-studies variance expressed in Fisher's $z$ through the use of $I^2$ as described in the main text.
 
-Figure 1 shows the bias in heterogeneity estimates in the absence of measurement unreliability in primary studies (i.e., reliability = 1).
+_Table B1_.
+```{r}
 
-```{r Figure 1, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/supplement/supplement_B_fig1.png")
+source("../code/01_compute_tau_fishers_z.r") #all computations were made here
+
+#Gives us the object 'cond'
+cond_table <- cond[,c(1, 3, 2, 4, 6)]
+
+knitr::kable(cond_table, booktabs = TRUE, col.names = c("$\\rho$", "N", "$\\tau_r$", "$I^2$", "$\\tau_z$"), escape = FALSE)
 ```
 
-_Figure 1._ Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1).
-The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
-Columns indicate the nominal true heterogeneity standard deviation $\tau$.
-Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
-
-As can be seen, all three models show a positive bias when heterogeneity is zero, as heterogeneity estimates are truncated at zero.
-However, we also see a positive bias when heterogeneity is above zero for the Pearson's correlation using the HV method (second row).
-This is due to a dependancy between effect size and sampling variance for Pearson's $r$ as normally computed.
-That is, the sampling variance is estimated as $\hat{\sigma}^2_i = \frac{(1-r_i^2)^2}{n_i - 1}$.
-We do not see this effect for the HS method or HV method using Fisher's $z$.
-In the HS method $r_i$ has been replaced by $\bar{r}$ and for Fisher's $z$ the sampling variance is approximated as $\hat{\sigma}^2_i = 1 / (n_i -3)$, in both cases removing the dependency between effect size and sampling variance.
-To demonstrate that this is indeed the case, Figure 2 demonstrates how the bias in Pearson's $r$ disappears if we reshuffle the sampling variances among primary studies so that the dependency between effect size and sampling variance disappears.
-
-```{r Figure 2, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/supplement/supplement_B_fig2.png")
-```
-
-_Figure 2._ Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$.
-The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 1e3 repetitions.
-Columns indicate the nominal true heterogeneity standard deviation $\tau$.
-Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+_Note._ $\rho$ = average effect size, N = (fixed) within-study sample size, $\tau_r$ = between-studies standard deviation in Pearson's $r$, $\tau_z$ between-studies standard deviation in Fisher's $z$.
+The $I^2$ statistic is here expressed as a proportion.

### 2023-06-14 code/05_plots.r

@@ -160,7 +160,7 @@ theme_bw()
 #ggsave("figures/manuscript/r_tau_0.02-0.08.png", width = 8.62, height = 9.93)
 
 #****************************************
-# supplement C plots
+# supplement C plots (supplemental main results)
 #****************************************
 
 
@@ -288,10 +288,10 @@ theme_bw()
 #ggsave("figures/supplement/supplement_C_fig3.png", width = 8.62, height = 9.93)
 
 #****************************************
-# Supplement B plots
+# Supplement A plots
 #****************************************
 
-# Figure B1. (z-r-hs, perfect reliability)
+# Figure A1. (z-r-hs, perfect reliability)
 #****************************************
 
 
@@ -344,10 +344,10 @@ xlab(expression("Average effect size "~mu)) +
 facet_grid(effect_type~nominal_tau, scales = "free", switch = "y") +
 theme_bw()
 
-#ggsave("figures/supplement/supplement_B_fig1.png", width = 8.62, height = 9.93)
+#ggsave("figures/supplement/supplement_A_fig1.png", width = 8.62, height = 9.93)
 
 
-# Figure B2
+# Figure A2
 #****************************************
 library(data.table)
 library(parallel)
@@ -467,4 +467,4 @@ facet_wrap(~nominal_tau) +
 ylab(expression("Between-studies standard deviation "~tau)) +
 xlab(expression("Average effect size "~mu))
 
-#ggsave("figures/supplement/supplement_B_fig2.png", width = 8.62, height = 9.93)
+#ggsave("figures/supplement/supplement_A_fig2.png", width = 8.62, height = 9.93)

### 2023-06-14 code/05_plots.r
moved fig C3 to main manuscript

@@ -159,10 +159,46 @@ theme_bw()
 
 #ggsave("figures/manuscript/r_tau_0.02-0.08.png", width = 8.62, height = 9.93)
 
+# Figure 3 (variable sample size)
+#****************************************
+dat_N <- dat[dat$k == 20 &
+             dat$reliability_sd == 0.15 &
+             dat$nominal_tau %in% c(0, 0.1, 0.15, 0.2) &
+             dat$method == "HV" &
+             dat$effect_type == "r",]
+
+
+dat_N$reliability_mean <- factor(dat_N$reliability_mean)
+dat_N$N <- factor(dat_N$N)
+
+# Truncated values (same as in Figure 2)
+es <- seq(0, 0.6, 0.1)
+tau <- c(0, 0.1, 0.15, 0.2)
+
+tau_trunc <- lapply(tau, compute_var_truncated, mu = es)
+tau_trunc <- unlist(tau_trunc)
+trunc_N <- data.frame(mu = es,
+                    nominal_tau = rep(tau, each = length(es)),
+                    true_tau = tau_trunc)
+
+
+
+ggplot(dat_N, aes(x = mu, y = tau_hat)) +
+geom_line(aes(linetype = reliability_mean), show.legend = TRUE) +
+scale_linetype_manual(values = 2:5) +
+geom_line(aes(x = mu, y = true_tau), data = trunc_N, linetype = 1) +
+guides(linetype = guide_legend(reverse = TRUE, title = "Mean Reliability")) +
+expand_limits(y = 0) +
+ylab(expression("Between-studies standard deviation "~tau)) +
+xlab(expression("Average effect size "~mu)) +
+facet_grid(N~nominal_tau, scales = "free") +
+theme_bw()
+
+#ggsave("figures/manuscript/sample_size.png", width = 8.62, height = 9.93)
+
 #****************************************
 # supplement C plots (supplemental main results)
 #****************************************
-
 
 # supplement C Figure 1 (variance in reliabilities)
 #****************************************
@@ -248,44 +284,6 @@ theme_bw()
 
 #ggsave("figures/supplement/supplement_C_fig2.png", width = 8.62, height = 9.93)
 
-
-# supplement C Figure 3 (variable sample size)
-#****************************************
-dat_c <- dat[dat$k == 20 &
-             dat$reliability_sd == 0.15 &
[…]

### 2023-06-14 code/03_supercomputer_simulator.r
updated code/03 to correspond with supplements

@@ -100,31 +100,32 @@ cond2 <- cond[cond$sample_size == 150 &
               cond$reliability_sd ==0.15 &
               cond$method == "HV" &
               cond$effect_type == "r",]
+
 cond2$true_tau2 <- rep(true_tau2, each = 4)
 
-# Supplement A1 (variance in reliabilities)
+#3) Figure 3 main manuscript (variable sample size)
 cond3 <- cond[cond$k == 20 &
-              cond$sample_size == 150 &
-              cond$reliability_mean < 1 &
-              cond$method == "HV" &
-              cond$effect_type == "r",]
-
-# Supplement A2 (variable K)
-cond4 <- cond[cond$sample_size == 150 &
-              cond$reliability_sd == 0.15 &
-              cond$method == "HV" &
-              cond$effect_type == "r",]
-
-# Supplement A3 (variable sample size)
-cond5 <- cond[cond$k == 20 &
               cond$reliability_sd == 0.15 &
               cond$method == "HV" &
               cond$effect_type == "r",]
 
 # Supplement B (Perfect reliability)
-cond6 <- cond[cond$k == 20 &
+cond4 <- cond[cond$k == 20 &
               cond$sample_size == 150 &
               cond$reliability_mean == 1,]
+
+# Supplement C1 (variance in reliabilities)
+cond5 <- cond[cond$k == 20 &
+              cond$sample_size == 150 &
+              cond$reliability_mean < 1 &
+              cond$method == "HV" &
+              cond$effect_type == "r",]
+
+# Supplement C2 (variable K)
+cond6 <- cond[cond$sample_size == 150 &
+              cond$reliability_sd == 0.15 &
+              cond$method == "HV" &
+              cond$effect_type == "r",]
 
 
 #****************************************

### 2023-06-14 code/04_in-text-values.r
created in-text values for Figure 4

@@ -89,6 +89,79 @@ underestimate_mu02 <- list(underestimate_min = underestimate_min,
                            underestimate_max = underestimate_max)
 
 
+## [1.4] sample size (N = 50 vs N = 200, mu = 0.2, tau >0)
+#****************************************
+
+dat <- read.csv("data/means_combined.csv")
+dat$tau_hat <- sqrt(dat$tau2_hat)
+dat$nominal_tau <- sqrt(dat$true_tau2)
+dat$k <- factor(dat$k, levels = c("5", "20", "40", "200"))
+dat$N <- factor(dat$sample_size, levels = c("50", "100", "150", "200"))
+
+dat_3 <- dat[dat$k == 20 &
+             dat$reliability_sd == 0.15 &
+             dat$nominal_tau %in% c(0.1, 0.15, 0.2) &
+             dat$sample_size %in% c(50, 200) &
+             dat$method == "HV" &
+             dat$effect_type == "r",]
+
+### [1.4.1] differences when  reliability is 0.8
+#****************************************
+
+dat_3b <- dat_3[dat_3$reliability_mean == 0.8 &
+              dat_3$mu == 0.2,]
+
+res2 <- split(dat_3b, dat_3b$nominal_tau)
+res2 <- lapply(res2, function(x) {
+    x[1, "tau_hat"] - x[2, "tau_hat"]
+})
+res2 <- abs(unlist(res2))
+
+max_bias_n50_0.8 <- round(max(res2), 2)
+
+### [1.4.2] difference reliability is 0.6
+#****************************************
+
+dat_3a <- dat_3[dat_3$reliability_mean == 0.6 &
+                dat_3$mu == 0.2,]
+
+res1 <- split(dat_3a, dat_3a$nominal_tau)
+res1 <- lapply(res1, function(x) {
+    x[1, "tau_hat"] - x[2, "tau_hat"]
+})
+res1 <- abs(unlist(res1))
+
+max_bias_n50_0.6 <- round(max(res1), 2)
+
+### [1.4.2] bias when mu = 0.2, tau = 0.15, N = 50, R =  0.6
+#****************************************
+dat_142 <- dat_3a[dat_3a$nominal_tau == 0.15 &
+                  dat_3a$N == 50,]
+
+net_bias_n50 <-  dat_142$tau_hat - dat_142$nominal_tau
+net_bias_n50  <- round(net_bias_n50, 3)
+
[…]

### 2023-06-16 manuscript/manuscript.qmd
writing now in qmd

--- 2023-05-25 manuscript/manuscript.qmd
@@
 ---
-format: docx
+title: "Unreliable heterogeneity: how measurement error obscures heterogeneity in meta-analyses in psychology"
+preprint: true
+author:
+  - name: "Anton Olsson-Collentine"
+    affiliation: 1
+    corresponding: true
+    email: anton@olssoncollentine.com
+
+  - name          : "Marjan Bakker"
+    affiliation   : "1"
+
+  - name          : "Jelte Wicherts"
+    affiliation   : "1"
+affiliation:
+  - code: 1
+    address: "Department of Methodology and Statistics, Tilburg School of Social and Behavioral Sciences, Tilburg University, the Netherlands."
+
+abstract: >
+  One of the main purposes of meta-analysis is to estimate heterogeneity of effect sizes.
+The quality of meta-analytic estimates depends in large part on the quality of evidence in primary studies, yet there exist broad concerns that proper measurement tends to be neglected in much of psychological primary research.
+Such inattention to measurement can also bias heterogeneity estimates.
+We performed a Monte-Carlo simulation study to examine the expected bias in heterogeneity estimates due to unreliability in primary studies.
+Although unreliability in primary studies can both inflate and suppress heterogeneity, our results indicate that under most circumstances unreliability in primary studies leads to a severe underestimate of heterogeneity in meta-analyses.
+For a typical effect size heterogeneity can be expected to be underestimated by 15-60%, or more if average reliability in primary studies is below 0.6.
+We observed a positive bias in heterogeneity estimates due to unreliability only under specific and arguably uncommon circumstances of (1) actual zero heterogeneity, particularly when mean effect sizes are large, or (2) combinations of very small true heterogeneity, large variance in primary study reliabilities, large mean effect sizes, and a limited number of primary studies.
+The severity of the underestimation of heterogeneity could lead to meta-analyses in psychology missing relevant moderators and overgeneralizing outcomes.
+Unreliability appears to bring truly different effects closer together, thereby obscuring true differences between studies that could be relevant for theory, practice, and future research efforts.
+Research on concrete guidance to applied meta-analysts is urgently needed, as sophisticated methods for correcting measurement unreliability such as meta-analytic structural equation modeling (MASEM) are only applicable in exceptional cases and corrections based on classical test theory come with caveats and strong assumptions.
+
+header-includes:
+   - \usepackage{setspace}
+   - \usepackage{lineno}
+   - \linenumbers
+
+csl: apa.csl
+
+output:
+  bookdown::pdf_book:
+    base_format: rticles::peerj_article # for using bookdown features like \@ref()
+    keep_tex: true
+  rticles::peerj_article: default
+
 ---
@@
 
-Estimating heterogeneity of effect sizes is usually considered the main purpose of meta-analysis in addition to estimating an average effect size.
-Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size's sensitivity to variability in study design, or more specifically to sensitivity to differences in four study factors: 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables (e.g., Campbell & Stanley, 2015).
-Heterogeneity is a primary outcome in meta-analysis because of several reasons.
+Estimating heterogeneity of effect sizes is usually considered the main purpose of meta-analysis, alongside estimating an average effect size.
+Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size’s sensitivity to variability in study design features related to the type of (1) sample, (2) setting, (3) treatment variable, and (4) measurement (e.g., Shadish, Cook, & Campbell, 2002).
+Heterogeneity is an important outcome in meta-analysis for several reasons.
 First, its presence signals the existence of moderators and so can be seen as an opportunity for theoretical development (Simons et al., 2017).
-Second, heterogeneity affects the implementation of research by indicating that an intervention may not be equally effective under all conditions or for all people.
-Third, the presence of heterogeneity changes the interpretation of the average effect sized derived in meta-analysis from the true population effect size (under homogeneity) to the average of true subpopulation effect sizes (under heterogeneity).
+Second, heterogeneity affects the implementation of research by indicating that an intervention may not be equally effective under all conditions or for everyone.
+Third, the presence of heterogeneity changes the interpretation of the average effect size derived in a meta-analysis from the true population effect size (under homogeneity) to the average of true subpopulation effect sizes (under heterogeneity).
 As such, estimating heterogeneity with minimal bias is of central importance to meta-analysis (for a more extensive discussion on the importance of heterogeneity, see Olsson Collentine et al., 2020; Simons, 2017).
 
-Meta-analytic estimates are highly dependent on the quality of data in primary studies.
+Meta-analytic estimates depend on the quality of data in primary studies.
 However, there is concern that researchers in psychology tend to neglect proper measurement (Flake & Fried, 2020) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021).
@@
 Such inattention to measurement can also bias heterogeneity estimates.
-We will focus here on reliability of measurement (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., Lord & Novick, 1968), which is the only measurement validity evidence researchers report consistently (Flake et al., 2017), and which can be corrected for in meta-analysis (Hunter & Schmidt; 2015).
+We will focus here on the reliability of measurement (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., Lord & Novick, 1968), which is the only  evidence on measurement quality that researchers report consistently (Flake et al., 2017), and which can be corrected for in meta-analysis (Hedges & Olkin, 1985; Hunter & Schmidt; 2015).
 
-Measurement error (imperfect reliability) is present in any empirical effect size estimate and systematically attenuates observed effect sizes compared to true underlying effect sizes.
-This affects heterogeneity estimates, which from a statistical perspective estimate the variance between true effect sizes.
-However, reliability is not necessarily accounted for in heterogeneity estimates, because ‘true’ underlying effect sizes can refer to two different entities.
-Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement artifacts (such as imperfect reliability, restriction of range, and dichotomization; Hunter & Schmidt, 2015).
-Whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for measurement artifacts is implicit only in the models of Hunter & Schmidt (2015).
+Measurement error (imperfect reliability) is present in any empirical effect size estimate and systematically attenuates observed effect sizes compared to true underlying effect sizes (Spearman, 1904).
+This affects heterogeneity estimates, which from a statistical perspective, estimate the variance between true effect sizes.
+However, reliability is not necessarily accounted for in heterogeneity estimates, because ‘true’ underlying effect sizes generally refer to two different entities: Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement error.
+Whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for measurement errors is explicitly modelled (alongside other measurement artifacts) only in the psychometric meta-analysis approach by Hunter & Schmidt (2015).
 This has consequences for heterogeneity estimates reported in many areas of psychology.
-Inattention to measurement reliability and how it affects heterogeneity estimates can lead researchers to misinterpret their meta-analytic average effect size estimate, ignore the presence of moderators or search for (and discover) non-existent moderators, and to implement research interventions inappropriately.
+Inattention to measurement reliability and how it affects heterogeneity estimates can lead meta-analysts to misinterpret their average effect size estimate, ignore the presence of theoretically or practically relevant moderators or search for (and discover) non-existent moderators, over-generalize outcomes, and implement research interventions inappropriately.
 
 Meta-analyses in psychology rarely correct for unreliability in primary study measurements.
-Wiernick and Dahlke (2020) report that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability.
-Similarly, Schmidt (2010) report that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts.
-The exception is the subfield of Industrial-Organizational studies where corrections tend to be more common (e.g., Cortina, 2003; Anguinis, 2011).
+Wiernick and Dahlke (2020) reported that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability.
+Similarly, Schmidt (2010) reported that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts.
+The exception is the subfield of industrial-organizational studies, where corrections tend to be more common (e.g., Cortina, 2003; Anguinis, 2011).
 
-One reason that few meta-analysts correct for unreliability could be that correcting for unreliability is generally controversial and has been so since it was proposed (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017).
+One reason that few meta-analysts correct for unreliability could be that correcting for unreliability has long been controversial (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017).
 On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017).
-Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1).
-Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (Sijtsma, 2008).
-Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated in psychology (e.g., Kvarven et al., 2019; Schäfer & Scharz, 2019; Scheel et al., 2020) due to selective reporting based on the significance of outcomes (Simmons et al., 2011) and publication bias (REF?), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
+Systematically correcting for unreliability may also lead to an overestimation of effect sizes, because reliability estimates are lower bound estimates of the true reliability (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1).
+Even worse, the most popular estimate of reliability in psychology (Cronbach’s Alpha; Flake et al., 2017), as well as improvements thereof, tend to underestimate this lower bound of reliability (Sijtsma, 2008).
+Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated in psychology (e.g., Kvarven et al., 2019; Schäfer & Scharz, 2019; Scheel et al., 2020) due to selective reporting based on the significance of outcomes (Simmons et al., 2011) and publication bias (Thornton & Lee, 2000), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
 On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures.
-As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest.
-From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015), although  treating disattenuated effect sizes as directly comparable to latent scores requires certain assumptions (Borsboom & Mellenbergh, 2002).
-In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses need to be aware of how this decision affects heterogeneity estimates.
+As such, neglecting measurement errors means that computed estimates do not correspond to the entity of interest.
+From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015), although treating disattenuated effect sizes as directly comparable to latent scores requires strong assumptions (Borsboom & Mellenbergh, 2002).
+In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses need to be aware of how unreliability of measurement affects heterogeneity estimates.
+Hunter & Schmidt (2015) assert that failure to control for variation in unreliability across studies results in a large overestimate of heterogeneity (p.
+159), but little systematic research has been done to study the effects of unreliability on estimating heterogeneity.
 
@@
 343).
-Imperfect reliability in primary studies imbues heterogeneity estimates with both a positive and a negative bias.
+Reliability is usually assumed to be known (but see Charles, 2005; Raju et al., 1991), Imperfect reliability in primary studies can both increase and suppress true heterogeneity.
 
-A positive bias can be expected in heterogeneity estimates of uncorrected effect sizes due to the variability in reliability between primary studies.
+Heterogeneity is inflated in heterogeneity estimates of uncorrected effect sizes due to the variability in reliability between primary studies.
 This is because the attenuation in effect size will then differ from study to study, which results in differences across studies in observed effect sizes beyond their true variability.
-This is most easily seen by assuming a common true effect size $\mu$ across studies (i.e., a fixed effect and no true heterogeneity) and infinite sample size within studies.
+This is most easily illustrated by assuming a common true effect size $\mu$ across studies (i.e., a fixed effect and no true heterogeneity) and infinite sample size within studies.
+
  Any observed effect size $\hat{\mu}$ can then be seen as the true effect size $\mu$ multiplied by a study’s reliability $R_i$ (assuming $R_i = R_{xx’} = R_{yy’}$).
 Hence, the distribution of observed effect sizes is the true effect size $\mu$ multiplied by the distribution of reliabilities across studies.
-That is, $\hat{mu} \sim \mu \times N(\bar{R}, \sigma_R) = N(\mu \bar{R}, \mu \sigma_R$, where $\bar{R}$ is the average reliability across studies and $\sigma_R$ its standard deviation.
+That is, $\hat{\mu} \sim \mu \times N(\bar{R}, \sigma_R) = N(\mu \bar{R}, \mu \sigma_R)$, where $\bar{R}$ is the average reliability across studies and $\sigma_R$ its standard deviation.
 Because reliability lies between 0 – 1, under these conditions imperfect reliability implies that the observed average effect size will be less than the true effect size by a factor $\bar{R}$ and estimated heterogeneity will be larger than true heterogeneity (i.e., zero) by the product of $\mu \times \sigma_R$ (if heterogeneity is expressed as standard deviation rather than variance).
-That is, variability in reliability across studies leads to a positive bias in heterogeneity that depends on the size of the true effect size.
+That is, variability in reliability across studies is a force that inflates heterogeneity depending on the size of the true effect size.
 Table 1 illustrates this effect for three studies with differing reliability and zero true heterogeneity.
@@
 
-*Variance in reliability leads to positive bias in heterogeneity estimates which increases with average effect size*
+*Variance in reliability inflates heterogeneity in combination with increased average effect size and leads to overestimated heterogeneity in a fixed effects model*
 
@@
 
-A negative bias can be expected in heterogeneity estimates of uncorrected effect sizes because in absolute terms larger effect sizes are more attenuated by a given reliability.
+Heterogeneity is suppressed in heterogeneity estimates of uncorrected effect sizes because in absolute terms larger effect sizes are more attenuated by a given reliability.
 That is, although the attenuation formula $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ implies that any effect size $\rho_{xy}$ will be attenuated by the same proportion given the same reliabilities $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$, this proportion corresponds to a larger absolute value for larger effect sizes.
 Consequently, in the presence of heterogeneity attenuation will move larger true effect sizes further towards zero than smaller ones, decreasing heterogeneity in observed effect sizes.
-This negative bias can be observed directly by assuming an average effect size $\theta$ with some heterogeneity $\tau^2$, infinite sample size within studies (i.e., no within-study sampling variance), and equal reliability across studies ($R = R_i = R_{xx'} = R_{yy'}$).
+This suppression can be observed directly by assuming an average effect size $\theta$ with some heterogeneity $\tau^2$, infinite sample size within studies (i.e., no within-study sampling variance), and equal reliability across studies ($R = R_i = R_{xx'} = R_{yy'}$).
 Any observed effect size $\hat{\theta_i}$ can then be seen as a study’s true effect size $\theta_i$ multiplied by the study reliability $R$.
@@
 This depiction is not entirely correct for Pearson's $r$ as it is bounded at {-1, 1}, but holds approximately if the mean effect size $\theta$ and heterogeneity $\tau^2$ are not too large.
-Hence, imperfect reliability in studies can be expected to create a negative bias in heterogeneity estimates whenever true heterogeneity is present.
+Hence, imperfect reliability in studies can be expected to suppress heterogeneity estimates whenever true heterogeneity is present.
 Consider an average reliability of 0.8 (the median reported reliability in psychology; Flake et al., 2017).
@@
 
-*Imperfect reliability leads to negative bias in heterogeneity estimates*
+*Imperfect reliability suppresses true heterogeneity*
 
@@
 
-It is difficult to predict the total effect of the negative and positive bias in heterogeneity due to unreliability, and existing literature tends to focus on the positive bias (e.g., Borenstein, 2009, p.
+It is difficult to predict if the total effect is a negative or positive bias in heterogeneity due to unreliability, and existing literature tends to focus on the inflation of heterogeneity (e.g., Borenstein, 2009, p.
 342; Card, 2015, p.
 126; Hunter & Schmidt, 2015; Wiernick & Dahlke, 2020).
-However, given the value of accurate  heterogeneity estimates for many research outcomes in psychology and that the negative bias could often be 20% or more, we consider it important for meta-analysts to have insight into the total bias in heterogeneity estimates that should be expected due to unreliability.
-This depends on average effect size, true heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies.
-Hence, we perform a Monte-Carlo simulation study to explore the expected bias in heterogeneity estimates due to unreliability in primary studies.
+However, given the value of accurate heterogeneity estimates for many research outcomes in psychology and that the suppression of heterogeneity could often be 20% or more, we consider it important for meta-analysts to have insight into the bias in heterogeneity estimates that should be expected due to unreliability in a wider range of settings.
+The effect of unreliability depends on average effect size, true heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies.
+Hence, here we perform a Monte-Carlo simulation study to explore the expected bias in heterogeneity estimates due to unreliability in primary studies.
+In our simulation, we consider reliabilities to be known and conforming to the strict assumptions of CTT to gain a better understanding of how measurement error biases heterogeneity estimates, while acknowledging that in meta-analytic practice, heterogeneity estimation is further challenged by selective reporting, sampling error, and estimation of reliabilities, which is particularly challenging when CTT’s strict assumptions are violated.
 
@@
 
-We focus on correlations as our effect size type in these simulations because correcting them for unreliability is straightforward and because they are the effect size type most commonly corrected in practice.
+We focus on correlations as our effect size type in these simulations because correcting them for unreliability is straightforward and because they are the effect size type most corrected in practice.
 However, unreliability attenuates all types of effect sizes and all can be corrected, either directly (standardized mean differences; Wiernick & Dahlke, 2020) or by first transforming them into correlations.
 Although effect sizes can generally be transformed into each other (e.g., Borenstein, 2009, p.
-46) this should be done with care as it can sometimes affect conclusions (e.g., we found a violation of monotonicity in heterogeneity estimates when converting effect sizes into correlations in a previous project; Olsson-Collentine, 2020, supplement A).
+46), this should be done with care as it can sometimes affect conclusions (e.g., we found a violation of monotonicity in heterogeneity estimates when converting effect sizes into correlations in a previous project; Olsson-Collentine, 2020, supplement A).
 
-Perhaps the most common way to meta-analyze correlations is to first transform them into Fisher $z$ correlations (Borenstein et al., 2009, p.
-41 - 43).
+Perhaps the most common way to meta-analyze correlations is to first transform them into Fisher z correlations (Borenstein et al., 2009, p.
+41 - 43) to normalize their distribution.
 The Fisher $z$ transformation is the inverse hyperbolic tangent of the product moment correlation and can be computed as $z = 0.5 \times ln\left(\frac{1+r}{1-r}\right)$ and its variance approximated as $V_z = 1 / (n -3)$.
-There is controversy in the use of the Fisher $z$ transformation for meta-analysis (e.g., Hunter & Schmidt, 2015, p.
-XX, ch.
-5 "Use of Fisher’s z in Meta-Analysis of Correlations"), largely related to whether Pearson's $r$ or Fisher's $z$ leads to less bias in average correlations (Schulze, 2007).
-However, estimating heterogeneity does not seem to have been the focus of most of the literature's discussion of the appropriateness of the Fisher $z$ transformation (Field, 2005; Hafdahl & Williams, 2009, Hunter & Schmidt, 2015).
-Hall and Brannick (2002) do report heterogeneity estimates but focus on the coverage of prediction intervals ('credibility intervals' in Hunter & Schmidt terminology) which combine the heterogeneity estimate and point estimate, and in Brannick et al., (2019) they examine results only after corrections for attenuation.
-Given the inconclusive evidence and that the Fisher $z$ transformation is often applied in practice, we use it as one of the effect size types in our simulations.
+There is controversy in the use of the Fisher z transformation for meta-analysis (e.g., Hunter & Schmidt, 2015, p.
+220 - 223), largely related to whether Pearson’s r or Fisher’s z leads to less bias in average correlations (Schulze, 2007).
+However, estimating heterogeneity does not seem to have been the focus of most of the literature’s discussion of the appropriateness of the Fisher z transformation (Field, 2005; Hafdahl & Williams, 2009, Hunter & Schmidt, 2015).
+Hall and Brannick (2002) do report heterogeneity estimates but focus on the coverage of prediction intervals (‘credibility intervals’ in Hunter & Schmidt terminology) which combine the heterogeneity estimate and point estimate, and Brannick et al.
+(2019) examined results only after corrections for attenuation.
+Given both this inconclusive evidence and the common use of the Fisher z transformation in practice, we use it as one of the effect size types in our simulations.
 
-The other effect size type we use is Pearson's $r$.
-Pearson's $r$ has the advantage that resulting heterogeneity estimates on this scale are more interpretable than on the Fisher $z$ scale.
-Although Fisher $z$ correlations can be back-transformed into Pearson's $r$ for interpretation, its heterogeneity cannot.
+We also use Pearson’s r.
+Pearson’s r has the advantage that it creates heterogeneity estimates on the same scale that are more interpretable than heterogeneity estimates on the Fisher z scale.
+Although Fisher z correlations can be back-transformed into Pearson’s r for interpretation, its heterogeneity cannot.
 A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (Hedges & Vevea, 1998).
 The width of the interval can then be used as an estimate of the heterogeneity.
-We prefer to compute comparable heterogeneity values for Pearson's $r$ and Fisher's $z$ (see the methods section 'Parameter values') and report the bias in heterogeneity estimates, as we believe this is more interpretable for applied meta-analysts.
+We prefer to compute comparable heterogeneity values for Pearson’s r and Fisher’s z (see the methods section ‘Parameter values’) and report the bias in heterogeneity estimates, as we believe this is more interpretable for applied meta-analysts.
 
-The disadvantage to using Pearson's $r$ when estimating heterogeneity is that 1) the effect size is bounded to {-1, 1} and 2) the sampling variance $\sigma^2_r = \frac{(1-\rho^2)^2}{n - 1}$ depends on the true effect size $\rho$.
-The bounded nature of the Pearson correlation means that large levels of heterogeneity will lead to substantial truncation of effect sizes if the average effect size is also large.
-That the sampling variance covaries with effect size leads to a small positive bias in heterogeneity estimates (supplement XX).
+The disadvantage to using Pearson’s r when estimating heterogeneity are that (1) the effect size is bounded to {-1, 1} and (2) its sampling variance $\sigma^2_r = \frac{(1-\rho^2)^2}{n - 1}$ depends on the true effect size $\rho$.
+The bounded nature of the Pearson correlation means that large levels of heterogeneity create substantial truncation of the distribution of effect sizes if the average effect size is also large.
+That the sampling variance covaries with effect size leads to a small positive bias in heterogeneity estimates (supplement A).
 This can be solved by replacing $\rho$ with the average correlation across studies $\bar{r}$ (Hunter & Schmidt, 2015), or using the Fisher $z$ transformation.
-Fisher $z$ correlations have the additional advantage that they are not bounded to {-1, 1}.
-A disadvantage to using Fisher $z$ correlations, in addition to interpretability of heterogeneity estimates, is that they require backtransforming into product-moment correlations before adding measurement error, and then transforming to Fisher $z$ again for meta-analysis.
-Especially for large Fisher $z$ values (which can happen if the average effect size and heterogeneity are large) this transformation process may introduce inaccuracies in the estimates.
-Given the common use of both effect size types and their mix of advantages and disadvantages we report results for both.
+Fisher z correlations have the additional advantage that they are not bound to {-1, 1}.
+A disadvantage to using Fisher z correlations, in addition to interpretability of heterogeneity estimates, is that they require back-transforming into product-moment correlations before adding measurement error, followed by another transformation to Fisher z for meta-analysis.
+Especially for large Fisher z values (which are common when both average effect size and heterogeneity are large) this transformation process may introduce inaccuracies in the estimates.
+Given the common use of both effect size types and their mix of advantages and disadvantages we report results for both Fisher’s z and Pearson’s r.
 
@@
 
-We apply two meta-analytic models to ensure that our results are not model-dependent: the Hedges and Vevea (1998) random-effects  model and the Hunter and Schmidt (2015) 'bare-bones' random-effects model.
+We apply two meta-analytic models to ensure that our results are not model-dependent: the Hedges and Vevea (1998) random-effects model and the Hunter and Schmidt (2015) ‘bare-bones’ random-effects model.
 The Hedges and Vevea (HV) model uses inverse variance weight computed as $w_i = 1 / (V_i + T^2)$ (e.g., Borenstein et al., 2010) where $V_i$ is the sampling variance of study $i$ and $T^2$ is the estimated between-studies variance.
@@
 The differences between these weights are relatively minor (Borenstein et al., 2010; Hunter and Schmidt, 2015, p.
-XX).
-Although the HV model is typically applied to Fisher $z$ transformed correlations (e.g., Borenstein et al., p.
+423-426).
+Although the HV model is typically applied to Fisher z transformed correlations (e.g., Borenstein et al., p.
 41-43), we apply it to both transformed and untransformed correlations to enable a more direct comparison with the HS model and facilitate interpretation.
-The HS model typically includes correcting for unreliability or other measurement artifacts.
-However, we implement it without corrections ('bare-bones'; Hunter & Schmidt, 2015) because this allows us to focus on the bias in heterogeneity estimates rather than the performance of corrections.
+The HS model typically includes corrections for unreliability or other measurement artifacts.
+However, we implement it without corrections (‘bare-bones’; Hunter & Schmidt, 2015) because this allows us to focus on the bias in heterogeneity estimates rather than the performance of corrections (e.g., see Zhang, 2022 for how well they perform in terms of mean estimates).
 
 Estimating heterogeneity in both the HV and bare-bones HS model essentially consists of comparing the observed variance in effect sizes with what would be expected from sampling error alone and then ascribing any excess observed variance to heterogeneity (Borenstein et al., 2010, Box 1; Hunter & Schmidt, 2015, p.
-XX).
+100).
 However, the process for doing so differs somewhat between methods.
 We apply the Restricted Maximum Likelihood (REML) estimator of heterogeneity when estimating heterogeneity in the Hedges and Vevea (1998) model, because it is the generally recommended heterogeneity estimator (Veroniki et al., 2016; Viechtbauer, 2005).
-Describing the REML estimator in detail is beyond the scope of this paper and we refer interested readers to the description in Veroniki et al.
-(2016).
-The Hunter and Schmidt (2015) model estimates heterogeneity as $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $ \frac{\Sigma[N_i \sigma^2_{ei}]}{\Sigma N_i}$ (p.
-87, 2nd edition) and $\sigma^2_{ei}$ is the usual sampling variance for the Pearson correlation but with $\rho$ replaced with the sample size weighted average correlation $\bar{r} = \frac{\Sigma n_i r_i}{\Sigma n_i}$ instead of $r_i$ as is common.
+Veroniki et al.
+(2016) describe the REML estimator in detail.
+The Hunter and Schmidt (2015) model estimates heterogeneity as $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $\frac{\Sigma [N_i \sigma^2_{ei}]}{\Sigma N_i}$ (Hunter & Schmidt, 2015, p.
+100) and $\sigma^2_{ei}$ is the usual sampling variance for the Pearson correlation but with $\rho$ replaced with the sample size weighted average correlation $\bar{r} = \frac{\Sigma n_i r_i}{\Sigma n_i}$ instead of $r_i$ as is common.
 A generalized version of the HS estimator for effect sizes other than correlations can be found in Viechtbauer (2005; 2015).
@@
 ```{r Figure 1, out.width= "100%"}
-knitr::include_graphics("../figures/flowchart-simulation-design.png")
+knitr::include_graphics("../figures/manuscript/flowchart-simulation-design.png")
 ```
 
-(FIGURE 1: a flowchart summarizing the simulation design)
+_Figure 1.
+Flowchart summarizing the simulation design.
+The flowchart shows the procedures followed to meta-anlyze results using the Hedge’s and Vevea model (see ‘Meta-Analytic Model’ header in methods section).
+The Hunter & Schmidt ‘bare-bones’ meta-analytic model followed the Pearson correlations procedure, but used sample size weights and replaced_ $r_i$ _in the sampling variance by $\bar{r}$ for meta-anlysis._
 
-When generating data for meta-analysis we proceeded as follows.
+We generated data for the meta-analyses as follows.
 First,  we sampled $i = 1, ..., k$ true study effect sizes $\theta_i$ from $Ntrunc(\theta, \tau^2)$.
 That is, a normal distribution with a mean of $\theta$ and a variance of $\tau^2$ that we truncated at $\{-1, 1\}$ to stay within the bounds of the product-moment correlation.
-We implemented the truncation in R using inverse transform sampling (link: functions.r).
-For Fisher' $z$ no truncation was applied.
+We implemented the effect size truncation in R using inverse transform sampling (link: functions.r).
+For Fisher’ z no truncation was applied.
+Throughout this text we will refer to three types of truncation: (1) effect size truncation (truncation of Pearson’s correlations between {-1, 1}), (2), reliability truncation (truncation of reliability between {0, 1}), and (3) heterogeneity truncation (truncation of heterogeneity estimates to be non-negative).
 
@@
 
-We ran our analyses across a range of within-study sample sizes $N$ and number of studies $K$.
-Within meta-analyses we used a fixed sample size across all studies such that $N = N_i$.
+We ran our analyses across a range of within-study sample sizes N and number of studies K.
+Within meta-analyses we used a fixed sample size across all studies such that N=N_i.
 We made this choice rather than treating sample size as a random variable (as done by e.g., Field, 2005; Brannick et al., 2019) to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance.
 We base our range of sample sizes on empirical estimates of typical sample size in correlational research in psychology.
-Fraley et al., report median sample sizes for between-person studies in 9 psychology journals between 2011 - 2018 which ranged between 69 - 496 depending on journal and year.
-Sassenber and Ditrich (2019) find a median sample size of 110 (interquartile range, 71 - 195) across 4 journals and years (2009, 2011, 2016, and 2018) in social psychology, and Bakker et al.
-(personal communication, 2023) find median sample sizes between 24 - 184 for 6 journals in psychology for the years 1995, 2006, 2019.
-Given these empirical findings we consider the following sample size values {50, 100, 150, 200}.
+Fraley et al.
+(2022) report median sample sizes for between-person studies in nine psychology journals between 2011 - 2018 which ranged between 69 - 496 depending on journal and year.
+Sassenber and Ditrich (2019) find a median sample size of 110 (interquartile range, 71 - 195) across four journals and years (2009, 2011, 2016, and 2018) in social psychology, and Bakker et al.
+(preliminary results; https://osf.io/zsjf4) find median sample sizes between 24 - 184 for six journals in psychology for the years 1995, 2006, 2019.
+Given these empirical findings, we consider the following sample size values {50, 100, 150, 200}.
 
-When focusing on the bias of heterogeneity estimate, the number of meta-analyzed studies, $K$, would not matter for an unbiased estimator.
-However, both the heterogeneity estimator we use, restricted maximum likelihood, and most other heterogeneity estimators are truncated at zero (Viechtbauer, 2005).
-Lower values for $K$ then results in a larger proportion of truncated estimates below zero because of increased sampling variance in $\tau^2$.
-As such, values of $K$ can affect bias, especially at lower levels of heterogeneity.
-We consider the following values of $K$: {5, 20, 40, 100}.
+When focusing on the bias of heterogeneity estimate, the number of meta-analyzed studies, K, would not matter for an unbiased estimator.
+However, both the heterogeneity estimator we used, restricted maximum likelihood, and most other heterogeneity estimators are truncated at zero (Viechtbauer, 2005).
+Lower values for K then result in a larger proportion of truncated variance estimates below zero because of increased sampling variance in $\tau^2$.
+As such, values of K can affect bias, especially at lower levels of heterogeneity.
+We consider the following values of K: {5, 20, 40, 200}.
 
 We vary the grand mean $\theta$ from 0 to 0.6 in steps of 0.1 to cover all realistic effect sizes and explore the boundaries of the interaction between effect size and reliability variance.
-For Fisher's $z$ we transformed these value into equivalent effect sizes at the Fisher's $z$ scale.
-For context to  these values consider the empirical estimates of typical correlational effect sizes provided by Nuijten et al.
-(2020) and Schäfer & Scharz (2019).
+For Fisher’s z, we transformed these values to the $z$ scale.
+For context to these values, we considered the empirical estimates of typical correlational effect sizes provided by Nuijten et al.
+(2020) and Schäfer & Schwarz (2019).
 Nuijten et al.
-performed a meta-meta-analysis on the fields of intelligence research and reports a median meta-analytic effect size of 0.24 for predictive validity and correlational studies.
+performed a meta-meta-analysis on the fields of intelligence research and reported a median meta-analytic correlation of 0.24 across 779 predictive validity and correlational studies.
 This estimate is likely positively biased due to publication bias and selective reporting.
-Schäfer & Scharz report a median $r$ of 0.16 amongst preregistered research in psychology in general, with an 'upper median' (i.e., the 83.35% quantile) of 0.41.
+Schäfer & Scharz reported a median $r$ of 0.16 amongst preregistered research in psychology in general, with an ‘upper median’ (i.e., the 83.35% quantile) of 0.41.
 This estimate is likely less affected by positive bias but should not be expected to be unbiased as there appears to be some positive bias also in preregistered research on average (Scheel, 2021).
@@
 Van Erp et al.
-(2017) provide empirical heterogeneity estimates from 747 meta-analyses in 61 article published in the journal Psychological Bulletin between 1990 - 2013.
-The median (uncorrected)Pearson's $r$ $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24).
-The point of this study is that empirical heterogeneity estimates may not be accurately estimated.
+(2017) provided empirical heterogeneity estimates from 747 meta-analyses in 61 articles published in the journal Psychological Bulletin between 1990 - 2013.
+The median (uncorrected) Pearson’s r $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24).
+The point of our study is that empirical heterogeneity estimates may not be accurately estimated.
 Nonetheless, we selected approximately the interquartile range of $\tau$-values reported in van Erp et al.
@@
 The values we selected were {0, 0.1, 0.15, 0.2}.
-We selected the largest heterogeneity level to avoid excessive truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1.
+We selected the largest heterogeneity level to avoid excessive effect size truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1.
 Our values are similar to those used by Brannick et al.
-(2019) in their simulation study on the performance of different heterogeneity estimators after corrections for unreliability (they used $\tau = $ 0, 0.08, 0.13, 0.2).
+(2019) in their simulation study on the performance of different heterogeneity estimators after corrections for unreliability (they used $\tau =$  0, 0.08, 0.13, 0.2).
 
-Fisher's $z$ is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson's $r$ into corresponding $\tau$-values on the Fisher's $z$ scale.
-To compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale.
+Fisher’s z is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson’s r into corresponding $\tau$-values on the Fisher’s z scale.
+To compare results between Fisher’s z and the product-moment correlation we defined heterogeneity on the $I^2$ scale.
 The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity (Higgins, 2002; 2003), and as such is measured from 0 - 100%.
@@
 The resulting $\tau$ values for Fisher $z$ were only minimally different from {0, 0.1, 0.15, 0.2}.
-For example, for $N = 50$ the corresponding Fisher $z$ $\tau$-values were {0, 0.1031263, 0.1566007, 0.2123514} and as $N$ increased heterogeneity values were more similar (see supplement XX).
+For example, for $N = 50$ the corresponding Fisher $z$ $\tau$-values were {0, 0.1031263, 0.1566007, 0.2123514} and as $N$ increased heterogeneity values were more similar (see supplement B).
 
@@
 Flake et al.
-(2017) report that based on 245 estimates of Cronbach's alpha in psychology the average estimate was 0.79 with a standard deviation of 0.13.
-The interquartile range was approximately 0.68 - 0.87 for studies wherein the associated scale lacked a reference and 0.79 - 0.88 for scales that did have a reference.
+(2017) documented 245 estimates of Cronbach’s Alpha in psychology and found an average of 0.79 and a standard deviation of 0.13.
+The interquartile range was approximately 0.68 – 0.87 for studies using (_ad hoc_) scales that lacked a reference to an earlier source and 0.79 – 0.88 for scales that did have a reference.
 Sánchez-Meca et al.
-(2003) report the reliability estimates based on five 'reliability generalization studies'.
-These five studies summarized reliability in 25 - 51 primary studies (184 total), and the mean reliability ranged from 0.767 to 0.891 with standard deviations ranging between 0.034 - 0.133.
-Given some likely positive bias in such empirical values (Hussey et al., 2023), we examined the following mean reliabilities $\{0.6, 0.7, 0.8, 0.9\}$, and standard deviations $\{0, 0.05, 0.1, 0.15\}$.
+(2003) reported the reliability estimates based on five so-called reliability generalization studies.
+These five studies summarized reliability in 25 – 51 primary studies (184 total), and the mean reliability ranged from 0.767 to 0.891 with standard deviations ranging between 0.034 – 0.133.
+Given some likely positive bias in such empirical values caused by selective reporting (Hussey et al., 2023), we examined the following mean reliabilities {0.6, 0.7, 0.8, 0.9}, and standard deviations {0, 0.05, 0.1, 0.15}.
 
@@
 
-The overall bias in heterogeneity estimates due to imperfect reliability in primary studies tended to be negative.
-We found that the positive bias due to variance in reliability across studies was less than the negative bias due to average unreliability across all levels of heterogeneity in our primary analysis.
-As such, we report here only the condition with maximum variance in reliability (SD = 0.15).
-Analyses with lower levels of reliability variance can be found in supplement XX.
-In addition, although a smaller number of studies in a meta-analysis (k) leads to a larger positive bias in the absence of heterogeneity, as would be expected, there was no practical difference in the bias once heterogeneity was above 0.1.
-As such, we report only results for k = 20 in the main manuscript.
-Analyses with other values for k (5, 40, 200) can be found in supplement XX.
-Finally, the effect on bias in heterogeneity estimates of varying within-study sample size was relatively small, and we similarly report only results for N = 150 in the main manuscript and present results for other sample sizes in supplement XX.
+Overall, the net bias in heterogeneity estimates due to imperfect reliability in primary studies was negative.
+We found that the inflation of heterogeneity due to variance in reliability across studies was less than the suppression due to average unreliability across all levels of heterogeneity in our primary analysis.
+For clarity of communication, we report here only the condition with maximum variance in reliability (SD = 0.15), as lower levels of variance will generate inflation between that observed for zero average effect size (i.e., zero inflation) and that generated under the maximum reliability variance.
+Consequently, analyses with smaller variances in reliabilities across primary studies entail more severe underestimates of heterogeneity for larger effect sizes than those presented here (supplement C).
+In addition, although a smaller number of studies in a meta-analysis (k) leads to a larger positive bias in the absence of heterogeneity, as would be expected, there was no practical difference in the bias once true heterogeneity exceeded $\tau = 0.1$ (Supplement C).
+Hence, we primarily report results for k = 20 in the main text, and only report results using other values for k (5, 40, 200) when considering heterogeneity below $\tau = 0.1$.
+Finally, for the sake of succinctness we similarly report only results for N = 150 in the main text, except when we discuss the effects of sample size on heterogeneity estimates (Figure 4).
 
-Figure 1 show the absolute and relative bias in heterogeneity estimates for different levels of heterogeneity (measured in $\tau$) and meta-analytic models.
-Each column of Figure 1 corresponds to one nominal level of heterogeneity ($\tau$ $\epsilon$ {0, 0.1, 0.15, 0.2}) and each row to a different analysis (HV with Fisher $z$, HV with Pearson's $r$, and HS).
-The black unbroken lines indicate actual heterogeneity, which may differ from nominal heterogeneity because of truncation or in the case of Fisher $z$ because the comparable heterogeneity levels were slightly different (see methods).
-The different dashed lines in Figure 1 shows the $\hat{\tau}$ estimates for four different levels of average reliability across studies, and the x-axis shows the average (superpopulation) effect size.
-As can be seen, the bias is marginally worse for Fisher $z$ than the other two methods (although Fisher $z$ is on a different scale) but the results are approximately the same for all three methods.
-As such, we will focus in the remainder of the results on Pearson's $r$ (row two).
+Figure 2 shows the net bias in heterogeneity estimates for different levels of heterogeneity (measured in $\tau$) and meta-analytic models.
+Each column of Figure 2 corresponds to one nominal level of actual heterogeneity ($\tau$ $\epsilon$ {0, 0.1, 0.15, 0.2}) and each row to a different analysis (HV with Fisher z, HV with Pearson’s r, and HS).
+The black solid lines indicate actual heterogeneity, which may differ from nominal heterogeneity because of effect size truncation or, in the case of Fisher z, because the comparable heterogeneity levels were slightly different (as discussed in the method section).
+The different dashed lines in Figure 2 show the $\tau$ ̂ estimates for four different levels of average reliability across primary studies, and the x-axis shows the average (superpopulation) effect size.
+As can be seen, the bias is marginally worse for Fisher z than the other two methods (although Fisher z is on a different scale), but the results are approximately the same for all three methods.
+As such, in the remainder of the results, we will focus on Pearson’s $r$ (row two).
 
 ```{r Figure 2, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/z-r-hs-plot.png")
+knitr::include_graphics("../figures/manuscript/z-r-hs-plot.png")
 ```
 
-_Figure 2._ EXPLANATION. Code to reproduce figure: LINK.
+_Figure 2.
+Unreliability in primary studies leads to a net negative bias when heterogeneity is positive.
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications.
+Columns indicate the nominal true heterogeneity standard deviation $\tau$.
+Due to truncation in Pearson’s r or translation to Fisher’s z these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Each dashed line corresponds to an average reliability in primary studies.
+Results are for standard deviation in reliability = 0.15, within-study sample size = 150, and 20 studies within each meta-analysis.
+Code to reproduce figure: LINK._
 
-As expected, for zero heterogeneity (leftmost column Figure 1) we see an overestimate of heterogeneity that increases with effect size.
-Although Pearson's $r$ has a small positive bias due to the dependance between effect size and sampling variance (supplement XX), more generally there are two sources of positive bias in the absence of heterogeneity.
-First, and the only one that applies when $\mu = 0$, there is positive bias due to the truncation of negative heterogeneity estimates (Viechtbauer, 2005).
-Second, there is bias due to the variance in reliability.
-For example, for $mu = 0.2$ (the closest value to the median correlation reported by Schäfer & Schwarz) and $\bar{R} = 0.8$ the bias for Pearson's $r$ compared to true zero heterogeneity is `r in_text$bias_tau_zero$mu_0.2`.
+As expected, for zero heterogeneity (leftmost column Figure 2) we see an overestimate of heterogeneity that increases with average effect size.
+Generally, there are two sources that explain the positive bias in the absence of true heterogeneity, although the use of Pearson’s r as the type of effect size additionally inflates heterogeneity estimates to a small degree for the HV model (row 2, Figure 2) due to the dependence between effect size and sampling variance (supplement A).
+First, and the only of these two upwardly biasing sources that applies when $\mu = 0$, the heterogeneity estimate is inflated due to the truncation of negative heterogeneity estimates (Viechtbauer, 2005).
+Second, there is inflation due to the variance in reliability across studies in the fixed effect scenario.
+For example, for $\mu = 0.2$ (the closest value to the median correlation reported by Schäfer & Schwarz, 2019) and $\bar{R} = 0.8$ the bias for Pearson's $r$ compared to true zero heterogeneity is `r in_text$bias_tau_zero$mu_0.2`.
 For $\mu = 0.4$ (the upper median reported by Schäfer & Schwarz) and $\bar{R} = 0.8$ the bias is `r in_text$bias_tau_zero$mu_0.4`.
-For the highest average reliability (0.9) a somewhat less inclined slope can be seen.
-This is because the standard deviation for this reliability level is severely truncated when the reliability is bounded at 1.
-The increase in positive bias with effect size can also be seen in the presence of heterogeneity (all other columns).
-However, it is superseded by the negative bias.
+For the highest average reliability (0.9) the slope is less inclined.
+This is because the standard deviation for this reliability level is severely truncated due to the reliability being bounded at 1.
+The upward force on estimations of heterogeneity with larger effect size can also be seen in the presence of heterogeneity (all other columns).
+However, this upward force is superseded by the suppression of heterogeneity that unreliability creates.
 
-For all heterogeneity levels in Figure 1 above zero we see an overall underestimate of heterogeneity due to unreliability in primary studies.
-This negative bias can be relatively severe, is worse with lower effect size and lower average reliability, and worsens in an absolute sense but improves in a relative sense as heterogeneity increases.
+For all true heterogeneity levels in Figure 2 above zero, we see an overall underestimate of heterogeneity due to unreliability in the primary studies.
+This negative bias can be relatively severe, is worse with smaller effect sizes and with lower average reliability, and worsens in an absolute sense as true heterogeneity increases.
 For example, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.1$, the bias is `r in_text$bias_increasing_tau$tau_0.1` for Pearson's $r$, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.15$ the bias is `r in_text$bias_increasing_tau$tau_0.15`, and for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.2$ the bias is `r in_text$bias_increasing_tau$tau_0.2`.
-For an effect size of $\mu = 0.2$ true heterogeneity between 0.1 - 0.2 is underestimated by between `r round(100*in_text$underestimate_mu02$underestimate_min$underestimate, 0)`% ($\tau = 0.2$, $\bar{R} = 0.9$) and `r round(100*in_text$underestimate_mu02$underestimate_max$underestimate, 0)`% ($\tau = 0.1$, $\bar{R} = 0.6$).
+For an effect size of $\mu = 0.2$, true heterogeneity between 0.1 - 0.2 is underestimated by between `r round(100*in_text$underestimate_mu02$underestimate_min$underestimate, 0)`% ($\tau = 0.2$, $\bar{R} = 0.9$) and `r round(100*in_text$underestimate_mu02$underestimate_max$underestimate, 0)`% ($\tau = 0.1$, $\bar{R} = 0.6$).
+In other words, the underestimation of true heterogeneity due to unreliability can be quite severe.
 
-To explore further under what conditions the positive and negative bias in heterogeneity estimates would reach an equilibrium, we added an analysis for several conditions with $\tau < 0.1$.
+To explore further under what conditions the forces of inflation and suppression of heterogeneity reach an equilibrium, we added an analysis for several conditions with $\tau$ < 0.1.
 We used the following values {0.02, 0.04, 0.06, 0.08}.
-Figure 2 presents the results of this analysis for Pearson's $r$.
-Because these lower heterogeneity levels can be expected to be affected by truncation to a larger extent (which results in a positive bias), we present results for four different numbers of studies in the meta-analysis (rows Figure 2, lower $k$ is associated with larger positive bias).
+Figure 3 presents the results of this analysis for Pearson’s r.
+Because these lower heterogeneity levels can be expected to be affected by heterogeneity truncation to a larger extent (which results in inflation), we present results for four different numbers of studies in the meta-analysis (rows Figure 3, lower k is associated with larger inflation of estimates).
 The standard deviation in reliabilities is again 0.15.
-Figure 2 shows that even when only little heterogeneity is present, it will generally be underestimated.
-If $\mu < 0.3$ then imperfect reliability leads to a negative bias in heterogeneity estimates for all $\tau < 0.06$ in Pearson's $r$.
-Even for lower levels of heterogeneity, heterogeneity will generally be positively biased only when average effect size is large or the number of meta-analyzed studies is small.
+Figure 3 shows that even when only little heterogeneity is present, it will generally be underestimated.
+If μ < 0.3 then imperfect reliability leads to a negative bias in heterogeneity estimates for all $\tau$ < 0.06 in Pearson’s r.
+Even for the lower levels of true heterogeneity, estimates of heterogeneity will generally be positively biased only when average effect size is large, or the number of meta-analyzed studies is small.
 
 ```{r Figure 3, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/r_tau_0.02-0.08.png")
+knitr::include_graphics("../figures/manuscript/r_tau_0.02-0.08.png")
 ```
 
-_Figure 3._ EXPLANATION. Code to reproduce figure: LINK.
+_Figure 3.
+Unreliability in primary studies typically leads to a net negative bias even for small degrees of heterogeneity.
+Effect size is Pearson’s r and rows indicate the number of studies in the meta-analysis.
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications.
+Columns indicate the nominal true heterogeneity standard deviation $\tau$.
+Due to effect size truncation in Pearson’s r these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Each dashed line corresponds to an average reliability in primary studies.
+Results are for standard deviation in reliability = 0.15, within-study sample size = 150.
+Code to reproduce figure: LINK._
+
+Finally, smaller sample sizes in primary studies tend to exacerbate the bias in observed heterogeneity estimates due to unreliability.
+In Figure 4, rows are the (fixed) sample size within primary studies when k = 20 and standard deviation in reliabilities is 0.15.
+When true heterogeneity is zero (leftmost column Figure 4), a smaller sample size tends to lead to more positive bias in estimated heterogeneity.
+When true heterogeneity is non-zero, smaller sample size corresponds with larger negative bias in heterogeneity estimates.
+For higher levels of reliability, differences due to sample size are small, but differences can be more concerning when average reliability is low.
+For example, consider N = 50 in our simulation compared with N = 200 (top versus bottom row Figure 4) when true heterogeneity lies between 0.1 – 0.2 and the average effect size is μ = 0.2.
+If average reliability is $\bar{R} = 0.8$, the lower sample size then leads to at most $\hat{\tau}$ =  `r in_text$bias_n50_0.8` additional bias in heterogeneity estimates in our simulations.
+However, when $\bar{R} = 0.6$, the additional bias can be as much as $\hat{\tau}$ = `r in_text[["bias_n50_0.6"]]` for an average sample size of 50 compared to 200, making up approximately one third of the total bias (`r in_text$net_bias_n50`) in this condition (N = 50, $\tau$ = 0.15).
+Generally, for an effect size of μ = 0.2, true heterogeneity between 0.1 – 0.2 is underestimated by between 16% ($\tau$ = 0.2, $\bar{R}= 0.9$) and 73% ($\tau$ = 0.1, $\bar{R} = 0.6$) in our simulations when sample size is 50, as compared to the 16-59% we observed when sample size was 150.
+
+```{r Figure 4, echo=FALSE, out.width = '100%'}
+knitr::include_graphics("../figures/manuscript/sample_size.png")
+```
+
+_Figure 4.
+A smaller sample size in primary studies tends to exacerbate the bias in observed heterogeneity estimates due to unreliability.
+The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications.
+Rows indicate the (fixed) sample size within studies and columns indicate nominal true between-studies standard deviation.
+Due to effect size truncation in Pearson's $r$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Each dashed line corresponds to an average reliability in primary studies.
+Results are for standard deviation in reliability = 0.15 and 20 studies within each meta-analysis._
+
+# Discussion
+
+We used Monte-Carlo simulations to study the total bias in meta-analytic heterogeneity estimates due to measurement unreliability in primary studies.
+Our results indicate that under most circumstances, uncorrected unreliability in primary studies leads to a severe underestimate of heterogeneity in meta-analyses.
+We observed a positive bias in heterogeneity estimates due to unreliability only under specific and arguably uncommon circumstances of (1) actual zero heterogeneity, particularly when mean effect sizes are large, or (2) combinations of very small true heterogeneity, large variance in primary study reliabilities, large mean effect sizes, and a limited number of primary studies.
+As expected, heterogeneity estimates generally improved with higher average reliability, and deteriorated with smaller average study sample size.
+Our results were approximately the same whether one used a classic random-effects meta-analytic model (Hedges & Vevea, 1998) or a ‘bare-bones’ psychometric meta-analytic model (Hunter & Schmidt, 2015), and whether one used Pearson’s r as the effect size or Fisher’s z, although the negative bias was slightly larger with Fisher’s z.
+The severity of the underestimation of heterogeneity could lead to meta-analyses in psychology missing relevant moderators and overgeneralizing outcomes.
+Unreliability appears to bring truly different effects closer together, thereby obscuring true differences between studies that could be relevant for theory, practice, and future research efforts.
+
+In meta-analyses, low reliabilities not only lower the average effect size estimate, but also artificially suppress heterogeneity estimates in most cases.
+For a typical effect size around r = 0.2 and substantial true heterogeneity ($\tau$>0.1) that is not uncommon in the meta-analytic literature (van Erp et al., 2017), we found expected heterogeneity to be underestimated by about 15% - 60% in our simulations primarily depending on average reliability in primary studies (higher reliability leads to a smaller underestimate), although the underestimate could be as high as 73% with low sample sizes.
+As true heterogeneity increases, the proportional underestimate decreases to some degree, but the underestimate in absolute terms will worsen.
+Variability in reliabilities across primary studies is a counter-acting force that inflates heterogeneity estimates as average effect size increases.
+Consequently, the underestimate of heterogeneity can become less severe as effect size increases, although the overall bias is likely to remain negative.
+Small degrees of true heterogeneity ($\tau$ < .08) can be overestimated under certain conditions, but this would typically require a combination of a small number of studies, large variance in study reliabilities, and a large average effect size.
+The severity  and generality of the underestimate in heterogeneity due to unreliability means that meta-analysts should not ignore the effect of reliability on their heterogeneity estimates and that research is needed to provide practical guidance to meta-analysts.
+
+If true heterogeneity is zero, there is a positive bias in observed heterogeneity estimates, but for even small degrees of true heterogeneity there tends to be a negative bias in observed estimates.
+As such, it is not possible to know whether a small degree of observed heterogeneity is due to overestimated true zero heterogeneity or underestimated small heterogeneity.
+However, for observed heterogeneity above $\tau > 0.025$ one can usually be confident that the estimate is an underestimate of true heterogeneity rather than an overestimate of zero heterogeneity if the average effect sizes is below approximately $r = 0.2$.
+
+Without any information on primary studies’ reliabilities, it would often be impossible to know whether a small observed heterogeneity is due to overestimated true zero heterogeneity or underestimated small heterogeneity.
+Moreover, it is certainly possible to observe a zero heterogeneity estimate under larger true heterogeneity levels, because the sampling variability in heterogeneity estimates is often substantial in practice (Ioannidis et al., 2007; Olsson-Collentine et al., 2020).
+Generally, distinguishing between zero and small heterogeneity in empirical meta-analyses is statistically challenging considering the typical sample sizes and numbers of studies in meta-analyses in psychology (Olsson-Collentine et al., 2020).
+The power to detect small true heterogeneity in most psychological meta-analyses is low.
+In our view, the distinction between zero and small heterogeneity in meta-analysis is primarily a hypothetical one as true zero heterogeneity is not expected in meta-analytic practice to begin with (see also Hedges, 1987), except perhaps in specific sets of very similar direct replications (Olsson Collentine et al., 2020).
+
+We found that using Pearson’s r or Fisher’s z as our effect size resulted in very similar degrees of bias in heterogeneity estimates due to uncorrected unreliability in primary studies.
+The underestimate is slightly worse for Fisher’s z than for Pearson’s r.
+The usage of Pearson’s r inflates heterogeneity estimates slightly (due to its bounded nature and effect size dependent sampling variance), which is apparently beneficial for heterogeneity estimates when not correcting for unreliability.
+However, the additional negative bias when using Fisher’s z is small and we consider the difference negligible in the light of the severe underestimates generated by unreliability.
+
+## Should meta-analysts correct for unreliability in primary studies?
+
+Our simulations show that heterogeneity estimates are biased when unreliability is ignored, so in principle one should correct for unreliability.
+However, even done perfectly this would not mean that heterogeneity estimates end up to be accurate, because few meta-analyses are able to include sufficient studies (van Erp et al., 2017; Davey et al., 2011) to measure heterogeneity without large uncertainty and wide confidence intervals (Ioannidis et al., 2007, Olsson-Collentine, 2020).
+In practice, attempting to correct for unreliability comes with many caveats and assumptions that must be considered.
+We first consider challenges with correcting for unreliability in the CTT framework presuming that CTT assumptions hold.
+Next, we consider potential concerns with the CTT assumptions and whether structural equation modeling (SEM) is a viable alternative in the meta-analytic context.
+
+There are several concerns with correcting for unreliability in the CTT framework even when its assumptions hold.
+First, reliability estimates are lower bound estimates, meaning that systematically correcting for unreliability can be expected to lead to overestimated mean effect sizes and subsequent overestimated heterogeneity.
+Unfortunately, the most popular reliability estimate (Cronbach’s Alpha; Flake et al., 2017; Hogan et al., 2002) is known to underestimate the lower bound even further than several alternatives, particularly when the strong assumptions underlying CTT are not met (Sijtsma, 2008).
+Second, CTT corrections of correlations as commonly applied (Hunter & Schmidt, 2015; but see Raju et al., 1991; Charles, 2005) do not consider that reliability estimates themselves are subject to substantial sampling error (Fan & Thompson, 2001) that add additional uncertainty to corrected effect sizes and subsequent estimates of heterogeneity.
+Third, there are several types of reliability estimates (e.g., "internal consistency”, test-retest, coefficient of equivalence) that do not necessarily derive from the same type of measurement error and may result in different reliability estimates (Dimitrov, 2002; Revelle & Condon, 2019).
+If the reliability estimate applied does not account for all relevant sources of measurement error, it will overestimate reliability and heterogeneity will remain underestimated despite unreliability corrections.
+Fourth, reliability estimates are sometimes not reported in primary studies (Flake et al., 2017).
+For meta-analysis, missing estimates are then often imputed based on average observed reliability or based on some ‘typical’ reliability (Hunter & Schmidt, 2015; Gnambs & Sengewald, 2023).
+However, what a ‘typical’ reliability should be is open to debate (LeBreton et al., 2014), and mean imputation leads to a biased pooled estimate, especially so if missing statistics are related to the value of the statistic (and it seems likely that non-reporting is more common when reliability is low).
+Mean imputation is seen by experts in missing data as “possibly the worst missing data handling method available” (Enders, 2010, p.
+43; see also van Buuren, 2018, p.
+12; Schafer, 2002, p.
+159).
+
+Correcting for unreliability in the CTT framework comes with several assumptions that may be hard to motivate in practice.
+First, most estimates of reliability in CTT assume unidimensionality (e.g., Borsboom & Mellenberh, 2002; Savalei & Reise, 2019), and violations of this assumption can severely bias reliability estimates (Savalei & Reise, 2019).
+Second, correcting for unreliability in the CTT framework assumes linear associations between scale and latent construct, which implies that corrections may result in biased estimates when ordinal scales (such as Likert-scales) are short (Zheng, 2022).
+Based on their simulations, Zheng (2022) recommends that ordinal scales should have at least 12 items (with $\geq$ 5 answer categories per item) to use Cronbach’s Alpha to correct for unreliability in meta-analysis.
+Problematically, Flake et al.
+(2017) report that the average scale length in psychology (excluding 1-item scales) is 4.69 items (SD = 6.35).
+Third, true scores $X$ and $Y$ are assumed to be normally distributed.
+Zheng (2022) demonstrated in simulation studies that when this assumption does not hold average effect size estimates will be inaccurate, especially when the average effect size is large (although they did not examine heterogeneity).
+Fourth, estimators such as Cronbach’s Alpha assume uncorrelated errors (e.g., Dimitrov, 2002; Teo & Fan, 2013) and essential tau-equivalence (i.e., items have equal factor loadings on a latent construct; e.g., Revelle & Condon, 2019).
+Correlation of errors can lead to a substantial overestimate of reliability (e.g., Dimitrov, 2002; Teo & Fan, 2013), and may occur in many measurement situations, such as when items on a scale refer to the same stimuli (e.g., multiple reading comprehension items referring to the same paragraph; Green & Yang, 2009).
+Overestimation of reliability would leave heterogeneity estimates with a negative bias remaining even after CTT correction.
+The assumption of essential tau-equivalence is empirically unrealistic, and violations lead to an underestimate of the reliability statistic (Sijtsma, 2009, p.
+111).
+This bias may often be small enough to be neglected (Green & Yang, 2009; Savale & Reise, 2019), but can interact with the assumption of normal distributed true scores and result in inaccurate estimates of average effect sizes even for ordinal scales with more than 12 items (Zheng, 2022).
+Extending the study of Zheng (2022) also to examine the effects on heterogeneity estimates would be a useful avenue of future research.
+
+When all of the CTT assumptions hold, and if practical concerns can be overcome, CTT corrections of unreliability appear to function well in meta-analysis of correlations (e.g., Hall & Brannick, 2002; Brannick et al., 2019; Hunter & Schmidt, 2015).
+However, research appears to be lacking on the consequences of CTT corrections for meta-analysis when assumptions do not hold (but see Zhang, 2022).
+Moreover, because latent constructs are defined by latent variable models (Borsboom & Mellenbergh, 2002), the strong assumptions of CTT regarding latent associations, factor loadings, uncorrelated errors, and the distribution of true scores may be unpalatable to many researchers.
+Modern latent models based on Structural Equation Modelling (SEM; e.g., Lei & Wu, 2007) may then be an option, as they permit delving into measurements and latent construct structures rather than relying on assumptions.
+
+Meta-analysis can be conceptualized as a SEM model (e.g., Cheung, 2015), which holds promise that the benefits of modern psychometric theory might be applied also to latent constructs summarized across multiple studies.
+In Meta-Analytic Structural Equation Modeling (MASEM; e.g., Cheung & Chan, 2005) unreliability in primary studies can be corrected in three ways (Gnambs & Sengewald, 2023): (1) by estimating a SEM model in each primary study and then pooling estimates (parameter-based MASEM: Cheung, 2015, p.
+241), (2) by directly modeling latent variables in the MASEM,  or (3) by first correcting correlations before introducing them into the MASEM framework (i.e., fundamentally identical to the CTT approach).
+The first option is limited because it is rarely the case that the same SEM model can be fit across multiple independent datasets, except in highly structured cases (e.g., Brunner et al., 2022).
+The second option is the preferable option, but requires item-level statistics rather than summary values, which are rarely available to meta-analysts.
+The third option, naturally, suffers the same challenges and assumptions as corrections under classical test theory.
+As such, the potential to use MASEM to take advantage of modern psychometric theory when correcting for unreliability is challenging in practice.
+That said, multi-lab replication projects are increasingly common in psychology (e.g., Klein et al., 2014, McCarthy et al., 2021) and consist of a set of direct replication studies where analysts have access to all data.
+As such, they are well-placed to take full advantage of MASEM for their analyses.
+
+In the end, meta-analysis depends on the quality of the primary studies that make up its data.
+As long as researchers in psychology underappreciate the importance of measurement, meta-analytic estimates will be poor.
+Meta-analysts that consider the challenges and assumptions of CTT corrections of unreliability acceptable should acknowledge the many caveats inherent in doing so and report average effect size and heterogeneity estimates based on both corrected and uncorrected estimates.
+Meta-analysts who consider CTT corrections untenable need to consider that heterogeneity estimates may be severely underestimated.
+In either case, meta-analysts should get used to extracting reliability information whenever available from primary studies.
+
+## Limitations
+
+Arguably the biggest challenge in estimating heterogeneity in meta- practice is the large degree of sampling variance in heterogeneity estimates that can be expected in almost any meta-analysis (Ioannidis et al., 2007; Olsson-Collentine et al., 2020).
+Consider a meta-analysis without measurement error of k =12 studies (the median number in psychology; van Erp et al., 2017) with N = 150 in all studies, $\tau = 0.17$ (the median observed $\tau$ for correlations in van Erp et al., 2017), and  $\mu = 0$.
+Across 10,000 replications the median width of the 95% confidence interval around the heterogeneity estimate (using default settings in metafor) is then 0.195 in Pearson’s $r$ units.
+In our study, we did not examine sampling variance in heterogeneity estimates or their root mean squared error.
+This is because although we apply different meta-analytic models and effect sizes, the purpose was not to compare the efficiency of these different modes of estimation.
+As the effect of unreliability in primary studies is simply to change the heterogeneity of effect sizes, the results of previous studies on the comparable performance of different heterogeneity estimators applies also here (e.g., Langan et al., 2017; Veroniki et al., 2016; Viechtbauer, 2005).
+That said, it would be useful to examine the variance of heterogeneity estimates in future research when the focus is on developing practical guidance for meta-analysts.
+Doing so could be helpful to gain a better idea of how likely a particular observed heterogeneity estimate is to belong to different distributions of true heterogeneity in the presence of measurement unreliability.
+As this would depend on N, K, mean reliability, variability in reliability, average effect size, observed heterogeneity, and the true heterogeneity levels being compared, implementing such an analysis would probably be best done on a case-by-case basis (e.g., through a webapp such as developed for sensitivity analyses of the effect of publication bias on heterogeneity by Augusteijn et al., 2017) .
+Given the challenges in correcting for unreliability, such a sensitivity analysis could be a useful tool to reason around the impact of unreliability for applied meta-analysts.
+An additional challenge when providing practical advice to meta-analysts is how to handle selective reporting of study reliabilities (Hussey et al., 2023) in primary studies, and the problems that p-hacking and publication bias create in estimating heterogeneity.
+We know from Augusteijn et al.
+(2017) that publication bias alone creates complex outcomes, but when meta-analysis is affected by unreliability and p-hacking true heterogeneity is additionally obscured.
+
+Our simulations come with some caveats.
+First, we assume reliabilities to be known.
+This is never the case, and any estimate of reliability is at best a lower bound estimate subject to sizeable sampling errors.
+By assuming reliabilities to be known, we were able to show the bias of observed heterogeneity estimates compared to the ideal case of no measurement error.
+However, this may mislead readers to believe that a simple correction for unreliability will leave them with an unbiased estimate.
+As discussed in the section ‘should meta-analysts correct for unreliability?’ correcting for reliability under realistic circumstances and using estimated reliability is not straightforward and whether the resulting estimate will be unbiased is contingent on several factors that remain to be studied.
+Examining the effect of corrections based on estimated reliabilities under realistic conditions in future studies could offer practical guidance to meta-analysts.
+
+Second, our simulations assumed equal reliability for the measures $X$ and $Y$ (i.e., $R_i = R_{xx'} = R_{yy'}$) within a study, which is not realistic.
+However, as we never use the reliabilities individually (i.e., only expressed as $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$) we do not expect major impact on our results.
+Consider that if $R_i = 0.7$ then $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}} = 0.7$ whereas if if $R_{xx'}$ and $R_{yy'}$ took other values from a distribution centered on 0.7 this value would be lower (e.g., if $R_{xx'} = 0.5$ and $R_{yy'} = 0.9$ then $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}} \approx 0.67$).
+As we use the results of $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ in our simulations to generating observed effect sizes from true effect sizes (i.e., $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$, where $\rho_{xy}$ is the true effect size after sampling error), letting the reliability for $X$ and $Y$ vary would lead to slightly larger attenuation of observed effect sizes on average.
+That is, letting the reliability of $X$ and $Y$ vary within studies would lead to larger negative bias in observed heterogeneity estimates than in our simulations, indicating that the effect might even be somewhat more severe than in our results.
+
+Third, we used a fixed sample size (N) across studies within a meta-analysis in our simulations, which is unrealistic.
+We showed that increasing the primary study sample size led to better heterogeneity estimates, especially for low reliability.
+There is large variability in how simulation studies focused on heterogeneity implement sample sizes (Langan et al., 2017).
+Most studies sample from a uniform distribution of sample sizes, but there are also studies that have used conditions with fixed sample sizes like us (Langan et al., 2018).
+There appears to be no systematic exploration of the effect of different distributions of sample sizes on heterogeneity estimation, and the effect of sample size is generally considered in view of its average value (e.g., Viechtbauer, 2005).
+The range of sample sizes in a meta-analysis appears to affect the performance of heterogeneity estimation (Langan et al., 2017), but we believe this relates only to the efficiency of heterogeneity estimates rather than bias, given a sufficiently large number of repetitions in the simulations.
+That said, implementing a realistic distribution of sample sizes would be important for research attempting to provide concrete guidance to meta-researchers on how to interpret observed heterogeneity in the presence of measurement error as variability in estimates then is important.
+Implementing a distribution of sample sizes based on relevant empirical literature (e.g., as done by Brannick et al., 2019) could be one way to increase realism in simulated sample sizes.
+
+Finally, we only considered bias in heterogeneity due to unreliability in primary studies, but there exist other measurement artifacts such as restriction of range, dichotomization, and more (Hunter & Schmidt, 2015).
+Although measurement error is the only measurement artifact that is always present, many of these other measurement artifacts may impact heterogeneity estimates when they are present.
+Based on Hunter & Schmidt’s arguments that measurement artifacts all tend to attenuate effect sizes in a similar way to unreliability (p.
+78-79), we expect that their net effect on heterogeneity estimates may similarly be a negative bias.
+The extent to which this prediction holds true (and whether these measurement artifacts can be corrected for in MASEM; Gnambs & Sengewalds, 2023) awaits further study.
+
+## Conclusion
+
+Imperfect measurement reliability in primary studies generally leads to a severe underestimate of observed meta-analytic heterogeneity.
+Unreliability may thereby obscure true differences between studies that could be relevant for theory, practice, and future research efforts.
+As few meta-analyses in psychology correct for unreliability in primary studies, heterogeneity is likely underestimated in a large proportion of meta-analyses in psychology.
+Yet, sophisticated methods for correcting measurement unreliability such as meta-analytic structural equation modeling (MASEM) are only applicable in exceptional cases and corrections based on classical test theory come with caveats and strong assumptions that are often unrealistic and currently do not fully consider insights from modern test theory.
+Accurate estimation of meta-analytic heterogeneity is difficult and will remain so unless measurement concerns (Flake & Fried, 2020; Flake et al., 2017; Kane et al., 2021) are taken seriously in primary research.
+The good news is that study designs (multi-lab replication studies) which can apply more sophisticated versions of MASEM are becoming increasingly common in psychology, allowing us to study how to best deal with measurement errors when estimating variation of true effect sizes in meta-analyses

### 2023-06-17 code/06_tables.qmd
floated tables for two column environment

--- 2023-05-21 code/tables.qmd
@@
                         "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .8",
-                        "SD(ES)")) %>%
+                        "SD(ES)"),
+                        table.env = 'table*') %>%
   kable_styling(position = "left") %>%
-  add_header_above(c(" " = 2, "Study 1"= 1, "Study 2" = 1, "Study 3" = 1)) %>%
-  add_header_above(c(" " = 2, "Observed Effect Sizes" = 3)) %>%
+  add_header_above(c(" " = 2, "Study 1"= 1, "Study 2" = 1, "Study 3" = 1, " " = 1 )) %>%
+  add_header_above(c(" " = 2, "Observed Effect Sizes" = 3, " " = 1)) %>%
   footnote(general = "Reproduced from Olsson Collentine et al.
@@
                "Study 3",
-               "SD(ES)")) %>%
+               "SD(ES)"),
+               table.env = 'table*') %>%
  kable_styling(position = "left") %>%
- add_header_above(c(" " = 1, "Observed Effect Sizes" = 3)) %>%
+ add_header_above(c(" " = 1, "Observed Effect Sizes" = 3, " " = 1)) %>%
  footnote(general = "The values under Study 1, 2 and 3 are observed effect sizes for that study given the measurement reliability $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ and the true effect size of that study (first row) when within-study sample size is infinite.
@@
              negative_bias_table = negative_bias_table),
-        "../manuscript/tables.RDS")
+        "./manuscript/tables.RDS")
 ```

### 2023-06-17 code/06_tables.qmd
table 2 now single column

--- 2023-05-21 code/tables.qmd
@@
 ---
-
-# Project: Unreliable heterogeneity
 
@@
 ```
-
-## Table 1
-### Variance in reliability leads to positive bias in heterogeneity estimates
 
@@
                         "$\\sqrt{R_{xx'}}\\times\\sqrt{R_{yy'}}$ = .8",
-                        "SD(ES)")) %>%
+                        "SD(ES)"),
+                        table.env = 'table*',
+                        label = "effect-size-heterogeneity",
+                        caption = "Variance in reliability inflates true heterogeneity") %>%
   kable_styling(position = "left") %>%
-  add_header_above(c(" " = 2, "Study 1"= 1, "Study 2" = 1, "Study 3" = 1)) %>%
-  add_header_above(c(" " = 2, "Observed Effect Sizes" = 3)) %>%
+  add_header_above(c(" " = 2, "Study 1"= 1, "Study 2" = 1, "Study 3" = 1, " " = 1 )) %>%
+  add_header_above(c(" " = 2, "Observed Effect Sizes" = 3, " " = 1)) %>%
   footnote(general = "Reproduced from Olsson Collentine et al.
@@
 ```
-
-## Table 2
-### Imperfect reliability leads to negative bias in heterogeneity estimates
 
@@
                "Study 3",
-               "SD(ES)")) %>%
+               "SD(ES)"),
+               label = "suppress-heterogeneity",
+               caption = "Imperfect reliability suppresses true heterogeneity") %>%
  kable_styling(position = "left") %>%
- add_header_above(c(" " = 1, "Observed Effect Sizes" = 3)) %>%
+ add_header_above(c(" " = 1, "Observed Effect Sizes" = 3, " " = 1)) %>%
  footnote(general = "The values under Study 1, 2 and 3 are observed effect sizes for that study given the measurement reliability $\\\\sqrt{R_{xx'}}\\\\times\\\\sqrt{R_{yy'}}$ and the true effect size of that study (first row) when within-study sample size is infinite.

### 2023-06-17 manuscript/manuscript.qmd
quarto formatting

@@
 title: "Unreliable heterogeneity: how measurement error obscures heterogeneity in meta-analyses in psychology"
-preprint: true
+date: today
+date-format: long
 author:
-  - name: "Anton Olsson-Collentine"
-    affiliation: 1
-    corresponding: true
+  - name: Anton Olsson-Collentine
+    url: https://example.com
+    orcid: 0000-0000-0000-0000
+    affiliations:
+      - id: tiu
+        name: "Department of Methodology and Statistics, Tilburg School of Social and Behavioral Sciences, Tilburg University, the Netherlands."
     email: anton@olssoncollentine.com
-
-  - name          : "Marjan Bakker"
-    affiliation   : "1"
-
-  - name          : "Jelte Wicherts"
-    affiliation   : "1"
-affiliation:
-  - code: 1
-    address: "Department of Methodology and Statistics, Tilburg School of Social and Behavioral Sciences, Tilburg University, the Netherlands."
-
-abstract: >
+  - name: Marjan Bakker
+    affiliations:
+      - ref: tiu
+  - name: Jelte Wicherts
+    affiliations:
+      - ref: tiu
+format:
+  bj-pdf:
+    classoption:
+      - twocolumn
+colorlinks: true
+link-citations: true
+keep-tex: false
+abstract: |
   One of the main purposes of meta-analysis is to estimate heterogeneity of effect sizes.
@@
 Research on concrete guidance to applied meta-analysts is urgently needed, as sophisticated methods for correcting measurement unreliability such as meta-analytic structural equation modeling (MASEM) are only applicable in exceptional cases and corrections based on classical test theory come with caveats and strong assumptions.
-
-header-includes:
-   - \usepackage{setspace}
-   - \usepackage{lineno}
-   - \linenumbers
-
+license: CC BY
+keywords:
+  - Research
+  - Cool Stuff
+citation:
+  type: article
+  citation-key: citationKeyNumber
+  status: draft
+cite-method: citeproc
 csl: apa.csl
-
-output:
-  bookdown::pdf_book:
-    base_format: rticles::peerj_article # for using bookdown features like \@ref()
-    keep_tex: true
-  rticles::peerj_article: default
-
+bibliography: references.bib
 ---
@@
 
-Table 1.
+```{r positive bias table}
 
-*Variance in reliability inflates heterogeneity in combination with increased average effect size and leads to overestimated heterogeneity in a fixed effects model*
-
-```{r positive bias table}
+#| tbl-pos: t
 
@@
 Table 2 illustrates this relationship for infinite sample size (i.e., no sampling variance) and three levels of reliability across three studies.
-
-Table 2.
-
-*Imperfect reliability suppresses true heterogeneity*
 
@@
 
+```{r Figure 1, out.width= "85%"}
+#| fig.env: figure*
+#| fig-pos: 't'
+#| fig-cap: "Figure 1.
+Flowchart summarizing the simulation design.
+The flowchart shows the procedures followed to meta-anlyze results using the Hedge’s and Vevea model (see ‘Meta-Analytic Model’ header in methods section).
+The Hunter & Schmidt ‘bare-bones’ meta-analytic model followed the Pearson correlations procedure, but used sample size weights and replaced $r_i$ in the sampling variance by r for meta-anlysis."
+
+knitr::include_graphics("../figures/manuscript/flowchart-simulation-design.png")
+```
+
 Estimating heterogeneity in both the HV and bare-bones HS model essentially consists of comparing the observed variance in effect sizes with what would be expected from sampling error alone and then ascribing any excess observed variance to heterogeneity (Borenstein et al., 2010, Box 1; Hunter & Schmidt, 2015, p.
@@
 Our design was broadly the same for product-moment correlations and Fisher's $z$, with some minor differences we highlight in the text below detailing the procedure.
-
-```{r Figure 1, out.width= "100%"}
-knitr::include_graphics("../figures/manuscript/flowchart-simulation-design.png")
-```
-
-_Figure 1.
-Flowchart summarizing the simulation design.
-The flowchart shows the procedures followed to meta-anlyze results using the Hedge’s and Vevea model (see ‘Meta-Analytic Model’ header in methods section).
-The Hunter & Schmidt ‘bare-bones’ meta-analytic model followed the Pearson correlations procedure, but used sample size weights and replaced_ $r_i$ _in the sampling variance by $\bar{r}$ for meta-anlysis._
 

### 2023-06-17 manuscript/manuscript.qmd
dynamic references

@@
 csl: apa.csl
-bibliography: references.bib
+bibliography: [unreliable-heterogeneity.bib]
 ---
@@
 Estimating heterogeneity of effect sizes is usually considered the main purpose of meta-analysis, alongside estimating an average effect size.
-Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size’s sensitivity to variability in study design features related to the type of (1) sample, (2) setting, (3) treatment variable, and (4) measurement (e.g., Shadish, Cook, & Campbell, 2002).
+Heterogeneity of effect sizes (henceforth referred to as heterogeneity) refers to an effect size’s sensitivity to variability in study design features related to the type of (1) sample, (2) setting, (3) treatment variable, and (4) measurement [e.g., @cook2002].
 Heterogeneity is an important outcome in meta-analysis for several reasons.
-First, its presence signals the existence of moderators and so can be seen as an opportunity for theoretical development (Simons et al., 2017).
+First, its presence signals the existence of moderators and so can be seen as an opportunity for theoretical development [@10.1177/1745691617708630].
 Second, heterogeneity affects the implementation of research by indicating that an intervention may not be equally effective under all conditions or for everyone.
 Third, the presence of heterogeneity changes the interpretation of the average effect size derived in a meta-analysis from the true population effect size (under homogeneity) to the average of true subpopulation effect sizes (under heterogeneity).
-As such, estimating heterogeneity with minimal bias is of central importance to meta-analysis (for a more extensive discussion on the importance of heterogeneity, see Olsson Collentine et al., 2020; Simons, 2017).
+As such, estimating heterogeneity with minimal bias is of central importance to meta-analysis [for a more extensive discussion on the importance of heterogeneity, see @10.1037/bul0000294; @10.1177/1745691617708630].
 
 Meta-analytic estimates depend on the quality of data in primary studies.
-However, there is concern that researchers in psychology tend to neglect proper measurement (Flake & Fried, 2020) to the extent that psychology can be said to be in a measurement crisis (Kane et al., 2021).
-For example, Flake et al.
-(2017) report that although latent constructs are widely examined in personality and social psychology, researchers barely report any of the validity evidence required to ascertain the extent to which implemented scales measure the constructs of interest.
+However, there is concern that researchers in psychology tend to neglect proper measurement [@10.1177/2515245920952393] to the extent that psychology can be said to be in a measurement crisis [@10.3758/s13428-021-01557-x].
+For example, @10.1177/1948550617693063 report that although latent constructs are widely examined in personality and social psychology, researchers barely report any of the validity evidence required to ascertain the extent to which implemented scales measure the constructs of interest.
 Such inattention to measurement can also bias heterogeneity estimates.
-We will focus here on the reliability of measurement (defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., Lord & Novick, 1968), which is the only  evidence on measurement quality that researchers report consistently (Flake et al., 2017), and which can be corrected for in meta-analysis (Hedges & Olkin, 1985; Hunter & Schmidt; 2015).
+We will focus here on the reliability of measurement [defined in classical test theory as the ratio of true score variance to measurement error variance; e.g., @lord1968], which is the only  evidence on measurement quality that researchers report consistently [@10.1177/1948550617693063], and which can be corrected for in meta-analysis [@hedges1985; @schmidt2015] .
 
-Measurement error (imperfect reliability) is present in any empirical effect size estimate and systematically attenuates observed effect sizes compared to true underlying effect sizes (Spearman, 1904).
+Measurement error (imperfect reliability) is present in any empirical effect size estimate and systematically attenuates observed effect sizes compared to true underlying effect sizes [@10.2307/1412159].
 This affects heterogeneity estimates, which from a statistical perspective, estimate the variance between true effect sizes.
 However, reliability is not necessarily accounted for in heterogeneity estimates, because ‘true’ underlying effect sizes generally refer to two different entities: Either effect sizes free from sampling error, or effect sizes free from both sampling error and measurement error.
-Whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for measurement errors is explicitly modelled (alongside other measurement artifacts) only in the psychometric meta-analysis approach by Hunter & Schmidt (2015).
+Whereas all meta-analytic models attempt to correct for sampling error in their estimates, correcting for measurement errors is explicitly modelled (alongside other measurement artifacts) only in the psychometric meta-analysis approach by @schmidt2015.
 This has consequences for heterogeneity estimates reported in many areas of psychology.
@@
 Meta-analyses in psychology rarely correct for unreliability in primary study measurements.
-Wiernick and Dahlke (2020) reported that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability.
-Similarly, Schmidt (2010) reported that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts.
-The exception is the subfield of industrial-organizational studies, where corrections tend to be more common (e.g., Cortina, 2003; Anguinis, 2011).
+@10.1177/2515245919885611 reported that among the 71 meta-analytic studies published in the journal Psychological Bulletin between 2016 and 2018, only 6/71 (8%) corrected for unreliability.
+Similarly, @10.1177/1745691610369339 reported that only 19/199 (10%) of the meta-analytic studies published in the same journal between 1978 - 2006 corrected for any measurement artifacts.
+The exception is the subfield of industrial-organizational studies, where corrections tend to be more common [e.g., @10.1177/0149206310377113; @10.1177/1094428103257358].
 
-One reason that few meta-analysts correct for unreliability could be that correcting for unreliability has long been controversial (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017).
-On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017).
+One reason that few meta-analysts correct for unreliability could be that correcting for unreliability has long been controversial [for arguments spanning the last century, see Table 2 of @10.1111/iops.12184].
+On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., @10.1016/0001-87918890043-7; @10.1111/iops.12184].
 Systematically correcting for unreliability may also lead to an overestimation of effect sizes, because reliability estimates are lower bound estimates of the true reliability (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1).
-Even worse, the most popular estimate of reliability in psychology (Cronbach’s Alpha; Flake et al., 2017), as well as improvements thereof, tend to underestimate this lower bound of reliability (Sijtsma, 2008).
-Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated in psychology (e.g., Kvarven et al., 2019; Schäfer & Scharz, 2019; Scheel et al., 2020) due to selective reporting based on the significance of outcomes (Simmons et al., 2011) and publication bias (Thornton & Lee, 2000), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
+Even worse, the most popular estimate of reliability in psychology [Cronbach’s Alpha; @10.1177/1948550617693063], as well as improvements thereof, tend to underestimate this lower bound of reliability [@10.1007/s11336-008-9101-0].
+Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated in psychology [e.g., @10.1038/s41562-019-0787-z; @10.3389/fpsyg.2019.00813; @10.1177/25152459211007467] due to selective reporting based on the significance of outcomes [@10.1177/0956797611417632] and publication bias [@10.1016/S0895-43569900161-4], so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer.
 On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures.
 As such, neglecting measurement errors means that computed estimates do not correspond to the entity of interest.
-From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015), although treating disattenuated effect sizes as directly comparable to latent scores requires strong assumptions (Borsboom & Mellenbergh, 2002).
+From this perspective, correcting for unreliability is desirable even if doing so is challenging [e.g., @10.1017/iop.2015.17; @schmidt2015], although treating disattenuated effect sizes as directly comparable to latent scores requires strong assumptions [@10.1016/S0160-28960200082-X].
 In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses need to be aware of how unreliability of measurement affects heterogeneity estimates.
-Hunter & Schmidt (2015) assert that failure to control for variation in unreliability across studies results in a large overestimate of heterogeneity (p.
+@schmidt2015 assert that failure to control for variation in unreliability across studies results in a large overestimate of heterogeneity (p.
 159), but little systematic research has been done to study the effects of unreliability on estimating heterogeneity.
@@
 For product-moment correlations the observed correlation can be computed as $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ where $r_{xy}$ is the observed correlation between the variables $X$ and $Y$, $\rho_{xy}$ is the true correlation (with or without sampling error) and $R_{xx'}$ and $R_{yy'}$ are the measurement reliabilities for $X$ and $Y$.
-As such, correcting for unreliability is straightforward for product-moment correlations, but can also be done for other effect size types (Wiernick and Dahlke, 2020).
+As such, correcting for unreliability is straightforward for product-moment correlations, but can also be done for other effect size types [@10.1177/2515245919885611].
 A corrected correlation should also have its sampling variance corrected.
-This is inflated by a factor equal to the square of the reliabilities, such that if $a = \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ then the corrected sampling variance $V_\rho = V_{r} / a^2$ (e.g., Borenstein, 2009, p.
-343).
-Reliability is usually assumed to be known (but see Charles, 2005; Raju et al., 1991), Imperfect reliability in primary studies can both increase and suppress true heterogeneity.
+This is inflated by a factor equal to the square of the reliabilities, such that if $a = \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ then the corrected sampling variance $V_\rho = V_{r} / a^2$ [e.g., @borenstein2009a, p.
+343].
+Reliability is usually assumed to be known [but see @10.1037/1082-989X.10.2.206; @10.1037/0021-9010.76.3.432].
+Imperfect reliability in primary studies can both increase and suppress true heterogeneity.
 
@@
 Hence, imperfect reliability in studies can be expected to suppress heterogeneity estimates whenever true heterogeneity is present.
-Consider an average reliability of 0.8 (the median reported reliability in psychology; Flake et al., 2017).
+Consider an average reliability of 0.8 [the median reported reliability in psychology; @10.1177/1948550617693063].
 Then, $0.8 \times N(\theta, \tau) = N(0.8\theta, 0.8 \tau)$, resulting in 20% less standard deviation among observed effect sizes than in true effect sizes.
@@
 
-It is difficult to predict if the total effect is a negative or positive bias in heterogeneity due to unreliability, and existing literature tends to focus on the inflation of heterogeneity (e.g., Borenstein, 2009, p.
-342; Card, 2015, p.
-126; Hunter & Schmidt, 2015; Wiernick & Dahlke, 2020).
+It is difficult to predict if the total effect is a negative or positive bias in heterogeneity due to unreliability, and existing literature tends to focus on the inflation of heterogeneity [e.g., @borenstein2009a, p.
+342; @card2015, p.
+126; @schmidt2015; @10.1177/2515245919885611].
 However, given the value of accurate heterogeneity estimates for many research outcomes in psychology and that the suppression of heterogeneity could often be 20% or more, we consider it important for meta-analysts to have insight into the bias in heterogeneity estimates that should be expected due to unreliability in a wider range of settings.
@@
 All code and data for this project are available at (OSF LINK).
-For our simulations and analyses we used R (REF) and we used the R package 'metafor' (REF) to perform meta-analyses and estimate heterogeneity.
-For parallel computing we took advantage of the R packages 'parallel' (REF) and 'parabar' (REF).
+For our simulations and analyses we used R `r paste(R.Version()$major, R.Version()$minor, sep = ".")` [@rcoreteam2020] and we used the R package 'metafor' [@10.18637/jss.v036.i03] to perform meta-analyses and estimate heterogeneity.
+For parallel computing we took advantage of the R package 'parabar' [@constantin2023] and for data handling 'data.table' [@dowle2021] and 'ggplot2' [@wickham2016].
 
@@
 We focus on correlations as our effect size type in these simulations because correcting them for unreliability is straightforward and because they are the effect size type most corrected in practice.
-However, unreliability attenuates all types of effect sizes and all can be corrected, either directly (standardized mean differences; Wiernick & Dahlke, 2020) or by first transforming them into correlations.
-Although effect sizes can generally be transformed into each other (e.g., Borenstein, 2009, p.
-46), this should be done with care as it can sometimes affect conclusions (e.g., we found a violation of monotonicity in heterogeneity estimates when converting effect sizes into correlations in a previous project; Olsson-Collentine, 2020, supplement A).
+However, unreliability attenuates all types of effect sizes and all can be corrected, either directly [standardized mean differences; @10.1177/2515245919885611] or by first transforming them into correlations.
+Although effect sizes can generally be transformed into each other [e.g., @borenstein2009a, p.
+46], this should be done with care as it can sometimes affect conclusions [e.g., we found a violation of monotonicity in heterogeneity estimates when converting effect sizes into correlations in a previous project, @10.1037/bul0000294, supplement A].
 
-Perhaps the most common way to meta-analyze correlations is to first transform them into Fisher z correlations (Borenstein et al., 2009, p.
-41 - 43) to normalize their distribution.
+Perhaps the most common way to meta-analyze correlations is to first transform them into Fisher z correlations [@borenstein2009a p.
+41 - 43] to normalize their distribution.
 The Fisher $z$ transformation is the inverse hyperbolic tangent of the product moment correlation and can be computed as $z = 0.5 \times ln\left(\frac{1+r}{1-r}\right)$ and its variance approximated as $V_z = 1 / (n -3)$.
-There is controversy in the use of the Fisher z transformation for meta-analysis (e.g., Hunter & Schmidt, 2015, p.
-220 - 223), largely related to whether Pearson’s r or Fisher’s z leads to less bias in average correlations (Schulze, 2007).
-However, estimating heterogeneity does not seem to have been the focus of most of the literature’s discussion of the appropriateness of the Fisher z transformation (Field, 2005; Hafdahl & Williams, 2009, Hunter & Schmidt, 2015).
-Hall and Brannick (2002) do report heterogeneity estimates but focus on the coverage of prediction intervals (‘credibility intervals’ in Hunter & Schmidt terminology) which combine the heterogeneity estimate and point estimate, and Brannick et al.
-(2019) examined results only after corrections for attenuation.
-Given both this inconclusive evidence and the common use of the Fisher z transformation in practice, we use it as one of the effect size types in our simulations.
+There is controversy in the use of the Fisher z transformation for meta-analysis [e.g., @schmidt2015, p.
+220 - 223], largely related to whether Pearson’s $r$ or Fisher’s $z$ leads to less bias in average correlations (Schulze, 2007).
+However, estimating heterogeneity does not seem to have been the focus of most of the literature’s discussion of the appropriateness of the Fisher z transformation [@10.1037/1082-989X.10.4.444; @10.1037/a0014697; @schmidt2015].
+@10.1037/0021-9010.87.2.377 do report heterogeneity estimates but focus on the coverage of prediction intervals (‘credibility intervals’ in Hunter & Schmidt terminology) which combine the heterogeneity estimate and point estimate, and @10.1177/1094428117741966 examined results only after corrections for attenuation.
+Given both this inconclusive evidence and the common use of the Fisher $z$ transformation in practice, we use it as one of the effect size types in our simulations.
 
-We also use Pearson’s r.
-Pearson’s r has the advantage that it creates heterogeneity estimates on the same scale that are more interpretable than heterogeneity estimates on the Fisher z scale.
-Although Fisher z correlations can be back-transformed into Pearson’s r for interpretation, its heterogeneity cannot.
-A workaround is to compute a prediction interval and convert this interval onto the product-moment scale (Hedges & Vevea, 1998).
+We also use Pearson's $r$.
+Pearson's $r$ has the advantage that it creates heterogeneity estimates on the same scale that are more interpretable than heterogeneity estimates on the Fisher z scale.
+Although Fisher $z$ correlations can be back-transformed into Pearson's $r$ for interpretation, its heterogeneity cannot.
+A workaround is to compute a prediction interval and convert this interval onto the product-moment scale [@10.1037/1082-989X.3.4.486].
 The width of the interval can then be used as an estimate of the heterogeneity.
-We prefer to compute comparable heterogeneity values for Pearson’s r and Fisher’s z (see the methods section ‘Parameter values’) and report the bias in heterogeneity estimates, as we believe this is more interpretable for applied meta-analysts.
+We prefer to compute comparable heterogeneity values for Pearson's $r$ and Fisher's $z$ (see the methods section ‘Parameter values’) and report the bias in heterogeneity estimates, as we believe this is more interpretable for applied meta-analysts.
 
-The disadvantage to using Pearson’s r when estimating heterogeneity are that (1) the effect size is bounded to {-1, 1} and (2) its sampling variance $\sigma^2_r = \frac{(1-\rho^2)^2}{n - 1}$ depends on the true effect size $\rho$.
+The disadvantage to using Pearson's $r$ when estimating heterogeneity are that (1) the effect size is bounded to {-1, 1} and (2) its sampling variance $\sigma^2_r = \frac{(1-\rho^2)^2}{n - 1}$ depends on the true effect size $\rho$.
 The bounded nature of the Pearson correlation means that large levels of heterogeneity create substantial truncation of the distribution of effect sizes if the average effect size is also large.
 That the sampling variance covaries with effect size leads to a small positive bias in heterogeneity estimates (supplement A).
-This can be solved by replacing $\rho$ with the average correlation across studies $\bar{r}$ (Hunter & Schmidt, 2015), or using the Fisher $z$ transformation.
-Fisher z correlations have the additional advantage that they are not bound to {-1, 1}.
-A disadvantage to using Fisher z correlations, in addition to interpretability of heterogeneity estimates, is that they require back-transforming into product-moment correlations before adding measurement error, followed by another transformation to Fisher z for meta-analysis.
-Especially for large Fisher z values (which are common when both average effect size and heterogeneity are large) this transformation process may introduce inaccuracies in the estimates.
-Given the common use of both effect size types and their mix of advantages and disadvantages we report results for both Fisher’s z and Pearson’s r.
+This can be solved by replacing $\rho$ with the average correlation across studies $\bar{r}$ [@schmidt2015], or using the Fisher $z$ transformation.
+Fisher $z$ correlations have the additional advantage that they are not bound to {-1, 1}.
+A disadvantage to using Fisher $z$ correlations, in addition to interpretability of heterogeneity estimates, is that they require back-transforming into product-moment correlations before adding measurement error, followed by another transformation to Fisher z for meta-analysis.
+Especially for large Fisher $z$ values (which are common when both average effect size and heterogeneity are large) this transformation process may introduce inaccuracies in the estimates.
+Given the common use of both effect size types and their mix of advantages and disadvantages we report results for both Fisher's $z$ and Pearson's $r$.
 
-## Meta-analytic model
+## Meta-analytic modelmet0000498
 
-We apply two meta-analytic models to ensure that our results are not model-dependent: the Hedges and Vevea (1998) random-effects model and the Hunter and Schmidt (2015) ‘bare-bones’ random-effects model.
-The Hedges and Vevea (HV) model uses inverse variance weight computed as $w_i = 1 / (V_i + T^2)$ (e.g., Borenstein et al., 2010) where $V_i$ is the sampling variance of study $i$ and $T^2$ is the estimated between-studies variance.
+We apply two meta-analytic models to ensure that our results are not model-dependent: the @10.1037/1082-989X.3.4.486 random-effects model and the @schmidt2015 ‘bare-bones’ random-effects model.
+The Hedges and Vevea (HV) model uses inverse variance weight computed as $w_i = 1 / (V_i + T^2)$ [e.g., @10.1002/jrsm.12] where $V_i$ is the sampling variance of study $i$ and $T^2$ is the estimated between-studies variance.
 The Hunter and Schmidt (HS) model weighs studies by sample size $w_i = N_i$.
-The differences between these weights are relatively minor (Borenstein et al., 2010; Hunter and Schmidt, 2015, p.
-423-426).
-Although the HV model is typically applied to Fisher z transformed correlations (e.g., Borenstein et al., p.
-41-43), we apply it to both transformed and untransformed correlations to enable a more direct comparison with the HS model and facilitate interpretation.
+The differences between these weights are relatively minor [@10.1002/jrsm.12; @schmidt2015, p.
+423-426].
+Although the HV model is typically applied to Fisher $z$ transformed correlations [e.g., @borenstein2009a, p.
+41-43], we apply it to both transformed and untransformed correlations to enable a more direct comparison with the HS model and facilitate interpretation.
 The HS model typically includes corrections for unreliability or other measurement artifacts.
-However, we implement it without corrections (‘bare-bones’; Hunter & Schmidt, 2015) because this allows us to focus on the bias in heterogeneity estimates rather than the performance of corrections (e.g., see Zhang, 2022 for how well they perform in terms of mean estimates).
+However, we implement it without corrections [‘bare-bones’, @schmidt2015] because this allows us to focus on the bias in heterogeneity estimates rather than the performance of corrections [e.g., see @10.1037/met0000498 for how well they perform in terms of mean estimates].
+
+Estimating heterogeneity in both the HV and bare-bones HS model essentially consists of comparing the observed variance in effect sizes with what would be expected from sampling error alone and then ascribing any excess observed variance to heterogeneity [@10.1002/jrsm.12, Box 1; @schmidt2015, p.
+100].
+However, the process for doing so differs somewhat between methods.
+We apply the Restricted Maximum Likelihood (REML) estimator of heterogeneity when estimating heterogeneity in the Hedges and Vevea (1998) model, because it is the generally recommended heterogeneity estimator [@10.1002/jrsm.1164; @10.3102/10769986030003261].
+@10.1002/jrsm.1164 describe the REML estimator in detail.
+The Hunter and Schmidt (2015) model estimates heterogeneity as $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $\frac{\Sigma [N_i \sigma^2_{ei}]}{\Sigma N_i}$ [@schmidt2015, p.
+100] and $\sigma^2_{ei}$ is the usual sampling variance for the Pearson correlation but with $\rho$ replaced with the sample size weighted average correlation $\bar{r} = \frac{\Sigma n_i r_i}{\Sigma n_i}$ instead of $r_i$ as is common.
+A generalized version of the HS estimator for effect sizes other than correlations was developed by Viechtbauer and colleagues [@10.3102/10769986030003261; @10.1037/met0000023].
+
+## Simulation study design
+
+Figure 1 provides an overview of the design of this simulation study.
+Our design was broadly the same for product-moment correlations and Fisher's $z$, with some minor differences we highlight in the text below detailing the procedure.
 
@@
 The flowchart shows the procedures followed to meta-anlyze results using the Hedge’s and Vevea model (see ‘Meta-Analytic Model’ header in methods section).
-The Hunter & Schmidt ‘bare-bones’ meta-analytic model followed the Pearson correlations procedure, but used sample size weights and replaced $r_i$ in the sampling variance by r for meta-anlysis."
+The Hunter & Schmidt ‘bare-bones’ meta-analytic model followed the Pearson correlations procedure, but used sample size weights and replaced $r_i$ in the sampling variance by $\\bar{r}$ for meta-anlysis."
 
@@
 ```
-
-Estimating heterogeneity in both the HV and bare-bones HS model essentially consists of comparing the observed variance in effect sizes with what would be expected from sampling error alone and then ascribing any excess observed variance to heterogeneity (Borenstein et al., 2010, Box 1; Hunter & Schmidt, 2015, p.
-100).
-However, the process for doing so differs somewhat between methods.
-We apply the Restricted Maximum Likelihood (REML) estimator of heterogeneity when estimating heterogeneity in the Hedges and Vevea (1998) model, because it is the generally recommended heterogeneity estimator (Veroniki et al., 2016; Viechtbauer, 2005).
-Veroniki et al.
-(2016) describe the REML estimator in detail.
-The Hunter and Schmidt (2015) model estimates heterogeneity as $\tau^2 = \sigma^2_r - \sigma^2_e$ where  $\sigma^2_r$ is the observed variance across correlations and $\sigma^2_e$ is the average sampling error variance across studies, computed as $\frac{\Sigma [N_i \sigma^2_{ei}]}{\Sigma N_i}$ (Hunter & Schmidt, 2015, p.
-100) and $\sigma^2_{ei}$ is the usual sampling variance for the Pearson correlation but with $\rho$ replaced with the sample size weighted average correlation $\bar{r} = \frac{\Sigma n_i r_i}{\Sigma n_i}$ instead of $r_i$ as is common.
-A generalized version of the HS estimator for effect sizes other than correlations can be found in Viechtbauer (2005; 2015).
-
-## Simulation study design
-
-Figure 1 provides an overview of the design of this simulation study.
-Our design was broadly the same for product-moment correlations and Fisher's $z$, with some minor differences we highlight in the text below detailing the procedure.
 
@@
 We implemented the effect size truncation in R using inverse transform sampling (link: functions.r).
-For Fisher’ z no truncation was applied.
+For Fisher’s $z$ no truncation was applied.
 Throughout this text we will refer to three types of truncation: (1) effect size truncation (truncation of Pearson’s correlations between {-1, 1}), (2), reliability truncation (truncation of reliability between {0, 1}), and (3) heterogeneity truncation (truncation of heterogeneity estimates to be non-negative).
@@
 Within meta-analyses we used a fixed sample size across all studies such that N=N_i.
-We made this choice rather than treating sample size as a random variable (as done by e.g., Field, 2005; Brannick et al., 2019) to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance.
+We made this choice rather than treating sample size as a random variable [as done by e.g., @10.1037/1082-989X.10.4.444; @10.1177/1094428117741966] to be able to observe the effect of changes in sample size on estimates, and because we are focused on the average heterogeneity estimates rather than its sampling variance.
 We base our range of sample sizes on empirical estimates of typical sample size in correlational research in psychology.
-Fraley et al.
-(2022) report median sample sizes for between-person studies in nine psychology journals between 2011 - 2018 which ranged between 69 - 496 depending on journal and year.
-Sassenber and Ditrich (2019) find a median sample size of 110 (interquartile range, 71 - 195) across four journals and years (2009, 2011, 2016, and 2018) in social psychology, and Bakker et al.
+@10.1177/25152459221120217 report median sample sizes for between-person studies in nine psychology journals between 2011 - 2018 which ranged between 69 - 496 depending on journal and year.
+@10.1177/2515245919838781a find a median sample size of 110 (interquartile range, 71 - 195) across four journals and years (2009, 2011, 2016, and 2018) in social psychology, and Bakker et al.
 (preliminary results; https://osf.io/zsjf4) find median sample sizes between 24 - 184 for six journals in psychology for the years 1995, 2006, 2019.
@@
 When focusing on the bias of heterogeneity estimate, the number of meta-analyzed studies, K, would not matter for an unbiased estimator.
-However, both the heterogeneity estimator we used, restricted maximum likelihood, and most other heterogeneity estimators are truncated at zero (Viechtbauer, 2005).
+However, both the heterogeneity estimator we used, restricted maximum likelihood, and most other heterogeneity estimators are truncated at zero [@10.3102/10769986030003261].
 Lower values for K then result in a larger proportion of truncated variance estimates below zero because of increased sampling variance in $\tau^2$.
@@
 We vary the grand mean $\theta$ from 0 to 0.6 in steps of 0.1 to cover all realistic effect sizes and explore the boundaries of the interaction between effect size and reliability variance.
-For Fisher’s z, we transformed these values to the $z$ scale.
-For context to these values, we considered the empirical estimates of typical correlational effect sizes provided by Nuijten et al.
-(2020) and Schäfer & Schwarz (2019).
-Nuijten et al.
-performed a meta-meta-analysis on the fields of intelligence research and reported a median meta-analytic correlation of 0.24 across 779 predictive validity and correlational studies.
+For Fisher's $z$, we transformed these values to the $z$ scale.
+For context to these values, we considered the empirical estimates of typical correlational effect sizes provided by @10.3390/jintelligence8040036 and @10.3389/fpsyg.2019.00813.
+@10.3390/jintelligence8040036 performed a meta-meta-analysis on the fields of intelligence research and reported a median meta-analytic correlation of 0.24 across 779 predictive validity and correlational studies.
 This estimate is likely positively biased due to publication bias and selective reporting.
-Schäfer & Scharz reported a median $r$ of 0.16 amongst preregistered research in psychology in general, with an ‘upper median’ (i.e., the 83.35% quantile) of 0.41.
-This estimate is likely less affected by positive bias but should not be expected to be unbiased as there appears to be some positive bias also in preregistered research on average (Scheel, 2021).
+@10.3389/fpsyg.2019.00813 reported a median $r$ of 0.16 amongst preregistered research in psychology in general, with an ‘upper median’ (i.e., the 83.35% quantile) of 0.41.
+This estimate is likely less affected by positive bias but should not be expected to be unbiased as there appears to be some positive bias also in preregistered research on average @10.1177/25152459211007467.
 
 We defined the between-studies standard deviation ($\tau$) to cover a wide range of variability in true effect sizes and empirical estimates reported in the psychological literature.
-Van Erp et al.
-(2017) provided empirical heterogeneity estimates from 747 meta-analyses in 61 articles published in the journal Psychological Bulletin between 1990 - 2013.
-The median (uncorrected) Pearson’s r $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24).
+@10.5334/jopd.33 provided empirical heterogeneity estimates from 747 meta-analyses in 61 articles published in the journal Psychological Bulletin between 1990 - 2013.
+The median (uncorrected) Pearson's $r$ $\tau$ value in these data was 0.17 (interquartile range 0.1 - 0.24).
 The point of our study is that empirical heterogeneity estimates may not be accurately estimated.
-Nonetheless, we selected approximately the interquartile range of $\tau$-values reported in van Erp et al.
-for the product-moment correlation, both to make sure we covered the empirically reported range of values and because we considered them to represent reasonable variability on the product-moment correlation scale.
+Nonetheless, we selected approximately the interquartile range of $\tau$-values reported in @10.5334/jopd.33 for the product-moment correlation, both to make sure we covered the empirically reported range of values and because we considered them to represent reasonable variability on the product-moment correlation scale.
 The values we selected were {0, 0.1, 0.15, 0.2}.
 We selected the largest heterogeneity level to avoid excessive effect size truncation for our maximum average effect size of 0.6: $\theta_i \sim N(\theta = 0.6, \tau = 0.2)$ implies approximately 95% of effect sizes will be below 1.
-Our values are similar to those used by Brannick et al.
-(2019) in their simulation study on the performance of different heterogeneity estimators after corrections for unreliability (they used $\tau =$  0, 0.08, 0.13, 0.2).
+Our values are similar to those used by @10.1177/1094428117741966 in their simulation study on the performance of different heterogeneity estimators after corrections for unreliability (they used $\tau =$  0, 0.08, 0.13, 0.2).
 
-Fisher’s z is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson’s r into corresponding $\tau$-values on the Fisher’s z scale.
-To compare results between Fisher’s z and the product-moment correlation we defined heterogeneity on the $I^2$ scale.
-The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity (Higgins, 2002; 2003), and as such is measured from 0 - 100%.
+Fisher's $z$ is measured at a different scale and it is not possible to directly convert $\tau$-values expressed in Pearson's $r$ into corresponding $\tau$-values on the Fisher's $z$ scale.
+To compare results between Fisher's $z$ and the product-moment correlation we defined heterogeneity on the $I^2$ scale.
+The $I^2$ index is a relative measure of heterogeneity describing the percentage of total variance that is due to heterogeneity [@10.1002/sim.1186; @10.1136/bmj.327.7414.557], and as such is measured from 0 - 100%.
 The $I^2$ index can be defined as $I^2 = \frac{\tau^2}{s^2 + \tau^2}$ where $s^2 = \frac{\Sigma w_i(k - 1)}{(\Sigma w_i)^2 - \Sigma w_i^2}$ and $w_i$ is the precision $w_i = 1 / \sigma^2$.
-This means that the $I^2$ value generally depends on both the number of studies $K$ and the sample size $N_i$ in these studies (Borenstein, 2017).
+This means that the $I^2$ value generally depends on both the number of studies $K$ and the sample size $N_i$ in these studies [@10.1002/jrsm.1230].
 However, because we keep $N_i$ fixed across studies, the value of $I^2$ depends only on $N = N_i$ and the heterogeneity.
@@
 Finally, we defined average reliability levels and their standard deviation based on empirical estimates from the literature.
-Flake et al.
-(2017) documented 245 estimates of Cronbach’s Alpha in psychology and found an average of 0.79 and a standard deviation of 0.13.
+@10.1177/1948550617693063 documented 245 estimates of Cronbach’s Alpha in psychology and found an average of 0.79 and a standard deviation of 0.13.
 The interquartile range was approximately 0.68 – 0.87 for studies using (_ad hoc_) scales that lacked a reference to an earlier source and 0.79 – 0.88 for scales that did have a reference.
-Sánchez-Meca et al.
-(2003) reported the reliability estimates based on five so-called reliability generalization studies.
+@10.1111/j.2044-8317.2012.02057.xa reported the reliability estimates based on five so-called reliability generalization studies.
 These five studies summarized reliability in 25 – 51 primary studies (184 total), and the mean reliability ranged from 0.767 to 0.891 with standard deviations ranging between 0.034 – 0.133.
-Given some likely positive bias in such empirical values caused by selective reporting (Hussey et al., 2023), we examined the following mean reliabilities {0.6, 0.7, 0.8, 0.9}, and standard deviations {0, 0.05, 0.1, 0.15}.
+Given some likely positive bias in such empirical values caused by selective reporting [@10.31234/osf.io/dm8xn], we examined the following mean reliabilities {0.6, 0.7, 0.8, 0.9}, and standard deviations {0, 0.05, 0.1, 0.15}.
 
@@
 Figure 2 shows the net bias in heterogeneity estimates for different levels of heterogeneity (measured in $\tau$) and meta-analytic models.
-Each column of Figure 2 corresponds to one nominal level of actual heterogeneity ($\tau$ $\epsilon$ {0, 0.1, 0.15, 0.2}) and each row to a different analysis (HV with Fisher z, HV with Pearson’s r, and HS).
+Each column of Figure 2 corresponds to one nominal level of actual heterogeneity ($\tau$ $\epsilon$ {0, 0.1, 0.15, 0.2}) and each row to a different analysis (HV with Fisher z, HV with Pearson's $r$, and HS).
 The black solid lines indicate actual heterogeneity, which may differ from nominal heterogeneity because of effect size truncation or, in the case of Fisher z, because the comparable heterogeneity levels were slightly different (as discussed in the method section).
@@
 As expected, for zero heterogeneity (leftmost column Figure 2) we see an overestimate of heterogeneity that increases with average effect size.
-Generally, there are two sources that explain the positive bias in the absence of true heterogeneity, although the use of Pearson’s r as the type of effect size additionally inflates heterogeneity estimates to a small degree for the HV model (row 2, Figure 2) due to the dependence between effect size and sampling variance (supplement A).
-First, and the only of these two upwardly biasing sources that applies when $\mu = 0$, the heterogeneity estimate is inflated due to the truncation of negative heterogeneity estimates (Viechtbauer, 2005).
+Generally, there are two sources that explain the positive bias in the absence of true heterogeneity, although the use of Pearson's $r$ as the type of effect size additionally inflates heterogeneity estimates to a small degree for the HV model (row 2, Figure 2) due to the dependence between effect size and sampling variance (supplement A).
+First, and the only of these two upwardly biasing sources that applies when $\mu = 0$, the heterogeneity estimate is inflated due to the truncation of negative heterogeneity estimates [@10.3102/10769986030003261].
 Second, there is inflation due to the variance in reliability across studies in the fixed effect scenario.
-For example, for $\mu = 0.2$ (the closest value to the median correlation reported by Schäfer & Schwarz, 2019) and $\bar{R} = 0.8$ the bias for Pearson's $r$ compared to true zero heterogeneity is `r in_text$bias_tau_zero$mu_0.2`.
+For example, for $\mu = 0.2$ [the closest value to the median correlation reported by @10.3389/fpsyg.2019.00813] and $\bar{R} = 0.8$ the bias for Pearson's $r$ compared to true zero heterogeneity is `r in_text$bias_tau_zero$mu_0.2`.
 For $\mu = 0.4$ (the upper median reported by Schäfer & Schwarz) and $\bar{R} = 0.8$ the bias is `r in_text$bias_tau_zero$mu_0.4`.
@@
 We used the following values {0.02, 0.04, 0.06, 0.08}.
-Figure 3 presents the results of this analysis for Pearson’s r.
+Figure 3 presents the results of this analysis for Pearson's $r$.
 Because these lower heterogeneity levels can be expected to be affected by heterogeneity truncation to a larger extent (which results in inflation), we present results for four different numbers of studies in the meta-analysis (rows Figure 3, lower k is associated with larger inflation of estimates).
@@
 Figure 3 shows that even when only little heterogeneity is present, it will generally be underestimated.
-If μ < 0.3 then imperfect reliability leads to a negative bias in heterogeneity estimates for all $\tau$ < 0.06 in Pearson’s r.
+If μ < 0.3 then imperfect reliability leads to a negative bias in heterogeneity estimates for all $\tau$ < 0.06 in Pearson's $r$.
 Even for the lower levels of true heterogeneity, estimates of heterogeneity will generally be positively biased only when average effect size is large, or the number of meta-analyzed studies is small.
@@
 As expected, heterogeneity estimates generally improved with higher average reliability, and deteriorated with smaller average study sample size.
-Our results were approximately the same whether one used a classic random-effects meta-analytic model (Hedges & Vevea, 1998) or a ‘bare-bones’ psychometric meta-analytic model (Hunter & Schmidt, 2015), and whether one used Pearson’s r as the effect size or Fisher’s z, although the negative bias was slightly larger with Fisher’s z.
+Our results were approximately the same whether one used a classic random-effects meta-analytic model [@10.1037/1082-989X.3.4.486] or a ‘bare-bones’ psychometric meta-analytic model [@schmidt2015], and whether one used Pearson's $r$ as the effect size or Fisher's $z$, although the negative bias was slightly larger with Fisher's $z$.
 The severity of the underestimation of heterogeneity could lead to meta-analyses in psychology missing relevant moderators and overgeneralizing outcomes.
@@
 In meta-analyses, low reliabilities not only lower the average effect size estimate, but also artificially suppress heterogeneity estimates in most cases.
-For a typical effect size around r = 0.2 and substantial true heterogeneity ($\tau$>0.1) that is not uncommon in the meta-analytic literature (van Erp et al., 2017), we found expected heterogeneity to be underestimated by about 15% - 60% in our simulations primarily depending on average reliability in primary studies (higher reliability leads to a smaller underestimate), although the underestimate could be as high as 73% with low sample sizes.
+For a typical effect size around r = 0.2 and substantial true heterogeneity ($\tau$ > 0.1) that is not uncommon in the meta-analytic literature [@10.5334/jopd.33], we found expected heterogeneity to be underestimated by about 15% - 60% in our simulations primarily depending on average reliability in primary studies (higher reliability leads to a smaller underestimate), although the underestimate could be as high as 73% with low sample sizes.
 As true heterogeneity increases, the proportional underestimate decreases to some degree, but the underestimate in absolute terms will worsen.
@@
 
-If true heterogeneity is zero, there is a positive bias in observed heterogeneity estimates, but for even small degrees of true heterogeneity there tends to be a negative bias in observed estimates.
-As such, it is not possible to know whether a small degree of observed heterogeneity is due to overestimated true zero heterogeneity or underestimated small heterogeneity.
-However, for observed heterogeneity above $\tau > 0.025$ one can usually be confident that the estimate is an underestimate of true heterogeneity rather than an overestimate of zero heterogeneity if the average effect sizes is below approximately $r = 0.2$.
+Without any information on primary studies’ reliabilities, it would often be impossible to know whether a small observed heterogeneity is due to overestimated true zero heterogeneity or underestimated small heterogeneity.
+Moreover, it is certainly possible to observe a zero heterogeneity estimate under larger true heterogeneity levels, because the sampling variability in heterogeneity estimates is often substantial in practice [@10.1136/bmj.39343.408449.80; @10.1037/bul0000294].
+Generally, distinguishing between zero and small heterogeneity in empirical meta-analyses is statistically challenging considering the typical sample sizes and numbers of studies in meta-analyses in psychology [@10.1037/bul0000294].
+The power to detect small true heterogeneity in most psychological meta-analyses is low.
+In our view, the distinction between zero and small heterogeneity in meta-analysis is primarily a hypothetical one as true zero heterogeneity is not expected in meta-analytic practice to begin with [see also @10.1037/0003-066X.42.5.443], except perhaps in specific sets of very similar direct replications [@10.1037/bul0000294].
 
-Without any information on primary studies’ reliabilities, it would often be impossible to know whether a small observed heterogeneity is due to overestimated true zero heterogeneity or underestimated small heterogeneity.
-Moreover, it is certainly possible to observe a zero heterogeneity estimate under larger true heterogeneity levels, because the sampling variability in heterogeneity estimates is often substantial in practice (Ioannidis et al., 2007; Olsson-Collentine et al., 2020).
-Generally, distinguishing between zero and small heterogeneity in empirical meta-analyses is statistically challenging considering the typical sample sizes and numbers of studies in meta-analyses in psychology (Olsson-Collentine et al., 2020).
-The power to detect small true heterogeneity in most psychological meta-analyses is low.
-In our view, the distinction between zero and small heterogeneity in meta-analysis is primarily a hypothetical one as true zero heterogeneity is not expected in meta-analytic practice to begin with (see also Hedges, 1987), except perhaps in specific sets of very similar direct replications (Olsson Collentine et al., 2020).
-
-We found that using Pearson’s r or Fisher’s z as our effect size resulted in very similar degrees of bias in heterogeneity estimates due to uncorrected unreliability in primary studies.
-The underestimate is slightly worse for Fisher’s z than for Pearson’s r.
-The usage of Pearson’s r inflates heterogeneity estimates slightly (due to its bounded nature and effect size dependent sampling variance), which is apparently beneficial for heterogeneity estimates when not correcting for unreliability.
-However, the additional negative bias when using Fisher’s z is small and we consider the difference negligible in the light of the severe underestimates generated by unreliability.
+We found that using Pearson's $r$ or Fisher's $z$ as our effect size resulted in very similar degrees of bias in heterogeneity estimates due to uncorrected unreliability in primary studies.
+The underestimate is slightly worse for Fisher's $z$ than for Pearson's $r$.
+The usage of Pearson's $r$ inflates heterogeneity estimates slightly (due to its bounded nature and effect size dependent sampling variance), which is apparently beneficial for heterogeneity estimates when not correcting for unreliability.
+However, the additional negative bias when using Fisher's $z$ is small and we consider the difference negligible in the light of the severe underestimates generated by unreliability.
 
@@
 Our simulations show that heterogeneity estimates are biased when unreliability is ignored, so in principle one should correct for unreliability.
-However, even done perfectly this would not mean that heterogeneity estimates end up to be accurate, because few meta-analyses are able to include sufficient studies (van Erp et al., 2017; Davey et al., 2011) to measure heterogeneity without large uncertainty and wide confidence intervals (Ioannidis et al., 2007, Olsson-Collentine, 2020).
+However, even done perfectly this would not mean that heterogeneity estimates end up to be accurate, because few meta-analyses are able to include sufficient studies [@10.5334/jopd.33; @10.1186/1471-2288-11-160] to measure heterogeneity without large uncertainty and wide confidence intervals [@10.1136/bmj.39343.408449.80; @10.1037/bul0000294].
 In practice, attempting to correct for unreliability comes with many caveats and assumptions that must be considered.
@@
 First, reliability estimates are lower bound estimates, meaning that systematically correcting for unreliability can be expected to lead to overestimated mean effect sizes and subsequent overestimated heterogeneity.
-Unfortunately, the most popular reliability estimate (Cronbach’s Alpha; Flake et al., 2017; Hogan et al., 2002) is known to underestimate the lower bound even further than several alternatives, particularly when the strong assumptions underlying CTT are not met (Sijtsma, 2008).
-Second, CTT corrections of correlations as commonly applied (Hunter & Schmidt, 2015; but see Raju et al., 1991; Charles, 2005) do not consider that reliability estimates themselves are subject to substantial sampling error (Fan & Thompson, 2001) that add additional uncertainty to corrected effect sizes and subsequent estimates of heterogeneity.
-Third, there are several types of reliability estimates (e.g., "internal consistency”, test-retest, coefficient of equivalence) that do not necessarily derive from the same type of measurement error and may result in different reliability estimates (Dimitrov, 2002; Revelle & Condon, 2019).
+Unfortunately, the most popular reliability estimate [Cronbach’s Alpha, @10.1177/1948550617693063; @10.1177/00131640021970691] is known to underestimate the lower bound even further than several alternatives, particularly when the strong assumptions underlying CTT are not met [@10.1007/s11336-008-9101-0].
+Second, CTT corrections of correlations as commonly applied [@schmidt2015; but see @10.1037/0021-9010.76.3.432; @10.1037/1082-989X.10.2.206] do not consider that reliability estimates themselves are subject to substantial sampling error [@10.1177/00131640121971365] that add additional uncertainty to corrected effect sizes and subsequent estimates of heterogeneity.
+Third, there are several types of reliability estimates (e.g., "internal consistency”, test-retest, coefficient of equivalence) that do not necessarily derive from the same type of measurement error and may result in different reliability estimates [@10.1177/001316402236878; @10.1037/pas0000754].
 If the reliability estimate applied does not account for all relevant sources of measurement error, it will overestimate reliability and heterogeneity will remain underestimated despite unreliability corrections.
-Fourth, reliability estimates are sometimes not reported in primary studies (Flake et al., 2017).
-For meta-analysis, missing estimates are then often imputed based on average observed reliability or based on some ‘typical’ reliability (Hunter & Schmidt, 2015; Gnambs & Sengewald, 2023).
-However, what a ‘typical’ reliability should be is open to debate (LeBreton et al., 2014), and mean imputation leads to a biased pooled estimate, especially so if missing statistics are related to the value of the statistic (and it seems likely that non-reporting is more common when reliability is low).
-Mean imputation is seen by experts in missing data as “possibly the worst missing data handling method available” (Enders, 2010, p.
-43; see also van Buuren, 2018, p.
-12; Schafer, 2002, p.
-159).
+Fourth, reliability estimates are sometimes not reported in primary studies [@10.1177/1948550617693063].
+For meta-analysis, missing estimates are then often imputed based on average observed reliability or based on some ‘typical’ reliability [@schmidt2015; @10.1027/2151-2604/a000511].
+However, what a ‘typical’ reliability should be is open to debate [@10.1111/iops.12184], and mean imputation leads to a biased pooled estimate, especially so if missing statistics are related to the value of the statistic (and it seems likely that non-reporting is more common when reliability is low).
+Mean imputation is seen by experts in missing data as “possibly the worst missing data handling method available” [@enders2010, p.
+43; see also @10.1201/9780429492259, p.
+12; @10.1037/1082-989X.7.2.147, p.
+159].
 
 Correcting for unreliability in the CTT framework comes with several assumptions that may be hard to motivate in practice.
-First, most estimates of reliability in CTT assume unidimensionality (e.g., Borsboom & Mellenberh, 2002; Savalei & Reise, 2019), and violations of this assumption can severely bias reliability estimates (Savalei & Reise, 2019).
-Second, correcting for unreliability in the CTT framework assumes linear associations between scale and latent construct, which implies that corrections may result in biased estimates when ordinal scales (such as Likert-scales) are short (Zheng, 2022).
-Based on their simulations, Zheng (2022) recommends that ordinal scales should have at least 12 items (with $\geq$ 5 answer categories per item) to use Cronbach’s Alpha to correct for unreliability in meta-analysis.
-Problematically, Flake et al.
-(2017) report that the average scale length in psychology (excluding 1-item scales) is 4.69 items (SD = 6.35).
+First, most estimates of reliability in CTT assume unidimensionality [e.g., @10.1016/S0160-28960200082-X; @10.1525/collabra.247], and violations of this assumption can severely bias reliability estimates [@10.1525/collabra.247].
+Second, correcting for unreliability in the CTT framework assumes linear associations between scale and latent construct, which implies that corrections may result in biased estimates when ordinal scales (such as Likert-scales) are short [@10.1037/met0000498].
+Based on their simulations, @10.1037/met0000498 recommends that ordinal scales should have at least 12 items (with $\geq$ 5 answer categories per item) to use Cronbach’s Alpha to correct for unreliability in meta-analysis.
+Problematically, @10.1177/1948550617693063 report that the average scale length in psychology (excluding 1-item scales) is 4.69 items (SD = 6.35).
 Third, true scores $X$ and $Y$ are assumed to be normally distributed.
-Zheng (2022) demonstrated in simulation studies that when this assumption does not hold average effect size estimates will be inaccurate, especially when the average effect size is large (although they did not examine heterogeneity).
-Fourth, estimators such as Cronbach’s Alpha assume uncorrelated errors (e.g., Dimitrov, 2002; Teo & Fan, 2013) and essential tau-equivalence (i.e., items have equal factor loadings on a latent construct; e.g., Revelle & Condon, 2019).
-Correlation of errors can lead to a substantial overestimate of reliability (e.g., Dimitrov, 2002; Teo & Fan, 2013), and may occur in many measurement situations, such as when items on a scale refer to the same stimuli (e.g., multiple reading comprehension items referring to the same paragraph; Green & Yang, 2009).
+@10.1037/met0000498 demonstrated in simulation studies that when this assumption does not hold average effect size estimates will be inaccurate, especially when the average effect size is large (although they did not examine heterogeneity).
+Fourth, estimators such as Cronbach’s Alpha assume uncorrelated errors [e.g., @10.1177/001316402236878; @10.1007/s40299-013-0075-z] and essential tau-equivalence [i.e., items have equal factor loadings on a latent construct, e.g., @10.1037/pas0000754].
+Correlation of errors can lead to a substantial overestimate of reliability [e.g., @10.1177/001316402236878; @10.1007/s40299-013-0075-z], and may occur in many measurement situations, such as when items on a scale refer to the same stimuli [e.g., multiple reading comprehension items referring to the same paragraph, @10.1007/s11336-008-9098-4].
 Overestimation of reliability would leave heterogeneity estimates with a negative bias remaining even after CTT correction.
-The assumption of essential tau-equivalence is empirically unrealistic, and violations lead to an underestimate of the reliability statistic (Sijtsma, 2009, p.
-111).
-This bias may often be small enough to be neglected (Green & Yang, 2009; Savale & Reise, 2019), but can interact with the assumption of normal distributed true scores and result in inaccurate estimates of average effect sizes even for ordinal scales with more than 12 items (Zheng, 2022).
-Extending the study of Zheng (2022) also to examine the effects on heterogeneity estimates would be a useful avenue of future research.
+The assumption of essential tau-equivalence is empirically unrealistic, and violations lead to an underestimate of the reliability statistic [@10.1007/s11336-008-9101-0, p.
+111].
+This bias may often be small enough to be neglected [@10.1007/s11336-008-9098-4; @10.1525/collabra.247], but can interact with the assumption of normal distributed true scores and result in inaccurate estimates of average effect sizes even for ordinal scales with more than 12 items [@10.1037/met0000498].
+Extending the study of @10.1037/met0000498 also to examine the effects on heterogeneity estimates would be a useful avenue of future research.
 
-When all of the CTT assumptions hold, and if practical concerns can be overcome, CTT corrections of unreliability appear to function well in meta-analysis of correlations (e.g., Hall & Brannick, 2002; Brannick et al., 2019; Hunter & Schmidt, 2015).
-However, research appears to be lacking on the consequences of CTT corrections for meta-analysis when assumptions do not hold (but see Zhang, 2022).
-Moreover, because latent constructs are defined by latent variable models (Borsboom & Mellenbergh, 2002), the strong assumptions of CTT regarding latent associations, factor loadings, uncorrelated errors, and the distribution of true scores may be unpalatable to many researchers.
-Modern latent models based on Structural Equation Modelling (SEM; e.g., Lei & Wu, 2007) may then be an option, as they permit delving into measurements and latent construct structures rather than relying on assumptions.
+When all of the CTT assumptions hold, and if practical concerns can be overcome, CTT corrections of unreliability appear to function well in meta-analysis of correlations [e.g., @10.1037/0021-9010.87.2.377; @10.1177/1094428117741966; @schmidt2015].
+However, research appears to be lacking on the consequences of CTT corrections for meta-analysis when assumptions do not hold [but see @10.1037/met0000498].
+Moreover, because latent constructs are defined by latent variable models [@10.1016/S0160-28960200082-X], the strong assumptions of CTT regarding latent associations, factor loadings, uncorrelated errors, and the distribution of true scores may be unpalatable to many researchers.
+Modern latent models based on Structural Equation Modelling [SEM, e.g., @10.1111/j.1745-3992.2007.00099.x] may then be an option, as they permit delving into measurements and latent construct structures rather than relying on assumptions.
 
-Meta-analysis can be conceptualized as a SEM model (e.g., Cheung, 2015), which holds promise that the benefits of modern psychometric theory might be applied also to latent constructs summarized across multiple studies.
-In Meta-Analytic Structural Equation Modeling (MASEM; e.g., Cheung & Chan, 2005) unreliability in primary studies can be corrected in three ways (Gnambs & Sengewald, 2023): (1) by estimating a SEM model in each primary study and then pooling estimates (parameter-based MASEM: Cheung, 2015, p.
-241), (2) by directly modeling latent variables in the MASEM,  or (3) by first correcting correlations before introducing them into the MASEM framework (i.e., fundamentally identical to the CTT approach).
+Meta-analysis can be conceptualized as a SEM model [e.g., @cheung2015], which holds promise that the benefits of modern psychometric theory might be applied also to latent constructs summarized across multiple studies.
+In Meta-Analytic Structural Equation Modeling [MASEM, e.g., @10.1037/1082-989X.10.1.40] unreliability in primary studies can be corrected in three ways [@10.1027/2151-2604/a000511]: (1) by estimating a SEM model in each primary study and then pooling estimates [parameter-based MASEM, @cheung2015, p.
+241], (2) by directly modeling latent variables in the MASEM,  or (3) by first correcting correlations before introducing them into the MASEM framework (i.e., fundamentally identical to the CTT approach).
 The first option is limited because it is rarely the case that the same SEM model can be fit across multiple independent datasets, except in highly structured cases (e.g., Brunner et al., 2022).
@@
 As such, the potential to use MASEM to take advantage of modern psychometric theory when correcting for unreliability is challenging in practice.
-That said, multi-lab replication projects are increasingly common in psychology (e.g., Klein et al., 2014, McCarthy et al., 2021) and consist of a set of direct replication studies where analysts have access to all data.
+That said, multi-lab replication projects are increasingly common in psychology [e.g., @10.1027/1864-9335/a000178; @10.1525/collabra.18738] and consist of a set of direct replication studies where analysts have access to all data.
 As such, they are well-placed to take full advantage of MASEM for their analyses.
@@
 Arguably the biggest challenge in estimating heterogeneity in meta- practice is the large degree of sampling variance in heterogeneity estimates that can be expected in almost any meta-analysis (Ioannidis et al., 2007; Olsson-Collentine et al., 2020).
-Consider a meta-analysis without measurement error of k =12 studies (the median number in psychology; van Erp et al., 2017) with N = 150 in all studies, $\tau = 0.17$ (the median observed $\tau$ for correlations in van Erp et al., 2017), and  $\mu = 0$.
+Consider a meta-analysis without measurement error of k =12 studies [the median number in psychology, @10.5334/jopd.33] with N = 150 in all studies, $\tau = 0.17$ [the median observed $\tau$ for correlations in @10.5334/jopd.33], and  $\mu = 0$.
 Across 10,000 replications the median width of the 95% confidence interval around the heterogeneity estimate (using default settings in metafor) is then 0.195 in Pearson’s $r$ units.
@@
 This is because although we apply different meta-analytic models and effect sizes, the purpose was not to compare the efficiency of these different modes of estimation.
-As the effect of unreliability in primary studies is simply to change the heterogeneity of effect sizes, the results of previous studies on the comparable performance of different heterogeneity estimators applies also here (e.g., Langan et al., 2017; Veroniki et al., 2016; Viechtbauer, 2005).
+As the effect of unreliability in primary studies is simply to change the heterogeneity of effect sizes, the results of previous studies on the comparable performance of different heterogeneity estimators applies also here [e.g., @10.1002/jrsm.1198; @10.1002/jrsm.1164; @10.3102/10769986030003261].
 That said, it would be useful to examine the variance of heterogeneity estimates in future research when the focus is on developing practical guidance for meta-analysts.
 Doing so could be helpful to gain a better idea of how likely a particular observed heterogeneity estimate is to belong to different distributions of true heterogeneity in the presence of measurement unreliability.
-As this would depend on N, K, mean reliability, variability in reliability, average effect size, observed heterogeneity, and the true heterogeneity levels being compared, implementing such an analysis would probably be best done on a case-by-case basis (e.g., through a webapp such as developed for sensitivity analyses of the effect of publication bias on heterogeneity by Augusteijn et al., 2017) .
+As this would depend on N, K, mean reliability, variability in reliability, average effect size, observed heterogeneity, and the true heterogeneity levels being compared, implementing such an analysis would probably be best done on a case-by-case basis [e.g., through a webapp such as developed for sensitivity analyses of the effect of publication bias on heterogeneity by @10.1037/met0000197] .
 Given the challenges in correcting for unreliability, such a sensitivity analysis could be a useful tool to reason around the impact of unreliability for applied meta-analysts.
-An additional challenge when providing practical advice to meta-analysts is how to handle selective reporting of study reliabilities (Hussey et al., 2023) in primary studies, and the problems that p-hacking and publication bias create in estimating heterogeneity.
-We know from Augusteijn et al.
-(2017) that publication bias alone creates complex outcomes, but when meta-analysis is affected by unreliability and p-hacking true heterogeneity is additionally obscured.
+An additional challenge when providing practical advice to meta-analysts is how to handle selective reporting of study reliabilities [@10.31234/osf.io/dm8xn]) in primary studies, and the problems that $p$-hacking and publication bias create in estimating heterogeneity.
+We know from @10.1037/met0000197 that publication bias alone creates complex outcomes, but when meta-analysis is affected by unreliability and $p$-hacking true heterogeneity is additionally obscured.
 
@@
 We showed that increasing the primary study sample size led to better heterogeneity estimates, especially for low reliability.
-There is large variability in how simulation studies focused on heterogeneity implement sample sizes (Langan et al., 2017).
-Most studies sample from a uniform distribution of sample sizes, but there are also studies that have used conditions with fixed sample sizes like us (Langan et al., 2018).
-There appears to be no systematic exploration of the effect of different distributions of sample sizes on heterogeneity estimation, and the effect of sample size is generally considered in view of its average value (e.g., Viechtbauer, 2005).
-The range of sample sizes in a meta-analysis appears to affect the performance of heterogeneity estimation (Langan et al., 2017), but we believe this relates only to the efficiency of heterogeneity estimates rather than bias, given a sufficiently large number of repetitions in the simulations.
+There is large variability in how simulation studies focused on heterogeneity implement sample sizes [@10.1002/jrsm.1198].
+Most studies sample from a uniform distribution of sample sizes, but there are also studies that have used conditions with fixed sample sizes like us [@10.1002/jrsm.1316].
+There appears to be no systematic exploration of the effect of different distributions of sample sizes on heterogeneity estimation, and the effect of sample size is generally considered in view of its average value [e.g., @10.3102/10769986030003261].
+The range of sample sizes in a meta-analysis appears to affect the performance of heterogeneity estimation [@10.1002/jrsm.1198], but we believe this relates only to the efficiency of heterogeneity estimates rather than bias, given a sufficiently large number of repetitions in the simulations.
 That said, implementing a realistic distribution of sample sizes would be important for research attempting to provide concrete guidance to meta-researchers on how to interpret observed heterogeneity in the presence of measurement error as variability in estimates then is important.
-Implementing a distribution of sample sizes based on relevant empirical literature (e.g., as done by Brannick et al., 2019) could be one way to increase realism in simulated sample sizes.
+Implementing a distribution of sample sizes based on relevant empirical literature [e.g., as done by @10.1177/1094428117741966] could be one way to increase realism in simulated sample sizes.
 
-Finally, we only considered bias in heterogeneity due to unreliability in primary studies, but there exist other measurement artifacts such as restriction of range, dichotomization, and more (Hunter & Schmidt, 2015).
+Finally, we only considered bias in heterogeneity due to unreliability in primary studies, but there exist other measurement artifacts such as restriction of range, dichotomization, and more [@schmidt2015].
 Although measurement error is the only measurement artifact that is always present, many of these other measurement artifacts may impact heterogeneity estimates when they are present.
-Based on Hunter & Schmidt’s arguments that measurement artifacts all tend to attenuate effect sizes in a similar way to unreliability (p.
+Based on @schmidt2015 's arguments that measurement artifacts all tend to attenuate effect sizes in a similar way to unreliability (p.
 78-79), we expect that their net effect on heterogeneity estimates may similarly be a negative bias.
-The extent to which this prediction holds true (and whether these measurement artifacts can be corrected for in MASEM; Gnambs & Sengewalds, 2023) awaits further study.
+The extent to which this prediction holds true [and whether these measurement artifacts can be corrected for in MASEM, @10.1027/2151-2604/a000511] awaits further study.
 
@@
 Yet, sophisticated methods for correcting measurement unreliability such as meta-analytic structural equation modeling (MASEM) are only applicable in exceptional cases and corrections based on classical test theory come with caveats and strong assumptions that are often unrealistic and currently do not fully consider insights from modern test theory.
-Accurate estimation of meta-analytic heterogeneity is difficult and will remain so unless measurement concerns (Flake & Fried, 2020; Flake et al., 2017; Kane et al., 2021) are taken seriously in primary research.
+Accurate estimation of meta-analytic heterogeneity is difficult and will remain so unless measurement concerns [@10.1177/2515245920952393; @10.1177/1948550617693063; @10.3758/s13428-021-01557-x] are taken seriously in primary research.
 The good news is that study designs (multi-lab replication studies) which can apply more sophisticated versions of MASEM are becoming increasingly common in psychology, allowing us to study how to best deal with measurement errors when estimating variation of true effect sizes in meta-analyses
+
+# References

### 2023-06-17 manuscript/manuscript.qmd
pdf figures

@@
 
-knitr::include_graphics("../figures/manuscript/flowchart-simulation-design.png")
+knitr::include_graphics("../figures/manuscript/flowchart-simulation-design.pdf")
 ```
@@
 ```{r Figure 2, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/manuscript/z-r-hs-plot.png")
+knitr::include_graphics("../figures/manuscript/z-r-hs-plot.pdf")
 ```
@@
 ```{r Figure 3, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/manuscript/r_tau_0.02-0.08.png")
+knitr::include_graphics("../figures/manuscript/r_tau_0.02-0.08.pdf")
 ```
@@
 ```{r Figure 4, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/manuscript/sample_size.png")
+knitr::include_graphics("../figures/manuscript/sample_size.pdf")
 ```

### 2023-06-17 code/05_plots.r

@@ -115,6 +115,7 @@ facet_grid(effect_type~nominal_tau, scales = "free", switch = "y") +
 theme_bw()
 
 #ggsave("figures/manuscript/z-r-hs-plot.png", width = 8.62, height = 9.93)
+#ggsave("figures/manuscript/z-r-hs-plot.pdf", width = 8.62, height = 9.93)
 
 
 # Figure 2 (r, N = 150, k variable, tau low)
@@ -158,6 +159,7 @@ theme_bw()
 
 
 #ggsave("figures/manuscript/r_tau_0.02-0.08.png", width = 8.62, height = 9.93)
+#ggsave("figures/manuscript/r_tau_0.02-0.08.pdf", width = 8.62, height = 9.93)
 
 # Figure 3 (variable sample size)
 #****************************************
@@ -195,6 +197,7 @@ facet_grid(N~nominal_tau, scales = "free") +
 theme_bw()
 
 #ggsave("figures/manuscript/sample_size.png", width = 8.62, height = 9.93)
+#ggsave("figures/manuscript/sample_size.pdf", width = 8.62, height = 9.93)
 
 #****************************************
 # supplement C plots (supplemental main results)
@@ -242,6 +245,7 @@ facet_grid(reliability_sd~nominal_tau, scales = "free") +
 theme_bw()
 
 #ggsave("figures/supplement/supplement_C_fig1.png", width = 8.62, height = 9.93)
+#ggsave("figures/supplement/supplement_C_fig1.pdf", width = 8.62, height = 9.93)
 
 
 # supplement C figure 2 (variable k)
@@ -283,6 +287,7 @@ facet_grid(k~nominal_tau, scales = "free") +
 theme_bw()
 
 #ggsave("figures/supplement/supplement_C_fig2.png", width = 8.62, height = 9.93)
+#ggsave("figures/supplement/supplement_C_fig2.pdf", width = 8.62, height = 9.93)
 
 
 #****************************************
@@ -343,6 +348,7 @@ facet_grid(effect_type~nominal_tau, scales = "free", switch = "y") +
 theme_bw()
 
 #ggsave("figures/supplement/supplement_A_fig1.png", width = 8.62, height = 9.93)
+ggsave("figures/supplement/supplement_A_fig1.pdf", width = 8.62, height = 9.93)
 
 
 # Figure A2
@@ -461,8 +467,10 @@ trunc <- data.frame(mu = es,
 ggplot(e) +
[…]

### 2023-06-17 manuscript/manuscript.qmd
updated flowchart to include hs

@@
 #| fig-pos: 't'
-#| fig-cap: "Figure 1.
-Flowchart summarizing the simulation design.
-The flowchart shows the procedures followed to meta-anlyze results using the Hedge’s and Vevea model (see ‘Meta-Analytic Model’ header in methods section).
-The Hunter & Schmidt ‘bare-bones’ meta-analytic model followed the Pearson correlations procedure, but used sample size weights and replaced $r_i$ in the sampling variance by $\\bar{r}$ for meta-anlysis."
+#| fig-cap: "Figure 1. Flowchart summarizing the simulation design."
 

### 2023-06-18 manuscript/manuscript.qmd
updated abstract

@@
 abstract: |
-  One of the main purposes of meta-analysis is to estimate heterogeneity of effect sizes.
-The quality of meta-analytic estimates depends in large part on the quality of evidence in primary studies, yet there exist broad concerns that proper measurement tends to be neglected in much of psychological primary research.
-Such inattention to measurement can also bias heterogeneity estimates.
-We performed a Monte-Carlo simulation study to examine the expected bias in heterogeneity estimates due to unreliability in primary studies.
-Although unreliability in primary studies can both inflate and suppress heterogeneity, our results indicate that under most circumstances unreliability in primary studies leads to a severe underestimate of heterogeneity in meta-analyses.
-For a typical effect size heterogeneity can be expected to be underestimated by 15-60%, or more if average reliability in primary studies is below 0.6.
-We observed a positive bias in heterogeneity estimates due to unreliability only under specific and arguably uncommon circumstances of (1) actual zero heterogeneity, particularly when mean effect sizes are large, or (2) combinations of very small true heterogeneity, large variance in primary study reliabilities, large mean effect sizes, and a limited number of primary studies.
-The severity of the underestimation of heterogeneity could lead to meta-analyses in psychology missing relevant moderators and overgeneralizing outcomes.
-Unreliability appears to bring truly different effects closer together, thereby obscuring true differences between studies that could be relevant for theory, practice, and future research efforts.
-Research on concrete guidance to applied meta-analysts is urgently needed, as sophisticated methods for correcting measurement unreliability such as meta-analytic structural equation modeling (MASEM) are only applicable in exceptional cases and corrections based on classical test theory come with caveats and strong assumptions.
+  Measurement error (imperfect reliability) is present in any empirical effect size estimate and systematically attenuates observed effect sizes compared to true underlying effect sizes.
+Yet there exist broad concerns that proper measurement tends to be neglected in much of psychological research.
+We examined how measurement error in primary studies affects meta-analytic heterogeneity estimates using Monte-Carlo simulations.
+Our results indicate that although measurement error in primary studies can both inflate and suppress heterogeneity, under most circumstances measurement error in primary studies leads to a severe underestimate of heterogeneity in meta-analysis.
+Our simulations showed expected heterogeneity to be underestimated by about 15% - 60% when considering a typical effect size around r = 0.2 and true heterogeneity levels that are common in the meta-analytic literature (τ>0.1, in Pearson’s r).
+The underestimate primarily depends on average reliability in primary studies (higher reliability leads to a smaller underestimate), but also worsens with smaller primary study sample sizes.
+We observed a positive bias in heterogeneity estimates due to measurement error only under specific and arguably uncommon circumstances of (1) actual zero heterogeneity, particularly when mean effect sizes are large, or (2) combinations of very small true heterogeneity, large variance in primary study reliabilities, large mean effect sizes, and a limited number of primary studies.
+Severe underestimates of heterogeneity due to measurement error may affect many meta-analyses in psychology and obscure true differences between studies that could be relevant for theory, practice, and future research efforts.
+Research on concrete guidance to applied meta-analysts is needed, as sophisticated methods for correcting measurement unreliability such as meta-analytic structural equation modeling (MASEM) are only applicable in exceptional cases and corrections based on classical test theory come with caveats and strong assumptions.
 license: CC BY

### 2023-06-18 manuscript/manuscript.qmd
fix flowchart + descriptive text

@@
 
-## Meta-analytic modelmet0000498
+## Meta-analytic model
 
@@
 
-```{r Figure 1, out.width= "85%"}
+```{r Figure 1, out.width= "90%"}
 #| fig.env: figure*
 #| fig-pos: 't'
-#| fig-cap: "Figure 1. Flowchart summarizing the simulation design."
+#| fig-cap: "Flowchart summarizing the simulation design.
+For product-moment correlations, option (1) corresponds to the Hedges and Vevea (1998) model and option (2) to the Schmidt and Hunter (2015) model (see methods section 'Meta-analytic model').
+$\\theta$ = average true effect size (i.e., without sampling or measurement error), $\\tau^2$ = variance of true effect sizes, $\\theta_i$ = study true effect size, $\\gamma_i$ = study effect size with sampling error $\\sigma_i^2$, $N_i$ = study sample size, $R_i$ = study reliability, $\\bar{R}$ = average reliability across studies, $\\delta$ = standard deviation of reliabilities across studies, $r_i$ = observed study effect size with sampling and measurement error, $\\bar{r}$ = average observed effect size across studies."
 

### 2023-06-18 manuscript/manuscript.qmd
osf links to figures/tables

--- 2023-06-18 manuscript/manuscript.qmd
@@
 
-```{r Figure 1, out.width= "85%"}
+```{r Figure 1, out.width= "90%"}
 #| fig.env: figure*
 #| fig-pos: 't'
-#| fig-cap: "Figure 1. Flowchart summarizing the simulation design."
+#| fig-cap: "Flowchart summarizing the simulation design.
+For product-moment correlations, option (1) corresponds to the Hedges and Vevea (1998) model and option (2) to the Schmidt and Hunter (2015) model (see methods section 'Meta-analytic model').
+$\\theta$ = average true effect size (i.e., without sampling or measurement error), $\\tau^2$ = variance of true effect sizes, $\\theta_i$ = study true effect size, $\\gamma_i$ = study effect size with sampling error $\\sigma_i^2$, $N_i$ = study sample size, $R_i$ = study reliability, $\\bar{R}$ = average reliability across studies, $\\delta$ = standard deviation of reliabilities across studies, $r_i$ = observed study effect size with sampling and measurement error, $\\bar{r}$ = average observed effect size across studies.
+This figure was created using the website draw.io."
 
@@
 
-```{r Figure 2, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/manuscript/z-r-hs-plot.pdf")
-```
+```{r Figure 2, echo=FALSE, out.width = '90%'}
 
-_Figure 2.
-Unreliability in primary studies leads to a net negative bias when heterogeneity is positive.
+#| fig-cap: "Unreliability in primary studies leads to a net negative bias when heterogeneity is positive.
 The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications.
@@
 Results are for standard deviation in reliability = 0.15, within-study sample size = 150, and 20 studies within each meta-analysis.
-Code to reproduce figure: LINK._
+Code to reproduce figure: osf.io/8ygwj."
+
+knitr::include_graphics("../figures/manuscript/z-r-hs-plot.pdf")
+```
 
@@
 
-```{r Figure 3, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/manuscript/r_tau_0.02-0.08.pdf")
-```
-
-_Figure 3.
-Unreliability in primary studies typically leads to a net negative bias even for small degrees of heterogeneity.
+```{r Figure 3, echo=FALSE, out.width = '90%'}
+#| fig-cap: "Unreliability in primary studies typically leads to a net negative bias even for small degrees of heterogeneity.
 Effect size is Pearson’s r and rows indicate the number of studies in the meta-analysis.
@@
 Results are for standard deviation in reliability = 0.15, within-study sample size = 150.
-Code to reproduce figure: LINK._
+Code to reproduce figure: osf.io/8ygwj."
+
+knitr::include_graphics("../figures/manuscript/r_tau_0.02-0.08.pdf")
+```
 
@@
 
-```{r Figure 4, echo=FALSE, out.width = '100%'}
-knitr::include_graphics("../figures/manuscript/sample_size.pdf")
-```
-
-_Figure 4.
-A smaller sample size in primary studies tends to exacerbate the bias in observed heterogeneity estimates due to unreliability.
+```{r Figure 4, echo=FALSE, out.width = '90%'}
+#| fig-cap: "A smaller sample size in primary studies tends to exacerbate the bias in observed heterogeneity estimates due to unreliability.
 The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 10,000 replications.
@@
 Each dashed line corresponds to an average reliability in primary studies.
-Results are for standard deviation in reliability = 0.15 and 20 studies within each meta-analysis._
+Results are for standard deviation in reliability = 0.15 and 20 studies within each meta-analysis.
+Code to reproduce figure: osf.io/8ygwj."
+
+knitr::include_graphics("../figures/manuscript/sample_size.pdf")
+```
 

### 2023-06-18 supplement/supplement_A.qmd
osf links to figures/tables
rendered supplements

@@
 
-```{r Figure A1, echo=FALSE, out.width = '100%'}
+```{r Figure A1, echo=FALSE, out.width = '90%'}
 knitr::include_graphics("../figures/supplement/supplement_A_fig1.png")
@@
 
-_Figure A1._ Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1).
+_Figure A1.
+Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1).
 The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
@@
 Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Code to reproduce figure: osf.io/8ygwj._
 
@@
 
-```{r Figure A2, echo=FALSE, out.width = '100%'}
+```{r Figure A2, echo=FALSE, out.width = '90%'}
 knitr::include_graphics("../figures/supplement/supplement_A_fig2.png")
@@
 
-_Figure A2._ Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$.
+_Figure A2.
+Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$.
 The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 1e3 repetitions.
@@
 Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).
+Code to reproduce figure: osf.io/8ygwj._

### 2023-06-18 supplement/supplement_C.qmd
osf links to figures/tables
rendered supplements

@@
 
-```{r Figure C1, echo=FALSE, out.width = '100%'}
+```{r Figure C1, echo=FALSE, out.width = '90%'}
 knitr::include_graphics("../figures/supplement/supplement_C_fig1.png")
@@
 
-_Figure C1._ Inflation of effect sizes decreases with lower reliability SD leading to larger net negative bias in heterogeneity, especially for large effect sizes and moderate heterogeneity.
+_Figure C1.
+Inflation of effect sizes decreases with lower reliability SD leading to larger net negative bias in heterogeneity, especially for large effect sizes and moderate heterogeneity.
 The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
@@
 Each dashed line correspond to an average reliability in primary studies.
+Code to reproduce figure: osf.io/8ygwj._
 
@@
 
-```{r Figure C2, echo=FALSE, out.width = '100%'}
+```{r Figure C2, echo=FALSE, out.width = '90%'}
 knitr::include_graphics("../figures/supplement/supplement_C_fig2.png")
@@
 
-_Figure C2._ Expected inflation of effect sizes decreases with larger number of meta-analyzed when heterogeneity is absent or small and average effect size is small.
+_Figure C2.
+Expected inflation of effect sizes decreases with larger number of meta-analyzed when heterogeneity is absent or small and average effect size is small.
 The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations.
@@
 Each dashed line correspond to an average reliability in primary studies.
+Code to reproduce figure: osf.io/8ygwj._

### 2023-06-18 code/05_plots.r
code that improved figure A2..

@@ -472,5 +472,5 @@ ylab(expression("Between-studies standard deviation "~tau)) +
 xlab(expression("Average effect size "~mu)) +
 theme_bw()
 
-#ggsave("figures/supplement/supplement_A_fig2.png", width = 8.62, height = 9.93)
-#ggsave("figures/supplement/supplement_A_fig2.pdf", width = 8.62, height = 9.93)
+#ggsave("figures/supplement/supplement_A_fig2.png", width = 8.62, height = 3.31)
+#ggsave("figures/supplement/supplement_A_fig2.pdf", width = 8.62, height = 3.31)

### 2023-06-18 manuscript/manuscript.qmd
link to OSF and zenodo

@@
 
+\flushbottom
+\thispagestyle{empty}
+
+\vspace*{\fill}
+
+\fbox{\parbox{\textwidth}{%
+All code and data for this project are available on the Open Science Framework (OSF) at https://osf.io/tg68 and permanently archived at Zenodo (doi.org/10.5281/zenodo.8052030).
+}}
+
+\clearpage
+
 Estimating heterogeneity of effect sizes is usually considered the main purpose of meta-analysis, alongside estimating an average effect size.
@@
 
+```{r positive bias table}
+
+#| tbl-pos: t
+
+if (doc_type == "docx") {
+  print("Table 1 approximately here. Cannot be printed nicely to Word, see tables.pdf")
+} else{
+  tables$positive_bias_table
+}
+
+```
+
 In classical test theory (CTT), reliability of measurement ($R_{xx}$) is defined as the proportion true score variance ($\sigma^2_T$) to observed score variance ($\sigma^2_X$) and is presumed to be constant for all participants in a sample $R_{xx} = \sigma^2_T / \sigma^2_X$.
@@
 This is most easily illustrated by assuming a common true effect size $\mu$ across studies (i.e., a fixed effect and no true heterogeneity) and infinite sample size within studies.
-
 Any observed effect size $\hat{\mu}$ can then be seen as the true effect size $\mu$ multiplied by a study’s reliability $R_i$ (assuming $R_i = R_{xx’} = R_{yy’}$).
@@
 Table 1 illustrates this effect for three studies with differing reliability and zero true heterogeneity.
-
-```{r positive bias table}
-
-#| tbl-pos: t
-
-if (doc_type == "docx") {
-  print("Table 1 approximately here. Cannot be printed nicely to Word, see tables.pdf")
-} else{
-  tables$positive_bias_table
-}
-
-```
 
@@
 
-All code and data for this project are available at (OSF LINK).
+All code and data for this project are available at the Open Science Framework (OSF) at https://osf.io/tg68 and permanently archived at Zenodo (doi.org/10.5281/zenodo.8052030).
 For our simulations and analyses we used R `r paste(R.Version()$major, R.Version()$minor, sep = ".")` [@rcoreteam2020] and we used the R package 'metafor' [@10.18637/jss.v036.i03] to perform meta-analyses and estimate heterogeneity.
@@
 ```{r Figure 2, echo=FALSE, out.width = '90%'}
-
+#| fig.env: figure*
+#| fig-pos: 't'
 #| fig-cap: "Unreliability in primary studies leads to a net negative bias when heterogeneity is positive.
@@
 ```{r Figure 3, echo=FALSE, out.width = '90%'}
+#| fig.env: figure*
+#| fig-pos: 't'
 #| fig-cap: "Unreliability in primary studies typically leads to a net negative bias even for small degrees of heterogeneity.
@@
 ```{r Figure 4, echo=FALSE, out.width = '90%'}
+#| fig.env: figure*
+#| fig-pos: 't'
 #| fig-cap: "A smaller sample size in primary studies tends to exacerbate the bias in observed heterogeneity estimates due to unreliability.

### 2023-06-18 code/06_tables.qmd
osf links to figures/tables
decreased font size Table 2

@@
 As true heterogeneity is absent, any SD(ES) values above zero represent bias.
-Code to reproduce table: LINK",
+Code to reproduce table: osf.io/f5eyc.",
   threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)
@@
                caption = "Imperfect reliability suppresses true heterogeneity") %>%
- kable_styling(position = "left") %>%
+ kable_styling(position = "left", font_size = 8) %>%
  add_header_above(c(" " = 1, "Observed Effect Sizes" = 3, " " = 1)) %>%
@@
 SD (ES) is the standard deviation of the observed effect sizes, equivalent to heterogeneity given infinite within-study sample sizes.
-Code to reproduce table: LINK",
+Code to reproduce table: osf.io/f5eyc.",
   threeparttable = TRUE, escape = FALSE, footnote_as_chunk = TRUE)

### 2023-06-18 code/05_plots.r
changed image size to less height

@@ -114,8 +114,8 @@ xlab(expression("Average effect size "~mu)) +
 facet_grid(effect_type~nominal_tau, scales = "free", switch = "y") +
 theme_bw()
 
-#ggsave("figures/manuscript/z-r-hs-plot.png", width = 8.62, height = 9.93)
-#ggsave("figures/manuscript/z-r-hs-plot.pdf", width = 8.62, height = 9.93)
+#ggsave("figures/manuscript/z-r-hs-plot.png", width = 6*1.4, height = 6)
+#ggsave("figures/manuscript/z-r-hs-plot.pdf", width = 6*1.4, height = 6)
 
 
 # Figure 2 (r, N = 150, k variable, tau low)
@@ -158,8 +158,8 @@ facet_grid(k~nominal_tau, scales = "free") +
 theme_bw()
 
 
-#ggsave("figures/manuscript/r_tau_0.02-0.08.png", width = 8.62, height = 9.93)
-#ggsave("figures/manuscript/r_tau_0.02-0.08.pdf", width = 8.62, height = 9.93)
+#ggsave("figures/manuscript/r_tau_0.02-0.08.png", width = 6*1.4, height = 6)
+#ggsave("figures/manuscript/r_tau_0.02-0.08.pdf", width = 6*1.4, height = 6)
 
 # Figure 3 (variable sample size)
 #****************************************
@@ -196,8 +196,8 @@ xlab(expression("Average effect size "~mu)) +
 facet_grid(N~nominal_tau, scales = "free") +
 theme_bw()
 
-#ggsave("figures/manuscript/sample_size.png", width = 8.62, height = 9.93)
-#ggsave("figures/manuscript/sample_size.pdf", width = 8.62, height = 9.93)
+#ggsave("figures/manuscript/sample_size.png", width = 6*1.4, height = 6)
+#ggsave("figures/manuscript/sample_size.pdf", width = 6*1.4, height = 6)
 
 #****************************************
 # supplement C plots (supplemental main results)
@@ -227,12 +227,6 @@ trunc_a <- data.frame(mu = es,
                     nominal_tau = rep(tau, each = length(es)),
                     true_tau = tau_trunc)
 
-trunc_a$true_tau <- trunc_a$nominal_tau
-trunc_a$tau_hat <- trunc_a$tau
-
-
-# NB! the object true_tau_line below comes from Figure 1 main manuscript section
-
 ggplot(dat_a, aes(x = mu, y = tau_hat)) +
[…]

### 2023-06-18 manuscript/manuscript.qmd
float placement

@@
 
+```{r positive bias table}
+
+#| tbl-pos: t
+
+if (doc_type == "docx") {
+  print("Table 1 approximately here. Cannot be printed nicely to Word, see tables.pdf")
+} else{
+  tables$positive_bias_table
+}
+
+```
+
 One reason that few meta-analysts correct for unreliability could be that correcting for unreliability has long been controversial [for arguments spanning the last century, see Table 2 of @10.1111/iops.12184].
@@
 ## Bias in heterogeneity estimates due to unreliability
-
-```{r positive bias table}
-
-#| tbl-pos: t
-
-if (doc_type == "docx") {
-  print("Table 1 approximately here. Cannot be printed nicely to Word, see tables.pdf")
-} else{
-  tables$positive_bias_table
-}
-
-```
 
@@
 
-All code and data for this project are available at the Open Science Framework (OSF) at https://osf.io/tg68 and permanently archived at Zenodo (doi.org/10.5281/zenodo.8052030).
+All code and data for this project are available at the Open Science Framework (OSF) at [osf.io/tg68](https://osf.io/tg68) and permanently archived at Zenodo ([doi.org/10.5281/zenodo.8052030](https://doi.org/10.5281/zenodo.8052030)).
 For our simulations and analyses we used R `r paste(R.Version()$major, R.Version()$minor, sep = ".")` [@rcoreteam2020] and we used the R package 'metafor' [@10.18637/jss.v036.i03] to perform meta-analyses and estimate heterogeneity.
@@
 
+```{r Figure 1, out.width= "100%"}
+#| fig.env: figure*
+#| fig-pos: 't'
+#| fig-cap: "Flowchart summarizing the simulation design.
+For product-moment correlations, option (1) corresponds to the Hedges and Vevea (1998) model and option (2) to the Schmidt and Hunter (2015) model (see methods section 'Meta-analytic model').
+$\\theta$ = average true effect size (i.e., without sampling or measurement error), $\\tau^2$ = variance of true effect sizes, $\\theta_i$ = study true effect size, $\\gamma_i$ = study effect size with sampling error $\\sigma_i^2$, $N_i$ = study sample size, $R_i$ = study reliability, $\\bar{R}$ = average reliability across studies, $\\delta$ = standard deviation of reliabilities across studies, $r_i$ = observed study effect size with sampling and measurement error, $\\bar{r}$ = average observed effect size across studies.
+This figure was created using the website draw.io."
+
+knitr::include_graphics("../figures/manuscript/flowchart-simulation-design.pdf")
+```
+
 Estimating heterogeneity in both the HV and bare-bones HS model essentially consists of comparing the observed variance in effect sizes with what would be expected from sampling error alone and then ascribing any excess observed variance to heterogeneity [@10.1002/jrsm.12, Box 1; @schmidt2015, p.
@@
 Our design was broadly the same for product-moment correlations and Fisher's $z$, with some minor differences we highlight in the text below detailing the procedure.
-
-```{r Figure 1, out.width= "90%"}
-#| fig.env: figure*
-#| fig-pos: 't'
-#| fig-cap: "Flowchart summarizing the simulation design.
-For product-moment correlations, option (1) corresponds to the Hedges and Vevea (1998) model and option (2) to the Schmidt and Hunter (2015) model (see methods section 'Meta-analytic model').
-$\\theta$ = average true effect size (i.e., without sampling or measurement error), $\\tau^2$ = variance of true effect sizes, $\\theta_i$ = study true effect size, $\\gamma_i$ = study effect size with sampling error $\\sigma_i^2$, $N_i$ = study sample size, $R_i$ = study reliability, $\\bar{R}$ = average reliability across studies, $\\delta$ = standard deviation of reliabilities across studies, $r_i$ = observed study effect size with sampling and measurement error, $\\bar{r}$ = average observed effect size across studies.
-This figure was created using the website draw.io."
-
-knitr::include_graphics("../figures/manuscript/flowchart-simulation-design.pdf")
-```
 
@@
 
-```{r Figure 2, echo=FALSE, out.width = '90%'}
+```{r Figure 2, echo=FALSE, out.width = '100%'}
 #| fig.env: figure*
@@
 
-For all true heterogeneity levels in Figure 2 above zero, we see an overall underestimate of heterogeneity due to unreliability in the primary studies.
-This negative bias can be relatively severe, is worse with smaller effect sizes and with lower average reliability, and worsens in an absolute sense as true heterogeneity increases.
-For example, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.1$, the bias is `r in_text$bias_increasing_tau$tau_0.1` for Pearson's $r$, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.15$ the bias is `r in_text$bias_increasing_tau$tau_0.15`, and for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.2$ the bias is `r in_text$bias_increasing_tau$tau_0.2`.
-For an effect size of $\mu = 0.2$, true heterogeneity between 0.1 - 0.2 is underestimated by between `r round(100*in_text$underestimate_mu02$underestimate_min$underestimate, 0)`% ($\tau = 0.2$, $\bar{R} = 0.9$) and `r round(100*in_text$underestimate_mu02$underestimate_max$underestimate, 0)`% ($\tau = 0.1$, $\bar{R} = 0.6$).
-In other words, the underestimation of true heterogeneity due to unreliability can be quite severe.
-
-To explore further under what conditions the forces of inflation and suppression of heterogeneity reach an equilibrium, we added an analysis for several conditions with $\tau$ < 0.1.
-We used the following values {0.02, 0.04, 0.06, 0.08}.
-Figure 3 presents the results of this analysis for Pearson's $r$.
-Because these lower heterogeneity levels can be expected to be affected by heterogeneity truncation to a larger extent (which results in inflation), we present results for four different numbers of studies in the meta-analysis (rows Figure 3, lower k is associated with larger inflation of estimates).
-The standard deviation in reliabilities is again 0.15.
-Figure 3 shows that even when only little heterogeneity is present, it will generally be underestimated.
-If μ < 0.3 then imperfect reliability leads to a negative bias in heterogeneity estimates for all $\tau$ < 0.06 in Pearson's $r$.
-Even for the lower levels of true heterogeneity, estimates of heterogeneity will generally be positively biased only when average effect size is large, or the number of meta-analyzed studies is small.
-
 ```{r Figure 3, echo=FALSE, out.width = '90%'}
@@
 
-Finally, smaller sample sizes in primary studies tend to exacerbate the bias in observed heterogeneity estimates due to unreliability.
-In Figure 4, rows are the (fixed) sample size within primary studies when k = 20 and standard deviation in reliabilities is 0.15.
-When true heterogeneity is zero (leftmost column Figure 4), a smaller sample size tends to lead to more positive bias in estimated heterogeneity.
-When true heterogeneity is non-zero, smaller sample size corresponds with larger negative bias in heterogeneity estimates.
-For higher levels of reliability, differences due to sample size are small, but differences can be more concerning when average reliability is low.
-For example, consider N = 50 in our simulation compared with N = 200 (top versus bottom row Figure 4) when true heterogeneity lies between 0.1 – 0.2 and the average effect size is μ = 0.2.
-If average reliability is $\bar{R} = 0.8$, the lower sample size then leads to at most $\hat{\tau}$ =  `r in_text$bias_n50_0.8` additional bias in heterogeneity estimates in our simulations.
-However, when $\bar{R} = 0.6$, the additional bias can be as much as $\hat{\tau}$ = `r in_text[["bias_n50_0.6"]]` for an average sample size of 50 compared to 200, making up approximately one third of the total bias (`r in_text$net_bias_n50`) in this condition (N = 50, $\tau$ = 0.15).
-Generally, for an effect size of μ = 0.2, true heterogeneity between 0.1 – 0.2 is underestimated by between 16% ($\tau$ = 0.2, $\bar{R}= 0.9$) and 73% ($\tau$ = 0.1, $\bar{R} = 0.6$) in our simulations when sample size is 50, as compared to the 16-59% we observed when sample size was 150.
+For all true heterogeneity levels in Figure 2 above zero, we see an overall underestimate of heterogeneity due to unreliability in the primary studies.
+This negative bias can be relatively severe, is worse with smaller effect sizes and with lower average reliability, and worsens in an absolute sense as true heterogeneity increases.
+For example, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.1$, the bias is `r in_text$bias_increasing_tau$tau_0.1` for Pearson's $r$, for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.15$ the bias is `r in_text$bias_increasing_tau$tau_0.15`, and for $\bar{R} = 0.8$, $\mu = 0.2$, $\tau = 0.2$ the bias is `r in_text$bias_increasing_tau$tau_0.2`.
+For an effect size of $\mu = 0.2$, true heterogeneity between 0.1 - 0.2 is underestimated by between `r round(100*in_text$underestimate_mu02$underestimate_min$underestimate, 0)`% ($\tau = 0.2$, $\bar{R} = 0.9$) and `r round(100*in_text$underestimate_mu02$underestimate_max$underestimate, 0)`% ($\tau = 0.1$, $\bar{R} = 0.6$).
+In other words, the underestimation of true heterogeneity due to unreliability can be quite severe.
 
-```{r Figure 4, echo=FALSE, out.width = '90%'}
+To explore further under what conditions the forces of inflation and suppression of heterogeneity reach an equilibrium, we added an analysis for several conditions with $\tau$ < 0.1.
+We used the following values {0.02, 0.04, 0.06, 0.08}.
+Figure 3 presents the results of this analysis for Pearson's $r$.
+Because these lower heterogeneity levels can be expected to be affected by heterogeneity truncation to a larger extent (which results in inflation), we present results for four different numbers of studies in the meta-analysis (rows Figure 3, lower k is associated with larger inflation of estimates).
+The standard deviation in reliabilities is again 0.15.
+Figure 3 shows that even when only little heterogeneity is present, it will generally be underestimated.
+If $\mu$ < 0.3 then imperfect reliability leads to a negative bias in heterogeneity estimates for all $\tau$ < 0.06 in Pearson's $r$.
+Even for the lower levels of true heterogeneity, estimates of heterogeneity will generally be positively biased only when average effect size is large, or the number of meta-analyzed studies is small.
+
+```{r Figure 4, echo=FALSE, out.width = '100%'}
 #| fig.env: figure*
@@
 ```
+
+Finally, smaller sample sizes in primary studies tend to exacerbate the bias in observed heterogeneity estimates due to unreliability.
+In Figure 4, rows are the (fixed) sample size within primary studies when k = 20 and standard deviation in reliabilities is 0.15.
+When true heterogeneity is zero (leftmost column Figure 4), a smaller sample size tends to lead to more positive bias in estimated heterogeneity.
+When true heterogeneity is non-zero, smaller sample size corresponds with larger negative bias in heterogeneity estimates.
+For higher levels of reliability, differences due to sample size are small, but differences can be more concerning when average reliability is low.
+For example, consider N = 50 in our simulation compared with N = 200 (top versus bottom row Figure 4) when true heterogeneity lies between 0.1 – 0.2 and the average effect size is μ = 0.2.
+If average reliability is $\bar{R} = 0.8$, the lower sample size then leads to at most $\hat{\tau}$ =  `r in_text$bias_n50_0.8` additional bias in heterogeneity estimates in our simulations.
+However, when $\bar{R} = 0.6$, the additional bias can be as much as $\hat{\tau}$ = `r in_text[["bias_n50_0.6"]]` for an average sample size of 50 compared to 200, making up approximately one third of the total bias (`r in_text$net_bias_n50`) in this condition (N = 50, $\tau$ = 0.15).
+Generally, for an effect size of μ = 0.2, true heterogeneity between 0.1 – 0.2 is underestimated by between 16% ($\tau$ = 0.2, $\bar{R}= 0.9$) and 73% ($\tau$ = 0.1, $\bar{R} = 0.6$) in our simulations when sample size is 50, as compared to the 16-59% we observed when sample size was 150.
 

### 2023-06-18 manuscript/unreliable-heterogeneity-manuscript.qmd
fixed figure 3 width

--- 2023-06-18 manuscript/manuscript.qmd
@@
 
-```{r Figure 3, echo=FALSE, out.width = '90%'}
+```{r Figure 3, echo=FALSE, out.width = '100%'}
 #| fig.env: figure*
